### 2021-03-17 bayes-sampler-prelim.Rmd
started pyro notebook

---
title: "Bayesian sampler working"
output: html_notebook
---

Analyses of the Bayesian sampler data

```{r}
library(tidyverse)
library(purrr)
library(rstan)
```

```{r}
## load all data for experiment 1
df_exp1 <- map_dfr(
  paste0("osfstorage-archive/Experiment 1/", list.files("osfstorage-archive/Experiment 1/")), 
  read_csv,
  col_types=cols()
  )

## load all data for experiment 2
df_exp2 <- map_dfr(
  paste0("osfstorage-archive/Experiment 2/", list.files("osfstorage-archive/Experiment 2/")), 
  read_csv,
  col_types=cols()
  )

# sanity check both data
length(unique(df_exp1$ID)) # should be 59 Ps
length(unique(df_exp2$ID)) # should be 84 Ps

```

# Plotting figure 6 (partial)

Quick sanity-check: let's recreate the top half of figure 6. Looks like things are all as expected.

```{r}
df_exp2 %>% 
  group_by(ID, querydetail, querytype) %>% 
  summarize(estimate=mean(estimate)) %>% 
  # filter(ID==1, block==2,)
  mutate(
    category = case_when(
      grepl("(windy|cloudy)", querydetail) ~ "{windy, cloudy}",
      grepl("(cold|rainy)", querydetail) ~ "{cold, rainy}",
      TRUE ~ "other"
  ) 
  ) %>% 
    filter(category!="other") %>% 
  mutate(querytype = ordered(querytype,
                             levels = c(
                               "A",
                                "B",
                                "notA",
                                "notB",
                                "AandB",
                                "notAandB",
                                "AandnotB",
                                "notAandnotB",
                                "AorB",
                                "notAorB",
                                "AornotB",
                                "notAornotB",
                                "AgB",
                                "notAgB",
                                "AgnotB",
                                "notAgnotB",
                               "BgA",
                                "notBgA",
                                "BgnotA",
                                "notBgnotA"
                             ))) %>% 
  group_by(querytype, category) %>% 
  mutate(estimate = estimate/100) %>% 
  summarize(
    M = mean(estimate),
    se = sd(estimate)/sqrt(n())
  ) %>% 
  ungroup() %>% 
  ggplot(aes(x=querytype, y=M, ymin=M-se*2, ymax=M+se*2)) +
  geom_bar(stat="identity", position="dodge", width = .5, fill="grey50") +
  geom_errorbar(width=.1) +
  facet_wrap(~category, ncol=1) +
  theme_bw(base_size=14) +
  theme(
    panel.grid=element_blank(),
    axis.text.x=element_text(angle=45, hjust=1)
    ) +
  ylim(0,1)
```
# The Bayesian sampler model in Stan

Bayesian sampler model is written in equation 15 as:

$$E[P_{BS}(A)] = P(A)\frac{N}{N + 2\beta} + \frac{\beta}{N+2\beta}$$
Which we should expect would be Beta distributed, given that it is itself a probability. This has the nice form of a linear regression equation, except that none of the terms in this model are actually observed! The big thing that's missing is $P(A)$. The authors' solution is to treat these as free parameters / unobserved latent variables, but to fix them with a constraint that e.g. in a 4-term conditional probability table the probabilities must sum to one, so there are only 3 degrees of freedom. (I think I can use a simplex variable in Stan to accomplish this.)

The actual observation is:

$$P_{BS}(A) = prop(A)\frac{N}{N + 2\beta} + \frac{\beta}{N+2\beta}$$
Where $prop(A)$ is the proportion of N samples that were A in the mental simulation.

To implement in a bayesian way we should go back to the original beta model:

$$P_{BS}(A) \sim Beta(\beta + S(A), \beta+F(A))$$
Which we can rewrite as:

$$P_{BS}(A) \sim Beta(\beta + \lambda(A)N, \beta+(1-\lambda(A))N)$$
Where $\lambda(A)$ is the proportion of successes in the mental simulation and is a function of the true underlying model probability, $p(A)$

$$\lambda(A) \sim Beta(p(A)*N, (1-p(A))*N)$$
To which we assign a uniform prior (or the multidimensional dirichlet equivalent):
$$p(A) \sim Beta(1,1)$$
They aren't looking to make an actual statistical model of the responses, so they essentially make use o the fact that the expectation of $\lambda(A)$ is $p(A)$, to ignore this bit when they are just trying to estimate the expectation.

This model might not sample well, as N and $\beta$ are not uniquely identifiable. But let's try anyhow.

```{r}
# first, code to create condition codes. This will be a pain :(

df_test <- df_exp2 %>% 
  group_by(ID, querytype, querydetail) %>% 
  summarize(estimate=mean(estimate)) %>%
  filter(ID==103) %>%
  # filter(ID==4, block==3) %>% 
  filter(grepl("(windy|cloudy)", querydetail)) %>% 
  # filter(querytype %in% c("AandB","AandnotB","notAandB","notAandnotB")) %>%
  ungroup() %>% 
  mutate(
    dummy = map(querytype, ~case_when(
      .x == "AandB" ~ list(1,0,0,0),
      .x == "AandnotB" ~ list(0,1,0,0),
      .x == "notAandB" ~ list(0,0,1,0),
      .x == "notAandnotB" ~ list(0,0,0,1),
      .x == "A" ~ list(1,1,0,0),
      .x == "B" ~ list(1,0,1,0),
      .x == "notA" ~ list(0,0,1,1),
      .x == "notB" ~ list(0,1,0,1),
      TRUE ~ list(0,0,0,0)
    )
  )
  ) %>% 
  filter(!grepl("g",querytype))

df_sim <- df_test %>% 
  mutate(
    AandBp = .3,
    AandnotBp = .1,
    notAandBp = .2,
    notAandnotBp = .4,
    subj_p = pmap_dbl(list(AandBp, AandnotBp, notAandBp, notAandnotBp, dummy), ~AandBp*dummy[[1]])
  )

standata <- list(y = df_test$estimate, x = df_test$dummy, N = nrow(df_test), k = 4)
```

```{stan output.var="model"}
// psuedostancode
data {
  int<lower=0> N;
  int<lower=0> K;
  vector[N] x[K];
  vector[N] y
  vector[K] alpha;
}
parameters {

  real beta;
  real<lower=0,upper=1> Nsamp;
  real p;
  simplex[K] theta[N];
}
model {
  beta ~ cauchy(0, 1); // prior on beta, constrain positive
  Nsamp ~ cauchy(0, 5); // prior on N, constrain positive
  for (n in 1:N) {
    theta[n] ~ dirichlet(alpha);
    p = theta[n]*x; // later add code to get condition combinations etc.
  }

  lambda ~ beta(p*Nsamp, (1-p)*Nsamp);
  y_hat ~ beta(beta + lambda*Nsamp, beta + (1-lambda)*Nsamp); // does not account for any other kinds of error
}

```

# other diagnostics and stuff

I'm not sure how to implement the PT+N model in this fashion. For simple events, the bayesian sampler essentially collapses into the PT+N model (eq. 10):

$$E[P_{PT+N}(A)] = (1-2d)P(A)+d$$
where $d = \frac{\beta}{N+2\beta} $.

The two models diverge only for conditional probabilities, $P(A|B)$. The Bayesian sampler treats these just like any other probability. The PT+N model posits a two-stage estimation algorithm with noise at both stages. It produces a somewhat nastier looking equation (13) that is a generalization of the basic equation (simplifies down to it when p(B)=1).

But accounting for individual-level noise will work differently.

## Simulating data

First I'll need to simulate some data. I'll start simulating for one participant, and I'll try to use the format of the final data as much as I can so things translate.

```{r}
## what does the sampler do to probabilities?

bayesian_sampler <- function(p,N,B){
  p*(N/(N+2*B)) + B/(N+2*B)
}

inv_bayesian_sampler <- function(est,N,B){
  (est - B/(N+2*B))/(N/(N+2*B))
}

x <- c(.77,.67,.5,.49)

norm_x <- x/sum(x)
bayesian_sampler(x, 10, 2) # median values
inv_bayesian_sampler(bayesian_sampler(x, 10, 2), 10, 2) # median values

sum(bayesian_sampler(c(.25,.25,.25,.25), 1, 100)) # extreme

```

There's really no way to get the kinds of raw responses participants are giving just via the bayesian sampler distortions. For instance, the sum of the BS estimates will never be over 2, but you can observe cases where that's true in the data. You also get cases where probabilities for one cell are over .5 and all sum to near 2 or above. An unbiased beta prior will always push estimates toward .5, so it can't explain these cases. Here's one example:

```{r}
# looking at the real data

df_exp2 %>% 
  group_by(ID, querytype, querydetail) %>% 
  summarize(estimate=mean(estimate)) %>%
  filter(ID==103) %>%
  # filter(ID==4, block==3) %>% 
  filter(grepl("(windy|cloudy)", querydetail)) %>% 
  filter(querytype %in% c("AandB","AandnotB","notAandB","notAandnotB")) %>%
  # filter(querytype %in% c("A","notA")) %>% 
  select(querytype, estimate)
```

Of course we could just normalize both according to their direct contrasts (e.g. A and ~A). If we do that the esimates from the joint distribution estimates are only modestly correlated with the marginal estimates (r ~ .3 to .5).

```{r message=FALSE}
df_exp2 %>% 
  group_by(ID, querytype, querydetail) %>% 
  summarize(estimate=mean(estimate)/100) %>%
  # filter(ID==1) %>%
  # filter(ID==4, block==3) %>% 
  filter(grepl("(windy|cloudy)", querydetail)) %>%
  select(-querydetail) %>% 
  spread(querytype,estimate) %>% 
  mutate(base1 = AandB + AandnotB + notAandB + notAandnotB) %>% 
  mutate(base2 = notA + A) %>%
  mutate(base3 = notB + B) %>%
  mutate_at(vars(-ID, -base1, base2, base3, -A, -B, -notA, -notB), ~./base1) %>% 
  mutate_at(vars(A, notA), ~./base2) %>% 
  mutate_at(vars(B, notB), ~./base3) %>% 
  mutate(
    est_A = AandB + AandnotB, 
    est_B = AandB + notAandB, 
    est_notA = notAandB + notAandnotB,
    est_notB = AandnotB + notAandnotB
  ) %>% 
  ungroup() %>% 
  select(A, B, notA, notB, est_A, est_B, est_notA, est_notB) %>% 
  cor() %>% 
  round(3)
```

So the participant-level estimates are pretty messy. The Bayesian sampler model is really only able to generate accurate predictions at the population level. But I want to measure at the participant level!

A lot of Ps are saying that the conjunction AandB is more probable than either A or B?  Over half it looks like. Say what you will about all the other things, there's really no story of probabilities that says this is an acceptable pattern of responses. A similar things is observed for A or B, where about half of the time the disjunction is rated less probable than one of its disjuncts.

On p7 the PT+N model gets this by augmenting the model to say that conjunction/disjunction probabilities are noisier, which pushes them toward .50. This could explain the conjunction fallacy, but only when the subjective probability of the is below .50, in which case it will be pushed toward .50. But we can see equally many cases where the observed estimate is above .50, which would suggest the subjective conjunction probability is even higher!

```{r}
df_exp2 %>% 
  group_by(ID, querytype, querydetail) %>% 
  summarize(estimate=mean(estimate)) %>%
  # filter(ID==103) %>%
  # filter(ID==4, block==3) %>% 
  filter(grepl("(windy|cloudy)", querydetail)) %>% 
  filter(querytype %in% c("A","B", "AandB","AandnotB","notAandB","notAandnotB", "AorB")) %>%
  # filter(querytype %in% c("A","notA")) %>% 
  select(querytype, estimate) %>% 
  spread(querytype, estimate) %>% 
  mutate(
    minAB = map2_dbl(A, B, ~min(.x,.y)),
    maxAB = map2_dbl(A, B, ~max(.x,.y))
    ) %>% 
  ggplot(aes(x=minAB, y = AandB)) + 
  geom_point() +
  geom_abline(slope=1,intercept=0) +
  geom_vline(xintercept=50, linetype="dashed") +
  theme_bw()
```

Ok so that seems bad but it's just more affirmation that this is not a model of individual probability ratings, but only population averages. The individual ratings are still perturbed by substantial noise that could make them do all kinds of things.

But it seems like the __amount of error__ at the individual level should be related to the sampling/noise properties. If my estimate is based only on a handful of samples, I'll give more extreme values that might deviate from the expectation more. So the sampling/noise process induces both a population-level bias and individual trial/block-level variance. Seems like that individual variance should be part of what motivates the model. But at the same time it doesn't really fly to just essentially downweight some participants b/c their responses are noisier. Or does it? Would also create trouble for the conjunction/disjunction trials b/c they would effectively contribute less to fit.

```{r}
# do that, can create a coding system to compute
# A = AandB + AandnotB
# AorB = AandB + AandnotB + notAandB + 0*notAandnotB
# AgB = AandB + AandnotB / (AandB + notAandB )

## like so ...
# x = (c1*p_j1 + c2*p_j2 + c3*p_j3 + c4*p_j4)/(g1*p_j1 + g2*p_j2 + g3*p_j3 + g4*p_j4)

# and can be turned into a matrix operation later

# x = (cvec*pvec)*(gvec*pvec)^-1

# then fit one real participant

# then expand to hierarchical version

# then fit it all!
```

```{r}
# Bayesian sampler model

# for a single person, avg across blocks:
# prior on N
# prior on beta
# prior on N' and beta' (for conjunction and disjunction)
# uninformative prior on unobserved true probabilities (simplex)

# parameters {
# simplex[K] theta[N];
# ...
# }
# model {
#   beta ~ halfnormal(0,1) // prior on beta, constrain positive
#   N ~ cauchy(0, 5) // prior on N, constrain positive
#   k ~ cauchy(0, 5) // prior on kappa constant, constrain to be positive
#   theta ~ dirichlet(something) // latent probabilities, possibly need a loop N times
#   x = // computed from theta[i] and trial data or matrix mult
#   d = beta/(N+2*beta)
#   mu = (1-2*d)*x + d
#   y ~ beta(mu*k, (1-mu)*k);
#   ...
# }

# for hierarchical model with many Ps, avg across blocks:
# group-level prior for N, beta, N', beta'
# no hierarchical structure for prior on unobserved true probabilities

# for hierarchical model with many Ps, individual trials:
# share observations across blocks, replace formula

# the actual psychological model is that the sampling can differ each time tho
# so that doesn't seem quite right?
```

This is making me wonder if you actually do need to do nested inference if you want to get meaningful individual-level predictions.

### 2021-03-24 bayes-sampler-prelim.Rmd
rearranging

@@
 $$P_{BS}(A) = prop(A)\frac{N}{N + 2\beta} + \frac{\beta}{N+2\beta}$$
-Where $prop(A)$ is the proportion of N samples that were A in the mental simulation.
+Where $prop(A)$ is the proportion of N samples that were A in the mental simulation for the query.
+The expectation of that proportion is the "true" subjective probability as encoded in their mental model.
 
@@
 
-$$P_{BS}(A) \sim Beta(\beta + \lambda(A)N, \beta+(1-\lambda(A))N)$$
-Where $\lambda(A)$ is the proportion of successes in the mental simulation and is a function of the true underlying model probability, $p(A)$
+$$P_{BS}(A) \sim Beta(\beta + \pi(A)N, \beta+(1-\pi(A))N)$$
+ok but in our regression
 
-$$\lambda(A) \sim Beta(p(A)*N, (1-p(A))*N)$$
+$$ y_A = E[P_{BS}(A)] $$
+
+Where $\pi(A)$ is the proportion of successes in a single run of the mental simulation and is a function of the true underlying model probability, $p(A)$
+
+$$\pi(A) \sim Beta(p(A)*N, (1-p(A))*N)$$
+
 To which we assign a uniform prior (or the multidimensional dirichlet equivalent):
 $$p(A) \sim Beta(1,1)$$
-They aren't looking to make an actual statistical model of the responses, so they essentially make use o the fact that the expectation of $\lambda(A)$ is $p(A)$, to ignore this bit when they are just trying to estimate the expectation.
+They aren't looking to make an actual statistical model of the responses, so they essentially make use of the fact that the expectation of $\pi(A)$ is $p(A)$, to ignore this bit when they are just trying to estimate the expectation.
+Later in Appendix E they claim the models are "discrete", which I'm not sure I understand.
+I guess they are saying that the model is really very literal about people reporting the mean of their distributions.
+The Bayesian sampler theory also seems to assume that the simulation is run only once and those same results are then used to by the participant to answer a series of questions.
+If there's only one set of samples drawn then there won't be any way to estimate $p(A)$ except as $E[p(A)] = E[\pi(A)]=\pi(A)$.
+But that doesn't mean there isn't that extra bit of noise there statistically speaking--actually I am very on the fence about this.
+Not sure who is right.
 
@@
     ) %>% 
-  ggplot(aes(x=minAB, y = AandB)) + 
+  ggplot(aes(x=A, y = AorB)) + 
   geom_point() +
@@
 ```
+Looks like people do better for disjunction than conjunction.
 

### 2021-03-24 model_helpers.py
switched to numpyro

@@ -1,0 +1,158 @@
+import jax.numpy as jnp
+import numpy as np
+
+trial_funcs = dict({
+    "AandB": lambda theta: jnp.matmul(theta, jnp.array([1.,0.,0.,0.])) ,
+    "AandnotB": lambda theta: jnp.matmul(theta, jnp.array([0.,1.,0.,0.])),
+    "notAandB": lambda theta: jnp.matmul(theta, jnp.array([0.,0.,1.,0.])),
+    "notAandnotB": lambda theta: jnp.matmul(theta, jnp.array([0.,0.,0.,1.])),
+    "A":lambda theta: jnp.matmul(theta, jnp.array([1.,1.,0.,0.])),
+    "B":lambda theta: jnp.matmul(theta, jnp.array([1.,0.,1.,0.])),
+    "notA":lambda theta: jnp.matmul(theta, jnp.array([0.,0.,1.,1.])),
+    "notB":lambda theta: jnp.matmul(theta, jnp.array([0.,1.,0.,1.])),
+    "AorB":lambda theta: jnp.matmul(theta, jnp.array([1.,1.,1.,0.])),
+    "AornotB":lambda theta: jnp.matmul(theta, jnp.array([1.,1.,0.,1.])),
+    "notAorB":lambda theta: jnp.matmul(theta, jnp.array([0.,1.,1.,1.])),
+    "notAornotB":lambda theta: jnp.matmul(theta, jnp.array([0.,1.,0.,1.])),
+    
+    "AgB": lambda theta: jnp.divide( jnp.matmul(theta, jnp.array([1.,0.,0.,0.])), jnp.matmul(theta, jnp.array([1.,0.,1.,0.])) ),
+    "notAgB": lambda theta: jnp.divide( jnp.matmul(theta, jnp.array([0.,0.,1.,0.])), jnp.matmul(theta, jnp.array([1.,0.,1.,0.])) ),
+    "AgnotB": lambda theta: jnp.divide( jnp.matmul(theta, jnp.array([0.,1.,0.,0.])), jnp.matmul(theta, jnp.array([0.,1.,0.,1.])) ),
+    "notAgnotB": lambda theta: jnp.divide( jnp.matmul(theta, jnp.array([0.,0.,0.,1.])), jnp.matmul(theta, jnp.array([0.,1.,0.,1.])) ),
[…]

### 2021-03-24 model_helpers.py
3 main models running

@@ -1,6 +1,58 @@
 import jax.numpy as jnp
 import numpy as np
 
+
+def is_conjdisj(trial):
+    conjdisj = set(["AorB","notAorB","AornotB","notAornotB", "AandB", "notAandB", "AandnotB", "notAandnotB"])
+    x = 1 if trial in conjdisj else 0
+    return x
+
+def load_data():
+    import glob
+    df = pd.concat([pd.read_csv(f) for f in glob.glob('osfstorage-archive/Experiment 2/*.csv')], ignore_index = True)
+
+    fix_id_dict = dict()
+    original_ids = list(np.unique(df.ID))
+
+    for i in range(0, len(original_ids)):
+        fix_id_dict[original_ids[i]] = i
+
+    df = (df >> 
+         filter_by(~X.querydetail.str.contains("warm|snowy")) >>
+          mutate(estimate = X.estimate/100.) >>
+          group_by(X.ID, X.querytype, X.querydetail) >>
+          summarize(estimate = np.mean(X.estimate)) >>
+          mutate(coldrainy = X.querydetail.str.contains("cold|rainy").astype("int")) >>
+          mutate(estimate = X.estimate.replace({0:.01, 1:.99})) >>
+          mutate(conjdisj_trial = X.querytype.apply(is_conjdisj)) >>
+          mutate(ID = X.ID.apply(lambda x: fix_id_dict[x]))
+         )
+    
+    return df
+
+def make_model_data(data):
+    df_test = data
+    # df_test = df >> filter_by(X.conjdisj_trial==0)
+    # df_test = df >> filter_by(X.coldrainy==1)
+
+    trial_list = df_test.querytype
+    X_num = jnp.stack([num_vecs[trial] for trial in trial_list])
+    X_denom = jnp.stack([denom_vecs[trial] for trial in trial_list])
+
+    X_data = {
+        "trial": trial_list,
+        "X_num": X_num,
+        "X_denom": X_denom,
+        "subj": jnp.array(list(df_test.ID)),
+        "conjdisj_trial": jnp.array(list(df_test.conjdisj_trial)),
+        "cond": jnp.array(list(df_test.coldrainy)).astype("int32")
+    #     "block": jnp.array(list(df_test.block))
+    }
+
+    y_data = jnp.array(df_test.estimate.to_numpy())
+    
+    return X_data, y_data
+
 trial_funcs = dict({
     "AandB": lambda theta: jnp.matmul(theta, jnp.array([1.,0.,0.,0.])) ,
[…]

### 2021-03-24 model_helpers.py
fixed PT+N, implemented hierarchial models

@@ -2,20 +2,14 @@ import jax.numpy as jnp
 import numpy as np
 
 
-def is_conjdisj(trial):
-    conjdisj = set(["AorB","notAorB","AornotB","notAornotB", "AandB", "notAandB", "AandnotB", "notAandnotB"])
-    x = 1 if trial in conjdisj else 0
-    return x
+### ------ Data processing
 
-def load_data():
+def load_query_avg_data():
     import glob
     df = pd.concat([pd.read_csv(f) for f in glob.glob('osfstorage-archive/Experiment 2/*.csv')], ignore_index = True)
 
-    fix_id_dict = dict()
     original_ids = list(np.unique(df.ID))
-
-    for i in range(0, len(original_ids)):
-        fix_id_dict[original_ids[i]] = i
+    fix_id_dict = {original_ids[i]:i for i in range(0, len(original_ids))}
 
     df = (df >> 
          filter_by(~X.querydetail.str.contains("warm|snowy")) >>
@@ -24,34 +18,40 @@ def load_data():
           summarize(estimate = np.mean(X.estimate)) >>
           mutate(coldrainy = X.querydetail.str.contains("cold|rainy").astype("int")) >>
           mutate(estimate = X.estimate.replace({0:.01, 1:.99})) >>
-          mutate(conjdisj_trial = X.querytype.apply(is_conjdisj)) >>
-          mutate(ID = X.ID.apply(lambda x: fix_id_dict[x]))
+#           mutate(conjdisj_trial = X.querytype.apply(is_conjdisj)) >>
+          mutate(ID = X.ID.apply(lambda x: fix_id_dict[x])) >>
+          ungroup()
          )
     
     return df
 
-def make_model_data(data):
-    df_test = data
-    # df_test = df >> filter_by(X.conjdisj_trial==0)
-    # df_test = df >> filter_by(X.coldrainy==1)
 
-    trial_list = df_test.querytype
-    X_num = jnp.stack([num_vecs[trial] for trial in trial_list])
-    X_denom = jnp.stack([denom_vecs[trial] for trial in trial_list])
+def make_model_data(data):
 
     X_data = {
-        "trial": trial_list,
-        "X_num": X_num,
-        "X_denom": X_denom,
-        "subj": jnp.array(list(df_test.ID)),
[…]

### 2021-06-08 costello_watts_condprob_test.R
adding paper exports

@@ -53,12 +53,145 @@ zhao %>%
 ## at least as good as theirs.
 
 ## ---------
-##  just showing the math works for the inequalities
-d = .2
-## inequality 9
-calc_ptn(.5, d) + calc_ptn(.25, d, .0) - calc_ptn(.75, d, .0)
 
-## inequality 13
-calc_ptn(.3, d, 0) - calc_ptn(.6, d)*calc_ptn(.5,d)
 
-         
+p_AB <- .8
+p_AnotB <- .1
+p_notAB <- .05
+p_notAnotB <- .05
+
+p_A <- p_AB + p_AnotB
+p_B <- p_AB + p_notAB
+p_AgB <- p_AB/p_B
+p_BgA <- p_AB/p_A
+
+##  just showing the math works for the equalities
+## if events are independent/equiprobable, then equalities all work with base
+## PT+N model. And even when events are mildly dependent, they are still approx right
+## p116, their prediction is about expectation, so could show this by simulation
+
+d = .24
+
+## identity 3 (= 0)
+identity_03 <- function(probs, d){
+  p_AB <- probs[,1]
+  p_AnotB <- probs[,2]
+  p_notAB <- probs[,3]
+  p_notAnotB <- probs[,4]
+  
+  p_A <- p_AB + p_AnotB
+  p_B <- p_AB + p_notAB
+  p_AgB <- p_AB/p_B
+  p_BgA <- p_AB/p_A
+  
+  calc_ptn(p_AgB,d)*calc_ptn(p_B, d) - calc_ptn(p_BgA, d)*calc_ptn(p_A,d)
+}
+
+## identity 8 (= 0)
+identity_08 <- function(probs, d){
+  p_AB <- probs[,1]
+  p_AnotB <- probs[,2]
+  p_notAB <- probs[,3]
+  p_notAnotB <- probs[,4]
+  
+  p_A <- p_AB + p_AnotB
+  p_B <- p_AB + p_notAB
+  p_AgB <- p_AB/p_B
+  p_BgA <- p_AB/p_A
+  p_BgnotA <- p_notAB/(1-p_A)
+  p_AgnotB <- p_AnotB/(1-p_B)
+  
[…]

### 2021-06-08 models.py

@@ -262,7 +262,7 @@ def bayesian_sampler_simple_mlm_d(data, y=None):
     conjdisj, not_conjdisj = jnp.array([is_conjdisj(i) for i in trial]), abs(1-jnp.array([is_conjdisj(i) for i in trial]))
 
     # population level parameters/priors
-    k = numpyro.sample("k", dist.HalfNormal(50)) # noise parameter
+    k = numpyro.sample("k", dist.HalfCauchy(20)) # noise parameter
 
     d_base_pop = numpyro.sample("d_base_pop", dist.Normal(-1.0, 1.0))
  #   d_delta_pop = numpyro.sample("d_delta_pop", dist.Normal(0, .5)) # bias toward lower values for non conj/disj trials
@@ -307,7 +307,7 @@ def bayesian_sampler_complex_mlm_d(data, y=None):
     conjdisj, not_conjdisj = jnp.array([is_conjdisj(i) for i in trial]), abs(1-jnp.array([is_conjdisj(i) for i in trial]))
 
     # population level parameters/priors
-    k = numpyro.sample("k", dist.HalfNormal(50)) # noise parameter
+    k = numpyro.sample("k", dist.HalfCauchy(20)) # noise parameter
 
     d_base_pop = numpyro.sample("d_base_pop", dist.Normal(-1.0, 1.0))
     d_delta_pop = numpyro.sample("d_delta_pop", dist.Normal(0, .5)) # bias toward lower values for non conj/disj trials
@@ -359,7 +359,7 @@ def bayesian_sampler_complex_mlm(data, y=None):
     conjdisj, not_conjdisj = jnp.array([is_conjdisj(i) for i in trial]), abs(1-jnp.array([is_conjdisj(i) for i in trial]))
 
     # population level parameters/priors
-    k = numpyro.sample("k", dist.HalfNormal(50)) # noise parameter
+    k = numpyro.sample("k", dist.HalfCauchy(20)) # noise parameter
     beta_pop = numpyro.sample("beta_pop", dist.Normal(0, 1)) # roughly uniform after summing with random effects + sigmoid()
     beta_sd = numpyro.sample("beta_sd", dist.HalfCauchy(1))
 
@@ -416,7 +416,7 @@ def PTN_simple_mlm(data, y=None):
     not_conditional = abs(1-conditional)
     
     # population level parameters/priors
-    k = numpyro.sample("k", dist.HalfNormal(50)) # noise parameter
[…]

### 2021-09-01 paper-rmd/paper-papaja.Rmd
added rmd files

---
title             : "Untitled probability judgment paper draft"
shorttitle        : "DRAFT"

author: 
  - name          : "Derek Powell"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    # address       : "Postal address"
    email         : "dmpowell@asu.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  # - name          : "Ernst-August Doelle"
  #   affiliation   : "1,2"
  #   role:
  #     - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Arizona State University, School of Social and Behavioral Sciences"
  # - id            : "2"
  #   institution   : "Konstanz Business School"

authornote: |
  Arizona State University, School of Social and Behavioral Sciences

abstract: |
  Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic. However, people’s explicit probability judgments often incoherently violate the laws of probability. Two recent proposals, the “Probability Theory plus Noise” [@costello.watts2014] and “Bayesian Sampler” [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic. The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity. Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the $LOO_{ic}$ information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020). My findings support the Bayesian Sampler’s account of conditional probability judgments, and I find the data are best explained by a “hybrid” model that can be seen as a modification of both theories.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("references.bib")
```

```{r load_packages, include=FALSE}
library(tidyverse)
library(kableExtra)
```

Bayesian theories and approaches offer a unified formal framework for cognitive science [@tenenbaum.etal2011] with remarkable successes across domains, including perception [e.g. @kersten.etal2004], memory [@anderson1991], language [@xu.tenenbaum2007], and reasoning [@lu.etal2012]. At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule. It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", “vaccines are safe,” or “this politician is trustworthy”.

Trouble starts as soon as we seek to measure beliefs. According to Bayesian theories of cognition and epistemology [@jaynes2003], the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities. So asking people to express beliefs in terms of probability seems only natural.

Unfortunately, people’s explicit probability judgments routinely violate the axioms of probability theory. For example, human probability judgments often exhibit the “conjunction fallacy”: people will often judge the conjunction of two events (e.g. “Tom Brady likes football and miniature horses”) as being more probable than one of the events in isolation (e.g. “Tom Brady likes miniature horses”), a plain and flagrant violation of probability theory [@tversky.kahneman1983]. Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or “unpacking” effects [@tversky.koehler1994], and a variety of other effects illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013]. Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.

Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic [@costello.watts2014; @zhu.etal2020]. Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by drawing a limited number of samples from probabilistic mental models [see also @chater.etal2020; @dasgupta.etal2017].

## Two probabilistic theories of probability judgment

Costello & Watts [-@costello.watts2014; -@costello.watts2016; -@costello.watts2018] proposed a theory of probability judgment they call the “Probability Theory plus Noise” theory (PT+N). In the PT+N model, mental “samples” are drawn from a probabilistic mental model of events and are then “read” with noise, so that some positive examples will be read as negative and some negative read as positive. This results in probability judgments that reflect probabilistic credences perturbed by noise. Under the simplest form of the PT+N model, the expected value of probability judgments is:

$$E[\hat{P}_{PT+N}(A)] = (1-2d)P(A) + d $$

where *d* represents the probability with which samples will be misread, or the amount of noise in the judgment process (where by assumption a maximum of 50% of samples can be misread, so that d is a number in the range [0, .50]). The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].

Meanwhile, Zhu, Sanborn, & Chater [-@zhu.etal2020] have proposed a Bayesian model of probability judgment they call the “Bayesian Sampler.” Under this model, probability judgment is itself seen as a process of Bayesian inference. To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event. Then, those “observed” samples are integrated with a prior over probabilities to produce a probability judgment. The effect of this prior is to regularize the judgments, pushing them away from extreme values. They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g. encourage taking large risks based on a small number of mental samples. However, this hedging comes at the cost of systematic incoherence and biases. Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts [@costello.watts2014; @costello.watts2016].

Under the simplest version of the Bayesian Sampler model, the expected value of a probability judgment is given by the following formula:

$$E[\hat{P}_{BS}(A)] = \frac{N}{N+2\beta}P(A) + \frac{\beta}{N+2\beta}$$

where *N* represents the number of mental samples drawn. The mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric beta distribution parameterized by $\beta$. Zhu and colleagues [-@zhu.etal2020] show that the *N* and $\beta$ parameters of their model can be related to the *d* parameter of the PT+N model via the following bridging formula:

$$d = \frac{\beta}{N+2\beta}$$

### Two accounts of conditional probability judgments

By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs [@chater.etal2020]. However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project. Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence. So, the conditioning of beliefs is at the heart of these theories.

According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments. By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models [@chater.etal2020].

In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian [@costello.watts2016]. According to the PT+N model, conditional probabilities P(A|B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B. The PT+N model predicts conditional probability estimates using the following equation:

$$ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} $$

The PT+N’s account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally. In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.

## Comparing the models

Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello & Watts’ (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments. Unfortunately, their results were somewhat equivocal. They fit both “simple” and “complex” versions of each model and computed a Bayesian Information Criteria score (BIC) for all models. The complex versions of these models introduce additional parameters d’ and N’ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments. These additional parameters are needed in order to explain conjunction fallacies [@costello.watts2017; @zhu.etal2020] . Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs [@zhu.etal2020, Supplementary materials].

```{r table1, echo=FALSE, message=F, warning=F}
t1_df <- read_csv("original-fits.csv")

t1_df %>% 
  select(-X1) %>% 
  mutate(model = gsub("BS", "Bayesian Sampler", model)) %>% 
  rename(Model = model) %>% 
  kbl(caption="Original model fitting results", digits=1, booktabs=T) %>% 
  kable_classic(full_width=F)
```

In both experiments, the simple version of the Bayesian Sampler scores best, but the complex version of the PT+N model comes in second (lower BIC scores are better). It is not obvious what conclusions should be drawn from these results. Of course, the simple Bayesian Sampler model appears to win by the numbers. But conjunction fallacies are an extremely robust empirical finding [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues (2020). So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory. For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together. This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.

### Accounting for Model Complexity

When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to “overfit” the data. Typically this correction comes in the form of a “complexity penalty,” penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014]. Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.

There are at least three challenges to accounting for model complexity in the comparison of the PT+N and Bayesian Sampler models. First, the models differ not only in the number of parameters but in the domain of those parameters. Zhu and colleagues (2020) restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1]. This restriction constrains the Bayesian Sampler model’s prior distribution to the class of “ignorance priors”. Via the bridging conditions (eq X), they show that this restricts the “noise level” for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$] whereas the PT+N model permits noise values between [0, 0.50].

Second, it is not immediately clear how the models’ structural differences impact their flexibility. To be sure, Zhu and colleagues’ (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities. But, what impact does the PT+N model’s treatment of conditional probabilities have on model complexity? Is this component of the model a sort of Ptolemaic epicycle, adding complexity to the model that should be penalized? Or, does it constitute a commitment to novel predictions that thereby constrain its flexibility? Determining model complexity a priori is not always straightforward when the models being compared differ structurally.

Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ (2020) use of “unpooled” models, where parameters are estimated independently for each individual participant. Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants. In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term. Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.

### The present work

Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison. A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as $WAIC$ and $LOO_{ic}$. These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data. Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017]. In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling. This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.

# Methods

## Data selection

Zhu, Sanborn, & Chater (2020) conducted two experiments to compare the PT+N and Bayesian sampler theories. These experiments asked participants to judge the probability of different events in various combinations. Following prior work by Costello and Watts [e.g. -@costello.watts2016; -@costello.watts2018], both experiments focused on the everyday events of different kinds of weather.

Experiment 1 asked about the events [icy, frosty] and [normal, typical] (e.g. “what is the probability that the weather in London is normal and not typical?”).  The authors’ goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated. Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment. This is especially clear, as the authors note, for disjunctive trials such as “normal or typical,” where “or typical” might not be read as a disjunction but rather an elaborative clause.  In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.

Experiment 2 focused on more moderately correlated events, [cold, rainy] and [windy, cloudy], that do not admit these misinterpretations. In addition, a third experimental condition asking about [warm, snowy] was also included in the experiment, but was dropped from the analyses reported in the paper. Exploring the raw responses from this condition reveals a substantial fraction of “zero” and “one” responses for certain trials. This may reflect a different response process than was intended. For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero—failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day. Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.

## Modeling

I implement both the Bayesian Sampler and PT+N models as Bayesian models in the probabilistic programming language Numpyro. All code and results are available as supplemental materials (https://github.com/derekpowell/bayesian-sampler).

### Bayesian implementations of the models

The PT+N model defines probability judgments as

$$P_{e}(A) = (1-2d)P(A) + d$$
$$P_e(A\land B) = (1-d^\prime)P(A \land B)+d^\prime$$
$$P_e(A\lor B) = (1-d^\prime)P(A \lor B)+d^\prime$$
$$P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}$$

In contrast, the Bayesian Sampler model defines probability judgments as

$$P_{e}(A) = \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta}$$
$$P_{e}(A \land B) = \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta}$$
$$P_{e}(A \lor B) = \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta}$$

Fixing d and d’ or N and N’ equal yields the “simple” variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.

Notice that for each model the probability judgments depend on underlying subjective probabilities, derived from a mental sampling process. These subjective probabilities are unobserved, and must be estimated as a latent variable. Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events ($A \land B$, $\neg A \land B$, $A \land \neg B$, $\neg A \land \neg B$).

Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate d, d’, N, N’, and $\beta$ parameters for each participant. Translated into a Bayesian Framework, we can write the PT+N model as:

$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
$$\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) $$
$$ d_j = \frac{expit(\delta_j)}{2}$$
$$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
$$ \delta_j \sim Normal(0,1)$$
$$ \Delta\delta_j \sim Normal(0,1)$$
$$ k \sim HalfCauchy(20)$$

The function $f_{PT+N}$ computes the expected probability estimate using the underlying subjective probability p, the noise parameters d and d’, and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details). Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters d and d’.

Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model. To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d’, rather than directly according to its $\beta$, N, and N’ parameters. I constrain d to [0, $1/3$] for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$. This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.

The Bayesian Sampler model is therefore:

$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
$$\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) $$
$$ d_j = \frac{expit(\delta_j)}{2}$$
$$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
$$ \delta_j \sim Normal(0,1)$$
$$ \Delta\delta_j \sim Normal(0,1)$$
$$ k \sim HalfCauchy(20)$$

Where the function $f_{BS}$ computes the expected probability estimate as prescribed by the Bayesian Sampler theory.

### Hierarchical implementations of the models

Both of these models can also be implemented as hierarchical models with partial pooling for the d and d’ parameters (implicitly, for N and N’ in the case of the Bayesian Sampler). This should prevent over-penalizing complex versions of the model, as well as potentially improving their overall predictive fit. The hierarchical implementation adds parameters for the population-level d and d’ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects. For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency [@papaspiliopoulos.etal2007].

$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
$$\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) $$
$$d_j = \frac{ expit(\delta_j )}{3}$$
$$ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}$$
$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
$$ \delta_{pop} \sim Normal(-1, 1)$$
$$ \Delta\delta_{pop} \sim Normal(0, .50)$$
$$ log(\sigma_{\delta}) \sim Normal(-1, 1)$$
$$ log(\sigma_{\Delta\delta}) \sim Normal(-1, 1)$$
$$\delta_j \sim Normal(\delta_{pop}, \sigma_{\delta})$$
$$\Delta\delta_j \sim Normal(\Delta\delta_{pop}, \sigma_{\Delta\delta})$$
$$k \sim HalfCauchy(20)$$

Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of $\beta > 1$. Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of “ignorance priors” [@zhu.etal2020]. However, it is also possible that people bring informative priors to the probability judgment task. Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted [see e.g., @fennell.baddeley2012]. If $\beta$ is unrestricted, allowed to fall in the domain [0, $\infty$] then the Bayesian Sampler model becomes more flexible, allowing for equivalent “noise” levels in the same [0, .50] range as the PT+N model. Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment. Thus, fitting this additional “hybrid” model allows for a complete comparison of the models along both of their differing dimensions.

# Results

I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $LOO_{ic}$ scores for each combination. Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler. For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$ [see @gelman.etal2014a]. Figure X below displays the estimated differences in $LOO_{ic}$ scores for each of the models as compared to the best-scoring model.

\begin{figure}[ht]
\centering
\includegraphics[width=6in]{plot_compare.png}
\caption{Model comparison results for data from Experiments 1 and 2. Error bars indicate standard errors of the estimates.}
\end{figure}

Data from Experiment 1 are only modestly informative with respect to discriminating among the models. However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the “hybrid” Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$] (again, lower scores for $LOO_{ic}$ are better).

Recall, this model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments. Comparing predictions for conditional probability judgments from the hybrid model and the best-fitting PT+N model, we see that the hybrid model better captures these judgments from Experiment 2 (trial-average-level $R^2$ = .85 vs .73; participant-level $R^2$ = .51 vs .45). These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.

Figure X shows the posterior distributions of the population-level d and d’ parameters inferred from the “hybrid” model. Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of “ignorance priors” in the Bayesian Sampler model. Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ d and d’ estimates also lie outside this range (13 of 59 for d, 26 of 59 for d’).

\begin{figure}[ht]
\centering
\includegraphics[width=4in]{plot_params.png}
\caption{Posterior density of population-level d and d' parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.}
\end{figure}

Among the remaining models, hierarchical implementations with partial pooling outperform the unpooled corresponding models—suggesting that unpool variants of the complex models were over-penalized in the original analysis by @zhu.etal2020. In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models. However, differences among the remaining three models are approximately within the error of the $LOO_{ic}$ estimates. As expected, the simple variants of the models have smaller penalty terms than the complex variants.

Compared to the hybrid model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d’. Evidently this structural component isn’t a complexifying epicycle, instead it is a prediction that constrains flexibility. However, it is one that ultimately leads to a worse-fitting model.

Perhaps surprisingly, the Bayesian Sampler with unrestricted $\beta$ actually receives a smaller penalty term than the more restricted version of the model. This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014]. Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is estimated from data [-@gelman.etal2014]. If the estimated value is some very large positive number, then the constraint won’t have been very informative. But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model’s penalty term will therefore be smaller. Here, it seems reasonable to conjecture that because the implied d and d’ estimated for this data under this model are both very near $1/3$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.

The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria. Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives [e.g.  @gronau.wagenmakers2019; @vehtari.etal2019]. Fortunately, conclusions from the comparisons here do not rest solely on differences between the models’ complexity.

```{r table2, echo=FALSE, message=F, warning=F}
t2_df <- read_csv("model-comparison-table.csv")

t2_df %>% 
  select(-X1) %>% 
  gather(measure, value, -Experiment, -model) %>% 
  pivot_wider(names_from = c(Experiment, measure), values_from=value,names_sort = T) %>% 
  rename(Model = model) %>%
  kbl(
    caption="Bayesian model comparison results", 
    digits=3, 
    booktabs=T,
    col.names = c("Model", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{trial}$", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{trial}$"),
    escape = F
    ) %>%
  add_header_above(c(" ", "Experiment 1"=4, "Experiment 2"=4)) %>% 
  kable_classic(full_width=F) %>% 
  kable_styling(latex_options="scale_down")
```

Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table X.

# Discussion

By a fair margin, the model best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] was a version of the Bayesian Sampler model without restriction on the range of its $\beta$ parameters. Alternatively, this “hybrid” model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments. Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task. In keeping with the larger theoretical framework of Bayesian cognitive science, this theory assumes that subjective probabilities underlie people’s probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process [@chater.etal2020; @zhu.etal2020].

Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect “ignorance priors,” symmetric Beta distributions with $\beta$ < 1. As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis. Nevertheless, the data suggest that people bring informative rather than “ignorance” priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $1/3$] range implied by the bridging conditions when beta is restricted < 1. One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments. If so, other experimental contexts might see parameter estimates consistent with an uninformative prior. In addition, the noise parameters also varied across individuals. Exploring the participant-level parameters reveals that some individuals’ implied d and d’ parameters were consistent with the class of ignorance priors (see supplemental materials). Further research should explore this heterogeneity within and across individuals. Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and N parameters are not uniquely identifiable from judgment data of the sort examined here.

One thing this model comparison has not decided, and likely *cannot* decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes. Given their tight connections via the bridging conditions [@zhu.etal2020], it may not be possible to draw decisive conclusions here. Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that “noise” might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].

Some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020]. Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying subjective probabilities to be inferred. Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision. Of course, Zhu and colleagues’ (2020) experiments were never designed for this purpose. Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g. @franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006]. One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes [e.g. @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].

\newpage

# References

### 2021-09-01 paper-rmd/paper-papaja.tex
added rmd files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Untitled probability judgment paper draft},
  pdfauthor={Derek Powell1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{DRAFT}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Untitled probability judgment paper draft}
\author{Derek Powell\textsuperscript{1}}
\date{}

\authornote{

Arizona State University, School of Social and Behavioral Sciences

The authors made the following contributions. Derek Powell: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Arizona State University, School of Social and Behavioral Sciences}

\abstract{
Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic. However, people's explicit probability judgments often incoherently violate the laws of probability. Two recent proposals, the ``Probability Theory plus Noise'' (Costello \& Watts, 2014) and ``Bayesian Sampler'' (Zhu, Sanborn, \& Chater, 2020) theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic. The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity. Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the \(LOO_{ic}\) information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020). My findings support the Bayesian Sampler's account of conditional probability judgments, and I find the data are best explained by a ``hybrid'' model that can be seen as a modification of both theories.
}

\begin{document}
\maketitle

Bayesian theories and approaches offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) with remarkable successes across domains, including perception (e.g. Kersten, Mamassian, \& Yuille, 2004), memory (Anderson, 1991), language (Xu \& Tenenbaum, 2007), and reasoning (Lu, Chen, \& Holyoak, 2012). At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule. It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow,'' ``vaccines are safe,'' or ``this politician is trustworthy.''

Trouble starts as soon as we seek to measure beliefs. According to Bayesian theories of cognition and epistemology (Jaynes, 2003), the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities. So asking people to express beliefs in terms of probability seems only natural.

Unfortunately, people's explicit probability judgments routinely violate the axioms of probability theory. For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory (Tversky \& Kahneman, 1983). Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects (Tversky \& Koehler, 1994), and a variety of other effects illustrating the incoherence of human probability judgments (for an accessible review, see Kahneman, 2013). Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.

Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020). Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by drawing a limited number of samples from probabilistic mental models (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).

\hypertarget{two-probabilistic-theories-of-probability-judgment}{%
\subsection{Two probabilistic theories of probability judgment}\label{two-probabilistic-theories-of-probability-judgment}}

Costello \& Watts (2014, 2016, 2018) proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N). In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative read as positive. This results in probability judgments that reflect probabilistic credences perturbed by noise. Under the simplest form of the PT+N model, the expected value of probability judgments is:

\[E[\hat{P}_{PT+N}(A)] = (1-2d)P(A) + d \]

where \emph{d} represents the probability with which samples will be misread, or the amount of noise in the judgment process (where by assumption a maximum of 50\% of samples can be misread, so that d is a number in the range {[}0, .50{]}). The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics (Costello \& Watts, 2014, 2016, 2017, 2018).

Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model of probability judgment they call the ``Bayesian Sampler.'' Under this model, probability judgment is itself seen as a process of Bayesian inference. To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event. Then, those ``observed'' samples are integrated with a prior over probabilities to produce a probability judgment. The effect of this prior is to regularize the judgments, pushing them away from extreme values. They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g.~encourage taking large risks based on a small number of mental samples. However, this hedging comes at the cost of systematic incoherence and biases. Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts (Costello \& Watts, 2014, 2016).

Under the simplest version of the Bayesian Sampler model, the expected value of a probability judgment is given by the following formula:

\[E[\hat{P}_{BS}(A)] = \frac{N}{N+2\beta}P(A) + \frac{\beta}{N+2\beta}\]

where \emph{N} represents the number of mental samples drawn. The mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric beta distribution parameterized by \(\beta\). Zhu and colleagues (2020) show that the \emph{N} and \(\beta\) parameters of their model can be related to the \emph{d} parameter of the PT+N model via the following bridging formula:

\[d = \frac{\beta}{N+2\beta}\]

\hypertarget{two-accounts-of-conditional-probability-judgments}{%
\subsubsection{Two accounts of conditional probability judgments}\label{two-accounts-of-conditional-probability-judgments}}

By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs (Chater et al., 2020). However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project. Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence. So, the conditioning of beliefs is at the heart of these theories.

According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments. By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models (Chater et al., 2020).

In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian (Costello \& Watts, 2016). According to the PT+N model, conditional probabilities P(A\textbar B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B. The PT+N model predicts conditional probability estimates using the following equation:

\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]

The PT+N's account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally. In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.

\hypertarget{comparing-the-models}{%
\subsection{Comparing the models}\label{comparing-the-models}}

Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments. Unfortunately, their results were somewhat equivocal. They fit both ``simple'' and ``complex'' versions of each model and computed a Bayesian Information Criteria score (BIC) for all models. The complex versions of these models introduce additional parameters d' and N' that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments. These additional parameters are needed in order to explain conjunction fallacies (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020) . Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs (Zhu, Sanborn, \& Chater, 2020, Supplementary materials).

\begin{table}

\caption{\label{tab:table1}Original model fitting results}
\centering
\begin{tabular}[t]{lrr}
\toprule
Model & Exp. 1 & Exp. 2\\
\midrule
Bayesian Sampler simple & 956.9 & -5371.9\\
Bayesian Sampler complex & 1174.4 & -5099.3\\
PT+N simple & 1257.5 & -4901.1\\
PT+N complex & 1039.7 & -5159.6\\
Bayesian Sampler avg. & 1065.6 & -5235.6\\
\addlinespace
PT+N avg. & 1148.6 & -5030.4\\
\bottomrule
\end{tabular}
\end{table}

In both experiments, the simple version of the Bayesian Sampler scores best, but the complex version of the PT+N model comes in second (lower BIC scores are better). It is not obvious what conclusions should be drawn from these results. Of course, the simple Bayesian Sampler model appears to win by the numbers. But conjunction fallacies are an extremely robust empirical finding (Mellers, Hertwig, \& Kahneman, 2001; Sides, Osherson, Bonini, \& Viale, 2002) and are clearly present in the data collected by Zhu and colleagues (2020). So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory. For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together. This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.

\hypertarget{accounting-for-model-complexity}{%
\subsubsection{Accounting for Model Complexity}\label{accounting-for-model-complexity}}

When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to ``overfit'' the data. Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data (Gelman, Hwang, \& Vehtari, 2014). Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models' complexity.

There are at least three challenges to accounting for model complexity in the comparison of the PT+N and Bayesian Sampler models. First, the models differ not only in the number of parameters but in the domain of those parameters. Zhu and colleagues (2020) restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}. This restriction constrains the Bayesian Sampler model's prior distribution to the class of ``ignorance priors.'' Via the bridging conditions (eq X), they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]} whereas the PT+N model permits noise values between {[}0, 0.50{]}.

Second, it is not immediately clear how the models' structural differences impact their flexibility. To be sure, Zhu and colleagues' (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities. But, what impact does the PT+N model's treatment of conditional probabilities have on model complexity? Is this component of the model a sort of Ptolemaic epicycle, adding complexity to the model that should be penalized? Or, does it constitute a commitment to novel predictions that thereby constrain its flexibility? Determining model complexity a priori is not always straightforward when the models being compared differ structurally.

Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues' (2020) use of ``unpooled'' models, where parameters are estimated independently for each individual participant. Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants. In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term. Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.

\hypertarget{the-present-work}{%
\subsubsection{The present work}\label{the-present-work}}

Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison. A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \(WAIC\) and \(LOO_{ic}\). These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data. Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017). In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling. This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-selection}{%
\subsection{Data selection}\label{data-selection}}

Zhu, Sanborn, \& Chater (2020) conducted two experiments to compare the PT+N and Bayesian sampler theories. These experiments asked participants to judge the probability of different events in various combinations. Following prior work by Costello and Watts (e.g. 2016, 2018), both experiments focused on the everyday events of different kinds of weather.

Experiment 1 asked about the events {[}icy, frosty{]} and {[}normal, typical{]} (e.g.~``what is the probability that the weather in London is normal and not typical?''). The authors' goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated. Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment. This is especially clear, as the authors note, for disjunctive trials such as ``normal or typical,'' where ``or typical'' might not be read as a disjunction but rather an elaborative clause. In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.

Experiment 2 focused on more moderately correlated events, {[}cold, rainy{]} and {[}windy, cloudy{]}, that do not admit these misinterpretations. In addition, a third experimental condition asking about {[}warm, snowy{]} was also included in the experiment, but was dropped from the analyses reported in the paper. Exploring the raw responses from this condition reveals a substantial fraction of ``zero'' and ``one'' responses for certain trials. This may reflect a different response process than was intended. For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero---failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day. Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.

\hypertarget{modeling}{%
\subsection{Modeling}\label{modeling}}

I implement both the Bayesian Sampler and PT+N models as Bayesian models in the probabilistic programming language Numpyro. All code and results are available as supplemental materials (\url{https://github.com/derekpowell/bayesian-sampler}).

\hypertarget{bayesian-implementations-of-the-models}{%
\subsubsection{Bayesian implementations of the models}\label{bayesian-implementations-of-the-models}}

The PT+N model defines probability judgments as

\[P_{e}(A) = (1-2d)P(A) + d\]
\[P_e(A\land B) = (1-d^\prime)P(A \land B)+d^\prime\]
\[P_e(A\lor B) = (1-d^\prime)P(A \lor B)+d^\prime\]
\[P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}\]

In contrast, the Bayesian Sampler model defines probability judgments as

\[P_{e}(A) = \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta}\]
\[P_{e}(A \land B) = \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta}\]
\[P_{e}(A \lor B) = \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta}\]

Fixing d and d' or N and N' equal yields the ``simple'' variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.

Notice that for each model the probability judgments depend on underlying subjective probabilities, derived from a mental sampling process. These subjective probabilities are unobserved, and must be estimated as a latent variable. Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (\(A \land B\), \(\neg A \land B\), \(A \land \neg B\), \(\neg A \land \neg B\)).

Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate d, d', N, N', and \(\beta\) parameters for each participant. Translated into a Bayesian Framework, we can write the PT+N model as:

\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
\[\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) \]
\[ d_j = \frac{expit(\delta_j)}{2}\]
\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
\[ \delta_j \sim Normal(0,1)\]
\[ \Delta\delta_j \sim Normal(0,1)\]
\[ k \sim HalfCauchy(20)\]

The function \(f_{PT+N}\) computes the expected probability estimate using the underlying subjective probability p, the noise parameters d and d', and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details). Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters d and d'.

Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model. To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its \(\beta\), N, and N' parameters. I constrain d to {[}0, \(1/3\){]} for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\). This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.

The Bayesian Sampler model is therefore:

\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
\[\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) \]
\[ d_j = \frac{expit(\delta_j)}{2}\]
\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
\[ \delta_j \sim Normal(0,1)\]
\[ \Delta\delta_j \sim Normal(0,1)\]
\[ k \sim HalfCauchy(20)\]

Where the function \(f_{BS}\) computes the expected probability estimate as prescribed by the Bayesian Sampler theory.

\hypertarget{hierarchical-implementations-of-the-models}{%
\subsubsection{Hierarchical implementations of the models}\label{hierarchical-implementations-of-the-models}}

Both of these models can also be implemented as hierarchical models with partial pooling for the d and d' parameters (implicitly, for N and N' in the case of the Bayesian Sampler). This should prevent over-penalizing complex versions of the model, as well as potentially improving their overall predictive fit. The hierarchical implementation adds parameters for the population-level d and d' as well as a parameter controlling the standard deviation of the distribution for the subject-level effects. For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (Papaspiliopoulos, Roberts, \& Sköld, 2007).

\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
\[\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) \]
\[d_j = \frac{ expit(\delta_j )}{3}\]
\[ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}\]
\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
\[ \delta_{pop} \sim Normal(-1, 1)\]
\[ \Delta\delta_{pop} \sim Normal(0, .50)\]
\[ log(\sigma_{\delta}) \sim Normal(-1, 1)\]
\[ log(\sigma_{\Delta\delta}) \sim Normal(-1, 1)\]
\[\delta_j \sim Normal(\delta_{pop}, \sigma_{\delta})\]
\[\Delta\delta_j \sim Normal(\Delta\delta_{pop}, \sigma_{\Delta\delta})\]
\[k \sim HalfCauchy(20)\]

Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of \(\beta > 1\). Restricting \(\beta\) to {[}0,1{]} restricts the prior distribution of the Bayesian sampler to the class of ``ignorance priors'' (Zhu, Sanborn, \& Chater, 2020). However, it is also possible that people bring informative priors to the probability judgment task. Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted (see e.g., Fennell \& Baddeley, 2012). If \(\beta\) is unrestricted, allowed to fall in the domain {[}0, \(\infty\){]} then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same {[}0, .50{]} range as the PT+N model. Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment. Thus, fitting this additional ``hybrid'' model allows for a complete comparison of the models along both of their differing dimensions.

\hypertarget{results}{%
\section{Results}\label{results}}

I fit each of the models specified above to data from Zhu et al's (2020) Experiment 1 and 2 and computed \(LOO_{ic}\) scores for each combination. Model posteriors were estimated using the Numpyro (Phan, Pradhan, \& Jankowiak, 2019) implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler. For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to \(\hat{R}\) (see Gelman et al., 2014). Figure X below displays the estimated differences in \(LOO_{ic}\) scores for each of the models as compared to the best-scoring model.

\begin{figure}[ht]
\centering
\includegraphics[width=6in]{plot_compare.png}
\caption{Model comparison results for data from Experiments 1 and 2. Error bars indicate standard errors of the estimates.}
\end{figure}

Data from Experiment 1 are only modestly informative with respect to discriminating among the models. However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the ``hybrid'' Bayesian Sampler model that removes the restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]} (again, lower scores for \(LOO_{ic}\) are better).

Recall, this model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments. Comparing predictions for conditional probability judgments from the hybrid model and the best-fitting PT+N model, we see that the hybrid model better captures these judgments from Experiment 2 (trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) = .51 vs .45). These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.

Figure X shows the posterior distributions of the population-level d and d' parameters inferred from the ``hybrid'' model. Using data from experiment 2, population-level estimates of d and d' are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model. Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' d and d' estimates also lie outside this range (13 of 59 for d, 26 of 59 for d').

\begin{figure}[ht]
\centering
\includegraphics[width=4in]{plot_params.png}
\caption{Posterior density of population-level d and d' parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.}
\end{figure}

Among the remaining models, hierarchical implementations with partial pooling outperform the unpooled corresponding models---suggesting that unpool variants of the complex models were over-penalized in the original analysis by Zhu, Sanborn, and Chater (2020). In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models. However, differences among the remaining three models are approximately within the error of the \(LOO_{ic}\) estimates. As expected, the simple variants of the models have smaller penalty terms than the complex variants.

Compared to the hybrid model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d'. Evidently this structural component isn't a complexifying epicycle, instead it is a prediction that constrains flexibility. However, it is one that ultimately leads to a worse-fitting model.

Perhaps surprisingly, the Bayesian Sampler with unrestricted \(\beta\) actually receives a smaller penalty term than the more restricted version of the model. This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data (see Gelman, Hwang, \& Vehtari, 2014). Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is estimated from data (2014). If the estimated value is some very large positive number, then the constraint won't have been very informative. But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model's penalty term will therefore be smaller. Here, it seems reasonable to conjecture that because the implied d and d' estimated for this data under this model are both very near \(1/3\), the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.

The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria. Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives (e.g. Gronau \& Wagenmakers, 2019; Vehtari, Simpson, Yao, \& Gelman, 2019). Fortunately, conclusions from the comparisons here do not rest solely on differences between the models' complexity.

\begin{table}

\caption{\label{tab:table2}Bayesian model comparison results}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
Model & $LOO_{ic}$ & Penalty & $r_{resp}$ & $r_{trial}$ & $LOO_{ic}$ & Penalty & $r_{resp}$ & $r_{trial}$\\
\midrule
Unrestricted BS MLM & -1178.1 & 149.2 & 0.725 & 0.821 & -3953.1 & 368.7 & 0.688 & 0.878\\
Bayesian Sampler complex MLM & -1167.7 & 145.6 & 0.723 & 0.816 & -3815.4 & 399.9 & 0.675 & 0.852\\
PT+N complex & -1163.5 & 133.6 & 0.722 & 0.821 & -3768.5 & 398.2 & 0.667 & 0.840\\
Bayesian Sampler complex & -1159.5 & 149.3 & 0.723 & 0.815 & -3733.2 & 453.6 & 0.679 & 0.848\\
PT+N complex MLM & -1157.0 & 147.1 & 0.722 & 0.820 & -3796.7 & 356.1 & 0.658 & 0.835\\
\addlinespace
Bayesian Sampler simple MLM & -1149.5 & 141.0 & 0.719 & 0.810 & -3794.7 & 381.0 & 0.667 & 0.839\\
Bayesian Sampler simple & -1138.1 & 146.8 & 0.718 & 0.806 & -3722.8 & 418.9 & 0.667 & 0.831\\
PT+N simple & -1051.8 & 132.6 & 0.695 & 0.793 & -3643.0 & 319.5 & 0.619 & 0.777\\
PT+N simple MLM & -1044.7 & 137.9 & 0.695 & 0.793 & -3643.8 & 305.3 & 0.617 & 0.772\\
Relative Freq. & -746.6 & 146.8 & 0.664 & 0.737 & -1287.7 & 424.6 & 0.516 & 0.639\\
\bottomrule
\end{tabular}}
\end{table}

Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table X.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

By a fair margin, the model best accounting for the experimental data from Zhu and colleagues (2020) was a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters. Alternatively, this ``hybrid'' model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments. Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task. In keeping with the larger theoretical framework of Bayesian cognitive science, this theory assumes that subjective probabilities underlie people's probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process (Chater et al., 2020; Zhu, Sanborn, \& Chater, 2020).

Findings from Zhu and colleagues' Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta\) \textless{} 1. As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis. Nevertheless, the data suggest that people bring informative rather than ``ignorance'' priors to the probability judgment task, indicated by estimated d and d' parameters outside the {[}1, \(1/3\){]} range implied by the bridging conditions when beta is restricted \textless{} 1. One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments. If so, other experimental contexts might see parameter estimates consistent with an uninformative prior. In addition, the noise parameters also varied across individuals. Exploring the participant-level parameters reveals that some individuals' implied d and d' parameters were consistent with the class of ignorance priors (see supplemental materials). Further research should explore this heterogeneity within and across individuals. Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the \(\beta\) and N parameters are not uniquely identifiable from judgment data of the sort examined here.

One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes. Given their tight connections via the bridging conditions (Zhu, Sanborn, \& Chater, 2020), it may not be possible to draw decisive conclusions here. Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that ``noise'' might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler (2020).

Some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs (Chater et al., 2020). Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying subjective probabilities to be inferred. Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision. Of course, Zhu and colleagues' (2020) experiments were never designed for this purpose. Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning (e.g. Franke et al., 2016; Griffiths \& Tenenbaum, 2006; Jern, Chang, \& Kemp, 2014). One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes (e.g. Cook \& Lewandowsky, 2016; Jern, Chang, \& Kemp, 2014; Powell, Weisman, \& Markman, 2018).

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-anderson1991}{}%
Anderson, J. R. (1991). The adaptive nature of human categorization. \emph{Psychological Review}, \emph{98}(3), 409--429. \url{https://doi.org/10.1037/0033-295X.98.3.409}

\leavevmode\hypertarget{ref-chater.etal2020}{}%
Chater, N., Zhu, J.-Q., Spicer, J., Sundh, J., León-Villagrá, P., \& Sanborn, A. (2020). Probabilistic {Biases Meet} the {Bayesian Brain}. \emph{Current Directions in Psychological Science}, \emph{29}(5), 506--512. \url{https://doi.org/10.1177/0963721420954801}

\leavevmode\hypertarget{ref-cook.lewandowsky2016}{}%
Cook, J., \& Lewandowsky, S. (2016). Rational {Irrationality}: {Modeling Climate Change Belief Polarization Using Bayesian Networks}. \emph{Topics in Cognitive Science}, \emph{8}(1), 160--179. \url{https://doi.org/10.1111/tops.12186}

\leavevmode\hypertarget{ref-costello.watts2014}{}%
Costello, F., \& Watts, P. (2014). Surprisingly rational: {Probability} theory plus noise explains biases in judgment. \emph{Psychological Review}, \emph{121}(3), 463--480. \url{https://doi.org/10.1037/a0037010}

\leavevmode\hypertarget{ref-costello.watts2016}{}%
Costello, F., \& Watts, P. (2016). People's conditional probability judgments follow probability theory (plus noise). \emph{Cognitive Psychology}, \emph{89}, 106--133. \url{https://doi.org/10.1016/j.cogpsych.2016.06.006}

\leavevmode\hypertarget{ref-costello.watts2017}{}%
Costello, F., \& Watts, P. (2017). Explaining {High Conjunction Fallacy Rates}: {The Probability Theory Plus Noise Account}. \emph{Journal of Behavioral Decision Making}, \emph{30}(2), 304--321. \url{https://doi.org/10.1002/bdm.1936}

\leavevmode\hypertarget{ref-costello.watts2018}{}%
Costello, F., \& Watts, P. (2018). Invariants in probabilistic reasoning. \emph{Cognitive Psychology}, \emph{100}, 1--16. \url{https://doi.org/10.1016/j.cogpsych.2017.11.003}

\leavevmode\hypertarget{ref-dasgupta.etal2017}{}%
Dasgupta, I., Schulz, E., \& Gershman, S. J. (2017). Where do hypotheses come from? \emph{Cognitive Psychology}, \emph{96}, 1--25. \url{https://doi.org/10.1016/j.cogpsych.2017.05.001}

\leavevmode\hypertarget{ref-fennell.baddeley2012}{}%
Fennell, J., \& Baddeley, R. (2012). Uncertainty plus prior equals rational bias: {An} intuitive {Bayesian} probability weighting function. \emph{Psychological Review}, \emph{119}(4), 878--887. \url{https://doi.org/10.1037/a0029346}

\leavevmode\hypertarget{ref-franke.etal2016}{}%
Franke, M., Dablander, F., Scholler, A., Bennett, E., Degen, J., Tessler, M. H., \ldots{} Goodman, N. D. (2016). What does the crowd believe? {A} hierarchical approach to estimating subjective beliefs from empirical data, 6.

\leavevmode\hypertarget{ref-gelman.etal2014a}{}%
Gelman, A., Carlin, J. B., Stern, H. S., Dunson, D. B., Vehtari, A., \& Rubin, D. B. (2014). \emph{Bayesian data analysis} (Third edition). {Boca Raton}: {CRC Press}.

\leavevmode\hypertarget{ref-gelman.etal2014}{}%
Gelman, A., Hwang, J., \& Vehtari, A. (2014). Understanding predictive information criteria for {Bayesian} models. \emph{Statistics and Computing}, \emph{24}(6), 997--1016. \url{https://doi.org/10.1007/s11222-013-9416-2}

\leavevmode\hypertarget{ref-griffiths.tenenbaum2006}{}%
Griffiths, T. L., \& Tenenbaum, J. B. (2006). Optimal {Predictions} in {Everyday Cognition}. \emph{Psychological Science}, \emph{17}(9), 767--773. \url{https://doi.org/10.1111/j.1467-9280.2006.01780.x}

\leavevmode\hypertarget{ref-gronau.wagenmakers2019}{}%
Gronau, Q. F., \& Wagenmakers, E.-J. (2019). Limitations of {Bayesian Leave}-{One}-{Out Cross}-{Validation} for {Model Selection}. \emph{Computational Brain \& Behavior}, \emph{2}(1), 1--11. \url{https://doi.org/10.1007/s42113-018-0011-7}

\leavevmode\hypertarget{ref-jaynes2003}{}%
Jaynes, E. T. (2003). \emph{Probability {Theory}: {The Logic} of {Science}}. (G. L. Bretthorst, Ed.). {Cambridge, UNITED KINGDOM}: {Cambridge University Press}.

\leavevmode\hypertarget{ref-jern.etal2014}{}%
Jern, A., Chang, K. K., \& Kemp, C. (2014). Belief polarization is not always irrational. \emph{Psychological Review}, \emph{121}(2), 206--224. \url{https://doi.org/10.1037/a0035941}

\leavevmode\hypertarget{ref-kahneman2013}{}%
Kahneman, D. (2013). \emph{Thinking, {Fast} and {Slow}} (1st edition). {New York}: {Farrar, Straus and Giroux}.

\leavevmode\hypertarget{ref-kersten.etal2004}{}%
Kersten, D., Mamassian, P., \& Yuille, A. (2004). Object {Perception} as {Bayesian Inference}. \emph{Annual Review of Psychology}, \emph{55}(1), 271--304. \url{https://doi.org/10.1146/annurev.psych.55.090902.142005}

\leavevmode\hypertarget{ref-lu.etal2012}{}%
Lu, H., Chen, D., \& Holyoak, K. J. (2012). Bayesian analogy with relational transformations. \emph{Psychological Review}, \emph{119}(3), 617--648. \url{https://doi.org/10.1037/a0028719}

\leavevmode\hypertarget{ref-mellers.etal2001}{}%
Mellers, B., Hertwig, R., \& Kahneman, D. (2001). Do {Frequency Representations Eliminate Conjunction Effects}? {An Exercise} in {Adversarial Collaboration}. \emph{Psychological Science}, \emph{12}(4), 269--275. \url{https://doi.org/10.1111/1467-9280.00350}

\leavevmode\hypertarget{ref-papaspiliopoulos.etal2007}{}%
Papaspiliopoulos, O., Roberts, G. O., \& Sköld, M. (2007). A {General Framework} for the {Parametrization} of {Hierarchical Models}. \emph{Statistical Science}, \emph{22}(1). \url{https://doi.org/10.1214/088342307000000014}

\leavevmode\hypertarget{ref-phan.etal2019}{}%
Phan, D., Pradhan, N., \& Jankowiak, M. (2019). Composable {Effects} for {Flexible} and {Accelerated Probabilistic Programming} in {NumPyro}. \emph{arXiv:1912.11554 {[}Cs, Stat{]}}. Retrieved from \url{http://arxiv.org/abs/1912.11554}

\leavevmode\hypertarget{ref-powell.etal2018}{}%
Powell, D., Weisman, K., \& Markman, E. M. (2018). Articulating lay theories through graphical models: {A} study of beliefs surrounding vaccination decisions, 6.

\leavevmode\hypertarget{ref-sides.etal2002}{}%
Sides, A., Osherson, D., Bonini, N., \& Viale, R. (2002). On the reality of the conjunction fallacy. \emph{Memory \& Cognition}, \emph{30}(2), 191--198. \url{https://doi.org/10.3758/BF03195280}

\leavevmode\hypertarget{ref-tenenbaum.etal2011}{}%
Tenenbaum, J. B., Kemp, C., Griffiths, T. L., \& Goodman, N. D. (2011). How to {Grow} a {Mind}: {Statistics}, {Structure}, and {Abstraction}. \emph{Science}, \emph{331}(6022), 1279--1285. \url{https://doi.org/10.1126/science.1192788}

\leavevmode\hypertarget{ref-tversky.kahneman1983}{}%
Tversky, A., \& Kahneman, D. (1983). Extensional {Versus Intuitive Reasoning}: {The Conjunction Fallacy} in {Probability Judgment}. \emph{Psychological Review}, \emph{90}(4), 23.

\leavevmode\hypertarget{ref-tversky.koehler1994}{}%
Tversky, A., \& Koehler, D. J. (1994). Support theory: {A} nonextensional representation of subjective probability. \emph{Psychological Review}, \emph{101}(4), 547--567. https://doi.org/\url{http://dx.doi.org.ezproxy1.lib.asu.edu/10.1037/0033-295X.101.4.547}

\leavevmode\hypertarget{ref-vehtari.etal2017}{}%
Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical {Bayesian} model evaluation using leave-one-out cross-validation and {WAIC}. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432. \url{https://doi.org/10.1007/s11222-016-9696-4}

\leavevmode\hypertarget{ref-vehtari.etal2019}{}%
Vehtari, A., Simpson, D. P., Yao, Y., \& Gelman, A. (2019). Limitations of {``{Limitations} of {Bayesian Leave}-one-out {Cross}-{Validation} for {Model Selection}.''} \emph{Computational Brain \& Behavior}, \emph{2}(1), 22--27. \url{https://doi.org/10.1007/s42113-018-0020-6}

\leavevmode\hypertarget{ref-xu.tenenbaum2007}{}%
Xu, F., \& Tenenbaum, J. B. (2007). Word learning as {Bayesian} inference. \emph{Psychological Review}, \emph{114}(2), 245--272. \url{https://doi.org/10.1037/0033-295X.114.2.245}

\leavevmode\hypertarget{ref-zhu.etal2020}{}%
Zhu, J.-Q., Sanborn, A. N., \& Chater, N. (2020). The {Bayesian} sampler: {Generic Bayesian} inference causes incoherence in human probability judgments. \emph{Psychological Review}, \emph{127}(5), 719--748. \url{https://doi.org/10.1037/rev0000190}

\end{CSLReferences}

\end{document}

### 2021-09-01 paper-rmd/paper.Rmd
added rmd files

--- 2021-09-01 paper-rmd/paper-papaja.Rmd
@@
 ---
-title             : "Untitled probability judgment paper draft"
-shorttitle        : "DRAFT"
-
-author: 
-  - name          : "Derek Powell"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    # address       : "Postal address"
-    email         : "dmpowell@asu.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Writing - Original Draft Preparation
-      - Writing - Review & Editing
-  # - name          : "Ernst-August Doelle"
-  #   affiliation   : "1,2"
-  #   role:
-  #     - Writing - Review & Editing
-
-affiliation:
-  - id            : "1"
-    institution   : "Arizona State University, School of Social and Behavioral Sciences"
-  # - id            : "2"
-  #   institution   : "Konstanz Business School"
-
-authornote: |
-  Arizona State University, School of Social and Behavioral Sciences
-
+title: A template for the *arxiv* style
+authors:
+  - name: Derek Powell
+    # thanks: Arizona State University
+    department: School of Social and Behavioral Sciences
+    affiliation: Arizona State University
+    location: Phoenix, AZ
+    email: dmpowell@asu.edu
 abstract: |
-  Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic.
-However, people’s explicit probability judgments often incoherently violate the laws of probability.
-Two recent proposals, the “Probability Theory plus Noise” [@costello.watts2014] and “Bayesian Sampler” [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
-The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity.
-Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the $LOO_{ic}$ information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020).
-My findings support the Bayesian Sampler’s account of conditional probability judgments, and I find the data are best explained by a “hybrid” model that can be seen as a modification of both theories.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "keywords"
-wordcount         : "X"
-
-bibliography      : ["references.bib"]
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
+  Enter the text of your abstract here.
+keywords:
+  - blah
+  - blee
+  - bloo
+  - these are optional and can be removed
+bibliography: references.bib
+biblio-style: unsrt
+output: 
+  rticles::arxiv_article:
+    keep_tex: true
 ---
-
-```{r setup, include = FALSE}
-library("papaja")
-r_refs("references.bib")
-```
 
@@
 
-Bayesian theories and approaches offer a unified formal framework for cognitive science [@tenenbaum.etal2011] with remarkable successes across domains, including perception [e.g.
-@kersten.etal2004], memory [@anderson1991], language [@xu.tenenbaum2007], and reasoning [@lu.etal2012].
-At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", “vaccines are safe,” or “this politician is trustworthy”.
+# Introduction
+
+Bayesian theories of cognition have had remarkable successes in explaining human reasoning and behavior across many domains (big cite).
+[The core of these theories is that people reason according to subjective mentally-represented degrees of belief, and they specify how they should be revised in light of evidence.
+It is somewhat embarrassing then that one area where these theories seem to fall down is in describing human "beliefs'' of the simple and everyday sort, such as beliefs like "it will rain tomorrow", “vaccines are safe,” or “this politician is trustworthy”.
 
 Trouble starts as soon as we seek to measure beliefs.
-According to Bayesian theories of cognition and epistemology [@jaynes2003], the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
+According to Bayesian theories of cognition and epistemology, the degree to which people believe in various propositions should reflect subjective mental probabilities.
 So asking people to express beliefs in terms of probability seems only natural.
@@
 “Tom Brady likes football and miniature horses”) as being more probable than one of the events in isolation (e.g.
-“Tom Brady likes miniature horses”), a plain and flagrant violation of probability theory [@tversky.kahneman1983].
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or “unpacking” effects [@tversky.koehler1994], and a variety of other effects illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013].
-Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
+“Tom Brady likes miniature horses”), a plain and flagrant violation of probability theory (cite some examples).
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies (e.g.
+XXX), “unpacking” effects (e.g.
+fox & tversky), and a variety of other effects illustrating the incoherence of human probability judgments [cite].
+Altogether these findings have led many researchers to abandon the notion that credences are represented as probabilities.
 
@@
 However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project.
-Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence.
-So, the conditioning of beliefs is at the heart of these theories.
-
 According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments.
-By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models [@chater.etal2020].
+Thus, the Bayesian sampler model does not assign any special status to conditional probability judgments and so fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other Bayesian mental models [@chater.etal2020].
 
@@
 $$ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} $$                                            
-
-The PT+N’s account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally.
-In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.
+Given that conditional inference and belief updating lies at the core of Bayesian cognitive theory, PT+N’s account of conditional probability judgment separates it from the Bayesian sampler theory not just quantitatively, but quite fundamentally.
+First, it denies any central importance of conditional probability judgments by pulling conditioning out of the mental model of an event and placing it into the probability judgment process.
+And perhaps more importantly, this non-bayesian account of conditional inference can no longer be considered “rational” at any stage under the PT+N model.
 
@@
 The complex versions of these models introduce additional parameters d’ and N’ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are needed in order to explain conjunction fallacies [@costello.watts2017; @zhu.etal2020] .
-Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs [@zhu.etal2020, Supplementary materials].
+These additional parameters are needed in order to explain conjunction fallacies (Costello & Watts, 2017; Zhu et al., 2020).
+Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs (Zhu et al, 2020, Supplementary materials).
 
@@
   kable_classic(full_width=F)
+
 ```
@@
 Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues (2020).
+But conjunction fallacies are an extremely robust empirical finding (e.g.
+cite) and are clearly present in the data collected by Zhu and colleagues (2020).
 So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
 For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.
+This approach somewhat favors the Bayesian Sampler theory overall, though they are rightfully cautious about drawing any strong conclusions.
 
@@
 
-When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to “overfit” the data.
-Typically this correction comes in the form of a “complexity penalty,” penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014].
 Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
-
-There are at least three challenges to accounting for model complexity in the comparison of the PT+N and Bayesian Sampler models.
+There are at least three challenges to accounting for model complexity in the comparison of these models.
 First, the models differ not only in the number of parameters but in the domain of those parameters.
 Zhu and colleagues (2020) restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1].
-This restriction constrains the Bayesian Sampler model’s prior distribution to the class of “ignorance priors”.
-Via the bridging conditions (eq X), they show that this restricts the “noise level” for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$] whereas the PT+N model permits noise values between [0, 0.50].
+This restriction constrains the Bayesian Sampler model’s prior distribution to the class of “ignorance priors”, symmetric beta distributions where $\beta < 1$.
+Via the bridging conditions (eq X), they show that this restricts the “noise level” for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $\frac{1}{3}$] whereas the PT+N model permits noise values between [0, 0.50].
 
-Second, it is not immediately clear how the models’ structural differences impact their flexibility.
+Second, it is not immediately clear how the structural differences between the models impact their flexibility.
+For the purposes of model comparison, the “complexity penalty” that models truly ought to pay is determined by their relative flexibility in accommodating different patterns of data, not simply the number or domain of their parameters (cite Gelman).
 To be sure, Zhu and colleagues’ (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
@@
 
-Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ (2020) use of “unpooled” models, where parameters are estimated independently for each individual participant.
-Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
+Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ (2020) use of unpooled models, where parameters are estimated independently for each individual participant.
+Adding a parameter for each participant may effectively over-penalize the “complex” variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
 In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison.
-A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as $WAIC$ and $LOO_{ic}$.
-These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
+A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as WAIC and LOO~ic~.
+These indices allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters [cite gelman].
 In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
@@
 
-Experiment 1 asked about the events [icy, frosty] and [normal, typical] (e.g.
+Experiment 1 asked about the events {icy, frosty} and {normal, typical} (e.g.
 “what is the probability that the weather in London is normal and not typical?”).
@@
 In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.
-
-Experiment 2 focused on more moderately correlated events, [cold, rainy] and [windy, cloudy], that do not admit these misinterpretations.
-In addition, a third experimental condition asking about [warm, snowy] was also included in the experiment, but was dropped from the analyses reported in the paper.
+Experiment 2 focused on more moderately correlated events, {cold, rainy} and {windy, cloudy}, that do not admit these misinterpretations.
+In addition, a third experimental condition asking about {warm, snowy} was also included in the experiment, but was dropped from the analyses reported in the paper.
 Exploring the raw responses from this condition reveals a substantial fraction of “zero” and “one” responses for certain trials.
@@
 These subjective probabilities are unobserved, and must be estimated as a latent variable.
-Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events ($A \land B$, $\neg A \land B$, $A \land \neg B$, $\neg A \land \neg B$).
+Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (A&B, \~A&B, A&\~B, \~A&\~B).
 
@@
 $$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) $$
+$$\mu_{ij} = f_{PT+N}(p_i, d_j, d_j) $$
 $$ d_j = \frac{expit(\delta_j)}{2}$$
 $$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
-$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
+$$ p_i \sim Dirichlet(\overrightarrow{1})$$
 $$ \delta_j \sim Normal(0,1)$$
@@
 To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d’, rather than directly according to its $\beta$, N, and N’ parameters.
-I constrain d to [0, $1/3$] for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
+I constrain d to [0, $\frac{1}{3}$] for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
@@
 $$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) $$
+$$\mu_{ij} = f_{BS}(p_i, d_j, d_j’) $$
 $$ d_j = \frac{expit(\delta_j)}{2}$$
 $$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
-$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
+$$ p_i \sim Dirichlet(\overrightarrow{1})$$
 $$ \delta_j \sim Normal(0,1)$$
@@
 The hierarchical implementation adds parameters for the population-level d and d’ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
-For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency [@papaspiliopoulos.etal2007].
+For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (citation?).
 
 $$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) $$
+$$\mu_{ij} = f_{BS}(p_i, d_j, d_j’) $$
 $$d_j = \frac{ expit(\delta_j )}{3}$$
 $$ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}$$
-$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
+$$ p_i \sim Dirichlet(\overrightarrow{1})$$
 $$ \delta_{pop} \sim Normal(-1, 1)$$
@@
 Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of $\beta > 1$.
-Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of “ignorance priors” [@zhu.etal2020].
+Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of “ignorance priors” (Zhu et al., 2020).
 However, it is also possible that people bring informative priors to the probability judgment task.
-Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted [see e.g., @fennell.baddeley2012].
+Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted (see e.g., Fennel & Baddeley, 2012).
 If $\beta$ is unrestricted, allowed to fall in the domain [0, $\infty$] then the Bayesian Sampler model becomes more flexible, allowing for equivalent “noise” levels in the same [0, .50] range as the PT+N model.
-Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
+Though it would seem a more fundamental change, this model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
 Thus, fitting this additional “hybrid” model allows for a complete comparison of the models along both of their differing dimensions.
@@
 
-I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $LOO_{ic}$ scores for each combination.
-Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
-For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$ [see @gelman.etal2014a].
-Figure X below displays the estimated differences in $LOO_{ic}$ scores for each of the models as compared to the best-scoring model.
+I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed LOO\_ic scores for each combination.
+Model posteriors were estimated using the Numpyro implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler (cite).
+For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$.
+Figure X below displays the estimated differences in LOO\_ic scores for each of the models as compared to the best-scoring model.
 
@@
 \centering
-\includegraphics[width=6in]{plot_compare.png}
-\caption{Model comparison results for data from Experiments 1 and 2. Error bars indicate standard errors of the estimates.}
+\includegraphics[width=5in]{plot_compare.png}
 \end{figure}
@@
 Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the “hybrid” Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$] (again, lower scores for $LOO_{ic}$ are better).
+However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the “hybrid” Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$].
 
@@
 Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of “ignorance priors” in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ d and d’ estimates also lie outside this range (13 of 59 for d, 26 of 59 for d’).
+Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of of individual participants’ d and d’ estimates also lie outside this range (13 of 59 for d, 26 of 59 for d’).
 
@@
 \includegraphics[width=4in]{plot_params.png}
-\caption{Posterior density of population-level d and d' parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.}
 \end{figure}
@@
 In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models.
-However, differences among the remaining three models are approximately within the error of the $LOO_{ic}$ estimates.
+However, differences among the remaining three models are approximately within the error of the LOO\_ic estimates.
 As expected, the simple variants of the models have smaller penalty terms than the complex variants.
@@
 
-Perhaps surprisingly, the Bayesian Sampler with unrestricted $\beta$ actually receives a smaller penalty term than the more restricted version of the model.
+Perhaps surprisingly, the Bayesian Sampler with unrestricted beta actually receives a smaller penalty term than the more restricted version of the model.
 This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
@@
 But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model’s penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied d and d’ estimated for this data under this model are both very near $1/3$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
+Here, it seems reasonable to conjecture that because the implied d and d’ estimated for this data under this model are both very near $\frac{1}{3}$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
 
@@
   add_header_above(c(" ", "Experiment 1"=4, "Experiment 2"=4)) %>% 
-  kable_classic(full_width=F) %>% 
-  kable_styling(latex_options="scale_down")
+  kable_classic(full_width=F)
+
 ```
@@
 
-By a fair margin, the model best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] was a version of the Bayesian Sampler model without restriction on the range of its $\beta$ parameters.
+By a fair margin, the model best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] was a version of the Bayesian Sampler model without restriction on the range of its Beta parameters.
 Alternatively, this “hybrid” model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
@@
 
-Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect “ignorance priors,” symmetric Beta distributions with $\beta$ < 1.
+Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect “ignorance priors,” symmetric beta distributions with $\beta$ < 1.
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people bring informative rather than “ignorance” priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $1/3$] range implied by the bridging conditions when beta is restricted < 1.
+Nevertheless, the data suggest that people bring informative rather than “ignorance” priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $\frac{1}{3}$] range implied by the bridging conditions when beta is restricted < 1.
 One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
@@
 Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g.
-@franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006].
+@jern.etal2014; @griffiths.tenenbaum2006; @franke.etal2016].
 One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes [e.g.
-@cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].
-
-\newpage
-
-# References
+@jern.etal2014; @powell.etal2018; @cook.lewandowsky2016].

### 2021-09-01 paper-rmd/paper.tex
added rmd files

--- 2021-09-01 paper-rmd/paper-papaja.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  english,
-  man,floatsintext]{apa6}
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Untitled probability judgment paper draft},
-  pdfauthor={Derek Powell1},
-  pdflang={en-EN},
-  pdfkeywords={keywords},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+\documentclass{article}
+
+\usepackage{arxiv}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{lipsum}
 \usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-% Manuscript styling
-\usepackage{upgreek}
-\captionsetup{font=singlespacing,justification=justified}
 
-% Table formatting
-\usepackage{longtable}
-\usepackage{lscape}
-% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
-\usepackage{multirow}		% Table styling
-\usepackage{tabularx}		% Control Column width
-\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
-\usepackage{threeparttablex}            % Lets threeparttable work with longtable
+\title{A template for the \emph{arxiv} style}
 
-% Create new environments so endfloat can handle them
-% \newenvironment{ltable}
-%   {\begin{landscape}\centering\begin{threeparttable}}
-%   {\end{threeparttable}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
+\author{
+    Derek Powell
+   \\
+    School of Social and Behavioral Sciences \\
+    Arizona State University \\
+  Phoenix, AZ \\
+  \texttt{\href{mailto:dmpowell@asu.edu}{\nolinkurl{dmpowell@asu.edu}}} \\
+  }
 
-% Enables adjusting longtable caption width to table width
-% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
-\makeatletter
-\newcommand\LastLTentrywidth{1em}
-\newlength\longtablewidth
-\setlength{\longtablewidth}{1in}
-\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
-
-% \setlength{\parindent}{0.5in}
-% \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% \usepackage{etoolbox}
-\makeatletter
-\patchcmd{\HyOrg@maketitle}
-  {\section{\normalfont\normalsize\abstractname}}
-  {\section*{\normalfont\normalsize\abstractname}}
-  {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
-  {\section{\protect\normalfont{\@title}}}
-  {\section*{\protect\normalfont{\@title}}}
-  {}{\typeout{Failed to patch title.}}
-\makeatother
-\shorttitle{DRAFT}
-\keywords{keywords\newline\indent Word count: X}
-\usepackage{csquotes}
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
-\ifluatex
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
+% Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
  {}
-\usepackage{calc}
+\usepackage{calc} % for calculating minipage widths
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-\title{Untitled probability judgment paper draft}
-\author{Derek Powell\textsuperscript{1}}
-\date{}
-
-\authornote{
-
-Arizona State University, School of Social and Behavioral Sciences
-
-The authors made the following contributions. Derek Powell: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
-
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Arizona State University, School of Social and Behavioral Sciences}
-
-\abstract{
-Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic.
-However, people's explicit probability judgments often incoherently violate the laws of probability.
-Two recent proposals, the ``Probability Theory plus Noise'' (Costello \& Watts, 2014) and ``Bayesian Sampler'' (Zhu, Sanborn, \& Chater, 2020) theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
-The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity.
-Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the \(LOO_{ic}\) information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020).
-My findings support the Bayesian Sampler's account of conditional probability judgments, and I find the data are best explained by a ``hybrid'' model that can be seen as a modification of both theories.
-}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-Bayesian theories and approaches offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) with remarkable successes across domains, including perception (e.g.
-Kersten, Mamassian, \& Yuille, 2004), memory (Anderson, 1991), language (Xu \& Tenenbaum, 2007), and reasoning (Lu, Chen, \& Holyoak, 2012).
-At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow,'' ``vaccines are safe,'' or ``this politician is trustworthy.''
+\def\tightlist{}
 
-Trouble starts as soon as we seek to measure beliefs.
-According to Bayesian theories of cognition and epistemology (Jaynes, 2003), the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
-So asking people to express beliefs in terms of probability seems only natural.
+\begin{abstract}
+Enter the text of your abstract here.
+\end{abstract}
 
-Unfortunately, people's explicit probability judgments routinely violate the axioms of probability theory.
-For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory (Tversky \& Kahneman, 1983).
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects (Tversky \& Koehler, 1994), and a variety of other effects illustrating the incoherence of human probability judgments (for an accessible review, see Kahneman, 2013).
-Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
+\keywords{
+    blah
+   \and
+    blee
+   \and
+    bloo
+   \and
+    these are optional and can be removed
+  }
 
-Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020).
-Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by drawing a limited number of samples from probabilistic mental models (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Bayesian theories of cognition have had remarkable successes in
+explaining human reasoning and behavior across many domains (big cite).
+{[}The core of these theories is that people reason according to
+subjective mentally-represented degrees of belief, and they specify how
+they should be revised in light of evidence. It is somewhat embarrassing
+then that one area where these theories seem to fall down is in
+describing human ``beliefs'\,' of the simple and everyday sort, such as
+beliefs like''it will rain tomorrow``, ``vaccines are safe,'' or ``this
+politician is trustworthy.''
+
+Trouble starts as soon as we seek to measure beliefs. According to
+Bayesian theories of cognition and epistemology, the degree to which
+people believe in various propositions should reflect subjective mental
+probabilities. So asking people to express beliefs in terms of
+probability seems only natural.
+
+Unfortunately, people's explicit probability judgments routinely violate
+the axioms of probability theory. For example, human probability
+judgments often exhibit the ``conjunction fallacy'': people will often
+judge the conjunction of two events (e.g.~``Tom Brady likes football and
+miniature horses'') as being more probable than one of the events in
+isolation (e.g.~``Tom Brady likes miniature horses''), a plain and
+flagrant violation of probability theory (cite some examples). Other
+demonstrations of the incoherence of probability judgments include
+disjunction fallacies (e.g.~XXX), ``unpacking'' effects (e.g.~fox \&
+tversky), and a variety of other effects illustrating the incoherence of
+human probability judgments {[}cite{]}. Altogether these findings have
+led many researchers to abandon the notion that credences are
+represented as probabilities.
+
+Recently however, two groups of researchers have proposed theories of
+human probability judgments that account for biases and apparent
+incoherence in these judgments while maintaining that mental credences
+are fundamentally probabilistic (Costello and Watts 2014; Zhu, Sanborn,
+and Chater 2020). Both of these theories build on the increasingly
+popular notion that a variety of human reasoning tasks are accomplished
+by drawing a limited number of samples from probabilistic mental models
+(see also Chater et al. 2020; Dasgupta, Schulz, and Gershman 2017).
 
 \hypertarget{two-probabilistic-theories-of-probability-judgment}{%
-\subsection{Two probabilistic theories of probability judgment}\label{two-probabilistic-theories-of-probability-judgment}}
+\subsection{Two probabilistic theories of probability
+judgment}\label{two-probabilistic-theories-of-probability-judgment}}
 
-Costello \& Watts (2014, 2016, 2018) proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
-In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative read as positive.
-This results in probability judgments that reflect probabilistic credences perturbed by noise.
-Under the simplest form of the PT+N model, the expected value of probability judgments is:
+Costello \& Watts (2014, 2016, 2018) proposed a theory of probability
+judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
+In the PT+N model, mental ``samples'' are drawn from a probabilistic
+mental model of events and are then ``read'' with noise, so that some
+positive examples will be read as negative and some negative read as
+positive. This results in probability judgments that reflect
+probabilistic credences perturbed by noise. Under the simplest form of
+the PT+N model, the expected value of probability judgments is:
 
@@
 
-where \emph{d} represents the probability with which samples will be misread, or the amount of noise in the judgment process (where by assumption a maximum of 50\% of samples can be misread, so that d is a number in the range {[}0, .50{]}).
-The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics (Costello \& Watts, 2014, 2016, 2017, 2018).
+where \emph{d} represents the probability with which samples will be
+misread, or the amount of noise in the judgment process (where by
+assumption a maximum of 50\% of samples can be misread, so that d is a
+number in the range {[}0, .50{]}). The PT+N theory provides a unified
+accounts for a wide variety of biases in probability judgment that were
+previously attributed to different types of heuristics (Costello and
+Watts 2014, 2016, 2017, 2018).
 
-Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model of probability judgment they call the ``Bayesian Sampler.'' Under this model, probability judgment is itself seen as a process of Bayesian inference.
-To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event.
-Then, those ``observed'' samples are integrated with a prior over probabilities to produce a probability judgment.
-The effect of this prior is to regularize the judgments, pushing them away from extreme values.
-They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g.~encourage taking large risks based on a small number of mental samples.
-However, this hedging comes at the cost of systematic incoherence and biases.
-Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts (Costello \& Watts, 2014, 2016).
+Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model
+of probability judgment they call the ``Bayesian Sampler.'' Under this
+model, probability judgment is itself seen as a process of Bayesian
+inference. To judge the probability of an event, a limited number of
+samples are again drawn from a mental model of the event. Then, those
+``observed'' samples are integrated with a prior over probabilities to
+produce a probability judgment. The effect of this prior is to
+regularize the judgments, pushing them away from extreme values. They
+argue this regularization of judgments is adaptive, in that it pushes
+judgments and decisions away from extremes that might, e.g.~encourage
+taking large risks based on a small number of mental samples. However,
+this hedging comes at the cost of systematic incoherence and biases.
+Like the PT+N model, their proposed model accounts for a wide array of
+biases in probability judgments, including the novel biases identified
+by Costello and Watts (Costello and Watts 2014, 2016).
 
-Under the simplest version of the Bayesian Sampler model, the expected value of a probability judgment is given by the following formula:
+Under the simplest version of the Bayesian Sampler model, the expected
+value of a probability judgment is given by the following formula:
 
@@
 
-where \emph{N} represents the number of mental samples drawn.
-The mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric beta distribution parameterized by \(\beta\).
-Zhu and colleagues (2020) show that the \emph{N} and \(\beta\) parameters of their model can be related to the \emph{d} parameter of the PT+N model via the following bridging formula:
+where \emph{N} represents the number of mental samples drawn. The
+mentally-observed samples are integrated with a Bayesian prior over
+probabilities, represented with a symmetric beta distribution
+parameterized by \(\beta\). Zhu and colleagues (2020) show that the
+\emph{N} and \(\beta\) parameters of their model can be related to the
+\emph{d} parameter of the PT+N model via the following bridging formula:
 
@@
 \hypertarget{two-accounts-of-conditional-probability-judgments}{%
-\subsubsection{Two accounts of conditional probability judgments}\label{two-accounts-of-conditional-probability-judgments}}
+\subsubsection{Two accounts of conditional probability
+judgments}\label{two-accounts-of-conditional-probability-judgments}}
 
-By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs (Chater et al., 2020).
-However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project.
-Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence.
-So, the conditioning of beliefs is at the heart of these theories.
+By explaining the incoherence of human probability judgments using
+coherent mental probabilities, both models have the potential to rescue
+the larger project of Bayesian cognitive science as applied to everyday
+beliefs (Chater et al. 2020). However, because they diverge in their
+treatment of conditional probability judgments, the two theories fit
+very differently into the larger Bayesian project. According to the
+Bayesian sampler model, conditioning is something that happens in the
+mental model of the events, not as part of the process of rendering
+probability judgments. Thus, the Bayesian sampler model does not assign
+any special status to conditional probability judgments and so fits
+neatly into the larger Bayesian project of cognitive science:
+probability judgments are simply another judgment process applied to the
+outputs of other Bayesian mental models (Chater et al. 2020).
 
-According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments.
-By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models (Chater et al., 2020).
+In contrast, the PT+N model presents a constructive account of
+conditional probability judgments that is fundamentally non-Bayesian
+(Costello and Watts 2016). According to the PT+N model, conditional
+probabilities P(A\textbar B) are estimated by a two-stage sampling
+procedure: first both events A and B are sampled with noise, and then a
+second noisy process computes the ratio of the events read as A and B
+over events read as B. The PT+N model predicts conditional probability
+estimates using the following equation:
 
-In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian (Costello \& Watts, 2016).
-According to the PT+N model, conditional probabilities P(A\textbar B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B.
-The PT+N model predicts conditional probability estimates using the following equation:
-
-\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]
-
-The PT+N's account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally.
-In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.
+\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]\\
+Given that conditional inference and belief updating lies at the core of
+Bayesian cognitive theory, PT+N's account of conditional probability
+judgment separates it from the Bayesian sampler theory not just
+quantitatively, but quite fundamentally. First, it denies any central
+importance of conditional probability judgments by pulling conditioning
+out of the mental model of an event and placing it into the probability
+judgment process. And perhaps more importantly, this non-bayesian
+account of conditional inference can no longer be considered
+``rational'' at any stage under the PT+N model.
 
@@
 
-Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments.
-Unfortunately, their results were somewhat equivocal.
-They fit both ``simple'' and ``complex'' versions of each model and computed a Bayesian Information Criteria score (BIC) for all models.
-The complex versions of these models introduce additional parameters d' and N' that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are needed in order to explain conjunction fallacies (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020) .
-Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs (Zhu, Sanborn, \& Chater, 2020, Supplementary materials).
+Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model
+against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as
+explanations for human probability judgments in two experiments.
+Unfortunately, their results were somewhat equivocal. They fit both
+``simple'' and ``complex'' versions of each model and computed a
+Bayesian Information Criteria score (BIC) for all models. The complex
+versions of these models introduce additional parameters d' and N' that
+allow for different patterns of judgments for conjunctive and
+disjunctive judgments as compared with simple probability judgments.
+These additional parameters are needed in order to explain conjunction
+fallacies (Costello \& Watts, 2017; Zhu et al., 2020). Table 1 below
+presents the total BIC scores computed for each model as originally fit,
+using the authors' original code and saved model outputs (Zhu et al,
+2020, Supplementary materials).
 
@@
 
-In both experiments, the simple version of the Bayesian Sampler scores best, but the complex version of the PT+N model comes in second (lower BIC scores are better).
-It is not obvious what conclusions should be drawn from these results.
-Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding (Mellers, Hertwig, \& Kahneman, 2001; Sides, Osherson, Bonini, \& Viale, 2002) and are clearly present in the data collected by Zhu and colleagues (2020).
-So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
-For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.
+In both experiments, the simple version of the Bayesian Sampler scores
+best, but the complex version of the PT+N model comes in second (lower
+BIC scores are better). It is not obvious what conclusions should be
+drawn from these results. Of course, the simple Bayesian Sampler model
+appears to win by the numbers. But conjunction fallacies are an
+extremely robust empirical finding (e.g.~cite) and are clearly present
+in the data collected by Zhu and colleagues (2020). So we might rule out
+the simple variants of the models on the grounds that they will fail to
+capture important qualitative features of the data, which would instead
+favor the PT+N theory. For their part, Zhu and colleagues (2020) elect
+to split the baby by averaging the simple and complex model scores
+together. This approach somewhat favors the Bayesian Sampler theory
+overall, though they are rightfully cautious about drawing any strong
+conclusions.
 
 \hypertarget{accounting-for-model-complexity}{%
-\subsubsection{Accounting for Model Complexity}\label{accounting-for-model-complexity}}
+\subsubsection{Accounting for Model
+Complexity}\label{accounting-for-model-complexity}}
 
-When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to ``overfit'' the data.
-Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data (Gelman, Hwang, \& Vehtari, 2014).
-Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models' complexity.
+Zhu and colleagues (2020) warn that BIC, which penalizes models based
+only on the number of parameters in the model, cannot fully account for
+the differences in the models' complexity. There are at least three
+challenges to accounting for model complexity in the comparison of these
+models. First, the models differ not only in the number of parameters
+but in the domain of those parameters. Zhu and colleagues (2020)
+restrict the range of the \(\beta\) parameter in their Bayesian Sampler
+model to fall within the interval {[}0,1{]}. This restriction constrains
+the Bayesian Sampler model's prior distribution to the class of
+``ignorance priors,'' symmetric beta distributions where \(\beta < 1\).
+Via the bridging conditions (eq X), they show that this restricts the
+``noise level'' for the Bayesian Sampler, represented in implied d under
+the PT+N model, to {[}0, \(\frac{1}{3}\){]} whereas the PT+N model
+permits noise values between {[}0, 0.50{]}.
 
-There are at least three challenges to accounting for model complexity in the comparison of the PT+N and Bayesian Sampler models.
-First, the models differ not only in the number of parameters but in the domain of those parameters.
-Zhu and colleagues (2020) restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}.
-This restriction constrains the Bayesian Sampler model's prior distribution to the class of ``ignorance priors.'' Via the bridging conditions (eq X), they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]} whereas the PT+N model permits noise values between {[}0, 0.50{]}.
+Second, it is not immediately clear how the structural differences
+between the models impact their flexibility. For the purposes of model
+comparison, the ``complexity penalty'' that models truly ought to pay is
+determined by their relative flexibility in accommodating different
+patterns of data, not simply the number or domain of their parameters
+(cite Gelman). To be sure, Zhu and colleagues' (2020) bridging condition
+makes it clear that the PT+N model is more flexible when it comes to
+predicting unconditional probabilities. But, what impact does the PT+N
+model's treatment of conditional probabilities have on model complexity?
+Is this component of the model a sort of Ptolemaic epicycle, adding
+complexity to the model that should be penalized? Or, does it constitute
+a commitment to novel predictions that thereby constrain its
+flexibility? Determining model complexity a priori is not always
+straightforward when the models being compared differ structurally.
 
-Second, it is not immediately clear how the models' structural differences impact their flexibility.
-To be sure, Zhu and colleagues' (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
-But, what impact does the PT+N model's treatment of conditional probabilities have on model complexity?
-Is this component of the model a sort of Ptolemaic epicycle, adding complexity to the model that should be penalized?
-Or, does it constitute a commitment to novel predictions that thereby constrain its flexibility?
-Determining model complexity a priori is not always straightforward when the models being compared differ structurally.
-
-Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues' (2020) use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
-Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
-In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
-Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
+Finally, one potential explanation for the relative weakness of the
+complex model versions is Zhu and colleagues' (2020) use of unpooled
+models, where parameters are estimated independently for each individual
+participant. Adding a parameter for each participant may effectively
+over-penalize the ``complex'' variants relative to the simple variants,
+especially if there is actually limited heterogeneity across
+participants. In contrast, a comparison of fully pooled variants of the
+simple and complex models (where a single population parameter is used
+for all participants) would require adding only one extra parameter to
+the penalty term. Partial pooling is a solution that balances between
+these extreme approaches, allowing for an accounting of heterogeneity
+without over-penalizing in cases where heterogeneity is low.
 
@@
 
-Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison.
-A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \(WAIC\) and \(LOO_{ic}\).
-These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017).
-In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.
+Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian
+data analysis framework that permits a more decisive (or at least
+alternative) comparison. A Bayesian analysis allows issues of model
+complexity to be addressed through comparisons of model fit based on
+information criteria, such as WAIC and LOO\textsubscript{ic}. These
+indices allow for complexity penalties that directly capture differences
+in the flexibility of the models, rather than being estimated based
+solely on the number of distinct parameters {[}cite gelman{]}. In
+addition, this framework also supports straightforward implementation of
+hierarchical versions of these models with partial pooling. This allows
+for information about model parameters to be shared across participants,
+resulting in an overall reduction in model complexity and a more
+realistic test of the models.
 
@@
 
-Zhu, Sanborn, \& Chater (2020) conducted two experiments to compare the PT+N and Bayesian sampler theories.
-These experiments asked participants to judge the probability of different events in various combinations.
-Following prior work by Costello and Watts (e.g.
-2016, 2018), both experiments focused on the everyday events of different kinds of weather.
+Zhu, Sanborn, \& Chater (2020) conducted two experiments to compare the
+PT+N and Bayesian sampler theories. These experiments asked participants
+to judge the probability of different events in various combinations.
+Following prior work by Costello and Watts (e.g. 2016, 2018), both
+experiments focused on the everyday events of different kinds of
+weather.
 
-Experiment 1 asked about the events {[}icy, frosty{]} and {[}normal, typical{]} (e.g.~``what is the probability that the weather in London is normal and not typical?'').
-The authors' goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated.
-Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment.
-This is especially clear, as the authors note, for disjunctive trials such as ``normal or typical,'' where ``or typical'' might not be read as a disjunction but rather an elaborative clause.
-In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.
-
-Experiment 2 focused on more moderately correlated events, {[}cold, rainy{]} and {[}windy, cloudy{]}, that do not admit these misinterpretations.
-In addition, a third experimental condition asking about {[}warm, snowy{]} was also included in the experiment, but was dropped from the analyses reported in the paper.
-Exploring the raw responses from this condition reveals a substantial fraction of ``zero'' and ``one'' responses for certain trials.
-This may reflect a different response process than was intended.
-For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero---failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day.
-Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.
+Experiment 1 asked about the events \{icy, frosty\} and \{normal,
+typical\} (e.g.~``what is the probability that the weather in London is
+normal and not typical?''). The authors' goal was to ask about highly
+correlated events, but the events used are perhaps nearly perfectly
+correlated. Because the terms used to describe these events are nearly
+synonymous, there is a concern about the interpretation of the
+statements evaluated in this experiment. This is especially clear, as
+the authors note, for disjunctive trials such as ``normal or typical,''
+where ``or typical'' might not be read as a disjunction but rather an
+elaborative clause. In light of these concerns, I excluded the
+disjunctive trials from Experiment 1 from my analyses. Experiment 2
+focused on more moderately correlated events, \{cold, rainy\} and
+\{windy, cloudy\}, that do not admit these misinterpretations. In
+addition, a third experimental condition asking about \{warm, snowy\}
+was also included in the experiment, but was dropped from the analyses
+reported in the paper. Exploring the raw responses from this condition
+reveals a substantial fraction of ``zero'' and ``one'' responses for
+certain trials. This may reflect a different response process than was
+intended. For instance, some participants may have engaged in deductive
+reasoning to judge that it is not possible for the weather to at once be
+warm and snowy, and therefore responded with zero---failing to properly
+consider that it is possible (at least logically) for it to be warm and
+snowy at different times within the same day. Given these potentially
+aberrant responses, I followed Zhu and colleagues (2020) in ignoring
+data from this condition.
 
@@
 
-I implement both the Bayesian Sampler and PT+N models as Bayesian models in the probabilistic programming language Numpyro.
-All code and results are available as supplemental materials (\url{https://github.com/derekpowell/bayesian-sampler}).
+I implement both the Bayesian Sampler and PT+N models as Bayesian models
+in the probabilistic programming language Numpyro. All code and results
+are available as supplemental materials
+(\url{https://github.com/derekpowell/bayesian-sampler}).
 
 \hypertarget{bayesian-implementations-of-the-models}{%
-\subsubsection{Bayesian implementations of the models}\label{bayesian-implementations-of-the-models}}
+\subsubsection{Bayesian implementations of the
+models}\label{bayesian-implementations-of-the-models}}
 
@@
 
-Fixing d and d' or N and N' equal yields the ``simple'' variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
+Fixing d and d' or N and N' equal yields the ``simple'' variant of each
+of the models, which treat conjunctive and disjunctive probability
+judgments identically to simple probability judgments.
 
-Notice that for each model the probability judgments depend on underlying subjective probabilities, derived from a mental sampling process.
-These subjective probabilities are unobserved, and must be estimated as a latent variable.
-Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (\(A \land B\), \(\neg A \land B\), \(A \land \neg B\), \(\neg A \land \neg B\)).
+Notice that for each model the probability judgments depend on
+underlying subjective probabilities, derived from a mental sampling
+process. These subjective probabilities are unobserved, and must be
+estimated as a latent variable. Here, they are estimated as a
+four-dimensional dirichlet distribution for each subject, representing
+the probability of the elementary events (A\&B, \textasciitilde A\&B,
+A\&\textasciitilde B, \textasciitilde A\&\textasciitilde B).
 
-Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate d, d', N, N', and \(\beta\) parameters for each participant.
+Zhu, Sanborn \& Chater (2020) implement completely unpooled models with
+separate d, d', N, N', and \(\beta\) parameters for each participant.
 Translated into a Bayesian Framework, we can write the PT+N model as:
@@
 \[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) \]
+\[\mu_{ij} = f_{PT+N}(p_i, d_j, d_j) \]
 \[ d_j = \frac{expit(\delta_j)}{2}\]
 \[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
-\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_j \sim Normal(0,1)\]
-\[ \Delta\delta_j \sim Normal(0,1)\]
+\[ p_i \sim Dirichlet(\overrightarrow{1})\]
+\[ \delta_j \sim Normal(0,1)\] \[ \Delta\delta_j \sim Normal(0,1)\]
 \[ k \sim HalfCauchy(20)\]
 
-The function \(f_{PT+N}\) computes the expected probability estimate using the underlying subjective probability p, the noise parameters d and d', and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
-Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters d and d'.
+The function \(f_{PT+N}\) computes the expected probability estimate
+using the underlying subjective probability p, the noise parameters d
+and d', and the relevant equation as defined by the PT+N theory (see
+supplemental materials for implementation details). Prior predictive
+checks were conducted for all models to select priors that would be
+uninformative or minimally informative on the scale of the model
+parameters d and d'.
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its \(\beta\), N, and N' parameters.
-I constrain d to {[}0, \(1/3\){]} for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
-This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
+Recall that Zhu and colleagues (2020) identified a bridging condition
+relating b and N in the Bayesian Sampler model to the d parameter of the
+PT+N model. To support direct comparisons of the models, I parameterize
+the Bayesian Sampler model according to the implied d and d', rather
+than directly according to its \(\beta\), N, and N' parameters. I
+constrain d to {[}0, \(\frac{1}{3}\){]} for the Bayesian Sampler model
+to reflect the assumption that \(\beta \in [0, 1]\). This allows the
+same priors to be used for the corresponding Bayesian Sampler and PT+N
+models, simplifying their comparison.
 
@@
 \[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) \]
+\[\mu_{ij} = f_{BS}(p_i, d_j, d_j’) \]
 \[ d_j = \frac{expit(\delta_j)}{2}\]
 \[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
-\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_j \sim Normal(0,1)\]
-\[ \Delta\delta_j \sim Normal(0,1)\]
+\[ p_i \sim Dirichlet(\overrightarrow{1})\]
+\[ \delta_j \sim Normal(0,1)\] \[ \Delta\delta_j \sim Normal(0,1)\]
 \[ k \sim HalfCauchy(20)\]
 
-Where the function \(f_{BS}\) computes the expected probability estimate as prescribed by the Bayesian Sampler theory.
+Where the function \(f_{BS}\) computes the expected probability estimate
+as prescribed by the Bayesian Sampler theory.
 
 \hypertarget{hierarchical-implementations-of-the-models}{%
-\subsubsection{Hierarchical implementations of the models}\label{hierarchical-implementations-of-the-models}}
+\subsubsection{Hierarchical implementations of the
+models}\label{hierarchical-implementations-of-the-models}}
 
-Both of these models can also be implemented as hierarchical models with partial pooling for the d and d' parameters (implicitly, for N and N' in the case of the Bayesian Sampler).
-This should prevent over-penalizing complex versions of the model, as well as potentially improving their overall predictive fit.
-The hierarchical implementation adds parameters for the population-level d and d' as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
-For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (Papaspiliopoulos, Roberts, \& Sköld, 2007).
+Both of these models can also be implemented as hierarchical models with
+partial pooling for the d and d' parameters (implicitly, for N and N' in
+the case of the Bayesian Sampler). This should prevent over-penalizing
+complex versions of the model, as well as potentially improving their
+overall predictive fit. The hierarchical implementation adds parameters
+for the population-level d and d' as well as a parameter controlling the
+standard deviation of the distribution for the subject-level effects.
+For ease of interpretation, the centered parameterization is shown
+below, although the actual models used a non-centered parameterization
+to improve sampling efficiency (citation?).
 
 \[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) \]
+\[\mu_{ij} = f_{BS}(p_i, d_j, d_j’) \]
 \[d_j = \frac{ expit(\delta_j )}{3}\]
 \[ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}\]
-\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
+\[ p_i \sim Dirichlet(\overrightarrow{1})\]
 \[ \delta_{pop} \sim Normal(-1, 1)\]
@@
 
-Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of \(\beta > 1\).
-Restricting \(\beta\) to {[}0,1{]} restricts the prior distribution of the Bayesian sampler to the class of ``ignorance priors'' (Zhu, Sanborn, \& Chater, 2020).
-However, it is also possible that people bring informative priors to the probability judgment task.
-Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted (see e.g., Fennell \& Baddeley, 2012).
-If \(\beta\) is unrestricted, allowed to fall in the domain {[}0, \(\infty\){]} then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same {[}0, .50{]} range as the PT+N model.
-Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
-Thus, fitting this additional ``hybrid'' model allows for a complete comparison of the models along both of their differing dimensions.
+Finally, I also explored fitting versions of the Bayesian Sampler model
+that allowed values of \(\beta > 1\). Restricting \(\beta\) to {[}0,1{]}
+restricts the prior distribution of the Bayesian sampler to the class of
+``ignorance priors'' (Zhu et al., 2020). However, it is also possible
+that people bring informative priors to the probability judgment task.
+Indeed, Zhu and colleagues (2020) acknowledge there are situations where
+an informative prior may be warranted (see e.g., Fennel \& Baddeley,
+2012). If \(\beta\) is unrestricted, allowed to fall in the domain {[}0,
+\(\infty\){]} then the Bayesian Sampler model becomes more flexible,
+allowing for equivalent ``noise'' levels in the same {[}0, .50{]} range
+as the PT+N model. Though it would seem a more fundamental change, this
+model may also be seen as a version of the PT+N theory that jettisons
+its constructive account of conditional probability judgment. Thus,
+fitting this additional ``hybrid'' model allows for a complete
+comparison of the models along both of their differing dimensions.
 
@@
 
-I fit each of the models specified above to data from Zhu et al's (2020) Experiment 1 and 2 and computed \(LOO_{ic}\) scores for each combination.
-Model posteriors were estimated using the Numpyro (Phan, Pradhan, \& Jankowiak, 2019) implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
-For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to \(\hat{R}\) (see Gelman et al., 2014).
-Figure X below displays the estimated differences in \(LOO_{ic}\) scores for each of the models as compared to the best-scoring model.
+I fit each of the models specified above to data from Zhu et al's (2020)
+Experiment 1 and 2 and computed LOO\_ic scores for each combination.
+Model posteriors were estimated using the Numpyro implementation of the
+No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler (cite).
+For each model, four MCMC chains of 2000 iterations were sampled after
+2000 iterations of warmup and all passed convergence tests according to
+\(\hat{R}\). Figure X below displays the estimated differences in
+LOO\_ic scores for each of the models as compared to the best-scoring
+model.
 
@@
 \centering
-\includegraphics[width=6in]{plot_compare.png}
-\caption{Model comparison results for data from Experiments 1 and 2. Error bars indicate standard errors of the estimates.}
+\includegraphics[width=5in]{plot_compare.png}
 \end{figure}
 
-Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the ``hybrid'' Bayesian Sampler model that removes the restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]} (again, lower scores for \(LOO_{ic}\) are better).
+Data from Experiment 1 are only modestly informative with respect to
+discriminating among the models. However, data from Experiment 2 reveal
+a clearer winner: a hierarchical implementation of the ``hybrid''
+Bayesian Sampler model that removes the restriction on \(\beta\),
+allowing it to range from {[}0, \(\infty\){]}.
 
-Recall, this model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the hybrid model and the best-fitting PT+N model, we see that the hybrid model better captures these judgments from Experiment 2 (trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) = .51 vs .45).
-These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
+Recall, this model may also be seen as a version of the PT+N theory that
+removes the original theory's constructive account of conditional
+probability judgments. Comparing predictions for conditional probability
+judgments from the hybrid model and the best-fitting PT+N model, we see
+that the hybrid model better captures these judgments from Experiment 2
+(trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) =
+.51 vs .45). These findings suggest that conditioning is better seen as
+part of the mental model than the probability judgment process.
 
-Figure X shows the posterior distributions of the population-level d and d' parameters inferred from the ``hybrid'' model.
-Using data from experiment 2, population-level estimates of d and d' are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' d and d' estimates also lie outside this range (13 of 59 for d, 26 of 59 for d').
+Figure X shows the posterior distributions of the population-level d and
+d' parameters inferred from the ``hybrid'' model. Using data from
+experiment 2, population-level estimates of d and d' are greater than
+\(\frac{1}{3}\), outside the range implied by the assumption of
+``ignorance priors'' in the Bayesian Sampler model. Parameters fit to
+the data from Experiment 1 are more consistent with this assumption,
+although a substantial proportion of of individual participants' d and
+d' estimates also lie outside this range (13 of 59 for d, 26 of 59 for
+d').
 
@@
 \includegraphics[width=4in]{plot_params.png}
-\caption{Posterior density of population-level d and d' parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.}
 \end{figure}
 
-Among the remaining models, hierarchical implementations with partial pooling outperform the unpooled corresponding models---suggesting that unpool variants of the complex models were over-penalized in the original analysis by Zhu, Sanborn, and Chater (2020).
-In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models.
-However, differences among the remaining three models are approximately within the error of the \(LOO_{ic}\) estimates.
-As expected, the simple variants of the models have smaller penalty terms than the complex variants.
+Among the remaining models, hierarchical implementations with partial
+pooling outperform the unpooled corresponding models---suggesting that
+unpool variants of the complex models were over-penalized in the
+original analysis by Zhu, Sanborn, and Chater (2020). In both
+experiments, the simple version of the PT+N model is clearly
+outperformed by the other competing models. However, differences among
+the remaining three models are approximately within the error of the
+LOO\_ic estimates. As expected, the simple variants of the models have
+smaller penalty terms than the complex variants.
 
-Compared to the hybrid model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d'.
-Evidently this structural component isn't a complexifying epicycle, instead it is a prediction that constrains flexibility.
-However, it is one that ultimately leads to a worse-fitting model.
+Compared to the hybrid model, the PT+N model with its constructive
+account of conditional probability judgments has a smaller penalty term
+despite having the same parameterization in terms of d and d'. Evidently
+this structural component isn't a complexifying epicycle, instead it is
+a prediction that constrains flexibility. However, it is one that
+ultimately leads to a worse-fitting model.
 
-Perhaps surprisingly, the Bayesian Sampler with unrestricted \(\beta\) actually receives a smaller penalty term than the more restricted version of the model.
-This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data (see Gelman, Hwang, \& Vehtari, 2014).
-Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is estimated from data (2014).
-If the estimated value is some very large positive number, then the constraint won't have been very informative.
-But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model's penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied d and d' estimated for this data under this model are both very near \(1/3\), the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
+Perhaps surprisingly, the Bayesian Sampler with unrestricted beta
+actually receives a smaller penalty term than the more restricted
+version of the model. This is at first unintuitive, but it illustrates
+how model complexity depends not only on the model and priors, but also
+the observed data (see Gelman, Hwang, and Vehtari 2014). Gelman and
+colleagues consider a case where a parameter is constrained to be
+positive and its value is estimated from data (2014). If the estimated
+value is some very large positive number, then the constraint won't have
+been very informative. But, if the estimated value is very close to
+zero, then the constraint that the parameter is positive will provide
+substantial information and the model's penalty term will therefore be
+smaller. Here, it seems reasonable to conjecture that because the
+implied d and d' estimated for this data under this model are both very
+near \(\frac{1}{3}\), the restriction results in posterior estimates of
+the linear parameters that are relatively far from the prior, which can
+result in a greater penalty.
 
-The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria.
-Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives (e.g.
-Gronau \& Wagenmakers, 2019; Vehtari, Simpson, Yao, \& Gelman, 2019).
-Fortunately, conclusions from the comparisons here do not rest solely on differences between the models' complexity.
+The dependence of complexity penalties on observed data may strike some
+as an undesirable feature of model comparison through information
+criteria. Indeed, it is worth acknowledging that principled model
+comparison is still an area of active inquiry, with differing
+perspectives (e.g. Gronau and Wagenmakers 2019; Vehtari et al. 2019).
+Fortunately, conclusions from the comparisons here do not rest solely on
+differences between the models' complexity.
 
@@
 \centering
-\resizebox{\linewidth}{!}{
 \begin{tabular}[t]{lrrrrrrrr}
@@
 \bottomrule
-\end{tabular}}
+\end{tabular}
 \end{table}
 
-Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table X.
+Finally, it is worth noting that these models provide quite strong
+overall fits to the data, not just for the query averages, but also for
+the trial averages across individual participants as seen from the
+correlations between predicted and observed responses in Table X.
 
@@
 
-By a fair margin, the model best accounting for the experimental data from Zhu and colleagues (2020) was a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters.
-Alternatively, this ``hybrid'' model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
-Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task.
-In keeping with the larger theoretical framework of Bayesian cognitive science, this theory assumes that subjective probabilities underlie people's probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process (Chater et al., 2020; Zhu, Sanborn, \& Chater, 2020).
+By a fair margin, the model best accounting for the experimental data
+from Zhu and colleagues (2020) was a version of the Bayesian Sampler
+model without restriction on the range of its Beta parameters.
+Alternatively, this ``hybrid'' model can also be seen as a variant of
+the PT+N model that removes its account of conditional probability
+judgments. Thus, what these findings indicate most clearly is that the
+Bayesian Sampler theory provides a superior account of conditional
+probability judgments in this experimental task. In keeping with the
+larger theoretical framework of Bayesian cognitive science, this theory
+assumes that subjective probabilities underlie people's probability
+judgments, and that conditional probability judgments are produced by
+Bayesian conditioning occurring in their mental models of the events in
+question, rather than as arising from the probability judgment process
+(Chater et al. 2020; Zhu, Sanborn, and Chater 2020).
 
-Findings from Zhu and colleagues' Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta\) \textless{} 1.
-As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people bring informative rather than ``ignorance'' priors to the probability judgment task, indicated by estimated d and d' parameters outside the {[}1, \(1/3\){]} range implied by the bridging conditions when beta is restricted \textless{} 1.
-One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-If so, other experimental contexts might see parameter estimates consistent with an uninformative prior.
-In addition, the noise parameters also varied across individuals.
-Exploring the participant-level parameters reveals that some individuals' implied d and d' parameters were consistent with the class of ignorance priors (see supplemental materials).
-Further research should explore this heterogeneity within and across individuals.
-Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the \(\beta\) and N parameters are not uniquely identifiable from judgment data of the sort examined here.
+Findings from Zhu and colleagues' Experiment 2 do cast doubt on their
+proposal that the priors of the Bayesian Sampler model should reflect
+``ignorance priors,'' symmetric beta distributions with \(\beta\)
+\textless{} 1. As a generic prior that would be used across contexts,
+this class of uninformative priors has an appealing rational basis.
+Nevertheless, the data suggest that people bring informative rather than
+``ignorance'' priors to the probability judgment task, indicated by
+estimated d and d' parameters outside the {[}1, \(\frac{1}{3}\){]} range
+implied by the bridging conditions when beta is restricted \textless{}
+1. One possibility is that people bring domain-specific priors to many
+judgment tasks, meaning that the most appropriate priors might be
+dictated by the context in which they make their judgments. If so, other
+experimental contexts might see parameter estimates consistent with an
+uninformative prior. In addition, the noise parameters also varied
+across individuals. Exploring the participant-level parameters reveals
+that some individuals' implied d and d' parameters were consistent with
+the class of ignorance priors (see supplemental materials). Further
+research should explore this heterogeneity within and across
+individuals. Unfortunately, pinning down specific components of the
+Bayesian Sampler model is a challenge, as the \(\beta\) and N parameters
+are not uniquely identifiable from judgment data of the sort examined
+here.
 
-One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given their tight connections via the bridging conditions (Zhu, Sanborn, \& Chater, 2020), it may not be possible to draw decisive conclusions here.
-Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that ``noise'' might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler (2020).
+One thing this model comparison has not decided, and likely
+\emph{cannot} decide, is whether the distortions of probability
+judgments are products of mental noise or of further reasoning
+processes. Given their tight connections via the bridging conditions
+(Zhu, Sanborn, and Chater 2020), it may not be possible to draw decisive
+conclusions here. Moreover, the theories may not be in any real
+competition over this point: Zhu and colleagues consider that ``noise''
+might give an algorithmic-level solution to the computation-level task
+defined by the Bayesian Sampler (2020).
 
-Some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs (Chater et al., 2020).
-Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying subjective probabilities to be inferred.
-Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
-Of course, Zhu and colleagues' (2020) experiments were never designed for this purpose.
-Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning (e.g.
-Franke et al., 2016; Griffiths \& Tenenbaum, 2006; Jern, Chang, \& Kemp, 2014).
-One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes (e.g.
-Cook \& Lewandowsky, 2016; Jern, Chang, \& Kemp, 2014; Powell, Weisman, \& Markman, 2018).
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+Some of the most interesting implications of these models go well beyond
+the probability judgment task itself: the models both support a
+probabilistic account of beliefs (Chater et al. 2020). Indeed, by
+representing the true subjective probabilities as a latent variable, the
+Bayesian modeling approach employed here allows those underlying
+subjective probabilities to be inferred. Examining the model posteriors
+reveals these estimates often come with considerable uncertainty, but at
+least for some participants they can be estimated with useful levels of
+precision. Of course, Zhu and colleagues' (2020) experiments were never
+designed for this purpose. Future research could explore how these
+estimates might be made more reliable, and how inferences about these
+mental probabilities might be integrated with other Bayesian models of
+reasoning (e.g. Jern, Chang, and Kemp 2014; Griffiths and Tenenbaum
+2006; Franke et al. 2016). One particularly promising direction could be
+to integrate these models with formal models of belief revision, which
+could shed new light on fundamental these cognitive processes (e.g.
+Jern, Chang, and Kemp 2014; Powell, Weisman, and Markman 2018; Cook and
+Lewandowsky 2016).
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-anderson1991}{}%
-Anderson, J. R. (1991). The adaptive nature of human categorization. \emph{Psychological Review}, \emph{98}(3), 409--429. \url{https://doi.org/10.1037/0033-295X.98.3.409}
-
 \leavevmode\hypertarget{ref-chater.etal2020}{}%
-Chater, N., Zhu, J.-Q., Spicer, J., Sundh, J., León-Villagrá, P., \& Sanborn, A.
-(2020).
-Probabilistic {Biases Meet} the {Bayesian Brain}.
-\emph{Current Directions in Psychological Science}, \emph{29}(5), 506--512.
-\url{https://doi.org/10.1177/0963721420954801}
+Chater, Nick, Jian-Qiao Zhu, Jake Spicer, Joakim Sundh, Pablo
+León-Villagrá, and Adam Sanborn. 2020. {``Probabilistic {Biases Meet}
+the {Bayesian Brain}.''} \emph{Current Directions in Psychological
+Science} 29 (5): 506--12.
+\url{https://doi.org/10.1177/0963721420954801}.
 
 \leavevmode\hypertarget{ref-cook.lewandowsky2016}{}%
-Cook, J., \& Lewandowsky, S.
-(2016).
-Rational {Irrationality}: {Modeling Climate Change Belief Polarization Using Bayesian Networks}.
-\emph{Topics in Cognitive Science}, \emph{8}(1), 160--179.
-\url{https://doi.org/10.1111/tops.12186}
+Cook, John, and Stephan Lewandowsky. 2016. {``Rational {Irrationality}:
+{Modeling Climate Change Belief Polarization Using Bayesian
+Networks}.''} \emph{Topics in Cognitive Science} 8 (1): 160--79.
+\url{https://doi.org/10.1111/tops.12186}.
 
 \leavevmode\hypertarget{ref-costello.watts2014}{}%
-Costello, F., \& Watts, P.
-(2014).
-Surprisingly rational: {Probability} theory plus noise explains biases in judgment.
-\emph{Psychological Review}, \emph{121}(3), 463--480.
-\url{https://doi.org/10.1037/a0037010}
+Costello, Fintan, and Paul Watts. 2014. {``Surprisingly Rational:
+{Probability} Theory Plus Noise Explains Biases in Judgment.''}
+\emph{Psychological Review} 121 (3): 463--80.
+\url{https://doi.org/10.1037/a0037010}.
 
 \leavevmode\hypertarget{ref-costello.watts2016}{}%
-Costello, F., \& Watts, P.
-(2016).
-People's conditional probability judgments follow probability theory (plus noise).
-\emph{Cognitive Psychology}, \emph{89}, 106--133.
-\url{https://doi.org/10.1016/j.cogpsych.2016.06.006}
+---------. 2016. {``People's Conditional Probability Judgments Follow
+Probability Theory (Plus Noise).''} \emph{Cognitive Psychology} 89
+(September): 106--33.
+\url{https://doi.org/10.1016/j.cogpsych.2016.06.006}.
 
 \leavevmode\hypertarget{ref-costello.watts2017}{}%
-Costello, F., \& Watts, P.
-(2017).
-Explaining {High Conjunction Fallacy Rates}: {The Probability Theory Plus Noise Account}.
-\emph{Journal of Behavioral Decision Making}, \emph{30}(2), 304--321.
-\url{https://doi.org/10.1002/bdm.1936}
+---------. 2017. {``Explaining {High Conjunction Fallacy Rates}: {The
+Probability Theory Plus Noise Account}.''} \emph{Journal of Behavioral
+Decision Making} 30 (2): 304--21.
+\url{https://doi.org/10.1002/bdm.1936}.
 
 \leavevmode\hypertarget{ref-costello.watts2018}{}%
-Costello, F., \& Watts, P. (2018). Invariants in probabilistic reasoning. \emph{Cognitive Psychology}, \emph{100}, 1--16. \url{https://doi.org/10.1016/j.cogpsych.2017.11.003}
+---------. 2018. {``Invariants in Probabilistic Reasoning.''}
+\emph{Cognitive Psychology} 100 (February): 1--16.
+\url{https://doi.org/10.1016/j.cogpsych.2017.11.003}.
 
 \leavevmode\hypertarget{ref-dasgupta.etal2017}{}%
-Dasgupta, I., Schulz, E., \& Gershman, S. J. (2017). Where do hypotheses come from? \emph{Cognitive Psychology}, \emph{96}, 1--25. \url{https://doi.org/10.1016/j.cogpsych.2017.05.001}
-
-\leavevmode\hypertarget{ref-fennell.baddeley2012}{}%
-Fennell, J., \& Baddeley, R.
-(2012).
-Uncertainty plus prior equals rational bias: {An} intuitive {Bayesian} probability weighting function.
-\emph{Psychological Review}, \emph{119}(4), 878--887.
-\url{https://doi.org/10.1037/a0029346}
+Dasgupta, Ishita, Eric Schulz, and Samuel J. Gershman. 2017. {``Where Do
+Hypotheses Come From?''} \emph{Cognitive Psychology} 96 (August): 1--25.
+\url{https://doi.org/10.1016/j.cogpsych.2017.05.001}.
 
 \leavevmode\hypertarget{ref-franke.etal2016}{}%
-Franke, M., Dablander, F., Scholler, A., Bennett, E., Degen, J., Tessler, M.
-H., \ldots{} Goodman, N.
-D.
-(2016).
-What does the crowd believe?
-{A} hierarchical approach to estimating subjective beliefs from empirical data, 6.
-
-\leavevmode\hypertarget{ref-gelman.etal2014a}{}%
-Gelman, A., Carlin, J. B., Stern, H. S., Dunson, D. B., Vehtari, A., \& Rubin, D. B. (2014). \emph{Bayesian data analysis} (Third edition). {Boca Raton}: {CRC Press}.
+Franke, Michael, Fabian Dablander, Anthea Scholler, Erin Bennett, Judith
+Degen, Michael Henry Tessler, Justine Kao, and Noah D Goodman. 2016.
+{``What Does the Crowd Believe? {A} Hierarchical Approach to Estimating
+Subjective Beliefs from Empirical Data,''} 6.
 
 \leavevmode\hypertarget{ref-gelman.etal2014}{}%
-Gelman, A., Hwang, J., \& Vehtari, A.
-(2014).
-Understanding predictive information criteria for {Bayesian} models.
-\emph{Statistics and Computing}, \emph{24}(6), 997--1016.
-\url{https://doi.org/10.1007/s11222-013-9416-2}
+Gelman, Andrew, Jessica Hwang, and Aki Vehtari. 2014. {``Understanding
+Predictive Information Criteria for {Bayesian} Models.''}
+\emph{Statistics and Computing} 24 (6): 997--1016.
+\url{https://doi.org/10.1007/s11222-013-9416-2}.
 
 \leavevmode\hypertarget{ref-griffiths.tenenbaum2006}{}%
-Griffiths, T.
-L., \& Tenenbaum, J.
-B.
-(2006).
-Optimal {Predictions} in {Everyday Cognition}.
-\emph{Psychological Science}, \emph{17}(9), 767--773.
-\url{https://doi.org/10.1111/j.1467-9280.2006.01780.x}
+Griffiths, Thomas L., and Joshua B. Tenenbaum. 2006. {``Optimal
+{Predictions} in {Everyday Cognition}.''} \emph{Psychological Science}
+17 (9): 767--73. \url{https://doi.org/10.1111/j.1467-9280.2006.01780.x}.
 
 \leavevmode\hypertarget{ref-gronau.wagenmakers2019}{}%
-Gronau, Q.
-F., \& Wagenmakers, E.-J.
-(2019).
-Limitations of {Bayesian Leave}-{One}-{Out Cross}-{Validation} for {Model Selection}.
-\emph{Computational Brain \& Behavior}, \emph{2}(1), 1--11.
-\url{https://doi.org/10.1007/s42113-018-0011-7}
-
-\leavevmode\hypertarget{ref-jaynes2003}{}%
-Jaynes, E. T. (2003). \emph{Probability {Theory}: {The Logic} of {Science}}. (G. L. Bretthorst, Ed.). {Cambridge, UNITED KINGDOM}: {Cambridge University Press}.
+Gronau, Quentin F., and Eric-Jan Wagenmakers. 2019. {``Limitations of
+{Bayesian Leave}-{One}-{Out Cross}-{Validation} for {Model
+Selection}.''} \emph{Computational Brain \& Behavior} 2 (1): 1--11.
+\url{https://doi.org/10.1007/s42113-018-0011-7}.
 
 \leavevmode\hypertarget{ref-jern.etal2014}{}%
-Jern, A., Chang, K. K., \& Kemp, C. (2014). Belief polarization is not always irrational. \emph{Psychological Review}, \emph{121}(2), 206--224. \url{https://doi.org/10.1037/a0035941}
-
-\leavevmode\hypertarget{ref-kahneman2013}{}%
-Kahneman, D. (2013). \emph{Thinking, {Fast} and {Slow}} (1st edition). {New York}: {Farrar, Straus and Giroux}.
-
-\leavevmode\hypertarget{ref-kersten.etal2004}{}%
-Kersten, D., Mamassian, P., \& Yuille, A.
-(2004).
-Object {Perception} as {Bayesian Inference}.
-\emph{Annual Review of Psychology}, \emph{55}(1), 271--304.
-\url{https://doi.org/10.1146/annurev.psych.55.090902.142005}
-
-\leavevmode\hypertarget{ref-lu.etal2012}{}%
-Lu, H., Chen, D., \& Holyoak, K. J. (2012). Bayesian analogy with relational transformations. \emph{Psychological Review}, \emph{119}(3), 617--648. \url{https://doi.org/10.1037/a0028719}
-
-\leavevmode\hypertarget{ref-mellers.etal2001}{}%
-Mellers, B., Hertwig, R., \& Kahneman, D.
-(2001).
-Do {Frequency Representations Eliminate Conjunction Effects}?
-{An Exercise} in {Adversarial Collaboration}.
-\emph{Psychological Science}, \emph{12}(4), 269--275.
-\url{https://doi.org/10.1111/1467-9280.00350}
-
-\leavevmode\hypertarget{ref-papaspiliopoulos.etal2007}{}%
-Papaspiliopoulos, O., Roberts, G.
-O., \& Sköld, M.
-(2007).
-A {General Framework} for the {Parametrization} of {Hierarchical Models}.
-\emph{Statistical Science}, \emph{22}(1).
-\url{https://doi.org/10.1214/088342307000000014}
-
-\leavevmode\hypertarget{ref-phan.etal2019}{}%
-Phan, D., Pradhan, N., \& Jankowiak, M.
-(2019).
-Composable {Effects} for {Flexible} and {Accelerated Probabilistic Programming} in {NumPyro}.
-\emph{arXiv:1912.11554 {[}Cs, Stat{]}}.
-Retrieved from \url{http://arxiv.org/abs/1912.11554}
+Jern, Alan, Kai-min K. Chang, and Charles Kemp. 2014. {``Belief
+Polarization Is Not Always Irrational.''} \emph{Psychological Review}
+121 (2): 206--24. \url{https://doi.org/10.1037/a0035941}.
 
 \leavevmode\hypertarget{ref-powell.etal2018}{}%
-Powell, D., Weisman, K., \& Markman, E. M. (2018). Articulating lay theories through graphical models: {A} study of beliefs surrounding vaccination decisions, 6.
-
-\leavevmode\hypertarget{ref-sides.etal2002}{}%
-Sides, A., Osherson, D., Bonini, N., \& Viale, R. (2002). On the reality of the conjunction fallacy. \emph{Memory \& Cognition}, \emph{30}(2), 191--198. \url{https://doi.org/10.3758/BF03195280}
-
-\leavevmode\hypertarget{ref-tenenbaum.etal2011}{}%
-Tenenbaum, J.
-B., Kemp, C., Griffiths, T.
-L., \& Goodman, N.
-D.
-(2011).
-How to {Grow} a {Mind}: {Statistics}, {Structure}, and {Abstraction}.
-\emph{Science}, \emph{331}(6022), 1279--1285.
-\url{https://doi.org/10.1126/science.1192788}
-
-\leavevmode\hypertarget{ref-tversky.kahneman1983}{}%
-Tversky, A., \& Kahneman, D. (1983). Extensional {Versus Intuitive Reasoning}: {The Conjunction Fallacy} in {Probability Judgment}. \emph{Psychological Review}, \emph{90}(4), 23.
-
-\leavevmode\hypertarget{ref-tversky.koehler1994}{}%
-Tversky, A., \& Koehler, D.
-J.
-(1994).
-Support theory: {A} nonextensional representation of subjective probability.
-\emph{Psychological Review}, \emph{101}(4), 547--567.
-https://doi.org/\url{http://dx.doi.org.ezproxy1.lib.asu.edu/10.1037/0033-295X.101.4.547}
-
-\leavevmode\hypertarget{ref-vehtari.etal2017}{}%
-Vehtari, A., Gelman, A., \& Gabry, J.
-(2017).
-Practical {Bayesian} model evaluation using leave-one-out cross-validation and {WAIC}.
-\emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
-\url{https://doi.org/10.1007/s11222-016-9696-4}
+Powell, Derek, Kara Weisman, and Ellen M Markman. 2018. {``Articulating
+Lay Theories Through Graphical Models: {A} Study of Beliefs Surrounding
+Vaccination Decisions,''} 6.
 
 \leavevmode\hypertarget{ref-vehtari.etal2019}{}%
-Vehtari, A., Simpson, D.
-P., Yao, Y., \& Gelman, A.
-(2019).
-Limitations of {``{Limitations} of {Bayesian Leave}-one-out {Cross}-{Validation} for {Model Selection}.''} \emph{Computational Brain \& Behavior}, \emph{2}(1), 22--27.
-\url{https://doi.org/10.1007/s42113-018-0020-6}
-
-\leavevmode\hypertarget{ref-xu.tenenbaum2007}{}%
-Xu, F., \& Tenenbaum, J. B. (2007). Word learning as {Bayesian} inference. \emph{Psychological Review}, \emph{114}(2), 245--272. \url{https://doi.org/10.1037/0033-295X.114.2.245}
+Vehtari, Aki, Daniel P. Simpson, Yuling Yao, and Andrew Gelman. 2019.
+{``Limitations of {`{Limitations} of {Bayesian Leave}-One-Out
+{Cross}-{Validation} for {Model Selection}'}.''} \emph{Computational
+Brain \& Behavior} 2 (1): 22--27.
+\url{https://doi.org/10.1007/s42113-018-0020-6}.
 
 \leavevmode\hypertarget{ref-zhu.etal2020}{}%
-Zhu, J.-Q., Sanborn, A.
-N., \& Chater, N.
-(2020).
-The {Bayesian} sampler: {Generic Bayesian} inference causes incoherence in human probability judgments.
-\emph{Psychological Review}, \emph{127}(5), 719--748.
-\url{https://doi.org/10.1037/rev0000190}
+Zhu, Jian-Qiao, Adam N. Sanborn, and Nick Chater. 2020. {``The
+{Bayesian} Sampler: {Generic Bayesian} Inference Causes Incoherence in
+Human Probability Judgments.''} \emph{Psychological Review} 127 (5):
+719--48. \url{https://doi.org/10.1037/rev0000190}.
 
@@
 
+\bibliographystyle{unsrt}
+\bibliography{references.bib}
+
 \end{document}

### 2021-09-01 model_helpers.py

@@ -2,6 +2,11 @@ import jax.numpy as jnp
 import numpy as np
 import arviz as az
 
+import re
+
+import siuba as s
+from siuba import _
+
 ### ------ Data processing
 
 def load_raw_data(x=1):
@@ -21,12 +26,18 @@ def load_raw_data(x=1):
     return df
 
 
+# def load_data_exp1():
+#     df = load_raw_data(1)
+#     df["condition"] = np.select([df.querydetail.str.contains("icy|frosty")],[0], default=0)
+# 
+#     return df
+
 def load_data_exp1():
     df = load_raw_data(1)
-    df["condition"] = np.select([df.querydetail.str.contains("icy|frosty")],[0], default=0)
+    # df["condition"] = np.select([df.querydetail.str.contains("icy|frosty")],[0], default=0)
+    df = df >> s.mutate(condition = s.if_else(df.querydetail.str.contains("icy|frosty"), 0, 1))
 
     return df
-
 
 def load_data_exp2():
     df = load_raw_data(2)
@@ -47,25 +58,27 @@ def load_data_exp2():
 
     return df
 
-def load_query_avg_data_exp2():
-
-    df = (load_data_exp2()
-          .groupby(["ID","condition","querytype","querydetail"], as_index=False)
-          .agg({"estimate":"mean"})
-          .assign(estimate = lambda x: x.estimate.replace({0:.01, 1:.99}))
-         )
-    
-    return df
 
 def load_query_avg_data_exp1():
+    df = (
+        load_data_exp1() >>
+        s.group_by("ID","condition","querytype", "querydetail") >>
+        s.summarize(estimate = np.mean(_.estimate)) >>
+        s.mutate(estimate =  _.estimate.replace({0:.01, 1:.99}))
+    )
 
-    df = (load_data_exp1()
-          .groupby(["ID", "condition", "querytype", "querydetail"], as_index=False)
-          .agg({"estimate":"mean"})
-          .assign(estimate = lambda x: x.estimate.replace({0:.01, 1:.99}))
-         )
-    
     return df
+
+
+def load_query_avg_data_exp2():
+    df = (
+        load_data_exp2() >>
+        s.group_by("ID","condition","querytype", "querydetail") >>
+        s.summarize(estimate = np.mean(_.estimate)) >>
+        s.mutate(estimate =  _.estimate.replace({0:.01, 1:.99}))
+    )
+
+    return df    
 
 
 def load_query_avg_data_all():

### 2021-09-02 bayes-sampler-prelim.Rmd
added plate diagrams

--- 2021-03-17 bayes-sampler-prelim.Rmd
@@
 library(purrr)
-library(rstan)
 ```
@@
 df_exp2 %>% 
-  group_by(ID, querydetail, querytype) %>% 
-  summarize(estimate=mean(estimate)) %>% 
+  filter(block==1) %>% 
   # filter(ID==1, block==2,)
@@
   ungroup() %>% 
-  ggplot(aes(x=querytype, y=M, ymin=M-se*2, ymax=M+se*2)) +
+  ggplot(aes(x=querytype, y=M, ymin=M-se, ymax=M+se)) +
   geom_bar(stat="identity", position="dodge", width = .5, fill="grey50") +
@@
 ```
-# The Bayesian sampler model in Stan
 
-Bayesian sampler model is written in equation 15 as:
+bayesian sampler model is written in equation 15 as:
 
 $$E[P_{BS}(A)] = P(A)\frac{N}{N + 2\beta} + \frac{\beta}{N+2\beta}$$
+
 Which we should expect would be Beta distributed, given that it is itself a probability.
@@
 in a 4-term conditional probability table the probabilities must sum to one, so there are only 3 degrees of freedom.
-(I think I can use a simplex variable in Stan to accomplish this.)
 
-The actual observation is:
+I think I can use a simplex variable in Stan to accomplish this.
 
-$$P_{BS}(A) = prop(A)\frac{N}{N + 2\beta} + \frac{\beta}{N+2\beta}$$
-Where $prop(A)$ is the proportion of N samples that were A in the mental simulation.
-
-To implement in a bayesian way we should go back to the original beta model:
-
-$$P_{BS}(A) \sim Beta(\beta + S(A), \beta+F(A))$$
-Which we can rewrite as:
-
-$$P_{BS}(A) \sim Beta(\beta + \lambda(A)N, \beta+(1-\lambda(A))N)$$
-Where $\lambda(A)$ is the proportion of successes in the mental simulation and is a function of the true underlying model probability, $p(A)$
-
-$$\lambda(A) \sim Beta(p(A)*N, (1-p(A))*N)$$
-To which we assign a uniform prior (or the multidimensional dirichlet equivalent):
-$$p(A) \sim Beta(1,1)$$
-They aren't looking to make an actual statistical model of the responses, so they essentially make use o the fact that the expectation of $\lambda(A)$ is $p(A)$, to ignore this bit when they are just trying to estimate the expectation.
-
-This model might not sample well, as N and $\beta$ are not uniquely identifiable. But let's try anyhow.
-
-```{r}
-# first, code to create condition codes. This will be a pain :(
-
-df_test <- df_exp2 %>% 
-  group_by(ID, querytype, querydetail) %>% 
-  summarize(estimate=mean(estimate)) %>%
-  filter(ID==103) %>%
-  # filter(ID==4, block==3) %>% 
-  filter(grepl("(windy|cloudy)", querydetail)) %>% 
-  # filter(querytype %in% c("AandB","AandnotB","notAandB","notAandnotB")) %>%
-  ungroup() %>% 
-  mutate(
-    dummy = map(querytype, ~case_when(
-      .x == "AandB" ~ list(1,0,0,0),
-      .x == "AandnotB" ~ list(0,1,0,0),
-      .x == "notAandB" ~ list(0,0,1,0),
-      .x == "notAandnotB" ~ list(0,0,0,1),
-      .x == "A" ~ list(1,1,0,0),
-      .x == "B" ~ list(1,0,1,0),
-      .x == "notA" ~ list(0,0,1,1),
-      .x == "notB" ~ list(0,1,0,1),
-      TRUE ~ list(0,0,0,0)
-    )
-  )
-  ) %>% 
-  filter(!grepl("g",querytype))
-
-df_sim <- df_test %>% 
-  mutate(
-    AandBp = .3,
-    AandnotBp = .1,
-    notAandBp = .2,
-    notAandnotBp = .4,
-    subj_p = pmap_dbl(list(AandBp, AandnotBp, notAandBp, notAandnotBp, dummy), ~AandBp*dummy[[1]])
-  )
-
-standata <- list(y = df_test$estimate, x = df_test$dummy, N = nrow(df_test), k = 4)
-```
-
-```{stan output.var="model"}
-// psuedostancode
-data {
-  int<lower=0> N;
-  int<lower=0> K;
-  vector[N] x[K];
-  vector[N] y
-  vector[K] alpha;
-}
-parameters {
-
-  real beta;
-  real<lower=0,upper=1> Nsamp;
-  real p;
-  simplex[K] theta[N];
-}
-model {
-  beta ~ cauchy(0, 1); // prior on beta, constrain positive
-  Nsamp ~ cauchy(0, 5); // prior on N, constrain positive
-  for (n in 1:N) {
-    theta[n] ~ dirichlet(alpha);
-    p = theta[n]*x; // later add code to get condition combinations etc.
-  }
-
-  lambda ~ beta(p*Nsamp, (1-p)*Nsamp);
-  y_hat ~ beta(beta + lambda*Nsamp, beta + (1-lambda)*Nsamp); // does not account for any other kinds of error
-}
-
-```
-
-# other diagnostics and stuff
-
-I'm not sure how to implement the PT+N model in this fashion. For simple events, the bayesian sampler essentially collapses into the PT+N model (eq. 10):
+Meanwhile the PT+N model has a bunch of different equations depending on the type of probability query. For a simple marginal probability it is (per equation 10):
 
 $$E[P_{PT+N}(A)] = (1-2d)P(A)+d$$
-where $d = \frac{\beta}{N+2\beta} $.
-
-The two models diverge only for conditional probabilities, $P(A|B)$.
-The Bayesian sampler treats these just like any other probability.
-The PT+N model posits a two-stage estimation algorithm with noise at both stages.
-It produces a somewhat nastier looking equation (13) that is a generalization of the basic equation (simplifies down to it when p(B)=1).
-
-But accounting for individual-level noise will work differently.
-
-## Simulating data
-
-First I'll need to simulate some data. I'll start simulating for one participant, and I'll try to use the format of the final data as much as I can so things translate.
-
-```{r}
-## what does the sampler do to probabilities?
-
-bayesian_sampler <- function(p,N,B){
-  p*(N/(N+2*B)) + B/(N+2*B)
-}
-
-inv_bayesian_sampler <- function(est,N,B){
-  (est - B/(N+2*B))/(N/(N+2*B))
-}
-
-x <- c(.77,.67,.5,.49)
-
-norm_x <- x/sum(x)
-bayesian_sampler(x, 10, 2) # median values
-inv_bayesian_sampler(bayesian_sampler(x, 10, 2), 10, 2) # median values
-
-sum(bayesian_sampler(c(.25,.25,.25,.25), 1, 100)) # extreme
-
-```
-
-There's really no way to get the kinds of raw responses participants are giving just via the bayesian sampler distortions.
-For instance, the sum of the BS estimates will never be over 2, but you can observe cases where that's true in the data.
-You also get cases where probabilities for one cell are over .5 and all sum to near 2 or above.
-An unbiased beta prior will always push estimates toward .5, so it can't explain these cases.
-Here's one example:
-
-```{r}
-# looking at the real data
-
-df_exp2 %>% 
-  group_by(ID, querytype, querydetail) %>% 
-  summarize(estimate=mean(estimate)) %>%
-  filter(ID==103) %>%
-  # filter(ID==4, block==3) %>% 
-  filter(grepl("(windy|cloudy)", querydetail)) %>% 
-  filter(querytype %in% c("AandB","AandnotB","notAandB","notAandnotB")) %>%
-  # filter(querytype %in% c("A","notA")) %>% 
-  select(querytype, estimate)
-```
-
-Of course we could just normalize both according to their direct contrasts (e.g.
-A and ~A).
-If we do that the esimates from the joint distribution estimates are only modestly correlated with the marginal estimates (r ~ .3 to .5).
-
-```{r message=FALSE}
-df_exp2 %>% 
-  group_by(ID, querytype, querydetail) %>% 
-  summarize(estimate=mean(estimate)/100) %>%
-  # filter(ID==1) %>%
-  # filter(ID==4, block==3) %>% 
-  filter(grepl("(windy|cloudy)", querydetail)) %>%
-  select(-querydetail) %>% 
-  spread(querytype,estimate) %>% 
-  mutate(base1 = AandB + AandnotB + notAandB + notAandnotB) %>% 
-  mutate(base2 = notA + A) %>%
-  mutate(base3 = notB + B) %>%
-  mutate_at(vars(-ID, -base1, base2, base3, -A, -B, -notA, -notB), ~./base1) %>% 
-  mutate_at(vars(A, notA), ~./base2) %>% 
-  mutate_at(vars(B, notB), ~./base3) %>% 
-  mutate(
-    est_A = AandB + AandnotB, 
-    est_B = AandB + notAandB, 
-    est_notA = notAandB + notAandnotB,
-    est_notB = AandnotB + notAandnotB
-  ) %>% 
-  ungroup() %>% 
-  select(A, B, notA, notB, est_A, est_B, est_notA, est_notB) %>% 
-  cor() %>% 
-  round(3)
-```
-
-So the participant-level estimates are pretty messy.
-The Bayesian sampler model is really only able to generate accurate predictions at the population level.
-But I want to measure at the participant level!
-
-A lot of Ps are saying that the conjunction AandB is more probable than either A or B?
-Over half it looks like.
-Say what you will about all the other things, there's really no story of probabilities that says this is an acceptable pattern of responses.
-A similar things is observed for A or B, where about half of the time the disjunction is rated less probable than one of its disjuncts.
-
-On p7 the PT+N model gets this by augmenting the model to say that conjunction/disjunction probabilities are noisier, which pushes them toward .50.
-This could explain the conjunction fallacy, but only when the subjective probability of the is below .50, in which case it will be pushed toward .50.
-But we can see equally many cases where the observed estimate is above .50, which would suggest the subjective conjunction probability is even higher!
-
-```{r}
-df_exp2 %>% 
-  group_by(ID, querytype, querydetail) %>% 
-  summarize(estimate=mean(estimate)) %>%
-  # filter(ID==103) %>%
-  # filter(ID==4, block==3) %>% 
-  filter(grepl("(windy|cloudy)", querydetail)) %>% 
-  filter(querytype %in% c("A","B", "AandB","AandnotB","notAandB","notAandnotB", "AorB")) %>%
-  # filter(querytype %in% c("A","notA")) %>% 
-  select(querytype, estimate) %>% 
-  spread(querytype, estimate) %>% 
-  mutate(
-    minAB = map2_dbl(A, B, ~min(.x,.y)),
-    maxAB = map2_dbl(A, B, ~max(.x,.y))
-    ) %>% 
-  ggplot(aes(x=minAB, y = AandB)) + 
-  geom_point() +
-  geom_abline(slope=1,intercept=0) +
-  geom_vline(xintercept=50, linetype="dashed") +
-  theme_bw()
-```
-
-Ok so that seems bad but it's just more affirmation that this is not a model of individual probability ratings, but only population averages.
-The individual ratings are still perturbed by substantial noise that could make them do all kinds of things.
-
-But it seems like the __amount of error__ at the individual level should be related to the sampling/noise properties.
-If my estimate is based only on a handful of samples, I'll give more extreme values that might deviate from the expectation more.
-So the sampling/noise process induces both a population-level bias and individual trial/block-level variance.
-Seems like that individual variance should be part of what motivates the model.
-But at the same time it doesn't really fly to just essentially downweight some participants b/c their responses are noisier.
-Or does it?
-Would also create trouble for the conjunction/disjunction trials b/c they would effectively contribute less to fit.
-
-```{r}
-# do that, can create a coding system to compute
-# A = AandB + AandnotB
-# AorB = AandB + AandnotB + notAandB + 0*notAandnotB
-# AgB = AandB + AandnotB / (AandB + notAandB )
-
-## like so ...
-# x = (c1*p_j1 + c2*p_j2 + c3*p_j3 + c4*p_j4)/(g1*p_j1 + g2*p_j2 + g3*p_j3 + g4*p_j4)
-
-# and can be turned into a matrix operation later
-
-# x = (cvec*pvec)*(gvec*pvec)^-1
-
-# then fit one real participant
-
-# then expand to hierarchical version
-
-# then fit it all!
-```
+Which is likewise a nice linear form but without any variables directly observed. The other equations are given in equations 11-13.
 
@@
 # uninformative prior on unobserved true probabilities (simplex)
-
-# parameters {
-# simplex[K] theta[N];
-# ...
-# }
-# model {
-#   beta ~ halfnormal(0,1) // prior on beta, constrain positive
-#   N ~ cauchy(0, 5) // prior on N, constrain positive
-#   k ~ cauchy(0, 5) // prior on kappa constant, constrain to be positive
-#   theta ~ dirichlet(something) // latent probabilities, possibly need a loop N times
-#   x = // computed from theta[i] and trial data or matrix mult
-#   d = beta/(N+2*beta)
-#   mu = (1-2*d)*x + d
-#   y ~ beta(mu*k, (1-mu)*k);
-#   ...
-# }
 
@@
 
-This is making me wonder if you actually do need to do nested inference if you want to get meaningful individual-level predictions.
+## Simulating data
+
+First I'll need to simulate some data. I'll start simulating for one participant, and I'll try to use the format of the final data as much as I can so things translate.
+
+```{r}
+# do that
+
+# then fit one real participant
+
+# then expand to hierarchical version
+
+# then fit it all!
+```

### 2021-09-02 paper-rmd/paper-papaja.Rmd
added plate diagrams

@@
 output            : papaja::apa6_pdf
+header-includes   :
+  - \usepackage{tikz}
+  - \usetikzlibrary{bayesnet}
 ---
@@
   kbl(caption="Original model fitting results", digits=1, booktabs=T) %>% 
-  kable_classic(full_width=F)
+  kable_classic(full_width=F) %>% 
+  row_spec(1, bold=TRUE)
 ```
@@
 
-$$P_{e}(A) = (1-2d)P(A) + d$$
-$$P_e(A\land B) = (1-d^\prime)P(A \land B)+d^\prime$$
-$$P_e(A\lor B) = (1-d^\prime)P(A \lor B)+d^\prime$$
-$$P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}$$
+\begin{align*}
+  P_{e}(A) &= (1-2d)P(A) + d \\
+  P_e(A\land B) &= (1-d^\prime)P(A \land B)+d^\prime \\
+  P_e(A\lor B) &= (1-d^\prime)P(A \lor B)+d^\prime \\
+  P_e(A|B) &= \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}
+\end{align*}
 
@@
 
-$$P_{e}(A) = \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta}$$
-$$P_{e}(A \land B) = \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta}$$
-$$P_{e}(A \lor B) = \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta}$$
+\begin{align*}
+  P_{e}(A) &= \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta} \\
+  P_{e}(A \land B) &= \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta} \\
+  P_{e}(A \lor B) &= \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta} 
+\end{align*}
 
@@
 
-$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) $$
-$$ d_j = \frac{expit(\delta_j)}{2}$$
-$$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
-$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
-$$ \delta_j \sim Normal(0,1)$$
-$$ \Delta\delta_j \sim Normal(0,1)$$
-$$ k \sim HalfCauchy(20)$$
+\begin{figure}
+  \begin{minipage}{.5\textwidth}
+    \flushright
+    \input{ptn_platediagram}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \flushleft
+    \small
+    \begin{align*}
+      y_{ij} &\sim beta\big(\mu_{ij} \kappa, (1-\mu_{ij}) \kappa \big) \\
+      \mu_{ij} &= f_{PT+N}(\overrightarrow{p_i}, d_j, d_j)  \\
+      d_j &= \frac{expit(\delta_j)}{2} \\
+      d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2} \\
+      \overrightarrow{p_i} &\sim Dirichlet(\overrightarrow{1}) \\
+      \delta_j &\sim Normal(0,1) \\
+      \Delta\delta_j &\sim Normal(0,1) \\
+      \kappa &\sim HalfCauchy(20)
+    \end{align*}
+  \end{minipage}
+  \caption{Complex PT+N model diagram and formula specifications.}
+\end{figure}
 
@@
 
-The Bayesian Sampler model is therefore:
+The Bayesian Sampler model is therefore identical to the PT+N model save for the following changes to $\mu_{ijk}$, $d$, and $d'$, as shown below:
 
-$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) $$
-$$ d_j = \frac{expit(\delta_j)}{2}$$
-$$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
-$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
-$$ \delta_j \sim Normal(0,1)$$
-$$ \Delta\delta_j \sim Normal(0,1)$$
-$$ k \sim HalfCauchy(20)$$
+\begin{align*}
+  \mu_{ij} &= f_{BS}(\overrightarrow{p_i}, d_j, d_j)  \\
+  d_j &= \frac{expit(\delta_j)}{3} \\
+  d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{3} \\
+\end{align*}
 
@@
 
-$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) $$
-$$d_j = \frac{ expit(\delta_j )}{3}$$
-$$ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}$$
-$$ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})$$
-$$ \delta_{pop} \sim Normal(-1, 1)$$
-$$ \Delta\delta_{pop} \sim Normal(0, .50)$$
-$$ log(\sigma_{\delta}) \sim Normal(-1, 1)$$
-$$ log(\sigma_{\Delta\delta}) \sim Normal(-1, 1)$$
-$$\delta_j \sim Normal(\delta_{pop}, \sigma_{\delta})$$
-$$\Delta\delta_j \sim Normal(\Delta\delta_{pop}, \sigma_{\Delta\delta})$$
-$$k \sim HalfCauchy(20)$$
+\begin{figure}
+  \begin{minipage}{.5\textwidth}
+    \flushright
+    \input{bs_platediagram_mlm}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \flushleft
+    \small
+    \input{bs_equation_mlm}
+  \end{minipage}
+  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications.}
+\end{figure}
 
@@
   kable_classic(full_width=F) %>% 
-  kable_styling(latex_options="scale_down")
+  kable_styling(latex_options="scale_down") %>% 
+  row_spec(1, bold=TRUE)
 ```

### 2021-09-02 paper-rmd/paper-papaja.tex
added plate diagrams

@@
   man,floatsintext]{apa6}
-\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
+\usepackage{amsmath}
 \usepackage{ifxetex,ifluatex}
@@
   \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \usepackage{csquotes}
+\usepackage{tikz}
+\usetikzlibrary{bayesnet}
 \ifxetex
@@
 \else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
+  \usepackage[shorthands=off,main=english]{babel}
 \fi
@@
 \midrule
-Bayesian Sampler simple & 956.9 & -5371.9\\
+\textbf{Bayesian Sampler simple} & \textbf{956.9} & \textbf{-5371.9}\\
 Bayesian Sampler complex & 1174.4 & -5099.3\\
@@
 
-\[P_{e}(A) = (1-2d)P(A) + d\]
-\[P_e(A\land B) = (1-d^\prime)P(A \land B)+d^\prime\]
-\[P_e(A\lor B) = (1-d^\prime)P(A \lor B)+d^\prime\]
-\[P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}\]
+\begin{align*}
+  P_{e}(A) &= (1-2d)P(A) + d \\
+  P_e(A\land B) &= (1-d^\prime)P(A \land B)+d^\prime \\
+  P_e(A\lor B) &= (1-d^\prime)P(A \lor B)+d^\prime \\
+  P_e(A|B) &= \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}
+\end{align*}
 
@@
 
-\[P_{e}(A) = \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta}\]
-\[P_{e}(A \land B) = \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta}\]
-\[P_{e}(A \lor B) = \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta}\]
+\begin{align*}
+  P_{e}(A) &= \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta} \\
+  P_{e}(A \land B) &= \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta} \\
+  P_{e}(A \lor B) &= \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta} 
+\end{align*}
 
@@
 
-\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{PT+N}(\overrightarrow{p_i}, d_j, d_j) \]
-\[ d_j = \frac{expit(\delta_j)}{2}\]
-\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
-\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_j \sim Normal(0,1)\]
-\[ \Delta\delta_j \sim Normal(0,1)\]
-\[ k \sim HalfCauchy(20)\]
+\begin{figure}
+  \begin{minipage}{.5\textwidth}
+    \flushright
+    \input{ptn_platediagram}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \flushleft
+    \small
+    \begin{align*}
+      y_{ij} &\sim beta\big(\mu_{ij} \kappa, (1-\mu_{ij}) \kappa \big) \\
+      \mu_{ij} &= f_{PT+N}(\overrightarrow{p_i}, d_j, d_j)  \\
+      d_j &= \frac{expit(\delta_j)}{2} \\
+      d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2} \\
+      \overrightarrow{p_i} &\sim Dirichlet(\overrightarrow{1}) \\
+      \delta_j &\sim Normal(0,1) \\
+      \Delta\delta_j &\sim Normal(0,1) \\
+      \kappa &\sim HalfCauchy(20)
+    \end{align*}
+  \end{minipage}
+  \caption{Complex PT+N model diagram and formula specifications.}
+\end{figure}
 
@@
 
-The Bayesian Sampler model is therefore:
+The Bayesian Sampler model is therefore identical to the PT+N model save for the following changes to \(\mu_{ijk}\), \(d\), and \(d'\), as shown below:
 
-\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) \]
-\[ d_j = \frac{expit(\delta_j)}{2}\]
-\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
-\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_j \sim Normal(0,1)\]
-\[ \Delta\delta_j \sim Normal(0,1)\]
-\[ k \sim HalfCauchy(20)\]
+\begin{align*}
+  \mu_{ij} &= f_{BS}(\overrightarrow{p_i}, d_j, d_j)  \\
+  d_j &= \frac{expit(\delta_j)}{3} \\
+  d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{3} \\
+\end{align*}
 
@@
 
-\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{BS}(\overrightarrow{p_i}, d_j, d_j’) \]
-\[d_j = \frac{ expit(\delta_j )}{3}\]
-\[ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}\]
-\[ \overrightarrow{p_i} \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_{pop} \sim Normal(-1, 1)\]
-\[ \Delta\delta_{pop} \sim Normal(0, .50)\]
-\[ log(\sigma_{\delta}) \sim Normal(-1, 1)\]
-\[ log(\sigma_{\Delta\delta}) \sim Normal(-1, 1)\]
-\[\delta_j \sim Normal(\delta_{pop}, \sigma_{\delta})\]
-\[\Delta\delta_j \sim Normal(\Delta\delta_{pop}, \sigma_{\Delta\delta})\]
-\[k \sim HalfCauchy(20)\]
+\begin{figure}
+  \begin{minipage}{.5\textwidth}
+    \flushright
+    \input{bs_platediagram_mlm}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \flushleft
+    \small
+    \input{bs_equation_mlm}
+  \end{minipage}
+  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications.}
+\end{figure}
 
@@
 \midrule
-Unrestricted BS MLM & -1178.1 & 149.2 & 0.725 & 0.821 & -3953.1 & 368.7 & 0.688 & 0.878\\
+\textbf{Unrestricted BS MLM} & \textbf{-1178.1} & \textbf{149.2} & \textbf{0.725} & \textbf{0.821} & \textbf{-3953.1} & \textbf{368.7} & \textbf{0.688} & \textbf{0.878}\\
 Bayesian Sampler complex MLM & -1167.7 & 145.6 & 0.723 & 0.816 & -3815.4 & 399.9 & 0.675 & 0.852\\

### 2021-09-02 paper-rmd/bs_platediagram_mlm.tex
added diag tex

\begin{tikzpicture}
      % Define nodes
      \node[obs]                               (y) {$y_{ijk}$};
      \node[det, above=of y] (mu) {$\mu_{ijk}$};
      \node[det, above=of mu, xshift=-1.2cm]  (d) {$d_j$};
      \node[det, above=of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
      \node[latent, above=of d]  (delta) {$\delta_j$};
      \node[latent, above=of d_prime]  (delta_delta) {$\Delta\delta_j$};
      \node[latent, right=2cm of y]  (kappa) {$\kappa$};
      \node[latent, left=1cm of mu]            (p) {$\vec{p_{jk}}$};
      \node[latent, above=of delta, xshift=-.5cm]  (delta_pop) {$\overline{\delta}$};
      \node[latent, above=of delta_delta, xshift=-.5cm]  (delta_delta_pop) {$\overline{\Delta\delta}$};
      \node[latent, above=of delta, xshift=.5cm]  (sigma_delta) {$\sigma_{\delta}$};
      \node[latent, above=of delta_delta, xshift=.5cm]  (sigma_delta_delta) {$\sigma_{\Delta\delta}$};
      % Connect the nodes
      \edge {mu, kappa} {y} ; %
      \edge {d, d_prime} {mu}
      \edge {delta} {d, d_prime}
      \edge {delta_delta} {d_prime}
      \edge {p} {mu}
      \edge {delta_pop} {delta}
      \edge {delta_delta_pop} {delta_delta}
      \edge {sigma_delta} {delta}
      \edge {sigma_delta_delta} {delta_delta}

      % Plates
      {
        \tikzset{plate caption/.append style={below=15pt of #1.south east}};
        \tikzstyle{plateIJK} = [align=center];
        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$ijk$} ;
     }
      {
        \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)} {$j$} ;
      }
      {
        \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$jk$};
      }

\end{tikzpicture}

### 2021-09-02 paper-rmd/ptn_platediagram.tex
added diag tex

--- 2021-09-02 paper-rmd/bs_platediagram_mlm.tex
@@
 \begin{tikzpicture}
+
      % Define nodes
@@
      \node[latent, left=1cm of mu]            (p) {$\vec{p_{jk}}$};
-      \node[latent, above=of delta, xshift=-.5cm]  (delta_pop) {$\overline{\delta}$};
-      \node[latent, above=of delta_delta, xshift=-.5cm]  (delta_delta_pop) {$\overline{\Delta\delta}$};
-      \node[latent, above=of delta, xshift=.5cm]  (sigma_delta) {$\sigma_{\delta}$};
-      \node[latent, above=of delta_delta, xshift=.5cm]  (sigma_delta_delta) {$\sigma_{\Delta\delta}$};
+
      % Connect the nodes
@@
      \edge {p} {mu}
-      \edge {delta_pop} {delta}
-      \edge {delta_delta_pop} {delta_delta}
-      \edge {sigma_delta} {delta}
-      \edge {sigma_delta_delta} {delta_delta}
 
@@
 
+   %   \plate {} {(mu)(y)(p)(mup.north west)(mup.south west)} {$M$} ;
+
    \end{tikzpicture}

### 2021-09-07 paper-rmd/paper-papaja.Rmd
added figs

@@
 ---
-title             : "Untitled probability judgment paper draft"
-shorttitle        : "DRAFT"
+title             : "Comparing probabilistic accounts of probability judgments"
+shorttitle        : "Comparing probability judgment accounts"
 
@@
 abstract: |
-  Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic.
-However, people’s explicit probability judgments often incoherently violate the laws of probability.
-Two recent proposals, the “Probability Theory plus Noise” [@costello.watts2014] and “Bayesian Sampler” [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
-The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity.
-Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the $LOO_{ic}$ information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020).
-My findings support the Bayesian Sampler’s account of conditional probability judgments, and I find the data are best explained by a “hybrid” model that can be seen as a modification of both theories.
+  Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic.
+However, one mark sometimes held against these theories is that people’s explicit probability judgments often violate the laws of probability.
+Two recent proposals, the "Probability Theory plus Noise" [@costello.watts2014] and "Bayesian Sampler" [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
+These theories fit quite differently into the larger project of Bayesian cognitive science, but their many similarities complicate comparisons of their predictive accuracy.
+In particular, comparing the models demands a careful accounting of model complexity.
+Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
+Compared the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a version of the Bayesian Sampler model that allows for informative priors.
 
@@
 
-Bayesian theories and approaches offer a unified formal framework for cognitive science [@tenenbaum.etal2011] with remarkable successes across domains, including perception [e.g.
+Bayesian theories and approaches offer a unified formal framework for cognitive science [@tenenbaum.etal2011] that has had remarkable successes across domains, including in perception [e.g.
 @kersten.etal2004], memory [@anderson1991], language [@xu.tenenbaum2007], and reasoning [@lu.etal2012].
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", “vaccines are safe,” or “this politician is trustworthy”.
+It is somewhat embarrassing then, that that these theories have often been accused of failing to describe human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", "vaccines are safe," or "this politician is trustworthy".
 
@@
 Unfortunately, people’s explicit probability judgments routinely violate the axioms of probability theory.
-For example, human probability judgments often exhibit the “conjunction fallacy”: people will often judge the conjunction of two events (e.g.
-“Tom Brady likes football and miniature horses”) as being more probable than one of the events in isolation (e.g.
-“Tom Brady likes miniature horses”), a plain and flagrant violation of probability theory [@tversky.kahneman1983].
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or “unpacking” effects [@tversky.koehler1994], and a variety of other effects illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013].
+For example, human probability judgments often exhibit the "conjunction fallacy": people will often judge the conjunction of two events (e.g.
+"Tom Brady likes football and miniature horses") as being more probable than one of the events in isolation (e.g.
+"Tom Brady likes miniature horses"), a plain and flagrant violation of probability theory [@tversky.kahneman1983].
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or "unpacking" effects [@tversky.koehler1994], and a variety of other effects illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013].
 Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
@@
 Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic [@costello.watts2014; @zhu.etal2020].
-Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by drawing a limited number of samples from probabilistic mental models [see also @chater.etal2020; @dasgupta.etal2017].
+Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental "sampling" from a probabilistic mental model [see also @chater.etal2020; @dasgupta.etal2017].
 
@@
 
-Costello & Watts [-@costello.watts2014; -@costello.watts2016; -@costello.watts2018] proposed a theory of probability judgment they call the “Probability Theory plus Noise” theory (PT+N).
-In the PT+N model, mental “samples” are drawn from a probabilistic mental model of events and are then “read” with noise, so that some positive examples will be read as negative and some negative read as positive.
-This results in probability judgments that reflect probabilistic credences perturbed by noise.
-Under the simplest form of the PT+N model, the expected value of probability judgments is:
+Costello & Watts [-@costello.watts2014; -@costello.watts2016; -@costello.watts2018] have proposed a theory of probability judgment they call the "Probability Theory plus Noise" theory (PT+N).
+In the PT+N model, mental "samples" are drawn from a probabilistic mental model of events and are then "read" with noise, so that some positive examples will be read as negative and some negative examples read as positive.
+The results are probability judgments reflecting probabilistic credences perturbed by noise.
+In their model, the probability a mental sample for an event $A$ is read as A is the probability that the sample truly is A, p(A), and that it is correctly read (1-d), plus the probability that the sample is not A, 1-P(A) and that it is incorrectly read (d), or:
+
+\begin{align*}
+  P(\text{read as A}) &= (1-d)P(A) + d(1-P(A)) \\
+  &= (1-2d)P(A) + d
+\end{align*}
+
+Thus under the simplest form of the PT+N model, the expected value of probability judgments is:
 
@@
 
-where *d* represents the probability with which samples will be misread, or the amount of noise in the judgment process (where by assumption a maximum of 50% of samples can be misread, so that d is a number in the range [0, .50]).
+where *d* represents the probability with which samples will be misread, or the amount of noise in the judgment process.
+By assumption, a maximum of 50% of samples can be misread, so that d is a number in the range [0, .50].
 The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
 
-Meanwhile, Zhu, Sanborn, & Chater [-@zhu.etal2020] have proposed a Bayesian model of probability judgment they call the “Bayesian Sampler.” Under this model, probability judgment is itself seen as a process of Bayesian inference.
+Meanwhile, Zhu, Sanborn, & Chater [-@zhu.etal2020] have proposed a Bayesian model of probability judgment they call the "Bayesian Sampler." Under this model, probability judgment is itself seen as a process of Bayesian inference.
 To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event.
-Then, those “observed” samples are integrated with a prior over probabilities to produce a probability judgment.
-The effect of this prior is to regularize the judgments, pushing them away from extreme values.
-They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g.
-encourage taking large risks based on a small number of mental samples.
-However, this hedging comes at the cost of systematic incoherence and biases.
-Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts [@costello.watts2014; @costello.watts2016].
+Then, those "observed" samples are integrated with a prior over probabilities to produce a probability judgment.
+The $N$ mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric Beta distribution parameterized by $\beta$.
+After observing $S(A)$ successes and $N - S(A)$ failures, this posterior is distributed $Beta(\beta + S(A), \beta + N - S(A))$.
+Zhu and colleagues [-@zhu.etal2020] assume that people report the mean of their posterior probability estimates.
+Since for any Beta distribution $x \sim Beta(a,b)$, $E[x] = \frac{a}{a+b}$, the probability estimate is thus a linear function of S, N, and $\beta$.
 
-Under the simplest version of the Bayesian Sampler model, the expected value of a probability judgment is given by the following formula:
+$$\hat{P}_{BS}(A) = \frac{S(A)}{N+2\beta} + \frac{\beta}{N+2\beta}$$
+
+The expected value of the estimate can then be written in terms of the expected number of successes, or $P(A) \cdot N$.
+Under the simplest version of the Bayesian Sampler model, this gives the following formula:
 
@@
 
-where *N* represents the number of mental samples drawn.
-The mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric beta distribution parameterized by $\beta$.
-Zhu and colleagues [-@zhu.etal2020] show that the *N* and $\beta$ parameters of their model can be related to the *d* parameter of the PT+N model via the following bridging formula:
+Like the PT+N model, the Bayesian Sampler model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts [@costello.watts2014; @costello.watts2016].
+In fact, important equivalencies can be drawn between the two models.
+Zhu and colleagues [-@zhu.etal2020] show that the $N$ and $\beta$ parameters of their model can be related to the $d$ parameter of the PT+N model via the following bridging formula:
 
 $$d = \frac{\beta}{N+2\beta}$$
+
+Thus, the effect of a Bayesian prior is nearly identical to the effect of noise in the PT+N model.
+But, rather than merely perturbing people's probability judgments, this prior can be seen as regularizing these judgments away from extreme values.
+Zhu and colleagues [-@zhu.etal2020] argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
+For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, a person incorporating no other information might conclude these are all certain victories and thereby be too willing to assume risk.
+A regularizing prior pushes these estimates away from extremes, thereby promoting better decision-making when mental samples are sparse.
+However, this hedging comes at the cost of systematic incoherence and biases.
 
@@
 By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs [@chater.etal2020].
-However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project.
+And the bridging condition identified by Zhu and colleagues [@zhu.etal2020] demonstrates that the theories make identical predictions in many cases.
+However, the two models diverge substantially in their treatment of conditional probability judgments.
 Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence.
-So, the conditioning of beliefs is at the heart of these theories.
+So, how they treat the conditioning of beliefs is at the heart of these theories.
 
@@
 
-$$ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} $$
+$$P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}$$
 
@@
 
-Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello & Watts’ (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments.
+Zhu, Sanborn, and Chater [-@zhu.etal2020] compared their Bayesian Sampler model against Costello & Watts’ (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments.
 Unfortunately, their results were somewhat equivocal.
-They fit both “simple” and “complex” versions of each model and computed a Bayesian Information Criteria score (BIC) for all models.
+They fit both "simple" and "complex" versions of each model and computed a Bayesian Information Criteria score (BIC) for all models.
 The complex versions of these models introduce additional parameters d’ and N’ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are needed in order to explain conjunction fallacies [@costello.watts2017; @zhu.etal2020] .
+These additional parameters are crucial to both models’ explanations of conjunction fallacies---key findings in the probability judgment literature [@costello.watts2017; @zhu.etal2020].
 Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs [@zhu.etal2020, Supplementary materials].
@@
 Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues (2020).
-So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
-For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.
+But conjunction fallacies are an extremely robust empirical finding [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues [-@zhu.etal2020].
+So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
+For their part, Zhu and colleagues [-@zhu.etal2020] elect to split the baby by averaging the simple and complex model scores together.
+This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
 
@@
 
-When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to “overfit” the data.
-Typically this correction comes in the form of a “complexity penalty,” penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014].
-Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
+When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to "overfit" the data.
+Typically this correction comes in the form of a "complexity penalty," penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014].
+Zhu and colleagues [-@zhu.etal2020] warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
 
@@
 First, the models differ not only in the number of parameters but in the domain of those parameters.
-Zhu and colleagues (2020) restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1].
-This restriction constrains the Bayesian Sampler model’s prior distribution to the class of “ignorance priors”.
-Via the bridging conditions (eq X), they show that this restricts the “noise level” for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$] whereas the PT+N model permits noise values between [0, 0.50].
+Zhu and colleagues [-@zhu.etal2020] assume that the Bayesian Sampler model’s prior distribution should reflect ignorance or a lack of information.
+A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey’s prior), Beta(0,0) (Haldene’s prior), or perhaps other  symmetric beta distributions Beta($\beta$,$\beta$) with $beta \leq 1$ [@jaynes2003].
+Thus, they restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1].
+Via the bridging conditions (eq X), they show that this restricts the "noise level" for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$].
+The Bayesian Sampler’s implied $d$ approaches the upper limit of $1/3$ as $N$ approaches $\infty$ and $\beta$ equals 1,  whereas the PT+N model permits noise values between [0, 0.50].
 
 Second, it is not immediately clear how the models’ structural differences impact their flexibility.
-To be sure, Zhu and colleagues’ (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
+To be sure, Zhu and colleagues’ [-@zhu.etal2020] bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
 But, what impact does the PT+N model’s treatment of conditional probabilities have on model complexity?
@@
 
-Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ (2020) use of “unpooled” models, where parameters are estimated independently for each individual participant.
+Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ [-@zhu.etal2020] use of "unpooled" models, where parameters are estimated independently for each individual participant.
 Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
@@
 
-Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison.
-A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as $WAIC$ and $LOO_{ic}$.
-These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
+Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as $WAIC$ and $LOO_{ic}$.
+Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
 Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
-In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.
+In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
+This allows for information about model parameters to be shared across participants, resulting in a potential reduction in model complexity and a more realistic test of the models.
 
@@
 
-Zhu, Sanborn, & Chater (2020) conducted two experiments to compare the PT+N and Bayesian sampler theories.
+Zhu, Sanborn, & Chater [-@chater.etal2020] conducted two experiments to compare the PT+N and Bayesian sampler theories.
 These experiments asked participants to judge the probability of different events in various combinations.
@@
 Experiment 1 asked about the events [icy, frosty] and [normal, typical] (e.g.
-“what is the probability that the weather in London is normal and not typical?”).
+"what is the probability that the weather in London is normal and not typical?").
 The authors’ goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated.
 Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment.
-This is especially clear, as the authors note, for disjunctive trials such as “normal or typical,” where “or typical” might not be read as a disjunction but rather an elaborative clause.
+This is especially clear, as the authors note, for disjunctive trials such as "normal or typical," where "or typical" might not be read as a disjunction but rather an elaborative clause.
 In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.
@@
 In addition, a third experimental condition asking about [warm, snowy] was also included in the experiment, but was dropped from the analyses reported in the paper.
-Exploring the raw responses from this condition reveals a substantial fraction of “zero” and “one” responses for certain trials.
+Exploring the raw responses from this condition reveals a substantial fraction of "zero" and "one" responses for certain trials.
 This may reflect a different response process than was intended.
 For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero—failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day.
-Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.
+Given these potentially aberrant responses, I followed Zhu and colleagues [-@zhu.etal2020] in ignoring data from this condition.
 
@@
 
-The PT+N model defines probability judgments as
+The PT+N model defines probability judgments as:
 
@@
 
-In contrast, the Bayesian Sampler model defines probability judgments as
+In contrast, the Bayesian Sampler model defines probability judgments as:
 
@@
 
-Fixing d and d’ or N and N’ equal yields the “simple” variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
+Fixing d and d’ or N and N’ equal yields the "simple" variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
 
@@
 These subjective probabilities are unobserved, and must be estimated as a latent variable.
-Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events ($A \land B$, $\neg A \land B$, $A \land \neg B$, $\neg A \land \neg B$).
+Here, they are represented with a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events ($A \land B$, $\neg A \land B$, $A \land \neg B$, $\neg A \land \neg B$).
 
 Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate d, d’, N, N’, and $\beta$ parameters for each participant.
-Translated into a Bayesian Framework, we can write the PT+N model as:
+Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 
@@
     \begin{align*}
-      y_{ij} &\sim beta\big(\mu_{ij} \kappa, (1-\mu_{ij}) \kappa \big) \\
-      \mu_{ij} &= f_{PT+N}(\overrightarrow{p_i}, d_j, d_j)  \\
+      y_{ijk} &\sim beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
+      \mu_{ijk} &= f_{PT+N}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
       d_j &= \frac{expit(\delta_j)}{2} \\
       d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2} \\
-      \overrightarrow{p_i} &\sim Dirichlet(\overrightarrow{1}) \\
+      \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
       \delta_j &\sim Normal(0,1) \\
@@
   \end{minipage}
-  \caption{Complex PT+N model diagram and formula specifications.}
+  \caption{Complex unpooled PT+N model diagram and formula specifications.
+Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters.
+Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
 \end{figure}
@@
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d’, rather than directly according to its $\beta$, N, and N’ parameters.
-I constrain d to [0, $1/3$] for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
+Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the *d* parameter of the PT+N model.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its $\beta$, N, and N' parameters.
+I constrain *d* to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
 
-The Bayesian Sampler model is therefore identical to the PT+N model save for the following changes to $\mu_{ijk}$, $d$, and $d'$, as shown below:
+The Bayesian Sampler model is therefore identical to the PT+N model save for thechanges to $\mu_{ijk}$, $d$, and $d'$ shown below:
 
 \begin{align*}
-  \mu_{ij} &= f_{BS}(\overrightarrow{p_i}, d_j, d_j)  \\
+  \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
   d_j &= \frac{expit(\delta_j)}{3} \\
@@
 Both of these models can also be implemented as hierarchical models with partial pooling for the d and d’ parameters (implicitly, for N and N’ in the case of the Bayesian Sampler).
-This should prevent over-penalizing complex versions of the model, as well as potentially improving their overall predictive fit.
+This should prevent over-penalizing complex versions of the models and potentially improve their overall predictive fit.
 The hierarchical implementation adds parameters for the population-level d and d’ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
 For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency [@papaspiliopoulos.etal2007].
+Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 
@@
   \end{minipage}
-  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications.}
+  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications.
+Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters.
+Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
 \end{figure}
@@
 Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of $\beta > 1$.
-Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of “ignorance priors” [@zhu.etal2020].
+Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of "ignorance priors" [@zhu.etal2020].
 However, it is also possible that people bring informative priors to the probability judgment task.
 Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted [see e.g., @fennell.baddeley2012].
-If $\beta$ is unrestricted, allowed to fall in the domain [0, $\infty$] then the Bayesian Sampler model becomes more flexible, allowing for equivalent “noise” levels in the same [0, .50] range as the PT+N model.
+If $\beta$ is unrestricted, allowed to fall in the domain [0, $\infty$] then the Bayesian Sampler model becomes more flexible, allowing for equivalent "noise" levels in the same [0, .50] range as the PT+N model.
+That is, through the bridging condition, the implied $d$ approaches .50 in the limit as $N$ and $\beta$ both approach $\infty$.
 Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
-Thus, fitting this additional “hybrid” model allows for a complete comparison of the models along both of their differing dimensions.
+Thus, fitting this additional unrestricted model allows for a complete comparison of the models along both of their differing dimensions.
 
@@
 I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $LOO_{ic}$ scores for each combination.
+Compared with BIC, $LOO_{ic}$ offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the "M-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
 Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
 For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$ [see @gelman.etal2014a].
-Figure X below displays the estimated differences in $LOO_{ic}$ scores for each of the models as compared to the best-scoring model.
+Figure 3 below displays the estimated differences in $LOO_{ic}$ scores for each of the models as compared to the best-scoring model.
 
@@
 Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the “hybrid” Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$] (again, lower scores for $LOO_{ic}$ are better).
+However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$] (again, lower scores for $LOO_{ic}$ are better).
 
 Recall, this model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the hybrid model and the best-fitting PT+N model, we see that the hybrid model better captures these judgments from Experiment 2 (trial-average-level $R^2$ = .85 vs .73; participant-level $R^2$ = .51 vs .45).
+Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from Experiment 2 (trial-average-level $R^2$ = .85 vs .73; participant-level $R^2$ = .51 vs .45).
 These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 
-Figure X shows the posterior distributions of the population-level d and d’ parameters inferred from the “hybrid” model.
-Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of “ignorance priors” in the Bayesian Sampler model.
+Figure X shows the posterior distributions of the population-level d and d’ parameters inferred from the unrestricted Bayesian Sampler model.
+Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
 Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ d and d’ estimates also lie outside this range (13 of 59 for d, 26 of 59 for d’).
+The finding that there are clear differences in d and d’ estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
 
@@
 
-Compared to the hybrid model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d’.
+Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d’.
 Evidently this structural component isn’t a complexifying epicycle, instead it is a prediction that constrains flexibility.
@@
 
-Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table X.
+Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
 
@@
 By a fair margin, the model best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] was a version of the Bayesian Sampler model without restriction on the range of its $\beta$ parameters.
-Alternatively, this “hybrid” model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
+Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
 Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task.
@@
 
-Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect “ignorance priors,” symmetric Beta distributions with $\beta$ < 1.
+Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta$ &lt; 1.
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people bring informative rather than “ignorance” priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $1/3$] range implied by the bridging conditions when beta is restricted < 1.
+Nevertheless, the data suggest that people may bring informative rather than "ignorance" priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $1/3$] range implied by the bridging conditions when beta is restricted &lt; 1.
 One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-If so, other experimental contexts might see parameter estimates consistent with an uninformative prior.
-In addition, the noise parameters also varied across individuals.
+This could potentially account for the differences in *d* and *d’* observed across experiments 1 and 2.
+In addition, these noise parameters also varied across individuals.
 Exploring the participant-level parameters reveals that some individuals’ implied d and d’ parameters were consistent with the class of ignorance priors (see supplemental materials).
@@
 Given their tight connections via the bridging conditions [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
-Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that “noise” might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].
+Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that "noise" might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].
 
-Some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020].
-Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying subjective probabilities to be inferred.
+Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020].
+Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying probabilities to be inferred.
 Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
-Of course, Zhu and colleagues’ (2020) experiments were never designed for this purpose.
+Of course, Zhu and colleagues’ [-@zhu.etal2020] experiments were never designed for this purpose.
 Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g.
 @franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006].
-One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes [e.g.
+One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes [e.g.
 @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].

### 2021-09-07 paper-rmd/paper-papaja.tex
added figs

@@
 \hypersetup{
-  pdftitle={Untitled probability judgment paper draft},
+  pdftitle={Comparing probabilistic accounts of probability judgments},
   pdfauthor={Derek Powell1},
@@
 \makeatother
-\shorttitle{DRAFT}
+\shorttitle{Comparing probability judgment accounts}
 \keywords{keywords\newline\indent Word count: X}
@@
 
-\title{Untitled probability judgment paper draft}
+\title{Comparing probabilistic accounts of probability judgments}
 \author{Derek Powell\textsuperscript{1}}
@@
 \abstract{
-Bayesian theories of cognitive science are centered around the idea that much of cognition is probabilistic.
-However, people's explicit probability judgments often incoherently violate the laws of probability.
+Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic.
+However, one mark sometimes held against these theories is that people's explicit probability judgments often violate the laws of probability.
 Two recent proposals, the ``Probability Theory plus Noise'' (Costello \& Watts, 2014) and ``Bayesian Sampler'' (Zhu, Sanborn, \& Chater, 2020) theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
-The two theories diverge significantly in their treatments of conditional probabilities, but comparing these theories poses challenges in accounting for model complexity.
-Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison through the \(LOO_{ic}\) information criteria and compared the fits of both models on data collected by Zhu and colleagues (2020).
-My findings support the Bayesian Sampler's account of conditional probability judgments, and I find the data are best explained by a ``hybrid'' model that can be seen as a modification of both theories.
+These theories fit quite differently into the larger project of Bayesian cognitive science, but their many similarities complicate comparisons of their predictive accuracy.
+In particular, comparing the models demands a careful accounting of model complexity.
+Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
+Compared the fits of both models on data collected by Zhu and colleagues (2020) I find the data are best explained by a version of the Bayesian Sampler model that allows for informative priors.
 }
@@
 
-Bayesian theories and approaches offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) with remarkable successes across domains, including perception (e.g.
+Bayesian theories and approaches offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) that has had remarkable successes across domains, including in perception (e.g.
 Kersten, Mamassian, \& Yuille, 2004), memory (Anderson, 1991), language (Xu \& Tenenbaum, 2007), and reasoning (Lu, Chen, \& Holyoak, 2012).
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that one area where these theories seem to fall down is in describing human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow,'' ``vaccines are safe,'' or ``this politician is trustworthy.''
+It is somewhat embarrassing then, that that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow,'' ``vaccines are safe,'' or ``this politician is trustworthy.''
 
@@
 Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020).
-Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by drawing a limited number of samples from probabilistic mental models (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).
+Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental ``sampling'' from a probabilistic mental model (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).
 
@@
 
-Costello \& Watts (2014, 2016, 2018) proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
-In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative read as positive.
-This results in probability judgments that reflect probabilistic credences perturbed by noise.
-Under the simplest form of the PT+N model, the expected value of probability judgments is:
+Costello \& Watts (2014, 2016, 2018) have proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
+In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative examples read as positive.
+The results are probability judgments reflecting probabilistic credences perturbed by noise.
+In their model, the probability a mental sample for an event \(A\) is read as A is the probability that the sample truly is A, p(A), and that it is correctly read (1-d), plus the probability that the sample is not A, 1-P(A) and that it is incorrectly read (d), or:
+
+\begin{align*}
+  P(\text{read as A}) &= (1-d)P(A) + d(1-P(A)) \\
+  &= (1-2d)P(A) + d
+\end{align*}
+
+Thus under the simplest form of the PT+N model, the expected value of probability judgments is:
 
@@
 
-where \emph{d} represents the probability with which samples will be misread, or the amount of noise in the judgment process (where by assumption a maximum of 50\% of samples can be misread, so that d is a number in the range {[}0, .50{]}).
+where \emph{d} represents the probability with which samples will be misread, or the amount of noise in the judgment process.
+By assumption, a maximum of 50\% of samples can be misread, so that d is a number in the range {[}0, .50{]}.
 The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics (Costello \& Watts, 2014, 2016, 2017, 2018).
@@
 Then, those ``observed'' samples are integrated with a prior over probabilities to produce a probability judgment.
-The effect of this prior is to regularize the judgments, pushing them away from extreme values.
-They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g.~encourage taking large risks based on a small number of mental samples.
-However, this hedging comes at the cost of systematic incoherence and biases.
-Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts (Costello \& Watts, 2014, 2016).
+The \(N\) mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric Beta distribution parameterized by \(\beta\).
+After observing \(S(A)\) successes and \(N - S(A)\) failures, this posterior is distributed \(Beta(\beta + S(A), \beta + N - S(A))\).
+Zhu and colleagues (2020) assume that people report the mean of their posterior probability estimates.
+Since for any Beta distribution \(x \sim Beta(a,b)\), \(E[x] = \frac{a}{a+b}\), the probability estimate is thus a linear function of S, N, and \(\beta\).
 
-Under the simplest version of the Bayesian Sampler model, the expected value of a probability judgment is given by the following formula:
+\[\hat{P}_{BS}(A) = \frac{S(A)}{N+2\beta} + \frac{\beta}{N+2\beta}\]
+
+The expected value of the estimate can then be written in terms of the expected number of successes, or \(P(A) \cdot N\).
+Under the simplest version of the Bayesian Sampler model, this gives the following formula:
 
@@
 
-where \emph{N} represents the number of mental samples drawn.
-The mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric beta distribution parameterized by \(\beta\).
-Zhu and colleagues (2020) show that the \emph{N} and \(\beta\) parameters of their model can be related to the \emph{d} parameter of the PT+N model via the following bridging formula:
+Like the PT+N model, the Bayesian Sampler model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts (Costello \& Watts, 2014, 2016).
+In fact, important equivalencies can be drawn between the two models.
+Zhu and colleagues (2020) show that the \(N\) and \(\beta\) parameters of their model can be related to the \(d\) parameter of the PT+N model via the following bridging formula:
 
 \[d = \frac{\beta}{N+2\beta}\]
+
+Thus, the effect of a Bayesian prior is nearly identical to the effect of noise in the PT+N model.
+But, rather than merely perturbing people's probability judgments, this prior can be seen as regularizing these judgments away from extreme values.
+Zhu and colleagues (2020) argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
+For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, a person incorporating no other information might conclude these are all certain victories and thereby be too willing to assume risk.
+A regularizing prior pushes these estimates away from extremes, thereby promoting better decision-making when mental samples are sparse.
+However, this hedging comes at the cost of systematic incoherence and biases.
 
@@
 By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs (Chater et al., 2020).
-However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project.
+And the bridging condition identified by Zhu and colleagues (Zhu, Sanborn, \& Chater, 2020) demonstrates that the theories make identical predictions in many cases.
+However, the two models diverge substantially in their treatment of conditional probability judgments.
 Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence.
-So, the conditioning of beliefs is at the heart of these theories.
+So, how they treat the conditioning of beliefs is at the heart of these theories.
 
@@
 
-\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]
+\[P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}\]
 
@@
 The complex versions of these models introduce additional parameters d' and N' that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are needed in order to explain conjunction fallacies (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020) .
+These additional parameters are crucial to both models' explanations of conjunction fallacies---key findings in the probability judgment literature (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020).
 Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs (Zhu, Sanborn, \& Chater, 2020, Supplementary materials).
@@
 But conjunction fallacies are an extremely robust empirical finding (Mellers, Hertwig, \& Kahneman, 2001; Sides, Osherson, Bonini, \& Viale, 2002) and are clearly present in the data collected by Zhu and colleagues (2020).
-So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
+So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
 For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are appropriately cautious about drawing any strong conclusions.
+This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
 
@@
 First, the models differ not only in the number of parameters but in the domain of those parameters.
-Zhu and colleagues (2020) restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}.
-This restriction constrains the Bayesian Sampler model's prior distribution to the class of ``ignorance priors.'' Via the bridging conditions (eq X), they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]} whereas the PT+N model permits noise values between {[}0, 0.50{]}.
+Zhu and colleagues (2020) assume that the Bayesian Sampler model's prior distribution should reflect ignorance or a lack of information.
+A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(beta \leq 1\) (Jaynes, 2003).
+Thus, they restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}.
+Via the bridging conditions (eq X), they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]}.
+The Bayesian Sampler's implied \(d\) approaches the upper limit of \(1/3\) as \(N\) approaches \(\infty\) and \(\beta\) equals 1, whereas the PT+N model permits noise values between {[}0, 0.50{]}.
 
@@
 
-Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison.
-A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \(WAIC\) and \(LOO_{ic}\).
-These indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
+Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \(WAIC\) and \(LOO_{ic}\).
+Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
 Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017).
-In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.
+In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
+This allows for information about model parameters to be shared across participants, resulting in a potential reduction in model complexity and a more realistic test of the models.
 
@@
 These subjective probabilities are unobserved, and must be estimated as a latent variable.
-Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (\(A \land B\), \(\neg A \land B\), \(A \land \neg B\), \(\neg A \land \neg B\)).
+Here, they are represented with a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (\(A \land B\), \(\neg A \land B\), \(A \land \neg B\), \(\neg A \land \neg B\)).
 
 Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate d, d', N, N', and \(\beta\) parameters for each participant.
-Translated into a Bayesian Framework, we can write the PT+N model as:
+Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 
@@
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model.
+Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the \emph{d} parameter of the PT+N model.
 To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its \(\beta\), N, and N' parameters.
-I constrain d to {[}0, \(1/3\){]} for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
+I constrain \emph{d} to \([0, 1/3]\) for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
 
-The Bayesian Sampler model is therefore identical to the PT+N model save for the following changes to \(\mu_{ijk}\), \(d\), and \(d'\), as shown below:
+The Bayesian Sampler model is therefore identical to the PT+N model save for thechanges to \(\mu_{ijk}\), \(d\), and \(d'\) shown below:
 
 \begin{align*}
-  \mu_{ij} &= f_{BS}(\overrightarrow{p_i}, d_j, d_j)  \\
+  \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
   d_j &= \frac{expit(\delta_j)}{3} \\
@@
 Both of these models can also be implemented as hierarchical models with partial pooling for the d and d' parameters (implicitly, for N and N' in the case of the Bayesian Sampler).
-This should prevent over-penalizing complex versions of the model, as well as potentially improving their overall predictive fit.
+This should prevent over-penalizing complex versions of the models and potentially improve their overall predictive fit.
 The hierarchical implementation adds parameters for the population-level d and d' as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
 For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (Papaspiliopoulos, Roberts, \& Sköld, 2007).
+Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 
@@
 If \(\beta\) is unrestricted, allowed to fall in the domain {[}0, \(\infty\){]} then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same {[}0, .50{]} range as the PT+N model.
+That is, through the bridging condition, the implied \(d\) approaches .50 in the limit as \(N\) and \(\beta\) both approach \(\infty\).
 Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
-Thus, fitting this additional ``hybrid'' model allows for a complete comparison of the models along both of their differing dimensions.
+Thus, fitting this additional unrestricted model allows for a complete comparison of the models along both of their differing dimensions.
 
@@
 I fit each of the models specified above to data from Zhu et al's (2020) Experiment 1 and 2 and computed \(LOO_{ic}\) scores for each combination.
+Compared with BIC, \(LOO_{ic}\) offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the ``M-open'' case; situations where we do not know if any of the models being compared are the ``true'' model (Vehtari, Simpson, Yao, \& Gelman, 2019).
 Model posteriors were estimated using the Numpyro (Phan, Pradhan, \& Jankowiak, 2019) implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
 For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to \(\hat{R}\) (see Gelman et al., 2014).
-Figure X below displays the estimated differences in \(LOO_{ic}\) scores for each of the models as compared to the best-scoring model.
+Figure 3 below displays the estimated differences in \(LOO_{ic}\) scores for each of the models as compared to the best-scoring model.
 
@@
 Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the ``hybrid'' Bayesian Sampler model that removes the restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]} (again, lower scores for \(LOO_{ic}\) are better).
+However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the Bayesian Sampler model that removes the restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]} (again, lower scores for \(LOO_{ic}\) are better).
 
 Recall, this model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the hybrid model and the best-fitting PT+N model, we see that the hybrid model better captures these judgments from Experiment 2 (trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) = .51 vs .45).
+Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from Experiment 2 (trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) = .51 vs .45).
 These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 
-Figure X shows the posterior distributions of the population-level d and d' parameters inferred from the ``hybrid'' model.
+Figure X shows the posterior distributions of the population-level d and d' parameters inferred from the unrestricted Bayesian Sampler model.
 Using data from experiment 2, population-level estimates of d and d' are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
 Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' d and d' estimates also lie outside this range (13 of 59 for d, 26 of 59 for d').
+The finding that there are clear differences in d and d' estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
 
@@
 
-Compared to the hybrid model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d'.
+Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d'.
 Evidently this structural component isn't a complexifying epicycle, instead it is a prediction that constrains flexibility.
@@
 By a fair margin, the model best accounting for the experimental data from Zhu and colleagues (2020) was a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters.
-Alternatively, this ``hybrid'' model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
+Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
 Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task.
@@
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people bring informative rather than ``ignorance'' priors to the probability judgment task, indicated by estimated d and d' parameters outside the {[}1, \(1/3\){]} range implied by the bridging conditions when beta is restricted \textless{} 1.
+Nevertheless, the data suggest that people may bring informative rather than ``ignorance'' priors to the probability judgment task, indicated by estimated d and d' parameters outside the {[}1, \(1/3\){]} range implied by the bridging conditions when beta is restricted \textless{} 1.
 One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-If so, other experimental contexts might see parameter estimates consistent with an uninformative prior.
-In addition, the noise parameters also varied across individuals.
+This could potentially account for the differences in \emph{d} and \emph{d'} observed across experiments 1 and 2.
+In addition, these noise parameters also varied across individuals.
 Exploring the participant-level parameters reveals that some individuals' implied d and d' parameters were consistent with the class of ignorance priors (see supplemental materials).
@@
 
-Some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs (Chater et al., 2020).
-Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying subjective probabilities to be inferred.
+Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs (Chater et al., 2020).
+Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying probabilities to be inferred.
 Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
@@
 Franke et al., 2016; Griffiths \& Tenenbaum, 2006; Jern, Chang, \& Kemp, 2014).
-One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes (e.g.
+One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes (e.g.
 Cook \& Lewandowsky, 2016; Jern, Chang, \& Kemp, 2014; Powell, Weisman, \& Markman, 2018).

### 2021-09-23 paper-rmd/paper-papaja.Rmd
paper edits

@@
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Compared the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a version of the Bayesian Sampler model that allows for informative priors.
+Compared the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
 
@@
 Thus, they restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1].
-Via the bridging conditions (eq X), they show that this restricts the "noise level" for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$].
+Via the bridging conditions, they show that this restricts the "noise level" for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$].
 The Bayesian Sampler’s implied $d$ approaches the upper limit of $1/3$ as $N$ approaches $\infty$ and $\beta$ equals 1,  whereas the PT+N model permits noise values between [0, 0.50].
@@
     \small
-    \begin{align*}
+    \begin{align}
       y_{ijk} &\sim beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
@@
       \kappa &\sim HalfCauchy(20)
-    \end{align*}
+    \end{align}
   \end{minipage}
@@
 
-The Bayesian Sampler model is therefore identical to the PT+N model save for thechanges to $\mu_{ijk}$, $d$, and $d'$ shown below:
+The Bayesian Sampler model is therefore identical to the PT+N model save for the changes to $\mu_{ijk}$, $d$, and $d'$ shown below:
 
@@
   d_j &= \frac{expit(\delta_j)}{3} \\
-  d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{3} \\
+  d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{3}
 \end{align*}
@@
 
-Figure X shows the posterior distributions of the population-level d and d’ parameters inferred from the unrestricted Bayesian Sampler model.
+Figure 4 shows the posterior distributions of the population-level d and d’ parameters inferred from the unrestricted Bayesian Sampler model.
 Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.

### 2021-09-23 paper-rmd/paper-papaja.tex
paper edits

@@
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Compared the fits of both models on data collected by Zhu and colleagues (2020) I find the data are best explained by a version of the Bayesian Sampler model that allows for informative priors.
+Compared the fits of both models on data collected by Zhu and colleagues (2020) I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
 }
@@
 Thus, they restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}.
-Via the bridging conditions (eq X), they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]}.
+Via the bridging conditions, they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]}.
 The Bayesian Sampler's implied \(d\) approaches the upper limit of \(1/3\) as \(N\) approaches \(\infty\) and \(\beta\) equals 1, whereas the PT+N model permits noise values between {[}0, 0.50{]}.
@@
 
-The Bayesian Sampler model is therefore identical to the PT+N model save for thechanges to \(\mu_{ijk}\), \(d\), and \(d'\) shown below:
+The Bayesian Sampler model is therefore identical to the PT+N model save for the changes to \(\mu_{ijk}\), \(d\), and \(d'\) shown below:
 
@@
 
-Figure X shows the posterior distributions of the population-level d and d' parameters inferred from the unrestricted Bayesian Sampler model.
+Figure 4 shows the posterior distributions of the population-level d and d' parameters inferred from the unrestricted Bayesian Sampler model.
 Using data from experiment 2, population-level estimates of d and d' are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.

### 2021-09-24 paper-rmd/paper-papaja.Rmd
updates to plots

@@
 t1_df %>% 
-  select(-X1) %>% 
+  select(-`...1`) %>% 
   mutate(model = gsub("BS", "Bayesian Sampler", model)) %>% 
@@
     \small
-    \begin{align}
+    \begin{align*}
       y_{ijk} &\sim beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
@@
       \kappa &\sim HalfCauchy(20)
-    \end{align}
+    \end{align*}
   \end{minipage}
@@
 t2_df %>% 
-  select(-X1) %>% 
+  select(-`...1`) %>% 
   gather(measure, value, -Experiment, -model) %>% 

### 2021-10-15 paper-rmd/paper-papaja.Rmd
reloo implemented

@@
     email         : "dmpowell@asu.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Writing - Original Draft Preparation
-      - Writing - Review & Editing
+    # role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+    #   - Conceptualization
+    #   - Writing - Original Draft Preparation
+    #   - Writing - Review & Editing
   # - name          : "Ernst-August Doelle"
@@
 authornote: |
-  Arizona State University, School of Social and Behavioral Sciences
+ This manuscript is pre-print that has not yet been peer reviewed.
 
@@
 
-keywords          : "keywords"
+keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
 wordcount         : "X"
@@
 
-Bayesian theories and approaches offer a unified formal framework for cognitive science [@tenenbaum.etal2011] that has had remarkable successes across domains, including in perception [e.g.
+Bayesian theories of cognition offer a unified formal framework for cognitive science [@tenenbaum.etal2011] that has had remarkable explanatory successes across domains, including in perception [e.g.
 @kersten.etal2004], memory [@anderson1991], language [@xu.tenenbaum2007], and reasoning [@lu.etal2012].
@@
 
-Trouble starts as soon as we seek to measure beliefs.
+Trouble starts as soon as we attempt to measure beliefs.
 According to Bayesian theories of cognition and epistemology [@jaynes2003], the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
@@
 "Tom Brady likes miniature horses"), a plain and flagrant violation of probability theory [@tversky.kahneman1983].
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or "unpacking" effects [@tversky.koehler1994], and a variety of other effects illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013].
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or "unpacking" effects [@tversky.koehler1994], and a variety of other findings illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013].
 Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
 
-Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic [@costello.watts2014; @zhu.etal2020].
+Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases in these judgments while maintaining that mental credences are fundamentally probabilistic [@costello.watts2014; @zhu.etal2020].
 Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental "sampling" from a probabilistic mental model [see also @chater.etal2020; @dasgupta.etal2017].
@@
 Costello & Watts [-@costello.watts2014; -@costello.watts2016; -@costello.watts2018] have proposed a theory of probability judgment they call the "Probability Theory plus Noise" theory (PT+N).
-In the PT+N model, mental "samples" are drawn from a probabilistic mental model of events and are then "read" with noise, so that some positive examples will be read as negative and some negative examples read as positive.
+In the PT+N model, mental "samples" are drawn from a probabilistic mental model of events and are then "read" with noise, so that some positive examples will be read as negative and some negative examples read as positive with some probability $d$.
 The results are probability judgments reflecting probabilistic credences perturbed by noise.
-In their model, the probability a mental sample for an event $A$ is read as A is the probability that the sample truly is A, p(A), and that it is correctly read (1-d), plus the probability that the sample is not A, 1-P(A) and that it is incorrectly read (d), or:
+In their model, the probability a mental sample for an event $A$ is read as A is the probability that the sample truly is $A$, $p(A)$, and that it is correctly read $(1-d)$, plus the probability that the sample is not $A$, $1-P(A)$ and that it is incorrectly read ($d$), or:
 
@@
 
-where *d* represents the probability with which samples will be misread, or the amount of noise in the judgment process.
-By assumption, a maximum of 50% of samples can be misread, so that d is a number in the range [0, .50].
-The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
+By assumption, a maximum of 50% of samples can be misread, so that $d$ is a number in the range $[0, 1/2]$.
+The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
 
@@
 Then, those "observed" samples are integrated with a prior over probabilities to produce a probability judgment.
-The $N$ mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric Beta distribution parameterized by $\beta$.
-After observing $S(A)$ successes and $N - S(A)$ failures, this posterior is distributed $Beta(\beta + S(A), \beta + N - S(A))$.
+This prior takes the form of a symmetric Beta distribution parameterized by $\beta$.
+After observing $S(A)$ successes and $N - S(A)$ failures, the posterior over probabilities is distributed $Beta(\beta + S(A), \beta + N - S(A))$.
 Zhu and colleagues [-@zhu.etal2020] assume that people report the mean of their posterior probability estimates.
@@
 Zhu and colleagues [-@zhu.etal2020] argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
-For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, a person incorporating no other information might conclude these are all certain victories and thereby be too willing to assume risk.
+For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, one might conclude these are all certain victories and thereby be too willing to assume risk.
 A regularizing prior pushes these estimates away from extremes, thereby promoting better decision-making when mental samples are sparse.
@@
 By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs [@chater.etal2020].
-And the bridging condition identified by Zhu and colleagues [@zhu.etal2020] demonstrates that the theories make identical predictions in many cases.
 However, the two models diverge substantially in their treatment of conditional probability judgments.
@@
 In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian [@costello.watts2016].
-According to the PT+N model, conditional probabilities P(A|B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B.
+According to the PT+N model, conditional probabilities $P(A|B)$ are estimated by a two-stage sampling procedure: first both events $A$ and $B$ are sampled with noise, and then a second noisy process computes the ratio of the events read as $A$ and $B$ over events read as $B$.
 The PT+N model predicts conditional probability estimates using the following equation:
@@
 
-The PT+N’s account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally.
-In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.
+The PT+N’s account of conditional probability judgment is distinctly non-Bayesian, separating it quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
 
@@
 Unfortunately, their results were somewhat equivocal.
-They fit both "simple" and "complex" versions of each model and computed a Bayesian Information Criteria score (BIC) for all models.
-The complex versions of these models introduce additional parameters d’ and N’ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
+They fit both "simple" and "complex" versions of each model, where the complex versions of these models introduce additional parameters $d'$ and $N'$ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
 These additional parameters are crucial to both models’ explanations of conjunction fallacies---key findings in the probability judgment literature [@costello.watts2017; @zhu.etal2020].
+They compared the fits of these models to data via Bayesian Information Criteria score.
 Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs [@zhu.etal2020, Supplementary materials].
@@
   rename(Model = model) %>% 
-  kbl(caption="Original model fitting results", digits=1, booktabs=T) %>% 
+  kbl(caption="Original model fitting results with best-fitting model in bold face.", digits=1, booktabs=T) %>% 
   kable_classic(full_width=F) %>% 
@@
 But conjunction fallacies are an extremely robust empirical finding [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues [-@zhu.etal2020].
-So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
+So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
 For their part, Zhu and colleagues [-@zhu.etal2020] elect to split the baby by averaging the simple and complex model scores together.
@@
 Typically this correction comes in the form of a "complexity penalty," penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014].
-Zhu and colleagues [-@zhu.etal2020] warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
+Zhu and colleagues [-@zhu.etal2020] warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
 
@@
 Zhu and colleagues [-@zhu.etal2020] assume that the Bayesian Sampler model’s prior distribution should reflect ignorance or a lack of information.
-A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey’s prior), Beta(0,0) (Haldene’s prior), or perhaps other  symmetric beta distributions Beta($\beta$,$\beta$) with $beta \leq 1$ [@jaynes2003].
-Thus, they restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1].
-Via the bridging conditions, they show that this restricts the "noise level" for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $1/3$].
-The Bayesian Sampler’s implied $d$ approaches the upper limit of $1/3$ as $N$ approaches $\infty$ and $\beta$ equals 1,  whereas the PT+N model permits noise values between [0, 0.50].
+A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey’s prior), Beta(0,0) (Haldene’s prior), or perhaps other symmetric beta distributions Beta($\beta$,$\beta$) with $\beta \in [0, 1]$ [@jaynes2003].
+Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to [0, $1/3$] (approaching $1/3$ as $N$ approaches $\infty$ and $\beta$ approaches 1), whereas the PT+N model permits noise values between $[0, 1/2]$.
 
 Second, it is not immediately clear how the models’ structural differences impact their flexibility.
-To be sure, Zhu and colleagues’ [-@zhu.etal2020] bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
+Zhu and colleagues’ [-@zhu.etal2020] bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
 But, what impact does the PT+N model’s treatment of conditional probabilities have on model complexity?
@@
 
-Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ [-@zhu.etal2020] use of "unpooled" models, where parameters are estimated independently for each individual participant.
-Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
+Finally, one potential explanation for the relative weakness of the "complex" model variants is Zhu and colleagues’ [-@zhu.etal2020] use of "unpooled" models, where parameters are estimated independently for each individual participant.
 In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
+If there is limited heterogeneity across individuals, then adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants.
 Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
@@
 Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
+Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
@@
 
-Fixing d and d’ or N and N’ equal yields the "simple" variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
+Fixing $d$ and $d'$ or $N$ and $N'$ equal yields the "simple" variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
 
@@
 
-Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate d, d’, N, N’, and $\beta$ parameters for each participant.
+Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate $d$, $d'$, $N$, $N'$, and $\beta$ parameters for each participant.
 Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 
-The function $f_{PT+N}$ computes the expected probability estimate using the underlying subjective probability p, the noise parameters d and d’, and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
-Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters d and d’.
+The function $f_{PT+N}$ computes the expected probability estimate using the underlying subjective probability $p$, the noise parameters $d$  and $d'$, and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
+Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters $d$  and $d'$.
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the *d* parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its $\beta$, N, and N' parameters.
-I constrain *d* to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
+Recall that Zhu and colleagues (2020) identified a bridging condition relating $\beta$ and $N$ in the Bayesian Sampler model to the $d$ parameter of the PT+N model.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and N' parameters.
+I constrain $d$ to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
@@
 
-Both of these models can also be implemented as hierarchical models with partial pooling for the d and d’ parameters (implicitly, for N and N’ in the case of the Bayesian Sampler).
+Both of these models can also be implemented as hierarchical models with partial pooling for the $d$ and $d'$ parameters (implicitly, for $N$ and $N'$ in the case of the Bayesian Sampler).
 This should prevent over-penalizing complex versions of the models and potentially improve their overall predictive fit.
-The hierarchical implementation adds parameters for the population-level d and d’ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
+The hierarchical implementation adds parameters for the population-level $d$  and $d'$ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
 For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency [@papaspiliopoulos.etal2007].
@@
 Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted [see e.g., @fennell.baddeley2012].
-If $\beta$ is unrestricted, allowed to fall in the domain [0, $\infty$] then the Bayesian Sampler model becomes more flexible, allowing for equivalent "noise" levels in the same [0, .50] range as the PT+N model.
-That is, through the bridging condition, the implied $d$ approaches .50 in the limit as $N$ and $\beta$ both approach $\infty$.
+If $\beta$ is unrestricted, allowed to fall in the domain $[0, \infty]$ then the Bayesian Sampler model becomes more flexible, allowing for equivalent "noise" levels in the same $[0, 1/2]$ range as the PT+N model.
+That is, through the bridging condition, the implied $d$ approaches $1/2$ in the limit as $N$ and $\beta$ both approach $\infty$.
 Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
@@
 I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $LOO_{ic}$ scores for each combination.
-Compared with BIC, $LOO_{ic}$ offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the "M-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
+Compared with BIC, $LOO_{ic}$ offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
 Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
@@
 \includegraphics[width=6in]{plot_compare.png}
-\caption{Model comparison results for data from Experiments 1 and 2. Error bars indicate standard errors of the estimates.}
+\caption{Model comparison results for data from Experiments 1 and 2.
+Error bars indicate two standard errors of the estimates.
+Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model \cite{sivula.etal2020}.}
 \end{figure}
 
-Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$] (again, lower scores for $LOO_{ic}$ are better).
+Data from Experiment 1 favor "complex" variants of the Bayesian Sampler model compared with the "simple" variants and all versions of the PT+N model.
+However, there is no clear winner.
+The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for $LOO_{ic}$ are better).
+That is, a Bayesian Sampler model without any restriction on $\beta$, allowing it to range from [0, $\infty$].
+However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\text{LOO}_{ic}$ scores within two standard errors of the difference, indicating that these models are also plausible.
 
+Data from Experiment 2 more decisively reveal a single winning model: the hierarchical "unrestricted" implementation of the Bayesian Sampler model allowing for informative priors.
 Recall, this model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from Experiment 2 (trial-average-level $R^2$ = .85 vs .73; participant-level $R^2$ = .51 vs .45).
+Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
 These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 
-Figure 4 shows the posterior distributions of the population-level d and d’ parameters inferred from the unrestricted Bayesian Sampler model.
-Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ d and d’ estimates also lie outside this range (13 of 59 for d, 26 of 59 for d’).
-The finding that there are clear differences in d and d’ estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
+Figure 4 shows the posterior distributions of the population-level $d$ and $d'$ parameters inferred from the unrestricted Bayesian Sampler model.
+In Experiment 2, population-level estimates of $d$  and $d'$ are greater than $\frac{1}{3}$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
+Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ $d$  and $d'$ estimates also lie outside this range (13 of 59 for d, 26 of 59 for $d'$).
+The finding that there are clear differences in $d$  and $d'$ estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
 
@@
 \includegraphics[width=4in]{plot_params.png}
-\caption{Posterior density of population-level d and d' parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.}
+\caption{Posterior density of population-level $d$  and $d'$ parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.
+Dashed line indicates theoretical maximum values for Bayesian Sampler model with uninformative priors.}
 \end{figure}
 
-Among the remaining models, hierarchical implementations with partial pooling outperform the unpooled corresponding models—suggesting that unpool variants of the complex models were over-penalized in the original analysis by @zhu.etal2020.
-In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models.
-However, differences among the remaining three models are approximately within the error of the $LOO_{ic}$ estimates.
-As expected, the simple variants of the models have smaller penalty terms than the complex variants.
-
-Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d’.
-Evidently this structural component isn’t a complexifying epicycle, instead it is a prediction that constrains flexibility.
+Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of $d$  and $d'$.
+Evidently this structural component isn’t a complexifying epicycle, instead it is better seen as a prediction that constrains flexibility.
 However, it is one that ultimately leads to a worse-fitting model.
 
-Perhaps surprisingly, the Bayesian Sampler with unrestricted $\beta$ actually receives a smaller penalty term than the more restricted version of the model.
-This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
-Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is estimated from data [-@gelman.etal2014].
+Perhaps surprisingly, the Bayesian Sampler with unrestricted $\beta$ actually receives a smaller penalty term than the more "restricted" version of the model for the data in Experiment 2.
+This is at first counterintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
+To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data [-@gelman.etal2014].
 If the estimated value is some very large positive number, then the constraint won’t have been very informative.
 But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model’s penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied d and d’ estimated for this data under this model are both very near $1/3$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
+Here, it seems reasonable to conjecture that because the implied $d$  and $d'$ estimated for this data under this model are both very near $1/3$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
 
@@
   kbl(
-    caption="Bayesian model comparison results", 
+    caption="Bayesian model comparison results with best scoring model in bold face.", 
     digits=3, 
@@
 
-Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta$ &lt; 1.
+Findings from  Zhu and colleagues’ experiments do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta \in [0, 1]$.
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people may bring informative rather than "ignorance" priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $1/3$] range implied by the bridging conditions when beta is restricted &lt; 1.
-One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-This could potentially account for the differences in *d* and *d’* observed across experiments 1 and 2.
-In addition, these noise parameters also varied across individuals.
-Exploring the participant-level parameters reveals that some individuals’ implied d and d’ parameters were consistent with the class of ignorance priors (see supplemental materials).
+Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated $d$  and $d'$ parameters outside the [1, $1/3$] range implied by "ignorance" priors.
+It could be that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
+This could potentially account for the differences in $d$ and $d'$ observed across Experiments 1 and 2.
+
+In addition, the implied $d$ and $d'$ noise parameters also varied across individuals.
+Of note, even in Experiment 2, some participant's implied $d$ and $d'$ parameters were consistent with the class of ignorance priors (see supplemental materials).
 Further research should explore this heterogeneity within and across individuals.
-Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and N parameters are not uniquely identifiable from judgment data of the sort examined here.
+Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and $N$ parameters are not uniquely identifiable from judgment data of the sort examined here.
 
 One thing this model comparison has not decided, and likely *cannot* decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given their tight connections via the bridging conditions [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
+Given the models' tight connections via the bridging condition [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
 Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that "noise" might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].
@@
 Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020].
-Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying probabilities to be inferred.
-Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
+Indeed, by representing the true subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
+Examining the model posteriors here reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
 Of course, Zhu and colleagues’ [-@zhu.etal2020] experiments were never designed for this purpose.
-Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g.
+Future research could explore how estimates of people's credences might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g.
 @franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006].

### 2021-10-15 paper-rmd/paper-papaja.tex
updates to plots
reloo implemented

@@
   man,floatsintext]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amsmath}
 \usepackage{ifxetex,ifluatex}
@@
   \usepackage{textcomp} % provide euro and other symbols
-  \usepackage{amssymb}
 \else % if luatex or xetex
@@
   pdflang={en-EN},
-  pdfkeywords={keywords},
+  pdfkeywords={probability judgments, Bayesian cognitive science, heuristics and biases},
   hidelinks,
@@
 \shorttitle{Comparing probability judgment accounts}
-\keywords{keywords\newline\indent Word count: X}
+\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: X}
 \usepackage{csquotes}
@@
 \else
-  \usepackage[shorthands=off,main=english]{babel}
+  \usepackage[main=english]{babel}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 \fi
@@
 
-Arizona State University, School of Social and Behavioral Sciences
-
-The authors made the following contributions. Derek Powell: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
+This manuscript is pre-print that has not yet been peer reviewed.
 
@@
 
-Bayesian theories and approaches offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) that has had remarkable successes across domains, including in perception (e.g.
+Bayesian theories of cognition offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) that has had remarkable explanatory successes across domains, including in perception (e.g.
 Kersten, Mamassian, \& Yuille, 2004), memory (Anderson, 1991), language (Xu \& Tenenbaum, 2007), and reasoning (Lu, Chen, \& Holyoak, 2012).
@@
 
-Trouble starts as soon as we seek to measure beliefs.
+Trouble starts as soon as we attempt to measure beliefs.
 According to Bayesian theories of cognition and epistemology (Jaynes, 2003), the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
@@
 For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory (Tversky \& Kahneman, 1983).
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects (Tversky \& Koehler, 1994), and a variety of other effects illustrating the incoherence of human probability judgments (for an accessible review, see Kahneman, 2013).
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects (Tversky \& Koehler, 1994), and a variety of other findings illustrating the incoherence of human probability judgments (for an accessible review, see Kahneman, 2013).
 Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
 
-Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases and apparent incoherence in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020).
+Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020).
 Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental ``sampling'' from a probabilistic mental model (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).
@@
 Costello \& Watts (2014, 2016, 2018) have proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
-In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative examples read as positive.
+In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative examples read as positive with some probability \(d\).
 The results are probability judgments reflecting probabilistic credences perturbed by noise.
-In their model, the probability a mental sample for an event \(A\) is read as A is the probability that the sample truly is A, p(A), and that it is correctly read (1-d), plus the probability that the sample is not A, 1-P(A) and that it is incorrectly read (d), or:
+In their model, the probability a mental sample for an event \(A\) is read as A is the probability that the sample truly is \(A\), \(p(A)\), and that it is correctly read \((1-d)\), plus the probability that the sample is not \(A\), \(1-P(A)\) and that it is incorrectly read (\(d\)), or:
 
@@
 
-where \emph{d} represents the probability with which samples will be misread, or the amount of noise in the judgment process.
-By assumption, a maximum of 50\% of samples can be misread, so that d is a number in the range {[}0, .50{]}.
-The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics (Costello \& Watts, 2014, 2016, 2017, 2018).
+By assumption, a maximum of 50\% of samples can be misread, so that \(d\) is a number in the range \([0, 1/2]\).
+The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions (Costello \& Watts, 2014, 2016, 2017, 2018).
 
@@
 Then, those ``observed'' samples are integrated with a prior over probabilities to produce a probability judgment.
-The \(N\) mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric Beta distribution parameterized by \(\beta\).
-After observing \(S(A)\) successes and \(N - S(A)\) failures, this posterior is distributed \(Beta(\beta + S(A), \beta + N - S(A))\).
+This prior takes the form of a symmetric Beta distribution parameterized by \(\beta\).
+After observing \(S(A)\) successes and \(N - S(A)\) failures, the posterior over probabilities is distributed \(Beta(\beta + S(A), \beta + N - S(A))\).
 Zhu and colleagues (2020) assume that people report the mean of their posterior probability estimates.
@@
 Zhu and colleagues (2020) argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
-For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, a person incorporating no other information might conclude these are all certain victories and thereby be too willing to assume risk.
+For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, one might conclude these are all certain victories and thereby be too willing to assume risk.
 A regularizing prior pushes these estimates away from extremes, thereby promoting better decision-making when mental samples are sparse.
@@
 By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs (Chater et al., 2020).
-And the bridging condition identified by Zhu and colleagues (Zhu, Sanborn, \& Chater, 2020) demonstrates that the theories make identical predictions in many cases.
 However, the two models diverge substantially in their treatment of conditional probability judgments.
@@
 In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian (Costello \& Watts, 2016).
-According to the PT+N model, conditional probabilities P(A\textbar B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B.
+According to the PT+N model, conditional probabilities \(P(A|B)\) are estimated by a two-stage sampling procedure: first both events \(A\) and \(B\) are sampled with noise, and then a second noisy process computes the ratio of the events read as \(A\) and \(B\) over events read as \(B\).
 The PT+N model predicts conditional probability estimates using the following equation:
@@
 
-The PT+N's account of conditional probability judgment is distinctly non-Bayesian, thereby separating it from the Bayesian Sampler theory quite fundamentally.
-In addition, drawing conditioning out of the mental models of events and into the conditional probability judgment process makes it unclear how the PT+N theory would interface with other probabilistic models of cognitive processes that invoke Bayes' rule.
+The PT+N's account of conditional probability judgment is distinctly non-Bayesian, separating it quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
 
@@
 Unfortunately, their results were somewhat equivocal.
-They fit both ``simple'' and ``complex'' versions of each model and computed a Bayesian Information Criteria score (BIC) for all models.
-The complex versions of these models introduce additional parameters d' and N' that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
+They fit both ``simple'' and ``complex'' versions of each model, where the complex versions of these models introduce additional parameters \(d'\) and \(N'\) that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
 These additional parameters are crucial to both models' explanations of conjunction fallacies---key findings in the probability judgment literature (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020).
+They compared the fits of these models to data via Bayesian Information Criteria score.
 Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs (Zhu, Sanborn, \& Chater, 2020, Supplementary materials).
@@
 
-\caption{\label{tab:table1}Original model fitting results}
+\caption{\label{tab:table1}Original model fitting results with best-fitting model in bold face.}
 \centering
@@
 But conjunction fallacies are an extremely robust empirical finding (Mellers, Hertwig, \& Kahneman, 2001; Sides, Osherson, Bonini, \& Viale, 2002) and are clearly present in the data collected by Zhu and colleagues (2020).
-So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
+So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
 For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
@@
 Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data (Gelman, Hwang, \& Vehtari, 2014).
-Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models' complexity.
+Zhu and colleagues (2020) warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the models' complexity.
 
@@
 Zhu and colleagues (2020) assume that the Bayesian Sampler model's prior distribution should reflect ignorance or a lack of information.
-A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(beta \leq 1\) (Jaynes, 2003).
-Thus, they restrict the range of the \(\beta\) parameter in their Bayesian Sampler model to fall within the interval {[}0,1{]}.
-Via the bridging conditions, they show that this restricts the ``noise level'' for the Bayesian Sampler, represented in implied d under the PT+N model, to {[}0, \(1/3\){]}.
-The Bayesian Sampler's implied \(d\) approaches the upper limit of \(1/3\) as \(N\) approaches \(\infty\) and \(\beta\) equals 1, whereas the PT+N model permits noise values between {[}0, 0.50{]}.
+A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(\beta \in [0, 1]\) (Jaynes, 2003).
+Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches \(\infty\) and \(\beta\) approaches 1), whereas the PT+N model permits noise values between \([0, 1/2]\).
 
 Second, it is not immediately clear how the models' structural differences impact their flexibility.
-To be sure, Zhu and colleagues' (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
+Zhu and colleagues' (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
 But, what impact does the PT+N model's treatment of conditional probabilities have on model complexity?
@@
 
-Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues' (2020) use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
-Adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
+Finally, one potential explanation for the relative weakness of the ``complex'' model variants is Zhu and colleagues' (2020) use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
 In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
+If there is limited heterogeneity across individuals, then adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants.
 Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
@@
 Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017).
+Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017).
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
@@
 
-Fixing d and d' or N and N' equal yields the ``simple'' variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
+Fixing \(d\) and \(d'\) or \(N\) and \(N'\) equal yields the ``simple'' variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
 
@@
 
-Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate d, d', N, N', and \(\beta\) parameters for each participant.
+Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate \(d\), \(d'\), \(N\), \(N'\), and \(\beta\) parameters for each participant.
 Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 
-The function \(f_{PT+N}\) computes the expected probability estimate using the underlying subjective probability p, the noise parameters d and d', and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
-Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters d and d'.
+The function \(f_{PT+N}\) computes the expected probability estimate using the underlying subjective probability \(p\), the noise parameters \(d\) and \(d'\), and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
+Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters \(d\) and \(d'\).
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the \emph{d} parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d', rather than directly according to its \(\beta\), N, and N' parameters.
-I constrain \emph{d} to \([0, 1/3]\) for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
+Recall that Zhu and colleagues (2020) identified a bridging condition relating \(\beta\) and \(N\) in the Bayesian Sampler model to the \(d\) parameter of the PT+N model.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied \(d\) and \(d'\), rather than directly according to its \(\beta\), \(N\), and N' parameters.
+I constrain \(d\) to \([0, 1/3]\) for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
@@
 
-Both of these models can also be implemented as hierarchical models with partial pooling for the d and d' parameters (implicitly, for N and N' in the case of the Bayesian Sampler).
+Both of these models can also be implemented as hierarchical models with partial pooling for the \(d\) and \(d'\) parameters (implicitly, for \(N\) and \(N'\) in the case of the Bayesian Sampler).
 This should prevent over-penalizing complex versions of the models and potentially improve their overall predictive fit.
-The hierarchical implementation adds parameters for the population-level d and d' as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
+The hierarchical implementation adds parameters for the population-level \(d\) and \(d'\) as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
 For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (Papaspiliopoulos, Roberts, \& Sköld, 2007).
@@
 Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted (see e.g., Fennell \& Baddeley, 2012).
-If \(\beta\) is unrestricted, allowed to fall in the domain {[}0, \(\infty\){]} then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same {[}0, .50{]} range as the PT+N model.
-That is, through the bridging condition, the implied \(d\) approaches .50 in the limit as \(N\) and \(\beta\) both approach \(\infty\).
+If \(\beta\) is unrestricted, allowed to fall in the domain \([0, \infty]\) then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same \([0, 1/2]\) range as the PT+N model.
+That is, through the bridging condition, the implied \(d\) approaches \(1/2\) in the limit as \(N\) and \(\beta\) both approach \(\infty\).
 Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
@@
 I fit each of the models specified above to data from Zhu et al's (2020) Experiment 1 and 2 and computed \(LOO_{ic}\) scores for each combination.
-Compared with BIC, \(LOO_{ic}\) offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the ``M-open'' case; situations where we do not know if any of the models being compared are the ``true'' model (Vehtari, Simpson, Yao, \& Gelman, 2019).
+Compared with BIC, \(LOO_{ic}\) offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model (Vehtari, Simpson, Yao, \& Gelman, 2019).
 Model posteriors were estimated using the Numpyro (Phan, Pradhan, \& Jankowiak, 2019) implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
@@
 \includegraphics[width=6in]{plot_compare.png}
-\caption{Model comparison results for data from Experiments 1 and 2. Error bars indicate standard errors of the estimates.}
+\caption{Model comparison results for data from Experiments 1 and 2.
+Error bars indicate two standard errors of the estimates.
+Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model \cite{sivula.etal2020}.}
 \end{figure}
 
-Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the Bayesian Sampler model that removes the restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]} (again, lower scores for \(LOO_{ic}\) are better).
+Data from Experiment 1 favor ``complex'' variants of the Bayesian Sampler model compared with the ``simple'' variants and all versions of the PT+N model.
+However, there is no clear winner.
+The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for \(LOO_{ic}\) are better).
+That is, a Bayesian Sampler model without any restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]}.
+However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting \(\beta \in [0,1]\)) have \(\text{LOO}_{ic}\) scores within two standard errors of the difference, indicating that these models are also plausible.
 
+Data from Experiment 2 more decisively reveal a single winning model: the hierarchical ``unrestricted'' implementation of the Bayesian Sampler model allowing for informative priors.
 Recall, this model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from Experiment 2 (trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) = .51 vs .45).
+Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
 These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 
-Figure 4 shows the posterior distributions of the population-level d and d' parameters inferred from the unrestricted Bayesian Sampler model.
-Using data from experiment 2, population-level estimates of d and d' are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' d and d' estimates also lie outside this range (13 of 59 for d, 26 of 59 for d').
-The finding that there are clear differences in d and d' estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
+Figure 4 shows the posterior distributions of the population-level \(d\) and \(d'\) parameters inferred from the unrestricted Bayesian Sampler model.
+In Experiment 2, population-level estimates of \(d\) and \(d'\) are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
+Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' \(d\) and \(d'\) estimates also lie outside this range (13 of 59 for d, 26 of 59 for \(d'\)).
+The finding that there are clear differences in \(d\) and \(d'\) estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
 
@@
 \includegraphics[width=4in]{plot_params.png}
-\caption{Posterior density of population-level d and d' parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.}
+\caption{Posterior density of population-level $d$  and $d'$ parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.
+Dashed line indicates theoretical maximum values for Bayesian Sampler model with uninformative priors.}
 \end{figure}
 
-Among the remaining models, hierarchical implementations with partial pooling outperform the unpooled corresponding models---suggesting that unpool variants of the complex models were over-penalized in the original analysis by Zhu, Sanborn, and Chater (2020).
-In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models.
-However, differences among the remaining three models are approximately within the error of the \(LOO_{ic}\) estimates.
-As expected, the simple variants of the models have smaller penalty terms than the complex variants.
-
-Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d'.
-Evidently this structural component isn't a complexifying epicycle, instead it is a prediction that constrains flexibility.
+Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of \(d\) and \(d'\).
+Evidently this structural component isn't a complexifying epicycle, instead it is better seen as a prediction that constrains flexibility.
 However, it is one that ultimately leads to a worse-fitting model.
 
-Perhaps surprisingly, the Bayesian Sampler with unrestricted \(\beta\) actually receives a smaller penalty term than the more restricted version of the model.
-This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data (see Gelman, Hwang, \& Vehtari, 2014).
-Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is estimated from data (2014).
+Perhaps surprisingly, the Bayesian Sampler with unrestricted \(\beta\) actually receives a smaller penalty term than the more ``restricted'' version of the model for the data in Experiment 2.
+This is at first counterintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data (see Gelman, Hwang, \& Vehtari, 2014).
+To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data (2014).
 If the estimated value is some very large positive number, then the constraint won't have been very informative.
 But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model's penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied d and d' estimated for this data under this model are both very near \(1/3\), the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
+Here, it seems reasonable to conjecture that because the implied \(d\) and \(d'\) estimated for this data under this model are both very near \(1/3\), the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
 
@@
 
-\caption{\label{tab:table2}Bayesian model comparison results}
+\caption{\label{tab:table2}Bayesian model comparison results with best scoring model in bold face.}
 \centering
@@
 \midrule
-\textbf{Unrestricted BS MLM} & \textbf{-1178.1} & \textbf{149.2} & \textbf{0.725} & \textbf{0.821} & \textbf{-3953.1} & \textbf{368.7} & \textbf{0.688} & \textbf{0.878}\\
-Bayesian Sampler complex MLM & -1167.7 & 145.6 & 0.723 & 0.816 & -3815.4 & 399.9 & 0.675 & 0.852\\
-PT+N complex & -1163.5 & 133.6 & 0.722 & 0.821 & -3768.5 & 398.2 & 0.667 & 0.840\\
-Bayesian Sampler complex & -1159.5 & 149.3 & 0.723 & 0.815 & -3733.2 & 453.6 & 0.679 & 0.848\\
-PT+N complex MLM & -1157.0 & 147.1 & 0.722 & 0.820 & -3796.7 & 356.1 & 0.658 & 0.835\\
+\textbf{Unrestricted BS hier.} & \textbf{-2163.2} & \textbf{296.4} & \textbf{0.884} & \textbf{0.964} & \textbf{-3953.3} & \textbf{368.4} & \textbf{0.688} & \textbf{0.878}\\
+Bayesian Sampler complex & -2134.4 & 285.1 & 0.883 & 0.961 & -3733.2 & 453.6 & 0.679 & 0.848\\
+Bayesian Sampler complex hier. & -2131.9 & 292.0 & 0.883 & 0.960 & -3815.6 & 399.5 & 0.675 & 0.852\\
+Bayesian Sampler simple hier. & -2053.9 & 271.8 & 0.874 & 0.953 & -3790.5 & 383.2 & 0.667 & 0.839\\
+Bayesian Sampler simple & -2051.3 & 273.7 & 0.874 & 0.952 & -3722.8 & 418.9 & 0.667 & 0.831\\
 \addlinespace
-Bayesian Sampler simple MLM & -1149.5 & 141.0 & 0.719 & 0.810 & -3794.7 & 381.0 & 0.667 & 0.839\\
-Bayesian Sampler simple & -1138.1 & 146.8 & 0.718 & 0.806 & -3722.8 & 418.9 & 0.667 & 0.831\\
-PT+N simple & -1051.8 & 132.6 & 0.695 & 0.793 & -3643.0 & 319.5 & 0.619 & 0.777\\
-PT+N simple MLM & -1044.7 & 137.9 & 0.695 & 0.793 & -3643.8 & 305.3 & 0.617 & 0.772\\
-Relative Freq. & -746.6 & 146.8 & 0.664 & 0.737 & -1287.7 & 424.6 & 0.516 & 0.639\\
+PT+N complex hier. & -1927.4 & 298.3 & 0.867 & 0.946 & -3795.3 & 356.2 & 0.658 & 0.835\\
+PT+N complex & -1922.8 & 264.5 & 0.864 & 0.941 & -3768.5 & 398.2 & 0.667 & 0.840\\
+PT+N simple & -1708.8 & 254.5 & 0.838 & 0.918 & -3643.0 & 319.5 & 0.619 & 0.777\\
+PT+N simple hier. & -1697.1 & 265.4 & 0.839 & 0.919 & -3643.8 & 305.3 & 0.617 & 0.772\\
+Relative Freq. & -1266.2 & 305.7 & 0.820 & 0.875 & -1287.7 & 424.6 & 0.516 & 0.639\\
 \bottomrule
@@
 
-Findings from Zhu and colleagues' Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta\) \textless{} 1.
+Findings from Zhu and colleagues' experiments do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta \in [0, 1]\).
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people may bring informative rather than ``ignorance'' priors to the probability judgment task, indicated by estimated d and d' parameters outside the {[}1, \(1/3\){]} range implied by the bridging conditions when beta is restricted \textless{} 1.
-One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-This could potentially account for the differences in \emph{d} and \emph{d'} observed across experiments 1 and 2.
-In addition, these noise parameters also varied across individuals.
-Exploring the participant-level parameters reveals that some individuals' implied d and d' parameters were consistent with the class of ignorance priors (see supplemental materials).
+Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated \(d\) and \(d'\) parameters outside the {[}1, \(1/3\){]} range implied by ``ignorance'' priors.
+It could be that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
+This could potentially account for the differences in \(d\) and \(d'\) observed across Experiments 1 and 2.
+
+In addition, the implied \(d\) and \(d'\) noise parameters also varied across individuals.
+Of note, even in Experiment 2, some participant's implied \(d\) and \(d'\) parameters were consistent with the class of ignorance priors (see supplemental materials).
 Further research should explore this heterogeneity within and across individuals.
-Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the \(\beta\) and N parameters are not uniquely identifiable from judgment data of the sort examined here.
+Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the \(\beta\) and \(N\) parameters are not uniquely identifiable from judgment data of the sort examined here.
 
 One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given their tight connections via the bridging conditions (Zhu, Sanborn, \& Chater, 2020), it may not be possible to draw decisive conclusions here.
+Given the models' tight connections via the bridging condition (Zhu, Sanborn, \& Chater, 2020), it may not be possible to draw decisive conclusions here.
 Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that ``noise'' might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler (2020).
@@
 Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs (Chater et al., 2020).
-Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying probabilities to be inferred.
-Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
+Indeed, by representing the true subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
+Examining the model posteriors here reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
 Of course, Zhu and colleagues' (2020) experiments were never designed for this purpose.
-Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning (e.g.
+Future research could explore how estimates of people's credences might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning (e.g.
 Franke et al., 2016; Griffiths \& Tenenbaum, 2006; Jern, Chang, \& Kemp, 2014).

### 2021-10-15 ic-compare.py

@@ -1,0 +1,421 @@
+from arviz.stats.stats_utils import ELPDData # modify to use relative import
+
+import warnings
+from copy import deepcopy
+from typing import List, Optional, Tuple, Union, Mapping, cast, Callable
+
+import numpy as np
+import pandas as pd
+import scipy.stats as st
+import xarray as xr
+from scipy.optimize import minimize
+from typing_extensions import Literal
+
+## set these back to relative imports
+from arviz.data import InferenceData, convert_to_dataset, convert_to_inference_data
+from arviz.rcparams import rcParams, ScaleKeyword, ICKeyword
+
+from arviz import loo, waic
+
+NO_GET_ARGS: bool = False
+try:
+    from typing_extensions import get_args
+except ImportError:
+    NO_GET_ARGS = True
+    
+
+def get_ic_type(ic_input):
+
+    try:
+        ic_input.loo
+        ic = "loo"
+    except:
+        try:
+            ic_input.waic
+            ic = "waic"
+        except:
+            ic = "unknown"
+    return(ic)
+
+
+def _ic_matrix(ics, ic_i):
+    """Store the previously computed pointwise predictive accuracy values (ics) in a 2D matrix."""
+    cols, _ = ics.shape
+    rows = len(ics[ic_i].iloc[0])
+    ic_i_val = np.zeros((rows, cols))
+
+    for idx, val in enumerate(ics.index):
+        ic = ics.loc[val][ic_i]
+
+        if len(ic) != rows:
+            raise ValueError("The number of observations should be the same across all models")
+
+        ic_i_val[:, idx] = ic
+
+    return rows, cols, ic_i_val
+
+
+def compare(
+    dataset_dict: Mapping[str, InferenceData],
+    ic: Optional[ICKeyword] = None,
+    method: Literal["stacking", "BB-pseudo-BMA", "pseudo-MA"] = "stacking",
+    b_samples: int = 1000,
+    alpha: float = 1,
+    seed=None,
+    scale: Optional[ScaleKeyword] = None,
+    var_name: Optional[str] = None,
+):
+    r"""Compare models based on PSIS-LOO `loo` or WAIC `waic` cross-validation.
+    LOO is leave-one-out (PSIS-LOO `loo`) cross-validation and
+    WAIC is the widely applicable information criterion.
+    Read more theory here - in a paper by some of the leading authorities
+    on model selection dx.doi.org/10.1111/1467-9868.00353
+    Parameters
+    ----------
[…]

### 2021-10-15 reloo-compare.py

@@ -1,0 +1,433 @@
+from arviz.stats.stats_utils import ELPDData # modify to use relative import
+
+import warnings
+from copy import deepcopy
+from typing import List, Optional, Tuple, Union, Mapping, cast, Callable
+
+import numpy as np
+import pandas as pd
+import scipy.stats as st
+import xarray as xr
+from scipy.optimize import minimize
+from typing_extensions import Literal
+
+## set these back to relative imports
+from arviz.data import InferenceData, convert_to_dataset, convert_to_inference_data
+from arviz.rcparams import rcParams, ScaleKeyword, ICKeyword
+
+from arviz import loo, waic
+
+NO_GET_ARGS: bool = False
+try:
+    from typing_extensions import get_args
+except ImportError:
+    NO_GET_ARGS = True
+    
+
+def get_ic_type(ic_input):
+
+    try:
+        ic_input.loo
+        ic = "loo"
+    except:
+        try:
+            ic_input.waic
+            ic = "waic"
+        except:
+            ic = "unknown"
+    return(ic)
+
+
+def _ic_matrix(ics, ic_i):
+    """Store the previously computed pointwise predictive accuracy values (ics) in a 2D matrix."""
+    cols, _ = ics.shape
+    rows = len(ics[ic_i].iloc[0])
+    ic_i_val = np.zeros((rows, cols))
+
+    for idx, val in enumerate(ics.index):
+        ic = ics.loc[val][ic_i]
+
+        if len(ic) != rows:
+            raise ValueError("The number of observations should be the same across all models")
+
+        ic_i_val[:, idx] = ic
+
+    return rows, cols, ic_i_val
+
+
+def compare(
+    dataset_dict: Mapping[str, InferenceData],
+    ic: Optional[ICKeyword] = None,
+    method: Literal["stacking", "BB-pseudo-BMA", "pseudo-MA"] = "stacking",
+    b_samples: int = 1000,
+    alpha: float = 1,
+    seed=None,
+    scale: Optional[ScaleKeyword] = None,
+    var_name: Optional[str] = None,
+):
+    r"""Compare models based on PSIS-LOO `loo` or WAIC `waic` cross-validation.
+    LOO is leave-one-out (PSIS-LOO `loo`) cross-validation and
+    WAIC is the widely applicable information criterion.
+    Read more theory here - in a paper by some of the leading authorities
+    on model selection dx.doi.org/10.1111/1467-9868.00353
+    Parameters
+    ----------
[…]

### 2021-10-18 paper-rmd/paper-papaja.Rmd
finished reloo

@@
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Compared the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
+Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
 
@@
 draft             : no
-
+biblio-style      : "apa"
 documentclass     : "apa6"
 classoption       : "man"
-output            : papaja::apa6_pdf
-header-includes   :
+output:
+  papaja::apa6_pdf:
+    citation_package: biblatex
+
+header-includes:
   - \usepackage{tikz}
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as $WAIC$ and $LOO_{ic}$.
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as ${\text{WAIC}}$ and $\text{LOO}_{ic}$.
 Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
@@
 
-Zhu, Sanborn, & Chater [-@chater.etal2020] conducted two experiments to compare the PT+N and Bayesian sampler theories.
+Zhu, Sanborn, & Chater [-@chater.etal2020] conducted two experiments to compare the PT+N and Bayesian Sampler theories.
 These experiments asked participants to judge the probability of different events in various combinations.
@@
 Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment.
-This is especially clear, as the authors note, for disjunctive trials such as "normal or typical," where "or typical" might not be read as a disjunction but rather an elaborative clause.
+This is especially clear, as the authors note, for disjunctive query trials such as "normal or typical," where "or typical" might not be read as a disjunction but rather an elaborative clause.
 In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.
@@
 
-I implement both the Bayesian Sampler and PT+N models as Bayesian models in the probabilistic programming language Numpyro.
+I implement several variants of the Bayesian Sampler and PT+N models in a Bayesian framework.
+These models were implemented in the probabilistic programming language Numpyro.
 All code and results are available as supplemental materials (https://github.com/derekpowell/bayesian-sampler).
@@
 Recall that Zhu and colleagues (2020) identified a bridging condition relating $\beta$ and $N$ in the Bayesian Sampler model to the $d$ parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and N' parameters.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and $N'$ parameters.
 I constrain $d$ to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
@@
 
-I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $LOO_{ic}$ scores for each combination.
-Compared with BIC, $LOO_{ic}$ offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
+I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $\text{LOO}_{ic}$ scores for each combination.
+Compared with BIC, $\text{LOO}_{ic}$ offers a more sophisticated estimation of model complexity and is more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
 Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
 For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$ [see @gelman.etal2014a].
-Figure 3 below displays the estimated differences in $LOO_{ic}$ scores for each of the models as compared to the best-scoring model.
+Figure 3 below displays the estimated differences in $\text{LOO}_{ic}$ scores for each of the models as compared to the best-scoring model.
 
@@
 \includegraphics[width=6in]{plot_compare.png}
-\caption{Model comparison results for data from Experiments 1 and 2.
+\caption[]{Model comparison results for data from Experiments 1 and 2.
 Error bars indicate two standard errors of the estimates.
-Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model \cite{sivula.etal2020}.}
+Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model (\cite{sivula.etal2020}).}
 \end{figure}
@@
 However, there is no clear winner.
-The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for $LOO_{ic}$ are better).
+The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for $\text{LOO}_{ic}$ are better).
 That is, a Bayesian Sampler model without any restriction on $\beta$, allowing it to range from [0, $\infty$].
-However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\text{LOO}_{ic}$ scores within two standard errors of the difference, indicating that these models are also plausible.
+However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\text{LOO}_{ic}$ scores within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
 
 Data from Experiment 2 more decisively reveal a single winning model: the hierarchical "unrestricted" implementation of the Bayesian Sampler model allowing for informative priors.
-Recall, this model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
-These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 
 Figure 4 shows the posterior distributions of the population-level $d$ and $d'$ parameters inferred from the unrestricted Bayesian Sampler model.
-In Experiment 2, population-level estimates of $d$  and $d'$ are greater than $\frac{1}{3}$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
+In Experiment 2, population-level estimates of $d$  and $d'$ are greater than $1/3$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
 Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ $d$  and $d'$ estimates also lie outside this range (13 of 59 for d, 26 of 59 for $d'$).
@@
 
+Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments.
+So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
+Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
+And comparing their prediction specifically for conditional probability judgments, the unrestricted Bayesian Sampler provides a better fit, especially for the modestly-correlated events of experiment 2 (Exp 1: response-level $r$ = .91 vs.
+.89; query-level $r$ = .98 vs.
+.97; Exp 2: response-level $r$ = .72 vs.
+.67; query-level $r$ = .92 vs.
+.85) These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of $d$  and $d'$.
@@
 ```{r table2, echo=FALSE, message=F, warning=F}
+### NOTE: be sure to rescale elppd_loo to deviance scale to report as loo_ic 10/15/21, 2:37 PM
+
 t2_df <- read_csv("model-comparison-table.csv")
@@
     booktabs=T,
-    col.names = c("Model", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{trial}$", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{trial}$"),
+    col.names = c("Model", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{query}$", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{query}$"),
     escape = F
@@
 
-Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
+Finally, it is worth noting that the best of these models provide quite strong overall fits to the data, not just for the query averages, but also for the query averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
 
@@
 Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
-Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task.
-In keeping with the larger theoretical framework of Bayesian cognitive science, this theory assumes that subjective probabilities underlie people’s probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process [@chater.etal2020; @zhu.etal2020].
+Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this task.
+In keeping with the larger theoretical framework of Bayesian cognitive science, the Bayesian Sampler theory assumes that subjective probabilities underlie people’s probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process [@chater.etal2020; @zhu.etal2020].
 
-Findings from  Zhu and colleagues’ experiments do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta \in [0, 1]$.
+Zhu and colleagues’ findings do cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta \in [0, 1]$.
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated $d$  and $d'$ parameters outside the [1, $1/3$] range implied by "ignorance" priors.
+Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated $d$  and $d'$ parameters outside the $[0, 1/3]$ range implied by "ignorance" priors.
 It could be that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
@@
 Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020].
-Indeed, by representing the true subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
+Indeed, by representing the "true" subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
 Examining the model posteriors here reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.

### 2021-10-18 paper-rmd/paper-papaja.tex
finished reloo

@@
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage[style=apa,]{biblatex}
+\addbibresource{references.bib}
 
@@
 However, one mark sometimes held against these theories is that people's explicit probability judgments often violate the laws of probability.
-Two recent proposals, the ``Probability Theory plus Noise'' (Costello \& Watts, 2014) and ``Bayesian Sampler'' (Zhu, Sanborn, \& Chater, 2020) theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
+Two recent proposals, the ``Probability Theory plus Noise'' \autocite{costello.watts2014} and ``Bayesian Sampler'' \autocite{zhu.etal2020} theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
 These theories fit quite differently into the larger project of Bayesian cognitive science, but their many similarities complicate comparisons of their predictive accuracy.
@@
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Compared the fits of both models on data collected by Zhu and colleagues (2020) I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
+Comparing the fits of both models on data collected by Zhu and colleagues \autocite*{zhu.etal2020} I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
 }
@@
 
-Bayesian theories of cognition offer a unified formal framework for cognitive science (Tenenbaum, Kemp, Griffiths, \& Goodman, 2011) that has had remarkable explanatory successes across domains, including in perception (e.g.
-Kersten, Mamassian, \& Yuille, 2004), memory (Anderson, 1991), language (Xu \& Tenenbaum, 2007), and reasoning (Lu, Chen, \& Holyoak, 2012).
+Bayesian theories of cognition offer a unified formal framework for cognitive science \autocite{tenenbaum.etal2011} that has had remarkable explanatory successes across domains, including in perception \autocite[e.g.][]{kersten.etal2004}, memory \autocite{anderson1991}, language \autocite{xu.tenenbaum2007}, and reasoning \autocite{lu.etal2012}.
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow,'' ``vaccines are safe,'' or ``this politician is trustworthy.''
+It is somewhat embarrassing then, that that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow'', ``vaccines are safe,'' or ``this politician is trustworthy''.
 
 Trouble starts as soon as we attempt to measure beliefs.
-According to Bayesian theories of cognition and epistemology (Jaynes, 2003), the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
+According to Bayesian theories of cognition and epistemology \autocite{jaynes2003}, the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
 So asking people to express beliefs in terms of probability seems only natural.
@@
 Unfortunately, people's explicit probability judgments routinely violate the axioms of probability theory.
-For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory (Tversky \& Kahneman, 1983).
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects (Tversky \& Koehler, 1994), and a variety of other findings illustrating the incoherence of human probability judgments (for an accessible review, see Kahneman, 2013).
+For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory \autocite{tversky.kahneman1983}.
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects \autocite{tversky.koehler1994}, and a variety of other findings illustrating the incoherence of human probability judgments \autocite[for an accessible review, see][]{kahneman2013}.
 Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
 
-Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases in these judgments while maintaining that mental credences are fundamentally probabilistic (Costello \& Watts, 2014; Zhu, Sanborn, \& Chater, 2020).
-Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental ``sampling'' from a probabilistic mental model (see also Chater et al., 2020; Dasgupta, Schulz, \& Gershman, 2017).
+Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases in these judgments while maintaining that mental credences are fundamentally probabilistic \autocite{costello.watts2014,zhu.etal2020}.
+Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental ``sampling'' from a probabilistic mental model \autocites[see also][]{chater.etal2020,dasgupta.etal2017}.
 
@@
 
-Costello \& Watts (2014, 2016, 2018) have proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
+Costello \& Watts \autocite*{costello.watts2014,costello.watts2016,costello.watts2018} have proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
 In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative examples read as positive with some probability \(d\).
@@
 By assumption, a maximum of 50\% of samples can be misread, so that \(d\) is a number in the range \([0, 1/2]\).
-The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions (Costello \& Watts, 2014, 2016, 2017, 2018).
+The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions \autocite{costello.watts2014,costello.watts2016,costello.watts2017,costello.watts2018}.
 
-Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model of probability judgment they call the ``Bayesian Sampler.'' Under this model, probability judgment is itself seen as a process of Bayesian inference.
+Meanwhile, Zhu, Sanborn, \& Chater \autocite*{zhu.etal2020} have proposed a Bayesian model of probability judgment they call the ``Bayesian Sampler.'' Under this model, probability judgment is itself seen as a process of Bayesian inference.
 To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event.
@@
 After observing \(S(A)\) successes and \(N - S(A)\) failures, the posterior over probabilities is distributed \(Beta(\beta + S(A), \beta + N - S(A))\).
-Zhu and colleagues (2020) assume that people report the mean of their posterior probability estimates.
+Zhu and colleagues \autocite*{zhu.etal2020} assume that people report the mean of their posterior probability estimates.
 Since for any Beta distribution \(x \sim Beta(a,b)\), \(E[x] = \frac{a}{a+b}\), the probability estimate is thus a linear function of S, N, and \(\beta\).
@@
 
-Like the PT+N model, the Bayesian Sampler model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts (Costello \& Watts, 2014, 2016).
+Like the PT+N model, the Bayesian Sampler model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts \autocite{costello.watts2014,costello.watts2016}.
 In fact, important equivalencies can be drawn between the two models.
-Zhu and colleagues (2020) show that the \(N\) and \(\beta\) parameters of their model can be related to the \(d\) parameter of the PT+N model via the following bridging formula:
+Zhu and colleagues \autocite*{zhu.etal2020} show that the \(N\) and \(\beta\) parameters of their model can be related to the \(d\) parameter of the PT+N model via the following bridging formula:
 
@@
 But, rather than merely perturbing people's probability judgments, this prior can be seen as regularizing these judgments away from extreme values.
-Zhu and colleagues (2020) argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
+Zhu and colleagues \autocite*{zhu.etal2020} argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
 For instance, consider someone estimating the probability that they can swim across a lake, outrun an animal, or win a hand of poker: if a mental simulation of these events produces two samples indicating success, one might conclude these are all certain victories and thereby be too willing to assume risk.
@@
 
-By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs (Chater et al., 2020).
+By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs \autocite{chater.etal2020}.
 However, the two models diverge substantially in their treatment of conditional probability judgments.
@@
 According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments.
-By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models (Chater et al., 2020).
+By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models \autocite{chater.etal2020}.
 
-In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian (Costello \& Watts, 2016).
+In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian \autocite{costello.watts2016}.
 According to the PT+N model, conditional probabilities \(P(A|B)\) are estimated by a two-stage sampling procedure: first both events \(A\) and \(B\) are sampled with noise, and then a second noisy process computes the ratio of the events read as \(A\) and \(B\) over events read as \(B\).
@@
 
-Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments.
+Zhu, Sanborn, and Chater \autocite*{zhu.etal2020} compared their Bayesian Sampler model against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments.
 Unfortunately, their results were somewhat equivocal.
 They fit both ``simple'' and ``complex'' versions of each model, where the complex versions of these models introduce additional parameters \(d'\) and \(N'\) that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are crucial to both models' explanations of conjunction fallacies---key findings in the probability judgment literature (Costello \& Watts, 2017; Zhu, Sanborn, \& Chater, 2020).
+These additional parameters are crucial to both models' explanations of conjunction fallacies---key findings in the probability judgment literature \autocite{costello.watts2017,zhu.etal2020}.
 They compared the fits of these models to data via Bayesian Information Criteria score.
-Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs (Zhu, Sanborn, \& Chater, 2020, Supplementary materials).
+Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors' original code and saved model outputs \autocite[Supplementary materials]{zhu.etal2020}.
 
@@
 Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding (Mellers, Hertwig, \& Kahneman, 2001; Sides, Osherson, Bonini, \& Viale, 2002) and are clearly present in the data collected by Zhu and colleagues (2020).
+But conjunction fallacies are an extremely robust empirical finding \autocite{mellers.etal2001,sides.etal2002} and are clearly present in the data collected by Zhu and colleagues \autocite*{zhu.etal2020}.
 So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
-For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
+For their part, Zhu and colleagues \autocite*{zhu.etal2020} elect to split the baby by averaging the simple and complex model scores together.
 This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
@@
 When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to ``overfit'' the data.
-Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data (Gelman, Hwang, \& Vehtari, 2014).
-Zhu and colleagues (2020) warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the models' complexity.
+Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data \autocite{gelman.etal2014}.
+Zhu and colleagues \autocite*{zhu.etal2020} warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the models' complexity.
 
@@
 First, the models differ not only in the number of parameters but in the domain of those parameters.
-Zhu and colleagues (2020) assume that the Bayesian Sampler model's prior distribution should reflect ignorance or a lack of information.
-A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(\beta \in [0, 1]\) (Jaynes, 2003).
+Zhu and colleagues \autocite*{zhu.etal2020} assume that the Bayesian Sampler model's prior distribution should reflect ignorance or a lack of information.
+A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(\beta \in [0, 1]\) \autocite{jaynes2003}.
 Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches \(\infty\) and \(\beta\) approaches 1), whereas the PT+N model permits noise values between \([0, 1/2]\).
@@
 Second, it is not immediately clear how the models' structural differences impact their flexibility.
-Zhu and colleagues' (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
+Zhu and colleagues' \autocite*{zhu.etal2020} bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
 But, what impact does the PT+N model's treatment of conditional probabilities have on model complexity?
@@
 
-Finally, one potential explanation for the relative weakness of the ``complex'' model variants is Zhu and colleagues' (2020) use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
+Finally, one potential explanation for the relative weakness of the ``complex'' model variants is Zhu and colleagues' \autocite*{zhu.etal2020} use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
 In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \(WAIC\) and \(LOO_{ic}\).
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \({\text{WAIC}}\) and \(\text{LOO}_{ic}\).
 Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters (Gelman, Hwang, \& Vehtari, 2014; Vehtari, Gelman, \& Gabry, 2017).
+Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters \autocite{gelman.etal2014,vehtari.etal2017}.
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
@@
 
-Zhu, Sanborn, \& Chater (2020) conducted two experiments to compare the PT+N and Bayesian sampler theories.
+Zhu, Sanborn, \& Chater \autocite*{chater.etal2020} conducted two experiments to compare the PT+N and Bayesian Sampler theories.
 These experiments asked participants to judge the probability of different events in various combinations.
-Following prior work by Costello and Watts (e.g.
-2016, 2018), both experiments focused on the everyday events of different kinds of weather.
+Following prior work by Costello and Watts \autocites*[e.g.][]{costello.watts2016,costello.watts2018}, both experiments focused on the everyday events of different kinds of weather.
 
@@
 For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero---failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day.
-Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.
+Given these potentially aberrant responses, I followed Zhu and colleagues \autocite*{zhu.etal2020} in ignoring data from this condition.
 
@@
 
-I implement both the Bayesian Sampler and PT+N models as Bayesian models in the probabilistic programming language Numpyro.
+I implement several variants of the Bayesian Sampler and PT+N models in a Bayesian framework.
+These models were implemented in the probabilistic programming language Numpyro.
 All code and results are available as supplemental materials (\url{https://github.com/derekpowell/bayesian-sampler}).
@@
 Recall that Zhu and colleagues (2020) identified a bridging condition relating \(\beta\) and \(N\) in the Bayesian Sampler model to the \(d\) parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied \(d\) and \(d'\), rather than directly according to its \(\beta\), \(N\), and N' parameters.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied \(d\) and \(d'\), rather than directly according to its \(\beta\), \(N\), and \(N'\) parameters.
 I constrain \(d\) to \([0, 1/3]\) for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
@@
 The hierarchical implementation adds parameters for the population-level \(d\) and \(d'\) as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
-For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (Papaspiliopoulos, Roberts, \& Sköld, 2007).
+For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency \autocite{papaspiliopoulos.etal2007}.
 Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of \(\beta > 1\).
-Restricting \(\beta\) to {[}0,1{]} restricts the prior distribution of the Bayesian sampler to the class of ``ignorance priors'' (Zhu, Sanborn, \& Chater, 2020).
+Restricting \(\beta\) to {[}0,1{]} restricts the prior distribution of the Bayesian sampler to the class of ``ignorance priors'' \autocite{zhu.etal2020}.
 However, it is also possible that people bring informative priors to the probability judgment task.
-Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted (see e.g., Fennell \& Baddeley, 2012).
+Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted \autocite[see e.g.,][]{fennell.baddeley2012}.
 If \(\beta\) is unrestricted, allowed to fall in the domain \([0, \infty]\) then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same \([0, 1/2]\) range as the PT+N model.
@@
 
-I fit each of the models specified above to data from Zhu et al's (2020) Experiment 1 and 2 and computed \(LOO_{ic}\) scores for each combination.
-Compared with BIC, \(LOO_{ic}\) offers more sophisticated methods for estimating model complexity, and are generally considered more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model (Vehtari, Simpson, Yao, \& Gelman, 2019).
-Model posteriors were estimated using the Numpyro (Phan, Pradhan, \& Jankowiak, 2019) implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
-For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to \(\hat{R}\) (see Gelman et al., 2014).
-Figure 3 below displays the estimated differences in \(LOO_{ic}\) scores for each of the models as compared to the best-scoring model.
+I fit each of the models specified above to data from Zhu et al's \autocite*{zhu.etal2020} Experiment 1 and 2 and computed \(\text{LOO}_{ic}\) scores for each combination.
+Compared with BIC, \(\text{LOO}_{ic}\) offers a more sophisticated estimation of model complexity and is more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model \autocite{vehtari.etal2019}.
+Model posteriors were estimated using the Numpyro \autocite{phan.etal2019} implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
+For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to \(\hat{R}\) \autocite[see][]{gelman.etal2014a}.
+Figure 3 below displays the estimated differences in \(\text{LOO}_{ic}\) scores for each of the models as compared to the best-scoring model.
 
@@
 However, there is no clear winner.
-The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for \(LOO_{ic}\) are better).
+The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for \(\text{LOO}_{ic}\) are better).
 That is, a Bayesian Sampler model without any restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]}.
-However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting \(\beta \in [0,1]\)) have \(\text{LOO}_{ic}\) scores within two standard errors of the difference, indicating that these models are also plausible.
+However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting \(\beta \in [0,1]\)) have \(\text{LOO}_{ic}\) scores within two standard errors of the difference, indicating that these models are also plausible \autocite{sivula.etal2020}.
 
 Data from Experiment 2 more decisively reveal a single winning model: the hierarchical ``unrestricted'' implementation of the Bayesian Sampler model allowing for informative priors.
-Recall, this model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
-These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 
 Figure 4 shows the posterior distributions of the population-level \(d\) and \(d'\) parameters inferred from the unrestricted Bayesian Sampler model.
-In Experiment 2, population-level estimates of \(d\) and \(d'\) are greater than \(\frac{1}{3}\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
+In Experiment 2, population-level estimates of \(d\) and \(d'\) are greater than \(1/3\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
 Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' \(d\) and \(d'\) estimates also lie outside this range (13 of 59 for d, 26 of 59 for \(d'\)).
@@
 
+Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments.
+So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
+Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
+These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of \(d\) and \(d'\).
@@
 Perhaps surprisingly, the Bayesian Sampler with unrestricted \(\beta\) actually receives a smaller penalty term than the more ``restricted'' version of the model for the data in Experiment 2.
-This is at first counterintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data (see Gelman, Hwang, \& Vehtari, 2014).
-To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data (2014).
+This is at first counterintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data \autocite[see][]{gelman.etal2014}.
+To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data \autocite*{gelman.etal2014}.
 If the estimated value is some very large positive number, then the constraint won't have been very informative.
@@
 The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria.
-Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives (e.g.
-Gronau \& Wagenmakers, 2019; Vehtari, Simpson, Yao, \& Gelman, 2019).
+Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives \autocites[e.g.][]{gronau.wagenmakers2019,vehtari.etal2019}.
 Fortunately, conclusions from the comparisons here do not rest solely on differences between the models' complexity.
@@
 
-Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
+Finally, it is worth noting that the best of these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
 
@@
 
-By a fair margin, the model best accounting for the experimental data from Zhu and colleagues (2020) was a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters.
+By a fair margin, the model best accounting for the experimental data from Zhu and colleagues \autocite*{zhu.etal2020} was a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters.
 Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
-Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task.
-In keeping with the larger theoretical framework of Bayesian cognitive science, this theory assumes that subjective probabilities underlie people's probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process (Chater et al., 2020; Zhu, Sanborn, \& Chater, 2020).
+Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this task.
+In keeping with the larger theoretical framework of Bayesian cognitive science, the Bayesian Sampler theory assumes that subjective probabilities underlie people's probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process \autocite{chater.etal2020,zhu.etal2020}.
 
-Findings from Zhu and colleagues' experiments do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta \in [0, 1]\).
+Zhu and colleagues' findings do cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta \in [0, 1]\).
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated \(d\) and \(d'\) parameters outside the {[}1, \(1/3\){]} range implied by ``ignorance'' priors.
+Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated \(d\) and \(d'\) parameters outside the \([0, 1/3]\) range implied by ``ignorance'' priors.
 It could be that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
@@
 One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given the models' tight connections via the bridging condition (Zhu, Sanborn, \& Chater, 2020), it may not be possible to draw decisive conclusions here.
-Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that ``noise'' might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler (2020).
+Given the models' tight connections via the bridging condition \autocite{zhu.etal2020}, it may not be possible to draw decisive conclusions here.
+Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that ``noise'' might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler \autocite*{zhu.etal2020}.
 
-Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs (Chater et al., 2020).
-Indeed, by representing the true subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
+Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs \autocite{chater.etal2020}.
+Indeed, by representing the ``true'' subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
 Examining the model posteriors here reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
-Of course, Zhu and colleagues' (2020) experiments were never designed for this purpose.
-Future research could explore how estimates of people's credences might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning (e.g.
-Franke et al., 2016; Griffiths \& Tenenbaum, 2006; Jern, Chang, \& Kemp, 2014).
-One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes (e.g.
-Cook \& Lewandowsky, 2016; Jern, Chang, \& Kemp, 2014; Powell, Weisman, \& Markman, 2018).
+Of course, Zhu and colleagues' \autocite*{zhu.etal2020} experiments were never designed for this purpose.
+Future research could explore how estimates of people's credences might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning \autocites[e.g.][]{franke.etal2016,jern.etal2014,griffiths.tenenbaum2006}.
+One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes \autocites[e.g.][]{cook.lewandowsky2016,jern.etal2014,powell.etal2018}.
 
@@
 \hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}}
 
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-anderson1991}{}%
-Anderson, J. R. (1991). The adaptive nature of human categorization. \emph{Psychological Review}, \emph{98}(3), 409--429. \url{https://doi.org/10.1037/0033-295X.98.3.409}
-
-\leavevmode\hypertarget{ref-chater.etal2020}{}%
-Chater, N., Zhu, J.-Q., Spicer, J., Sundh, J., León-Villagrá, P., \& Sanborn, A.
-(2020).
-Probabilistic {Biases Meet} the {Bayesian Brain}.
-\emph{Current Directions in Psychological Science}, \emph{29}(5), 506--512.
-\url{https://doi.org/10.1177/0963721420954801}
-
-\leavevmode\hypertarget{ref-cook.lewandowsky2016}{}%
-Cook, J., \& Lewandowsky, S.
-(2016).
-Rational {Irrationality}: {Modeling Climate Change Belief Polarization Using Bayesian Networks}.
-\emph{Topics in Cognitive Science}, \emph{8}(1), 160--179.
-\url{https://doi.org/10.1111/tops.12186}
-
-\leavevmode\hypertarget{ref-costello.watts2014}{}%
-Costello, F., \& Watts, P.
-(2014).
-Surprisingly rational: {Probability} theory plus noise explains biases in judgment.
-\emph{Psychological Review}, \emph{121}(3), 463--480.
-\url{https://doi.org/10.1037/a0037010}
-
-\leavevmode\hypertarget{ref-costello.watts2016}{}%
-Costello, F., \& Watts, P.
-(2016).
-People's conditional probability judgments follow probability theory (plus noise).
-\emph{Cognitive Psychology}, \emph{89}, 106--133.
-\url{https://doi.org/10.1016/j.cogpsych.2016.06.006}
-
-\leavevmode\hypertarget{ref-costello.watts2017}{}%
-Costello, F., \& Watts, P.
-(2017).
-Explaining {High Conjunction Fallacy Rates}: {The Probability Theory Plus Noise Account}.
-\emph{Journal of Behavioral Decision Making}, \emph{30}(2), 304--321.
-\url{https://doi.org/10.1002/bdm.1936}
-
-\leavevmode\hypertarget{ref-costello.watts2018}{}%
-Costello, F., \& Watts, P. (2018). Invariants in probabilistic reasoning. \emph{Cognitive Psychology}, \emph{100}, 1--16. \url{https://doi.org/10.1016/j.cogpsych.2017.11.003}
-
-\leavevmode\hypertarget{ref-dasgupta.etal2017}{}%
-Dasgupta, I., Schulz, E., \& Gershman, S. J. (2017). Where do hypotheses come from? \emph{Cognitive Psychology}, \emph{96}, 1--25. \url{https://doi.org/10.1016/j.cogpsych.2017.05.001}
-
-\leavevmode\hypertarget{ref-fennell.baddeley2012}{}%
-Fennell, J., \& Baddeley, R.
-(2012).
-Uncertainty plus prior equals rational bias: {An} intuitive {Bayesian} probability weighting function.
-\emph{Psychological Review}, \emph{119}(4), 878--887.
-\url{https://doi.org/10.1037/a0029346}
-
-\leavevmode\hypertarget{ref-franke.etal2016}{}%
-Franke, M., Dablander, F., Scholler, A., Bennett, E., Degen, J., Tessler, M.
-H., \ldots{} Goodman, N.
-D.
-(2016).
-What does the crowd believe?
-{A} hierarchical approach to estimating subjective beliefs from empirical data, 6.
-
-\leavevmode\hypertarget{ref-gelman.etal2014a}{}%
-Gelman, A., Carlin, J. B., Stern, H. S., Dunson, D. B., Vehtari, A., \& Rubin, D. B. (2014). \emph{Bayesian data analysis} (Third edition). {Boca Raton}: {CRC Press}.
-
-\leavevmode\hypertarget{ref-gelman.etal2014}{}%
-Gelman, A., Hwang, J., \& Vehtari, A.
-(2014).
-Understanding predictive information criteria for {Bayesian} models.
-\emph{Statistics and Computing}, \emph{24}(6), 997--1016.
-\url{https://doi.org/10.1007/s11222-013-9416-2}
-
-\leavevmode\hypertarget{ref-griffiths.tenenbaum2006}{}%
-Griffiths, T.
-L., \& Tenenbaum, J.
-B.
-(2006).
-Optimal {Predictions} in {Everyday Cognition}.
-\emph{Psychological Science}, \emph{17}(9), 767--773.
-\url{https://doi.org/10.1111/j.1467-9280.2006.01780.x}
-
-\leavevmode\hypertarget{ref-gronau.wagenmakers2019}{}%
-Gronau, Q.
-F., \& Wagenmakers, E.-J.
-(2019).
-Limitations of {Bayesian Leave}-{One}-{Out Cross}-{Validation} for {Model Selection}.
-\emph{Computational Brain \& Behavior}, \emph{2}(1), 1--11.
-\url{https://doi.org/10.1007/s42113-018-0011-7}
-
-\leavevmode\hypertarget{ref-jaynes2003}{}%
-Jaynes, E. T. (2003). \emph{Probability {Theory}: {The Logic} of {Science}}. (G. L. Bretthorst, Ed.). {Cambridge, UNITED KINGDOM}: {Cambridge University Press}.
-
-\leavevmode\hypertarget{ref-jern.etal2014}{}%
-Jern, A., Chang, K. K., \& Kemp, C. (2014). Belief polarization is not always irrational. \emph{Psychological Review}, \emph{121}(2), 206--224. \url{https://doi.org/10.1037/a0035941}
-
-\leavevmode\hypertarget{ref-kahneman2013}{}%
-Kahneman, D. (2013). \emph{Thinking, {Fast} and {Slow}} (1st edition). {New York}: {Farrar, Straus and Giroux}.
-
-\leavevmode\hypertarget{ref-kersten.etal2004}{}%
-Kersten, D., Mamassian, P., \& Yuille, A.
-(2004).
-Object {Perception} as {Bayesian Inference}.
-\emph{Annual Review of Psychology}, \emph{55}(1), 271--304.
-\url{https://doi.org/10.1146/annurev.psych.55.090902.142005}
-
-\leavevmode\hypertarget{ref-lu.etal2012}{}%
-Lu, H., Chen, D., \& Holyoak, K. J. (2012). Bayesian analogy with relational transformations. \emph{Psychological Review}, \emph{119}(3), 617--648. \url{https://doi.org/10.1037/a0028719}
-
-\leavevmode\hypertarget{ref-mellers.etal2001}{}%
-Mellers, B., Hertwig, R., \& Kahneman, D.
-(2001).
-Do {Frequency Representations Eliminate Conjunction Effects}?
-{An Exercise} in {Adversarial Collaboration}.
-\emph{Psychological Science}, \emph{12}(4), 269--275.
-\url{https://doi.org/10.1111/1467-9280.00350}
-
-\leavevmode\hypertarget{ref-papaspiliopoulos.etal2007}{}%
-Papaspiliopoulos, O., Roberts, G.
-O., \& Sköld, M.
-(2007).
-A {General Framework} for the {Parametrization} of {Hierarchical Models}.
-\emph{Statistical Science}, \emph{22}(1).
-\url{https://doi.org/10.1214/088342307000000014}
-
-\leavevmode\hypertarget{ref-phan.etal2019}{}%
-Phan, D., Pradhan, N., \& Jankowiak, M.
-(2019).
-Composable {Effects} for {Flexible} and {Accelerated Probabilistic Programming} in {NumPyro}.
-\emph{arXiv:1912.11554 {[}Cs, Stat{]}}.
-Retrieved from \url{http://arxiv.org/abs/1912.11554}
-
-\leavevmode\hypertarget{ref-powell.etal2018}{}%
-Powell, D., Weisman, K., \& Markman, E. M. (2018). Articulating lay theories through graphical models: {A} study of beliefs surrounding vaccination decisions, 6.
-
-\leavevmode\hypertarget{ref-sides.etal2002}{}%
-Sides, A., Osherson, D., Bonini, N., \& Viale, R. (2002). On the reality of the conjunction fallacy. \emph{Memory \& Cognition}, \emph{30}(2), 191--198. \url{https://doi.org/10.3758/BF03195280}
-
-\leavevmode\hypertarget{ref-tenenbaum.etal2011}{}%
-Tenenbaum, J.
-B., Kemp, C., Griffiths, T.
-L., \& Goodman, N.
-D.
-(2011).
-How to {Grow} a {Mind}: {Statistics}, {Structure}, and {Abstraction}.
-\emph{Science}, \emph{331}(6022), 1279--1285.
-\url{https://doi.org/10.1126/science.1192788}
-
-\leavevmode\hypertarget{ref-tversky.kahneman1983}{}%
-Tversky, A., \& Kahneman, D. (1983). Extensional {Versus Intuitive Reasoning}: {The Conjunction Fallacy} in {Probability Judgment}. \emph{Psychological Review}, \emph{90}(4), 23.
-
-\leavevmode\hypertarget{ref-tversky.koehler1994}{}%
-Tversky, A., \& Koehler, D.
-J.
-(1994).
-Support theory: {A} nonextensional representation of subjective probability.
-\emph{Psychological Review}, \emph{101}(4), 547--567.
-https://doi.org/\url{http://dx.doi.org.ezproxy1.lib.asu.edu/10.1037/0033-295X.101.4.547}
-
-\leavevmode\hypertarget{ref-vehtari.etal2017}{}%
-Vehtari, A., Gelman, A., \& Gabry, J.
-(2017).
-Practical {Bayesian} model evaluation using leave-one-out cross-validation and {WAIC}.
-\emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
-\url{https://doi.org/10.1007/s11222-016-9696-4}
-
-\leavevmode\hypertarget{ref-vehtari.etal2019}{}%
-Vehtari, A., Simpson, D.
-P., Yao, Y., \& Gelman, A.
-(2019).
-Limitations of {``{Limitations} of {Bayesian Leave}-one-out {Cross}-{Validation} for {Model Selection}.''} \emph{Computational Brain \& Behavior}, \emph{2}(1), 22--27.
-\url{https://doi.org/10.1007/s42113-018-0020-6}
-
-\leavevmode\hypertarget{ref-xu.tenenbaum2007}{}%
-Xu, F., \& Tenenbaum, J. B. (2007). Word learning as {Bayesian} inference. \emph{Psychological Review}, \emph{114}(2), 245--272. \url{https://doi.org/10.1037/0033-295X.114.2.245}
-
-\leavevmode\hypertarget{ref-zhu.etal2020}{}%
-Zhu, J.-Q., Sanborn, A.
-N., \& Chater, N.
-(2020).
-The {Bayesian} sampler: {Generic Bayesian} inference causes incoherence in human probability judgments.
-\emph{Psychological Review}, \emph{127}(5), 719--748.
-\url{https://doi.org/10.1037/rev0000190}
-
-\end{CSLReferences}
+\printbibliography
 

### 2021-10-27 paper-rmd/paper-papaja.Rmd
initial submission

@@
     institution   : "Arizona State University, School of Social and Behavioral Sciences"
-  # - id            : "2"
-  #   institution   : "Konstanz Business School"
 
 authornote: |
- This manuscript is pre-print that has not yet been peer reviewed.
+ This manuscript is pre-print that has not yet been peer reviewed (October 27, 2021). This study was not preregistered.
 
 abstract: |
- Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic.
-However, one mark sometimes held against these theories is that people’s explicit probability judgments often violate the laws of probability.
+ Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but one mark sometimes held against these theories is that people’s explicit probability judgments often violate the laws of probability.
 Two recent proposals, the "Probability Theory plus Noise" [@costello.watts2014] and "Bayesian Sampler" [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
@@
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
+Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a modified version of the Bayesian Sampler model under which people may hold informative priors about probabilities.
 
@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "X"
+wordcount         : "4617"
 
@@
   - \usetikzlibrary{bayesnet}
+  - \mathchardef\hyph="2D
 ---
@@
 Bayesian theories of cognition offer a unified formal framework for cognitive science [@tenenbaum.etal2011] that has had remarkable explanatory successes across domains, including in perception [e.g.
-@kersten.etal2004], memory [@anderson1991], language [@xu.tenenbaum2007], and reasoning [@lu.etal2012].
+@kersten.etal2004], memory [e.g.
+@anderson1991], language [e.g.
+@xu.tenenbaum2007], and reasoning [e.g.
+@lu.etal2012].
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
@@
 
-Unfortunately, people’s explicit probability judgments routinely violate the axioms of probability theory.
+Unfortunately, people’s explicit probability judgments routinely violate the most basic axioms of probability theory.
 For example, human probability judgments often exhibit the "conjunction fallacy": people will often judge the conjunction of two events (e.g.
@@
 By assumption, a maximum of 50% of samples can be misread, so that $d$ is a number in the range $[0, 1/2]$.
-The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
+The PT+N theory provides a unified account for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
 
@@
 Then, those "observed" samples are integrated with a prior over probabilities to produce a probability judgment.
-This prior takes the form of a symmetric Beta distribution parameterized by $\beta$.
+This prior takes the form of a symmetric Beta distribution, $Beta(\beta, \beta)$.
 After observing $S(A)$ successes and $N - S(A)$ failures, the posterior over probabilities is distributed $Beta(\beta + S(A), \beta + N - S(A))$.
 Zhu and colleagues [-@zhu.etal2020] assume that people report the mean of their posterior probability estimates.
-Since for any Beta distribution $x \sim Beta(a,b)$, $E[x] = \frac{a}{a+b}$, the probability estimate is thus a linear function of S, N, and $\beta$.
+For any Beta distribution $x \sim Beta(a,b)$, $E[x] = \frac{a}{a+b}$.
+So, the expected probability estimate is a linear function of S, N, and $\beta$.
 
@@
 Bayesian cognitive theories are fundamentally theories of inductive reasoning: Bayes' rule describes how existing beliefs should be updated conditional on the observation of different kinds of evidence.
-So, how they treat the conditioning of beliefs is at the heart of these theories.
+So, treatment of the conditioning of beliefs is at the heart of these theories.
 
@@
 
-The PT+N’s account of conditional probability judgment is distinctly non-Bayesian, separating it quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
+This non-Bayesian account of conditional probabiliity judgments separates the PT+N theory quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
 
@@
 So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
-For their part, Zhu and colleagues [-@zhu.etal2020] elect to split the baby by averaging the simple and complex model scores together.
+For their part, Zhu and colleagues [-@zhu.etal2020] chose to average the simple and complex model scores together.
 This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as ${\text{WAIC}}$ and $\text{LOO}_{ic}$.
-Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as the widely-applicable information criterion (${\text{WAIC}}$) and Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$).
+These approaches measure the ability of a model to capture new unseen data by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data.
+Importantly, these the methods for computing the $\widehat{\text{elpd}}$ penalize model complexity in a way that directly captures differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
@@
 
-The PT+N model defines probability judgments as:
+The PT+N model defines expected probability judgments ($P_e$) as:
 
@@
 
-In contrast, the Bayesian Sampler model defines probability judgments as:
+In contrast, the Bayesian Sampler model defines expected probability judgments as:
 
@@
   P_{e}(A \land B) &= \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta} \\
-  P_{e}(A \lor B) &= \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta} 
+  P_{e}(A \lor B) &= \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta} \\
+  P_{e}(A|B) &= \frac{N}{N + 2 \beta}P(A|B) + \frac{\beta}{N+2 \beta}
 \end{align*}
@@
     \begin{align*}
-      y_{ijk} &\sim beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
+      y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
       \mu_{ijk} &= f_{PT+N}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
-      d_j &= \frac{expit(\delta_j)}{2} \\
-      d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2} \\
+      d_j &= \frac{1}{2} \ \text{logistic}(\delta_j) \\
+      d_j’ &= \frac{1}{2} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
       \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
@@
       \Delta\delta_j &\sim Normal(0,1) \\
-      \kappa &\sim HalfCauchy(20)
+      \kappa &\sim \mathit{Half \hyph Cauchy}(20)
     \end{align*}
@@
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating $\beta$ and $N$ in the Bayesian Sampler model to the $d$ parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and $N'$ parameters.
-I constrain $d$ to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
+Recall that Zhu and colleagues [-@zhu.etal2020] identified a bridging condition relating $\beta$ and $N$ in the Bayesian Sampler model to the $d$ parameter of the PT+N model.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and $N'$ parameters.^[Strictly speaking, under the original form of the Bayesian sampler model, $N$ and $N'$ are discrete parameters representing the number of distinct independent samples drawn.
+Given a particular implied $d$, this could create constraints on the possible values of $d'$, assuming $\beta$ is held constant.
+However, Zhu and colleagues [-@zhu.etal2020] also consider the possibility that people draw non-independent mental samples, in which case $N$ and $N'$ would represent the _effective number of samples_, accounting for their autocorrelation.
+In this case, we could treat this effective number of samples as a continuous quantity, and therefore imagine there are no clear constraints on $d$ and $d'$ except the stipulation that $d \leq d'$.] I constrain $d$ to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
@@
   \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
-  d_j &= \frac{expit(\delta_j)}{3} \\
-  d_j’ &= \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{3}
+  d_j &= \frac{1}{3} \ \text{logistic}(\delta_j) \\
+  d_j’ &= \frac{1}{3} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big)
 \end{align*}
@@
 
-I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed $\text{LOO}_{ic}$ scores for each combination.
-Compared with BIC, $\text{LOO}_{ic}$ offers a more sophisticated estimation of model complexity and is more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
+I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and estimated the expected log predictive density with PSIS-LOO ($\widehat{\text{elpd}}_{\text{LOO}}$) for each combination.
+Compared with BIC, $\widehat{\text{elpd}}_{\text{LOO}}$ offers a more sophisticated estimation of model complexity and is more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
 Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
 For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$ [see @gelman.etal2014a].
-Figure 3 below displays the estimated differences in $\text{LOO}_{ic}$ scores for each of the models as compared to the best-scoring model.
+Figure 3 below displays the estimated differences in $\widehat{\text{elpd}}_{\text{LOO}}$ scores for each of the models as compared to the best-scoring model.
 
@@
 Data from Experiment 1 favor "complex" variants of the Bayesian Sampler model compared with the "simple" variants and all versions of the PT+N model.
-However, there is no clear winner.
-The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for $\text{LOO}_{ic}$ are better).
-That is, a Bayesian Sampler model without any restriction on $\beta$, allowing it to range from [0, $\infty$].
-However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\text{LOO}_{ic}$ scores within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
+However, there is no clear single winner.
+As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task, that is, a Bayesian Sampler model allowing $\beta \in [0, \infty]$ (greater values of $\widehat{\text{elpd}}_{\text{LOO}}$ are better).
+However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\widehat{\text{elpd}}_{\text{LOO}}$ scores within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
 
@@
 Figure 4 shows the posterior distributions of the population-level $d$ and $d'$ parameters inferred from the unrestricted Bayesian Sampler model.
-In Experiment 2, population-level estimates of $d$  and $d'$ are greater than $1/3$, outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ $d$  and $d'$ estimates also lie outside this range (13 of 59 for d, 26 of 59 for $d'$).
+In Experiment 2, population-level estimates of $d'$ are greater than $1/3$, as are a substantial number of participant-level estimates for $d$ (37 of 83).
+These values fall outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
+Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants’ $d$  and $d'$ estimates also lie outside this range (11 of 59 for $d$, 18 of 59 for $d'$).
 The finding that there are clear differences in $d$  and $d'$ estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
@@
 Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
-And comparing their prediction specifically for conditional probability judgments, the unrestricted Bayesian Sampler provides a better fit, especially for the modestly-correlated events of experiment 2 (Exp 1: response-level $r$ = .91 vs.
+And comparing their prediction specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level $r$ = .91 vs.
 .89; query-level $r$ = .98 vs.
@@
 .67; query-level $r$ = .92 vs.
-.85) These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
+.85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
+
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of $d$  and $d'$.
@@
 @gronau.wagenmakers2019; @vehtari.etal2019].
-Fortunately, conclusions from the comparisons here do not rest solely on differences between the models’ complexity.
 
@@
     booktabs=T,
-    col.names = c("Model", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{query}$", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{query}$"),
+    col.names = c("Model", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "Penalty","$r_{resp}$", "$r_{query}$", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "Penalty","$r_{resp}$", "$r_{query}$"),
     escape = F
@@
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated $d$  and $d'$ parameters outside the $[0, 1/3]$ range implied by "ignorance" priors.
-It could be that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-This could potentially account for the differences in $d$ and $d'$ observed across Experiments 1 and 2.
+Nevertheless, many participants' estimated $d$ and $d'$ parameters fell outside the $[0, 1/3]$ range implied by "ignorance" priors.
+These data suggest that people may instead bring informative priors to some probability judgment tasks.
+Given the different values of $d$ and $d'$ across Experiments 1 and 2, it seems likely that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
 
 In addition, the implied $d$ and $d'$ noise parameters also varied across individuals.
-Of note, even in Experiment 2, some participant's implied $d$ and $d'$ parameters were consistent with the class of ignorance priors (see supplemental materials).
+Of note, some participant's implied $d$ and $d'$ parameters _were_ consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
 Further research should explore this heterogeneity within and across individuals.
@@
 One thing this model comparison has not decided, and likely *cannot* decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given the models' tight connections via the bridging condition [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
+Given the two models' tight connections via the bridging condition [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
 Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that "noise" might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].
@@
 @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].
-
-\newpage
-
-# References

### 2021-10-27 paper-rmd/paper-papaja.tex
initial submission

@@
 \shorttitle{Comparing probability judgment accounts}
-\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: X}
+\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: 4617}
 \usepackage{csquotes}
@@
 \usetikzlibrary{bayesnet}
+\mathchardef\hyph="2D
 \ifxetex
@@
 
-This manuscript is pre-print that has not yet been peer reviewed.
+This manuscript is pre-print that has not yet been peer reviewed (October 27, 2021). This study was not preregistered.
 
@@
 \abstract{
-Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic.
-However, one mark sometimes held against these theories is that people's explicit probability judgments often violate the laws of probability.
+Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but one mark sometimes held against these theories is that people's explicit probability judgments often violate the laws of probability.
 Two recent proposals, the ``Probability Theory plus Noise'' \autocite{costello.watts2014} and ``Bayesian Sampler'' \autocite{zhu.etal2020} theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
@@
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Comparing the fits of both models on data collected by Zhu and colleagues \autocite*{zhu.etal2020} I find the data are best explained by a modified version of the Bayesian Sampler model that allows for informative priors.
+Comparing the fits of both models on data collected by Zhu and colleagues \autocite*{zhu.etal2020} I find the data are best explained by a modified version of the Bayesian Sampler model under which people may hold informative priors about probabilities.
 }
@@
 
-Bayesian theories of cognition offer a unified formal framework for cognitive science \autocite{tenenbaum.etal2011} that has had remarkable explanatory successes across domains, including in perception \autocite[e.g.][]{kersten.etal2004}, memory \autocite{anderson1991}, language \autocite{xu.tenenbaum2007}, and reasoning \autocite{lu.etal2012}.
+Bayesian theories of cognition offer a unified formal framework for cognitive science \autocite{tenenbaum.etal2011} that has had remarkable explanatory successes across domains, including in perception \autocite[e.g.][]{kersten.etal2004}, memory \autocite[e.g.][]{anderson1991}, language \autocite[e.g.][]{xu.tenenbaum2007}, and reasoning \autocite[e.g.][]{lu.etal2012}.
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
@@
 
-Unfortunately, people's explicit probability judgments routinely violate the axioms of probability theory.
+Unfortunately, people's explicit probability judgments routinely violate the most basic axioms of probability theory.
 For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory \autocite{tversky.kahneman1983}.
@@
 By assumption, a maximum of 50\% of samples can be misread, so that \(d\) is a number in the range \([0, 1/2]\).
-The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions \autocite{costello.watts2014,costello.watts2016,costello.watts2017,costello.watts2018}.
+The PT+N theory provides a unified account for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions \autocite{costello.watts2014,costello.watts2016,costello.watts2017,costello.watts2018}.
 
@@
 Then, those ``observed'' samples are integrated with a prior over probabilities to produce a probability judgment.
-This prior takes the form of a symmetric Beta distribution parameterized by \(\beta\).
+This prior takes the form of a symmetric Beta distribution, \(Beta(\beta, \beta)\).
 After observing \(S(A)\) successes and \(N - S(A)\) failures, the posterior over probabilities is distributed \(Beta(\beta + S(A), \beta + N - S(A))\).
 Zhu and colleagues \autocite*{zhu.etal2020} assume that people report the mean of their posterior probability estimates.
-Since for any Beta distribution \(x \sim Beta(a,b)\), \(E[x] = \frac{a}{a+b}\), the probability estimate is thus a linear function of S, N, and \(\beta\).
+For any Beta distribution \(x \sim Beta(a,b)\), \(E[x] = \frac{a}{a+b}\).
+So, the expected probability estimate is a linear function of S, N, and \(\beta\).
 
@@
 
-The PT+N's account of conditional probability judgment is distinctly non-Bayesian, separating it quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
+This non-Bayesian account of conditional probabiliity judgments separates the PT+N theory quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
 
@@
 So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
-For their part, Zhu and colleagues \autocite*{zhu.etal2020} elect to split the baby by averaging the simple and complex model scores together.
+For their part, Zhu and colleagues \autocite*{zhu.etal2020} chose to average the simple and complex model scores together.
 This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as \({\text{WAIC}}\) and \(\text{LOO}_{ic}\).
-Both of these indices approximate leave-one-out cross validation, thereby offering a measure of the ability of a model to capture new unseen data.
-Importantly, these indices also allow for complexity penalties that directly capture differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters \autocite{gelman.etal2014,vehtari.etal2017}.
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as the widely-applicable information criterion (\({\text{WAIC}}\)) and Pareto smoothed importance sampling approximate leave-one-out cross validation (\(\text{PSIS-LOO}\)).
+These approaches measure the ability of a model to capture new unseen data by estimating the expected log predictive density (\(\widehat{\text{elpd}}\)) of the model, or the expected probability of new unseen data.
+Importantly, these the methods for computing the \(\widehat{\text{elpd}}\) penalize model complexity in a way that directly captures differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters \autocite{gelman.etal2014,vehtari.etal2017}.
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
@@
 Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment.
-This is especially clear, as the authors note, for disjunctive trials such as ``normal or typical,'' where ``or typical'' might not be read as a disjunction but rather an elaborative clause.
+This is especially clear, as the authors note, for disjunctive query trials such as ``normal or typical,'' where ``or typical'' might not be read as a disjunction but rather an elaborative clause.
 In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.
@@
 
-The PT+N model defines probability judgments as:
+The PT+N model defines expected probability judgments (\(P_e\)) as:
 
@@
 
-Recall that Zhu and colleagues (2020) identified a bridging condition relating \(\beta\) and \(N\) in the Bayesian Sampler model to the \(d\) parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied \(d\) and \(d'\), rather than directly according to its \(\beta\), \(N\), and \(N'\) parameters.
-I constrain \(d\) to \([0, 1/3]\) for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
+Recall that Zhu and colleagues \autocite*{zhu.etal2020} identified a bridging condition relating \(\beta\) and \(N\) in the Bayesian Sampler model to the \(d\) parameter of the PT+N model.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied \(d\) and \(d'\), rather than directly according to its \(\beta\), \(N\), and \(N'\) parameters.\footnote{Strictly speaking, under the original form of the Bayesian sampler model, \(N\) and \(N'\) are discrete parameters representing the number of distinct independent samples drawn.
+Given a particular implied \(d\), this could create constraints on the possible values of \(d'\), assuming \(\beta\) is held constant.
+However, Zhu and colleagues \autocite*{zhu.etal2020} also consider the possibility that people draw non-independent mental samples, in which case \(N\) and \(N'\) would represent the \emph{effective number of samples}, accounting for their autocorrelation.
+In this case, we could treat this effective number of samples as a continuous quantity, and therefore imagine there are no clear constraints on \(d\) and \(d'\) except the stipulation that \(d \leq d'\).} I constrain \(d\) to \([0, 1/3]\) for the Bayesian Sampler model to reflect the assumption that \(\beta \in [0, 1]\).
 This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
@@
 
-I fit each of the models specified above to data from Zhu et al's \autocite*{zhu.etal2020} Experiment 1 and 2 and computed \(\text{LOO}_{ic}\) scores for each combination.
-Compared with BIC, \(\text{LOO}_{ic}\) offers a more sophisticated estimation of model complexity and is more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model \autocite{vehtari.etal2019}.
+I fit each of the models specified above to data from Zhu et al's \autocite*{zhu.etal2020} Experiment 1 and 2 and estimated the expected log predictive density with PSIS-LOO (\(\widehat{\text{elpd}}_{\text{LOO}}\)) for each combination.
+Compared with BIC, \(\widehat{\text{elpd}}_{\text{LOO}}\) offers a more sophisticated estimation of model complexity and is more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model \autocite{vehtari.etal2019}.
 Model posteriors were estimated using the Numpyro \autocite{phan.etal2019} implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
 For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to \(\hat{R}\) \autocite[see][]{gelman.etal2014a}.
-Figure 3 below displays the estimated differences in \(\text{LOO}_{ic}\) scores for each of the models as compared to the best-scoring model.
+Figure 3 below displays the estimated differences in \(\widehat{\text{elpd}}_{\text{LOO}}\) scores for each of the models as compared to the best-scoring model.
 
@@
 Data from Experiment 1 favor ``complex'' variants of the Bayesian Sampler model compared with the ``simple'' variants and all versions of the PT+N model.
-However, there is no clear winner.
-The beset-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task (again, lower scores for \(\text{LOO}_{ic}\) are better).
-That is, a Bayesian Sampler model without any restriction on \(\beta\), allowing it to range from {[}0, \(\infty\){]}.
-However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting \(\beta \in [0,1]\)) have \(\text{LOO}_{ic}\) scores within two standard errors of the difference, indicating that these models are also plausible \autocite{sivula.etal2020}.
+However, there is no clear single winner.
+As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task, that is, a Bayesian Sampler model allowing \(\beta \in [0, \infty]\) (greater values of \(\widehat{\text{elpd}}_{\text{LOO}}\) are better).
+However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting \(\beta \in [0,1]\)) have \(\widehat{\text{elpd}}_{\text{LOO}}\) scores within two standard errors of the difference, indicating that these models are also plausible \autocite{sivula.etal2020}.
 
@@
 Figure 4 shows the posterior distributions of the population-level \(d\) and \(d'\) parameters inferred from the unrestricted Bayesian Sampler model.
-In Experiment 2, population-level estimates of \(d\) and \(d'\) are greater than \(1/3\), outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' \(d\) and \(d'\) estimates also lie outside this range (13 of 59 for d, 26 of 59 for \(d'\)).
+In Experiment 2, population-level estimates of \(d'\) are greater than \(1/3\), as are a substantial number of participant-level estimates for \(d\) (37 of 83).
+These values fall outside the range implied by the assumption of ``ignorance priors'' in the Bayesian Sampler model.
+Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of individual participants' \(d\) and \(d'\) estimates also lie outside this range (11 of 59 for \(d\), 18 of 59 for \(d'\)).
 The finding that there are clear differences in \(d\) and \(d'\) estimated across experiments suggest that the mental sampling processes producing estimates vary in the different conditions, either in terms of the number of samples that are drawn, the noise in reading those samples, or the form of the prior distribution assumed by participants in each context.
@@
 Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
-These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
+And comparing their prediction specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level \(r\) = .91 vs.~.89; query-level \(r\) = .98 vs.~.97; Exp 2: response-level \(r\) = .72 vs.~.67; query-level \(r\) = .92 vs.~.85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
+
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of \(d\) and \(d'\).
@@
 Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives \autocites[e.g.][]{gronau.wagenmakers2019,vehtari.etal2019}.
-Fortunately, conclusions from the comparisons here do not rest solely on differences between the models' complexity.
 
@@
 \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
-Model & $LOO_{ic}$ & Penalty & $r_{resp}$ & $r_{trial}$ & $LOO_{ic}$ & Penalty & $r_{resp}$ & $r_{trial}$\\
+Model & $\widehat{\text{elpd}}_{\text{LOO}}$ & Penalty & $r_{resp}$ & $r_{query}$ & $\widehat{\text{elpd}}_{\text{LOO}}$ & Penalty & $r_{resp}$ & $r_{query}$\\
 \midrule
-\textbf{Unrestricted BS hier.} & \textbf{-2163.2} & \textbf{296.4} & \textbf{0.884} & \textbf{0.964} & \textbf{-3953.3} & \textbf{368.4} & \textbf{0.688} & \textbf{0.878}\\
-Bayesian Sampler complex & -2134.4 & 285.1 & 0.883 & 0.961 & -3733.2 & 453.6 & 0.679 & 0.848\\
-Bayesian Sampler complex hier. & -2131.9 & 292.0 & 0.883 & 0.960 & -3815.6 & 399.5 & 0.675 & 0.852\\
-Bayesian Sampler simple hier. & -2053.9 & 271.8 & 0.874 & 0.953 & -3790.5 & 383.2 & 0.667 & 0.839\\
-Bayesian Sampler simple & -2051.3 & 273.7 & 0.874 & 0.952 & -3722.8 & 418.9 & 0.667 & 0.831\\
+\textbf{Unrestricted BS hier.} & \textbf{1118.7} & \textbf{259.3} & \textbf{0.884} & \textbf{0.964} & \textbf{1978.6} & \textbf{366.4} & \textbf{0.688} & \textbf{0.878}\\
+Bayesian Sampler complex hier. & 1088.3 & 269.6 & 0.883 & 0.960 & 1912.3 & 395.0 & 0.675 & 0.852\\
+Bayesian Sampler complex & 1087.8 & 264.6 & 0.883 & 0.961 & 1876.9 & 443.4 & 0.679 & 0.848\\
+Bayesian Sampler simple & 1045.6 & 253.8 & 0.874 & 0.952 & 1861.2 & 419.1 & 0.667 & 0.831\\
+Bayesian Sampler simple hier. & 1039.6 & 259.1 & 0.874 & 0.953 & 1900.7 & 377.8 & 0.667 & 0.839\\
 \addlinespace
-PT+N complex hier. & -1927.4 & 298.3 & 0.867 & 0.946 & -3795.3 & 356.2 & 0.658 & 0.835\\
-PT+N complex & -1922.8 & 264.5 & 0.864 & 0.941 & -3768.5 & 398.2 & 0.667 & 0.840\\
-PT+N simple & -1708.8 & 254.5 & 0.838 & 0.918 & -3643.0 & 319.5 & 0.619 & 0.777\\
-PT+N simple hier. & -1697.1 & 265.4 & 0.839 & 0.919 & -3643.8 & 305.3 & 0.617 & 0.772\\
-Relative Freq. & -1266.2 & 305.7 & 0.820 & 0.875 & -1287.7 & 424.6 & 0.516 & 0.639\\
+PT+N complex hier. & 993.0 & 268.9 & 0.867 & 0.946 & 1902.4 & 351.5 & 0.658 & 0.835\\
+PT+N complex & 966.1 & 259.8 & 0.864 & 0.941 & 1886.9 & 395.5 & 0.667 & 0.840\\
+PT+N simple hier. & 864.0 & 250.0 & 0.839 & 0.919 & 1821.9 & 305.3 & 0.617 & 0.772\\
+PT+N simple & 863.5 & 245.5 & 0.838 & 0.918 & 1821.5 & 319.5 & 0.619 & 0.777\\
+Relative Freq. & 649.3 & 289.4 & 0.820 & 0.875 & 643.8 & 424.6 & 0.516 & 0.639\\
 \bottomrule
@@
 
-Finally, it is worth noting that the best of these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
+Finally, it is worth noting that the best of these models provide quite strong overall fits to the data, not just for the query averages, but also for the query averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
 
@@
 As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people may bring informative priors to the probability judgment task, indicated by estimated \(d\) and \(d'\) parameters outside the \([0, 1/3]\) range implied by ``ignorance'' priors.
-It could be that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-This could potentially account for the differences in \(d\) and \(d'\) observed across Experiments 1 and 2.
+Nevertheless, many participants' estimated \(d\) and \(d'\) parameters fell outside the \([0, 1/3]\) range implied by ``ignorance'' priors.
+These data suggest that people may instead bring informative priors to some probability judgment tasks.
+Given the different values of \(d\) and \(d'\) across Experiments 1 and 2, it seems likely that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
 
 In addition, the implied \(d\) and \(d'\) noise parameters also varied across individuals.
-Of note, even in Experiment 2, some participant's implied \(d\) and \(d'\) parameters were consistent with the class of ignorance priors (see supplemental materials).
+Of note, some participant's implied \(d\) and \(d'\) parameters \emph{were} consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
 Further research should explore this heterogeneity within and across individuals.
@@
 One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given the models' tight connections via the bridging condition \autocite{zhu.etal2020}, it may not be possible to draw decisive conclusions here.
+Given the two models' tight connections via the bridging condition \autocite{zhu.etal2020}, it may not be possible to draw decisive conclusions here.
 Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that ``noise'' might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler \autocite*{zhu.etal2020}.
@@
 
-\newpage
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
 \printbibliography

### 2021-10-27 paper-rmd/paper-papaja.Rmd
initial submission 2

@@
 authornote: |
- This manuscript is pre-print that has not yet been peer reviewed (October 27, 2021). This study was not preregistered.
+ This manuscript has not yet been peer-reviewed. This study was not preregistered.
 
@@
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that that these theories have often been accused of failing to describe human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", "vaccines are safe," or "this politician is trustworthy".
+It is somewhat embarrassing then, that these theories have often been accused of failing to describe human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", "vaccines are safe," or "this politician is trustworthy".
 

### 2021-10-27 paper-rmd/paper-papaja.tex
initial submission 2

@@
 
-This manuscript is pre-print that has not yet been peer reviewed (October 27, 2021). This study was not preregistered.
+This manuscript has not yet been peer-reviewed. This study was not preregistered.
 
@@
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow'', ``vaccines are safe,'' or ``this politician is trustworthy''.
+It is somewhat embarrassing then, that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow'', ``vaccines are safe,'' or ``this politician is trustworthy''.
 

### 2021-11-05 paper-rmd/paper-papaja.Rmd
more presubmission updates

@@
 
-### Accounting for Model Complexity
+### Predictive generalization and model complexity
 
-When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to "overfit" the data.
+A chief goal of scientific models is to make accurate predictions about future observations.
+When comparing and selecting models based on their predictive accuracy, it is crucial to consider how the models will perform on new, as-yet-unseen data.
+The traditional approach to this issue is to compute the fit of the model to the observed data, but to then correct for the potential for these models to "overfit" these data.
 Typically this correction comes in the form of a "complexity penalty," penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014].
-Zhu and colleagues [-@zhu.etal2020] warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
+Zhu and colleagues [-@zhu.etal2020] warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the competing models’ complexity [also see @piantadosi2018].
 
@@
 A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey’s prior), Beta(0,0) (Haldene’s prior), or perhaps other symmetric beta distributions Beta($\beta$,$\beta$) with $\beta \in [0, 1]$ [@jaynes2003].
-Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to [0, $1/3$] (approaching $1/3$ as $N$ approaches $\infty$ and $\beta$ approaches 1), whereas the PT+N model permits noise values between $[0, 1/2]$.
+Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values between $[0, 1/2]$.
 
@@
 
-Finally, one potential explanation for the relative weakness of the "complex" model variants is Zhu and colleagues’ [-@zhu.etal2020] use of "unpooled" models, where parameters are estimated independently for each individual participant.
+Third, one potential explanation for the relative weakness of the "complex" model variants is Zhu and colleagues’ [-@zhu.etal2020] use of "unpooled" models, where parameters are estimated independently for each individual participant.
 In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
@@
 
+Finally, it is worth recognizing that formal measures of model complexity cannot be expected to perfectly track notions of simplicity or elegance in scientific explanation [for some related discussions, see @kuhn1977; @sober2002; @piantadosi2018].
+For instance, even if the PT+N model's account of conditional probability judgments constrains its flexibility empirically, it seems clear this added component makes it more complex as a putative scientific explanation.
+
 ### The present work
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as the widely-applicable information criterion (${\text{WAIC}}$) and Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$).
-These approaches measure the ability of a model to capture new unseen data by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data.
-Importantly, these the methods for computing the $\widehat{\text{elpd}}$ penalize model complexity in a way that directly captures differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters [@gelman.etal2014; @vehtari.etal2017].
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based newer information criteria, such as the widely-applicable information criterion (${\text{WAIC}}$) and Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$).
+In particular, rather than estimating model fit and then penalizing for model complexity, $\text{PSIS-LOO}$ estimates out-of-sample prediction performance directly by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data [@gelman.etal2014; @vehtari.etal2017].
+From these calculations, an estimate of model complexity ($\hat{p}_{\text{LOO}}$) can also be derived.
+
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in a potential reduction in model complexity and a more realistic test of the models.
+This allows for information about model parameters to be shared across participants, resulting in a potential improvements out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
 
@@
 Both of these models can also be implemented as hierarchical models with partial pooling for the $d$ and $d'$ parameters (implicitly, for $N$ and $N'$ in the case of the Bayesian Sampler).
-This should prevent over-penalizing complex versions of the models and potentially improve their overall predictive fit.
+This partial pooling can help to regularize parameter estimates and improve out-of-sample predictive performance.
+In addition, partial pooling effectively reduces model complexity, and could support more realistic comparison between the "simple" and "complex" variants of the models.
+
 The hierarchical implementation adds parameters for the population-level $d$ and $d'$ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
@@
 If $\beta$ is unrestricted, allowed to fall in the domain $[0, \infty]$ then the Bayesian Sampler model becomes more flexible, allowing for equivalent "noise" levels in the same $[0, 1/2]$ range as the PT+N model.
-That is, through the bridging condition, the implied $d$ approaches $1/2$ in the limit as $N$ and $\beta$ both approach $\infty$.
+That is, through the bridging condition, the implied $d$ approaches $1/2$ in the limit as $N \to 1$ and $\beta \to \infty$.
 Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
@@
 
-Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments.
+Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory excising its constructive account of conditional probability judgments.
 So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
 Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
-And comparing their prediction specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level $r$ = .91 vs.
-.89; query-level $r$ = .98 vs.
+And comparing their predictions specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level $r$ = .91 vs.
+.89 (_p_ < .05); query-level $r$ = .98 vs.
 .97; Exp 2: response-level $r$ = .72 vs.
-.67; query-level $r$ = .92 vs.
+.67 (_p_ < .01); query-level $r$ = .92 vs.
 .85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
 
+Table 2 also presents estimates of $\hat{p}_{\text{LOO}}$, the estimate of the effective number of parameters for each model.
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of $d$  and $d'$.
-Evidently this structural component isn’t a complexifying epicycle, instead it is better seen as a prediction that constrains flexibility.
-However, it is one that ultimately leads to a worse-fitting model.
+Although its special treatment of conditional probability judgments makes it more complex as a putative scientific explanation, this structural component appears to actually constrain its predictive flexibility.
+However, this constraint leads to a worse-fitting model.
 
 Perhaps surprisingly, the Bayesian Sampler with unrestricted $\beta$ actually receives a smaller penalty term than the more "restricted" version of the model for the data in Experiment 2.
-This is at first counterintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
+This is at first counterintuitive.
+However, model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
 To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data [-@gelman.etal2014].
@@
 But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model’s penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied $d$  and $d'$ estimated for this data under this model are both very near $1/3$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
+Here, it seems reasonable to conjecture that because the implied $d$  and $d'$ estimated from Experiment 2's data are both very near $1/3$ under this model, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
 
-The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria.
+The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria like $\text{PSIS-LOO}$.
 Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives [e.g.
 @gronau.wagenmakers2019; @vehtari.etal2019].
+Fortunately, in this case, comparisons between the models do not rest solely on differences in estimated model complexity.
 
@@
     booktabs=T,
-    col.names = c("Model", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "Penalty","$r_{resp}$", "$r_{query}$", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "Penalty","$r_{resp}$", "$r_{query}$"),
+    col.names = c("Model", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$","$r_{resp}$", "$r_{query}$", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$","$r_{resp}$", "$r_{query}$"),
     escape = F
@@
 
-By a fair margin, the model best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] was a version of the Bayesian Sampler model without restriction on the range of its $\beta$ parameters.
+By a fair margin, the models best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] were implementations of the Bayesian Sampler theory, with the best single model being a version of the Bayesian Sampler model without restriction on the range of its $\beta$ parameters.
 Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
@@
 In addition, the implied $d$ and $d'$ noise parameters also varied across individuals.
-Of note, some participant's implied $d$ and $d'$ parameters _were_ consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
+Of note, some participants' implied $d$ and $d'$ parameters _were_ consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
 Further research should explore this heterogeneity within and across individuals.

### 2021-11-05 paper-rmd/paper-papaja.tex
more presubmission updates

@@
 
-\hypertarget{accounting-for-model-complexity}{%
-\subsubsection{Accounting for Model Complexity}\label{accounting-for-model-complexity}}
+\hypertarget{predictive-generalization-and-model-complexity}{%
+\subsubsection{Predictive generalization and model complexity}\label{predictive-generalization-and-model-complexity}}
 
-When comparing and selecting models based on their predictive accuracy, it is necessary to correct for the potential for these models to ``overfit'' the data.
+A chief goal of scientific models is to make accurate predictions about future observations.
+When comparing and selecting models based on their predictive accuracy, it is crucial to consider how the models will perform on new, as-yet-unseen data.
+The traditional approach to this issue is to compute the fit of the model to the observed data, but to then correct for the potential for these models to ``overfit'' these data.
 Typically this correction comes in the form of a ``complexity penalty,'' penalizing models in proportion to their flexibility for accommodating different patterns of data \autocite{gelman.etal2014}.
-Zhu and colleagues \autocite*{zhu.etal2020} warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the models' complexity.
+Zhu and colleagues \autocite*{zhu.etal2020} warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the competing models' complexity \autocite[also see][]{piantadosi2018}.
 
@@
 A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(\beta \in [0, 1]\) \autocite{jaynes2003}.
-Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches \(\infty\) and \(\beta\) approaches 1), whereas the PT+N model permits noise values between \([0, 1/2]\).
+Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches 1), whereas the PT+N model permits noise values between \([0, 1/2]\).
 
@@
 
-Finally, one potential explanation for the relative weakness of the ``complex'' model variants is Zhu and colleagues' \autocite*{zhu.etal2020} use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
+Third, one potential explanation for the relative weakness of the ``complex'' model variants is Zhu and colleagues' \autocite*{zhu.etal2020} use of ``unpooled'' models, where parameters are estimated independently for each individual participant.
 In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
@@
 Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
+
+Finally, it is worth recognizing that formal measures of model complexity cannot be expected to perfectly track notions of simplicity or elegance in scientific explanation \autocites[for some related discussions, see][]{kuhn1977,sober2002,piantadosi2018}.
+For instance, even if the PT+N model's account of conditional probability judgments constrains its flexibility empirically, it seems clear this added component makes it more complex as a putative scientific explanation.
 
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as the widely-applicable information criterion (\({\text{WAIC}}\)) and Pareto smoothed importance sampling approximate leave-one-out cross validation (\(\text{PSIS-LOO}\)).
-These approaches measure the ability of a model to capture new unseen data by estimating the expected log predictive density (\(\widehat{\text{elpd}}\)) of the model, or the expected probability of new unseen data.
-Importantly, these the methods for computing the \(\widehat{\text{elpd}}\) penalize model complexity in a way that directly captures differences in the flexibility of the models, instead of being estimated based solely on the number of distinct parameters \autocite{gelman.etal2014,vehtari.etal2017}.
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based newer information criteria, such as the widely-applicable information criterion (\({\text{WAIC}}\)) and Pareto smoothed importance sampling approximate leave-one-out cross validation (\(\text{PSIS-LOO}\)).
+In particular, rather than estimating model fit and then penalizing for model complexity, \(\text{PSIS-LOO}\) estimates out-of-sample prediction performance directly by estimating the expected log predictive density (\(\widehat{\text{elpd}}\)) of the model, or the expected probability of new unseen data \autocite{gelman.etal2014,vehtari.etal2017}.
+From these calculations, an estimate of model complexity (\(\hat{p}_{\text{LOO}}\)) can also be derived.
+
 In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in a potential reduction in model complexity and a more realistic test of the models.
+This allows for information about model parameters to be shared across participants, resulting in a potential improvements out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
 
@@
 Both of these models can also be implemented as hierarchical models with partial pooling for the \(d\) and \(d'\) parameters (implicitly, for \(N\) and \(N'\) in the case of the Bayesian Sampler).
-This should prevent over-penalizing complex versions of the models and potentially improve their overall predictive fit.
+This partial pooling can help to regularize parameter estimates and improve out-of-sample predictive performance.
+In addition, partial pooling effectively reduces model complexity, and could support more realistic comparison between the ``simple'' and ``complex'' variants of the models.
+
 The hierarchical implementation adds parameters for the population-level \(d\) and \(d'\) as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
@@
 If \(\beta\) is unrestricted, allowed to fall in the domain \([0, \infty]\) then the Bayesian Sampler model becomes more flexible, allowing for equivalent ``noise'' levels in the same \([0, 1/2]\) range as the PT+N model.
-That is, through the bridging condition, the implied \(d\) approaches \(1/2\) in the limit as \(N\) and \(\beta\) both approach \(\infty\).
+That is, through the bridging condition, the implied \(d\) approaches \(1/2\) in the limit as \(N \to 1\) and \(\beta \to \infty\).
 Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
@@
 
-Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory that removes the original theory's constructive account of conditional probability judgments.
+Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory excising its constructive account of conditional probability judgments.
 So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
 Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
-And comparing their prediction specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level \(r\) = .91 vs.~.89; query-level \(r\) = .98 vs.~.97; Exp 2: response-level \(r\) = .72 vs.~.67; query-level \(r\) = .92 vs.~.85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
+And comparing their predictions specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level \(r\) = .91 vs.~.89 (\emph{p} \textless{} .05); query-level \(r\) = .98 vs.~.97; Exp 2: response-level \(r\) = .72 vs.~.67 (\emph{p} \textless{} .01); query-level \(r\) = .92 vs.~.85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
 
+Table 2 also presents estimates of \(\hat{p}_{\text{LOO}}\), the estimate of the effective number of parameters for each model.
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of \(d\) and \(d'\).
-Evidently this structural component isn't a complexifying epicycle, instead it is better seen as a prediction that constrains flexibility.
-However, it is one that ultimately leads to a worse-fitting model.
+Although its special treatment of conditional probability judgments makes it more complex as a putative scientific explanation, this structural component appears to actually constrain its predictive flexibility.
+However, this constraint leads to a worse-fitting model.
 
 Perhaps surprisingly, the Bayesian Sampler with unrestricted \(\beta\) actually receives a smaller penalty term than the more ``restricted'' version of the model for the data in Experiment 2.
-This is at first counterintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data \autocite[see][]{gelman.etal2014}.
+This is at first counterintuitive.
+However, model complexity depends not only on the model and priors, but also the observed data \autocite[see][]{gelman.etal2014}.
 To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data \autocite*{gelman.etal2014}.
@@
 But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model's penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied \(d\) and \(d'\) estimated for this data under this model are both very near \(1/3\), the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
+Here, it seems reasonable to conjecture that because the implied \(d\) and \(d'\) estimated from Experiment 2's data are both very near \(1/3\) under this model, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
 
-The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria.
+The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria like \(\text{PSIS-LOO}\).
 Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives \autocites[e.g.][]{gronau.wagenmakers2019,vehtari.etal2019}.
+Fortunately, in this case, comparisons between the models do not rest solely on differences in estimated model complexity.
 
@@
 \cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
-Model & $\widehat{\text{elpd}}_{\text{LOO}}$ & Penalty & $r_{resp}$ & $r_{query}$ & $\widehat{\text{elpd}}_{\text{LOO}}$ & Penalty & $r_{resp}$ & $r_{query}$\\
+Model & $\widehat{\text{elpd}}_{\text{LOO}}$ & $\hat{p}_{\text{LOO}}$ & $r_{resp}$ & $r_{query}$ & $\widehat{\text{elpd}}_{\text{LOO}}$ & $\hat{p}_{\text{LOO}}$ & $r_{resp}$ & $r_{query}$\\
 \midrule
@@
 
-By a fair margin, the model best accounting for the experimental data from Zhu and colleagues \autocite*{zhu.etal2020} was a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters.
+By a fair margin, the models best accounting for the experimental data from Zhu and colleagues \autocite*{zhu.etal2020} were implementations of the Bayesian Sampler theory, with the best single model being a version of the Bayesian Sampler model without restriction on the range of its \(\beta\) parameters.
 Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
@@
 In addition, the implied \(d\) and \(d'\) noise parameters also varied across individuals.
-Of note, some participant's implied \(d\) and \(d'\) parameters \emph{were} consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
+Of note, some participants' implied \(d\) and \(d'\) parameters \emph{were} consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
 Further research should explore this heterogeneity within and across individuals.

### 2021-11-07 paper-rmd/paper-papaja.Rmd
submitted

@@
 authornote: |
- This manuscript has not yet been peer-reviewed. This study was not preregistered.
+ This manuscript has not yet been peer-reviewed. This study reports secondary data analyses and was not preregistered.
 
 abstract: |
- Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but one mark sometimes held against these theories is that people’s explicit probability judgments often violate the laws of probability.
+ Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but people’s explicit probability judgments often violate the laws of probability.
 Two recent proposals, the "Probability Theory plus Noise" [@costello.watts2014] and "Bayesian Sampler" [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "4617"
+wordcount         : "4826"
 
@@
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that these theories have often been accused of failing to describe human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", "vaccines are safe," or "this politician is trustworthy".
+It is somewhat embarrassing then, that these theories have often been accused of failing to describe human "beliefs" of the simple and everyday sort, such as beliefs like "it will rain tomorrow", "vaccines are safe," or "this politician is trustworthy" [@chater.etal2020].
 
@@
 
-Costello & Watts [-@costello.watts2014; -@costello.watts2016; -@costello.watts2018] have proposed a theory of probability judgment they call the "Probability Theory plus Noise" theory (PT+N).
+Costello and Watts [-@costello.watts2014; -@costello.watts2016; -@costello.watts2018] have proposed a theory of probability judgment they call the "Probability Theory plus Noise" theory (PT+N).
 In the PT+N model, mental "samples" are drawn from a probabilistic mental model of events and are then "read" with noise, so that some positive examples will be read as negative and some negative examples read as positive with some probability $d$.
@@
 They fit both "simple" and "complex" versions of each model, where the complex versions of these models introduce additional parameters $d'$ and $N'$ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are crucial to both models’ explanations of conjunction fallacies---key findings in the probability judgment literature [@costello.watts2017; @zhu.etal2020].
+These additional parameters are crucial to both models’ explanations of conjunction and disjunction fallacies---key findings in the probability judgment literature [@costello.watts2017; @zhu.etal2020].
 They compared the fits of these models to data via Bayesian Information Criteria score.
@@
 Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues [-@zhu.etal2020].
+But conjunction and disjunction fallacies are extremely robust empirical findings [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues [-@zhu.etal2020].
 So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
 For their part, Zhu and colleagues [-@zhu.etal2020] chose to average the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
+This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing strong conclusions in favor of their theory.
 
@@
 A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey’s prior), Beta(0,0) (Haldene’s prior), or perhaps other symmetric beta distributions Beta($\beta$,$\beta$) with $\beta \in [0, 1]$ [@jaynes2003].
-Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values between $[0, 1/2]$.
+Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to fall within [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values in $[0, 1/2]$.
 
@@
 
-In addition, the Bayesian framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in a potential improvements out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
+In addition, the Bayesian framework supports straightforward implementation of hierarchical versions of these models with partial pooling.
+This allows for information about model parameters to be shared across participants, resulting in potential improvements to out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
 
@@
 Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate $d$, $d'$, $N$, $N'$, and $\beta$ parameters for each participant.
+Although hierarchical models with partial pooling might be expected to better account for the data and offer a better test of the models, for consistency and comparison with Zhu et al.'s [@-zhu.etal2020] analyses, I first estimated implementations of these unpooled models.
 Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 
-Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of $\beta > 1$.
+Finally, I also explored fitting a hierarchical version of the Bayesian Sampler model that allowed values of $\beta > 1$.
 Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of "ignorance priors" [@zhu.etal2020].
@@
 I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and estimated the expected log predictive density with PSIS-LOO ($\widehat{\text{elpd}}_{\text{LOO}}$) for each combination.
-Compared with BIC, $\widehat{\text{elpd}}_{\text{LOO}}$ offers a more sophisticated estimation of model complexity and is more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
+Compared with BIC, $\widehat{\text{elpd}}_{\text{LOO}}$ offers a more sophisticated account of model complexity and is more appropriate in the "$\mathcal{M}$-open" case; situations where we do not know if any of the models being compared are the "true" model [@vehtari.etal2019].
 Model posteriors were estimated using the Numpyro [@phan.etal2019] implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
@@
 
-Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory excising its constructive account of conditional probability judgments.
+Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory that excises its constructive account of conditional probability judgments.
 So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
@@
 Nevertheless, many participants' estimated $d$ and $d'$ parameters fell outside the $[0, 1/3]$ range implied by "ignorance" priors.
-These data suggest that people may instead bring informative priors to some probability judgment tasks.
-Given the different values of $d$ and $d'$ across Experiments 1 and 2, it seems likely that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
+Given the different values of $d$ and $d'$ across Experiments 1 and 2, it seems likely that people bring informative, domain-specific priors to some judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
 
@@
 
-One thing this model comparison has not decided, and likely *cannot* decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
+One thing this model comparison has not decided, and likely *cannot* decide, is whether distortions of probability judgments are products of mental noise or of further reasoning processes.
 Given the two models' tight connections via the bridging condition [@zhu.etal2020], it may not be possible to draw decisive conclusions here.

### 2021-11-07 paper-rmd/paper-papaja.tex
submitted

@@
 \shorttitle{Comparing probability judgment accounts}
-\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: 4617}
+\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: 4826}
 \usepackage{csquotes}
@@
 
-This manuscript has not yet been peer-reviewed. This study was not preregistered.
+This manuscript has not yet been peer-reviewed. This study reports secondary data analyses and was not preregistered.
 
@@
 \abstract{
-Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but one mark sometimes held against these theories is that people's explicit probability judgments often violate the laws of probability.
+Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but people's explicit probability judgments often violate the laws of probability.
 Two recent proposals, the ``Probability Theory plus Noise'' \autocite{costello.watts2014} and ``Bayesian Sampler'' \autocite{zhu.etal2020} theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
@@
 At the heart of the Bayesian project is the idea that cognition is fundamentally probabilistic: that people reason according to subjective degrees of belief which follow the laws of probability and, in particular, that they are revised in light of evidence according to Bayes' Rule.
-It is somewhat embarrassing then, that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow'', ``vaccines are safe,'' or ``this politician is trustworthy''.
+It is somewhat embarrassing then, that these theories have often been accused of failing to describe human ``beliefs'' of the simple and everyday sort, such as beliefs like ``it will rain tomorrow'', ``vaccines are safe,'' or ``this politician is trustworthy'' \autocite{chater.etal2020}.
 
@@
 
-Costello \& Watts \autocite*{costello.watts2014,costello.watts2016,costello.watts2018} have proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
+Costello and Watts \autocite*{costello.watts2014,costello.watts2016,costello.watts2018} have proposed a theory of probability judgment they call the ``Probability Theory plus Noise'' theory (PT+N).
 In the PT+N model, mental ``samples'' are drawn from a probabilistic mental model of events and are then ``read'' with noise, so that some positive examples will be read as negative and some negative examples read as positive with some probability \(d\).
@@
 They fit both ``simple'' and ``complex'' versions of each model, where the complex versions of these models introduce additional parameters \(d'\) and \(N'\) that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are crucial to both models' explanations of conjunction fallacies---key findings in the probability judgment literature \autocite{costello.watts2017,zhu.etal2020}.
+These additional parameters are crucial to both models' explanations of conjunction and disjunction fallacies---key findings in the probability judgment literature \autocite{costello.watts2017,zhu.etal2020}.
 They compared the fits of these models to data via Bayesian Information Criteria score.
@@
 Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding \autocite{mellers.etal2001,sides.etal2002} and are clearly present in the data collected by Zhu and colleagues \autocite*{zhu.etal2020}.
+But conjunction and disjunction fallacies are extremely robust empirical findings \autocite{mellers.etal2001,sides.etal2002} and are clearly present in the data collected by Zhu and colleagues \autocite*{zhu.etal2020}.
 So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
 For their part, Zhu and colleagues \autocite*{zhu.etal2020} chose to average the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing any strong conclusions in favor of their theory.
+This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing strong conclusions in favor of their theory.
 
@@
 A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey's prior), Beta(0,0) (Haldene's prior), or perhaps other symmetric beta distributions Beta(\(\beta\),\(\beta\)) with \(\beta \in [0, 1]\) \autocite{jaynes2003}.
-Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches 1), whereas the PT+N model permits noise values between \([0, 1/2]\).
+Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to fall within {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches 1), whereas the PT+N model permits noise values in \([0, 1/2]\).
 
@@
 Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate \(d\), \(d'\), \(N\), \(N'\), and \(\beta\) parameters for each participant.
+Although hierarchical models with partial pooling might be expected to better account for the data and offer a better test of the models, for consistency and comparison with Zhu et al.'s {[}@-zhu.etal2020{]} analyses, I first estimated implementations of these unpooled models.
 Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 
-Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of \(\beta > 1\).
+Finally, I also explored fitting a hierarchical version of the Bayesian Sampler model that allowed values of \(\beta > 1\).
 Restricting \(\beta\) to {[}0,1{]} restricts the prior distribution of the Bayesian sampler to the class of ``ignorance priors'' \autocite{zhu.etal2020}.
@@
 I fit each of the models specified above to data from Zhu et al's \autocite*{zhu.etal2020} Experiment 1 and 2 and estimated the expected log predictive density with PSIS-LOO (\(\widehat{\text{elpd}}_{\text{LOO}}\)) for each combination.
-Compared with BIC, \(\widehat{\text{elpd}}_{\text{LOO}}\) offers a more sophisticated estimation of model complexity and is more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model \autocite{vehtari.etal2019}.
+Compared with BIC, \(\widehat{\text{elpd}}_{\text{LOO}}\) offers a more sophisticated account of model complexity and is more appropriate in the ``\(\mathcal{M}\)-open'' case; situations where we do not know if any of the models being compared are the ``true'' model \autocite{vehtari.etal2019}.
 Model posteriors were estimated using the Numpyro \autocite{phan.etal2019} implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler.
@@
 Nevertheless, many participants' estimated \(d\) and \(d'\) parameters fell outside the \([0, 1/3]\) range implied by ``ignorance'' priors.
-These data suggest that people may instead bring informative priors to some probability judgment tasks.
-Given the different values of \(d\) and \(d'\) across Experiments 1 and 2, it seems likely that people bring domain-specific priors to judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
+Given the different values of \(d\) and \(d'\) across Experiments 1 and 2, it seems likely that people bring informative, domain-specific priors to some judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
 
@@
 
-One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
+One thing this model comparison has not decided, and likely \emph{cannot} decide, is whether distortions of probability judgments are products of mental noise or of further reasoning processes.
 Given the two models' tight connections via the bridging condition \autocite{zhu.etal2020}, it may not be possible to draw decisive conclusions here.

### 2022-07-19 paper-rmd/paper-papaja.Rmd
working on rounding

@@
     corresponding : yes    # Define only one corresponding author
-    # address       : "Postal address"
+    address       : "4701 W Thunderbird Rd, Phoenix AZ 85306"
     email         : "dmpowell@asu.edu"
@@
 authornote: |
- This manuscript has not yet been peer-reviewed. This study reports secondary data analyses and was not preregistered.
+ This manuscript has not yet been peer-reviewed. This study reports secondary analyses of data a and was not preregistered.
 
@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "4826"
+wordcount         : "5088"
 
@@
 "Tom Brady likes miniature horses"), a plain and flagrant violation of probability theory [@tversky.kahneman1983].
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or "unpacking" effects [@tversky.koehler1994], and a variety of other findings illustrating the incoherence of human probability judgments [for an accessible review, see @kahneman2013].
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or "unpacking" effects [@tversky.koehler1994], and a number of others [for an accessible review, see @kahneman2013].
 Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
@@
 Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases in these judgments while maintaining that mental credences are fundamentally probabilistic [@costello.watts2014; @zhu.etal2020].
-Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental "sampling" from a probabilistic mental model [see also @chater.etal2020; @dasgupta.etal2017].
+Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental "sampling" from a probabilistic mental model [see also @chater.etal2020; @dasgupta.etal2017].[^1]
+
+[^1]: It is worth noting that other non-sampling based approaches have been proposed to account for distortions in people's use of explicit probabilities in decision-making [e.g.
+@zhang.maloney2012; @zhang.etal2020].
+Further theorizing might extend these accounts to also describe the generation of probability estimates, so that a probabilistic account of beliefs might not rest entirely on the assumption of sampling from mental models.
 
@@
 
-Thus, the effect of a Bayesian prior is nearly identical to the effect of noise in the PT+N model.
+Thus, in many cases the effect of a Bayesian prior is identical to the effect of noise in the PT+N model (at least in expectation).
 But, rather than merely perturbing people's probability judgments, this prior can be seen as regularizing these judgments away from extreme values.
@@
 Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to fall within [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values in $[0, 1/2]$.
+$d$ values in ther ange $[1/3, 1/2]$ imply $\beta$ values greater than 1, corresponding to an informative prior over probabilities.
 
@@
 Experiment 1 asked about the events [icy, frosty] and [normal, typical] (e.g.
-"what is the probability that the weather in London is normal and not typical?").
+"what is the probability that the weather in England is normal and not typical?").
 The authors’ goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated.
@@
 Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate $d$, $d'$, $N$, $N'$, and $\beta$ parameters for each participant.
-Although hierarchical models with partial pooling might be expected to better account for the data and offer a better test of the models, for consistency and comparison with Zhu et al.'s [@-zhu.etal2020] analyses, I first estimated implementations of these unpooled models.
+Although hierarchical models with partial pooling might be expected to better account for the data and offer a better test of the models, for consistency and comparison with Zhu et al.'s [-@zhu.etal2020] analyses, I first estimated implementations of these unpooled models.
 Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
+This is true for both query-level averages as well as for individual participants' trial responses.
 And comparing their predictions specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level $r$ = .91 vs.
@@
 
-Zhu and colleagues’ findings do cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta \in [0, 1]$.
-As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
+Zhu and colleagues’ experiments cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta \in [0, 1]$.
+As a generic prior that would be used across contexts, this class of uninformative priors has clear appeal.
 Nevertheless, many participants' estimated $d$ and $d'$ parameters fell outside the $[0, 1/3]$ range implied by "ignorance" priors.
-Given the different values of $d$ and $d'$ across Experiments 1 and 2, it seems likely that people bring informative, domain-specific priors to some judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
+Given the different values of $d$ and $d'$ across Experiments 1 and 2, it seems likely that people bring informative priors to at least some probability judgment tasks.
+
+How might these theories explain the difference in implied $d$ parameters between Experiments 1 and 2?
+Under the Bayesian Sampler theory, it could be that people bring informative and domain-specific priors to the probability judgment task, so that participants' priors differed across the different contexts of Experiments 1 and 2.
+However, these experimental contexts are highly similar, so it is not entirely clear why this should be the case.
+Alternatively, it could be that participants held identical informative priors in both contexts, but that they were able to more effectively draw mental samples for the queries presented in Experiment 1 than in Experiment 2 (resulting in lower $d$).
+Similarly, under a noise-based account, it could be that the mental samples are read more noisily in Experiment 2 than Experiment 1.
+Unfortunately, teasing apart these different accounts is likely to prove challenging.
+First, the same sort of factors that could be expected to influence the number of samples could also plausibly influence "noise" (e.g.
+task complexity, expertise).
+Moreover, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and $N$ parameters are not uniquely identifiable from this kind of judgment data alone.
 
 In addition, the implied $d$ and $d'$ noise parameters also varied across individuals.
-Of note, some participants' implied $d$ and $d'$ parameters _were_ consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
+Of note, some participants' implied $d$ and $d'$ parameters _were_ consistent with the class of ignorance priors, even in Experiment 2.
 Further research should explore this heterogeneity within and across individuals.
-Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and $N$ parameters are not uniquely identifiable from judgment data of the sort examined here.
 
@@
 @franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006].
+For instance, people's responses in various reasoning tasks are often explicitly related to inferred subjective mental probabilities (e.g.
+@jern.etal2014), so accounting for biases in those reports may permit more rigorous model testing.
 One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes [e.g.
 @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].
+
+\pagebreak
+
+# Declarations
+
+### Funding
+
+No funding was obtained for this study.
+
+### Conflict of Interests
+
+The author declares no conflicts of interest.
+
+### Availability of Data and Materials
+
+This paper presents secondary analyses of data. The datasets generated and/or analysed during the current study are available at https://osf.io/mgcxj/files/.
+
+### Code availability
+
+All analysis code is available at https://github.com/derekpowell/bayesian-sampler and at https://osf.io/bpkjf/.
+
+### Author contributions
+
+Derek Powell is the sole author of this manuscript.
+
+### Ethical approval
+
+Not applicable.
+
+### Consent to participate
+
+Not applicable.
+
+### Consent to publish
+
+Not applicable.

### 2022-07-19 paper-rmd/paper-papaja.tex
working on rounding

@@
 
+% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
+% See https://github.com/crsh/papaja/issues/292
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
+
 % \usepackage{etoolbox}
@@
 \makeatother
-\shorttitle{Comparing probability judgment accounts}
-\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: 4826}
+
+\usepackage{xpatch}
+\makeatletter
+\xapptocmd\appendix
+  {\xapptocmd\section
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {}{\InnerPatchFailed}%
+  }
+{}{\PatchFailed}
+\keywords{probability judgments, Bayesian cognitive science, heuristics and biases\newline\indent Word count: 5088}
 \usepackage{csquotes}
@@
 
+\shorttitle{Comparing probability judgment accounts}
+
 \authornote{
 
-This manuscript has not yet been peer-reviewed. This study reports secondary data analyses and was not preregistered.
+This manuscript has not yet been peer-reviewed. This study reports secondary analyses of data a and was not preregistered.
+
+Correspondence concerning this article should be addressed to Derek Powell, 4701 W Thunderbird Rd, Phoenix AZ 85306. E-mail: \href{mailto:dmpowell@asu.edu}{\nolinkurl{dmpowell@asu.edu}}
 
@@
 
-\abstract{
+\abstract{%
 Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but people's explicit probability judgments often violate the laws of probability.
@@
 For example, human probability judgments often exhibit the ``conjunction fallacy'': people will often judge the conjunction of two events (e.g.~``Tom Brady likes football and miniature horses'') as being more probable than one of the events in isolation (e.g.~``Tom Brady likes miniature horses''), a plain and flagrant violation of probability theory \autocite{tversky.kahneman1983}.
-Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects \autocite{tversky.koehler1994}, and a variety of other findings illustrating the incoherence of human probability judgments \autocite[for an accessible review, see][]{kahneman2013}.
+Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or ``unpacking'' effects \autocite{tversky.koehler1994}, and a number of others \autocite[for an accessible review, see][]{kahneman2013}.
 Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
@@
 Recently however, two groups of researchers have proposed theories of human probability judgments that account for biases in these judgments while maintaining that mental credences are fundamentally probabilistic \autocite{costello.watts2014,zhu.etal2020}.
-Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental ``sampling'' from a probabilistic mental model \autocites[see also][]{chater.etal2020,dasgupta.etal2017}.
+Both of these theories build on the increasingly popular notion that a variety of human reasoning tasks are accomplished by a limited process of mental ``sampling'' from a probabilistic mental model \autocites[see also][]{chater.etal2020,dasgupta.etal2017}.\footnote{It is worth noting that other non-sampling based approaches have been proposed to account for distortions in people's use of explicit probabilities in decision-making \autocites[e.g.][]{zhang.maloney2012,zhang.etal2020}.
+Further theorizing might extend these accounts to also describe the generation of probability estimates, so that a probabilistic account of beliefs might not rest entirely on the assumption of sampling from mental models.}
 
@@
 Via the bridging conditions, they show that assuming \(\beta \in [0, 1]\) restricts the ``noise level'' for the Bayesian Sampler, represented in implied \(d\) under the PT+N model, to fall within {[}0, \(1/3\){]} (approaching \(1/3\) as \(N\) approaches 1), whereas the PT+N model permits noise values in \([0, 1/2]\).
+\(d\) values in ther ange \([1/3, 1/2]\) imply \(\beta\) values greater than 1, corresponding to an informative prior over probabilities.
 
@@
 
-Experiment 1 asked about the events {[}icy, frosty{]} and {[}normal, typical{]} (e.g.~``what is the probability that the weather in London is normal and not typical?'').
+Experiment 1 asked about the events {[}icy, frosty{]} and {[}normal, typical{]} (e.g.~``what is the probability that the weather in England is normal and not typical?'').
 The authors' goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated.
@@
 Zhu, Sanborn \& Chater (2020) implement completely unpooled models with separate \(d\), \(d'\), \(N\), \(N'\), and \(\beta\) parameters for each participant.
-Although hierarchical models with partial pooling might be expected to better account for the data and offer a better test of the models, for consistency and comparison with Zhu et al.'s {[}@-zhu.etal2020{]} analyses, I first estimated implementations of these unpooled models.
+Although hierarchical models with partial pooling might be expected to better account for the data and offer a better test of the models, for consistency and comparison with Zhu et al.'s \autocite*{zhu.etal2020} analyses, I first estimated implementations of these unpooled models.
 Figure 1 displays the translation of the PT+N model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
@@
 Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
+This is true for both query-level averages as well as for individual participants' trial responses.
 And comparing their predictions specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level \(r\) = .91 vs.~.89 (\emph{p} \textless{} .05); query-level \(r\) = .98 vs.~.97; Exp 2: response-level \(r\) = .72 vs.~.67 (\emph{p} \textless{} .01); query-level \(r\) = .92 vs.~.85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
@@
 
-Zhu and colleagues' findings do cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta \in [0, 1]\).
-As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
+Zhu and colleagues' experiments cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect ``ignorance priors,'' symmetric Beta distributions with \(\beta \in [0, 1]\).
+As a generic prior that would be used across contexts, this class of uninformative priors has clear appeal.
 Nevertheless, many participants' estimated \(d\) and \(d'\) parameters fell outside the \([0, 1/3]\) range implied by ``ignorance'' priors.
-Given the different values of \(d\) and \(d'\) across Experiments 1 and 2, it seems likely that people bring informative, domain-specific priors to some judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
+Given the different values of \(d\) and \(d'\) across Experiments 1 and 2, it seems likely that people bring informative priors to at least some probability judgment tasks.
+
+How might these theories explain the difference in implied \(d\) parameters between Experiments 1 and 2?
+Under the Bayesian Sampler theory, it could be that people bring informative and domain-specific priors to the probability judgment task, so that participants' priors differed across the different contexts of Experiments 1 and 2.
+However, these experimental contexts are highly similar, so it is not entirely clear why this should be the case.
+Alternatively, it could be that participants held identical informative priors in both contexts, but that they were able to more effectively draw mental samples for the queries presented in Experiment 1 than in Experiment 2 (resulting in lower \(d\)).
+Similarly, under a noise-based account, it could be that the mental samples are read more noisily in Experiment 2 than Experiment 1.
+Unfortunately, teasing apart these different accounts is likely to prove challenging.
+First, the same sort of factors that could be expected to influence the number of samples could also plausibly influence ``noise'' (e.g.~task complexity, expertise).
+Moreover, pinning down specific components of the Bayesian Sampler model is a challenge, as the \(\beta\) and \(N\) parameters are not uniquely identifiable from this kind of judgment data alone.
 
 In addition, the implied \(d\) and \(d'\) noise parameters also varied across individuals.
-Of note, some participants' implied \(d\) and \(d'\) parameters \emph{were} consistent with the class of ignorance priors, even in Experiment 2 (see supplemental materials).
+Of note, some participants' implied \(d\) and \(d'\) parameters \emph{were} consistent with the class of ignorance priors, even in Experiment 2.
 Further research should explore this heterogeneity within and across individuals.
-Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the \(\beta\) and \(N\) parameters are not uniquely identifiable from judgment data of the sort examined here.
 
@@
 Future research could explore how estimates of people's credences might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning \autocites[e.g.][]{franke.etal2016,jern.etal2014,griffiths.tenenbaum2006}.
+For instance, people's responses in various reasoning tasks are often explicitly related to inferred subjective mental probabilities (e.g.
+\textcite{jern.etal2014}), so accounting for biases in those reports may permit more rigorous model testing.
 One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes \autocites[e.g.][]{cook.lewandowsky2016,jern.etal2014,powell.etal2018}.
+
+\pagebreak
+
+\hypertarget{declarations}{%
+\section{Declarations}\label{declarations}}
+
+\hypertarget{funding}{%
+\subsubsection{Funding}\label{funding}}
+
+No funding was obtained for this study.
+
+\hypertarget{conflict-of-interests}{%
+\subsubsection{Conflict of Interests}\label{conflict-of-interests}}
+
+The author declares no conflicts of interest.
+
+\hypertarget{availability-of-data-and-materials}{%
+\subsubsection{Availability of Data and Materials}\label{availability-of-data-and-materials}}
+
+This paper presents secondary analyses of data. The datasets generated and/or analysed during the current study are available at \url{https://osf.io/mgcxj/files/}.
+
+\hypertarget{code-availability}{%
+\subsubsection{Code availability}\label{code-availability}}
+
+All analysis code is available at \url{https://github.com/derekpowell/bayesian-sampler} and at \url{https://osf.io/bpkjf/}.
+
+\hypertarget{author-contributions}{%
+\subsubsection{Author contributions}\label{author-contributions}}
+
+Derek Powell is the sole author of this manuscript.
+
+\hypertarget{ethical-approval}{%
+\subsubsection{Ethical approval}\label{ethical-approval}}
+
+Not applicable.
+
+\hypertarget{consent-to-participate}{%
+\subsubsection{Consent to participate}\label{consent-to-participate}}
+
+Not applicable.
+
+\hypertarget{consent-to-publish}{%
+\subsubsection{Consent to publish}\label{consent-to-publish}}
+
+Not applicable.
 

### 2022-07-19 model_helpers.py

@@ -58,6 +58,30 @@ def load_data_exp2():
 
     return df
 
+def load_data_exp2_trials():
+    df = load_raw_data(2)
+    
+    df["condition"] = np.select(
+    [
+        df.querydetail.str.contains("windy|cloudy"), 
+        df.querydetail.str.contains("cold|rainy"),
+        df.querydetail.str.contains("warm|snowy")
+
+    ], 
+    [
+        0,
+        1,
+        2
+    ], 
+    default=0 )
+    
+    original_ids = list(np.unique(df.ID))
+    fix_id_dict = {original_ids[i]:i for i in range(0, len(original_ids))}
+    
+    df = df.assign(ID = df.ID.apply(lambda x: fix_id_dict[x]))
+
+    return df
+
 
 def load_query_avg_data_exp1():
     df = (
@@ -98,17 +122,31 @@ def load_query_avg_data_all():
     
 
 def make_model_data(data):
+    # for averaged data
 
     X_data = {
         "trial": data.querytype,
         "subj": jnp.array(list(data.ID)),
-        "cond": jnp.array(list(data.condition), dtype="int32"),
+        "cond": jnp.array(list(data.condition), dtype="int32")
     }
 
     y_data = jnp.array(data.estimate.to_numpy())
     
     return X_data, y_data
 
+def make_model_data_trials(data):
+    # for trial-level (non-averaged) data
+
+    X_data = {
+        "trial": data.querytype,
+        "subj": jnp.array(list(data.ID)),
+        "cond": jnp.array(list(data.condition), dtype="int32"),
+        "block": jnp.array(list(data.block), dtype="int32"),
+    }
+
+    y_data = jnp.array(data.estimate.to_numpy())
+    
+    return X_data, y_data
 
 ### -------- Simulation functions
 
@@ -135,8 +173,8 @@ trial_funcs = dict({
     "notB":lambda theta: jnp.matmul(theta, jnp.array([0.,1.,0.,1.])),
     "AorB":lambda theta: jnp.matmul(theta, jnp.array([1.,1.,1.,0.])),
     "AornotB":lambda theta: jnp.matmul(theta, jnp.array([1.,1.,0.,1.])),
-    "notAorB":lambda theta: jnp.matmul(theta, jnp.array([0.,1.,1.,1.])),
[…]

### 2022-09-16 lib/models.py
more trial-level updates

@@ -929,7 +929,72 @@ def ptn_simplecond_mlm_trial_level_disc(data, y=None):
         return(yhat)
     
     
-def ptn_simplecond_mlm_trial_level_disc2(data, y=None):
+# def ptn_simplecond_mlm_trial_level_disc2(data, y=None):
+    
+#     # parameterized in terms of d and d' for comparison of model fit
+
+#     # Data processing
+#     trial, subj, cond = data["trial"], data["subj"], data["cond"]
+#     n_Ps, n_conds = np.unique(subj).shape[0], np.unique(cond).shape[0] 
+    
+#     # setup "design matrix" (of sorts)
+#     X_num, X_denom = jnp.stack([num_vecs[i] for i in trial]), jnp.stack([denom_vecs[i] for i in trial])
+#     conjdisj, not_conjdisj = jnp.array([is_conjdisj(i) for i in trial]), abs(1-jnp.array([is_conjdisj(i) for i in trial]))
+
+#     # population level parameters/priors
+# #     k = numpyro.sample("k", dist.HalfCauchy(20)) # noise parameter
+#     rnd_policy = numpyro.sample("rnd_policy", dist.Dirichlet(jnp.ones(5)))
+    
+#     d_base_pop = numpyro.sample("d_base_pop", dist.Normal(-1.0, 1.0))
+#     d_delta_pop = numpyro.sample("d_delta_pop", dist.Normal(0, .5)) # bias toward lower values for non conj/disj trials
+#     d_base_sd = numpyro.sample("d_base_sd", dist.LogNormal(-1., 1.)) # was halfcauchy(1)
+#     d_delta_sd = numpyro.sample("d_delta_sd", dist.LogNormal(-1., 1.)) # approx uniform altogether we hope
+
+#     # subject-level parameters/priors 
+#     with numpyro.plate("subj", n_Ps):
+#         d_bases = numpyro.sample("d_base_r", dist.Normal(0, 1))
+#         d_deltas = numpyro.sample("d_delta_r", dist.Normal(0, 1))
+#         ks = numpyro.sample("k", dist.HalfCauchy(20)) # noise parameter
+        
+#     # subject/query-level parameters/priors
+#     with numpyro.plate("cond", n_Ps*n_conds):
+#         thetas = numpyro.sample("theta", dist.Dirichlet(jnp.ones(4)))
+    
+#     d_lin = (d_base_pop + 
[…]

### 2022-09-16 lib/simdata.py

@@ -23,15 +23,18 @@ def p_to_ptn(p, d):
 
 def sim_bayes_samp_trial(prob, beta, N):
     # p = dist.Beta(prob*N, (1-prob)*N).sample(PRNGKey(0), )
-    p = np.random.beta(prob*N, (1-prob)*N, size=len(prob))
+    p = np.random.beta(1+prob*N, 1+(1-prob)*N, size=len(prob)) # updated 9/13/22, 3:48 PM
     res = (p * N + beta)/(N + 2.*beta)
     return np.asarray(res)
 
 
 def sim_ptn_samp_trial(prob, d, N):
     # will round N
-    S = np.random.binomial(np.round(N).astype("int"), (1-2*d)*prob + d)
-    res = S/np.round(N).astype("int")
+#     S = np.random.binomial(np.round(N).astype("int"), (1-2*d)*prob + d)
+#     res = S/np.round(N).astype("int")
+    # updated 9/13/22, 3:49 PM
+    prob = (1-2*d)*prob + d
+    res = np.random.beta(1+prob*N, 1+(1-prob)*N, size=len(prob))
     return np.asarray(res)
 
 

### 2022-09-16 lib/helpers.py
plotting updates

@@ -3,7 +3,7 @@ import jax.numpy as jnp
 import numpy as np
 import arviz as az
 import pandas as pd
-
+import seaborn as sns
 import re
 
 import siuba as s

### 2022-09-16 lib/icc.py

@@ -18,12 +18,438 @@ from arviz.rcparams import rcParams, ScaleKeyword, ICKeyword
 
 from arviz import loo, waic
 
+from jax import random
+import numpyro
+from numpyro.infer import MCMC, NUTS, Predictive
+
+# NO_GET_ARGS: bool = False
+# try:
+#     from typing_extensions import get_args
+# except ImportError:
+#     NO_GET_ARGS = True
+
+
+# def get_ic_type(ic_input):
+
+#     try:
+#         ic_input.loo
+#         ic = "loo"
+#     except:
+#         try:
+#             ic_input.waic
+#             ic = "waic"
+#         except:
+#             ic = "unknown"
+#     return(ic)
+
+
+# def _ic_matrix(ics, ic_i):
+#     """Store the previously computed pointwise predictive accuracy values (ics) in a 2D matrix."""
+#     cols, _ = ics.shape
+#     rows = len(ics[ic_i].iloc[0])
+#     ic_i_val = np.zeros((rows, cols))
+
+#     for idx, val in enumerate(ics.index):
+#         ic = ics.loc[val][ic_i]
+
+#         if len(ic) != rows:
+#             raise ValueError("The number of observations should be the same across all models")
+
+#         ic_i_val[:, idx] = ic
+
+#     return rows, cols, ic_i_val
+
+
+# def compare(
+#     dataset_dict: Mapping[str, InferenceData],
+#     ic: Optional[ICKeyword] = None,
+#     method: Literal["stacking", "BB-pseudo-BMA", "pseudo-MA"] = "stacking",
+#     b_samples: int = 1000,
+#     alpha: float = 1,
+#     seed=None,
+#     scale: Optional[ScaleKeyword] = None,
+#     var_name: Optional[str] = None,
+# ):
+#     r"""Compare models based on PSIS-LOO `loo` or WAIC `waic` cross-validation.
+#     LOO is leave-one-out (PSIS-LOO `loo`) cross-validation and
+#     WAIC is the widely applicable information criterion.
+#     Read more theory here - in a paper by some of the leading authorities
+#     on model selection dx.doi.org/10.1111/1467-9868.00353
+#     Parameters
+#     ----------
+#     dataset_dict: dict[str] -> InferenceData
+#         A dictionary of model names and InferenceData objects
+#     ic: str, optional
+#         Information Criterion (PSIS-LOO `loo` or WAIC `waic`) used to compare models. Defaults to
+#         ``rcParams["stats.information_criterion"]``.
[…]

### 2022-09-16 create-paper-figures.Rmd
created revised manuscript

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(latex2exp)
library(reticulate)
use_condaenv("bsampler")
```

```{python}
# import numpyro
# numpyro.util.set_host_device_count(4)
# import numpyro.distributions as dist
# from numpyro.infer import MCMC, NUTS, Predictive, DiscreteHMCGibbs

from jax import numpy as jnp
from jax import random

import numpy as np
import pandas as pd
# import seaborn as sns
import arviz as az

# from dfply import *
from siuba import _
import siuba as s

import pickle

from lib.helpers import *
from lib.models import *
from lib.icc import *

```

```{python}
df1 = load_query_avg_data_exp1() # see data_helpers.py
df1 = df1[~df1.querytype.str.contains('or')]
# df1 = df1[df1["condition"]!=2] # filter out "warm/snowy" as per paper
df1.head()
```

```{python}
df2 = load_query_avg_data_exp2() # see data_helpers.py
df2 = df2[df2["condition"]!=2] # filter out "warm/snowy" as per paper
df2.head()
```

```{python}
fits = dict()
fits["exp1"] = dict()
fits["exp2"] = dict()

fits["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/az_data_rf.p", "rb"))
fits["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/az_data_bs_simple.p", "rb"))
fits["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/az_data_bs.p", "rb"))
fits["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/az_data_bs_mlm_simple.p", "rb"))
fits["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/az_data_bs_mlm.p", "rb"))
fits["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/az_data_ptn.p", "rb"))
fits["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/az_data_ptn_simple.p", "rb"))
fits["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm.p", "rb"))
fits["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simple.p", "rb"))
fits["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond.p", "rb"))

fits["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/az_data_rf.p", "rb"))
fits["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/az_data_bs_simple.p", "rb"))
fits["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/az_data_bs.p", "rb"))
fits["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/az_data_bs_mlm_simple.p", "rb"))
fits["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/az_data_bs_mlm.p", "rb"))
fits["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/az_data_ptn.p", "rb"))
fits["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/az_data_ptn_simple.p", "rb"))
fits["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm.p", "rb"))
fits["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simple.p", "rb"))
fits["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simplecond.p", "rb"))

loos = dict()
loos["exp1"] = dict()
loos["exp2"] = dict()

loos["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/loo_rf.p", "rb"))
loos["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/loo_bs_simple.p", "rb"))
loos["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/loo_bs.p", "rb"))
loos["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/loo_bs_mlm_simple.p", "rb"))
loos["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/loo_bs_mlm.p", "rb"))
loos["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/loo_ptn.p", "rb"))
loos["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/loo_ptn_simple.p", "rb"))
loos["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/loo_ptn_mlm.p", "rb"))
loos["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simple.p", "rb"))
loos["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simplecond.p", "rb"))

loos["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/loo_rf.p", "rb"))
loos["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/loo_bs_simple.p", "rb"))
loos["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/loo_bs.p", "rb"))
loos["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/loo_bs_mlm_simple.p", "rb"))
loos["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/loo_bs_mlm.p", "rb"))
loos["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/loo_ptn.p", "rb"))
loos["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/loo_ptn_simple.p", "rb"))
loos["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/loo_ptn_mlm.p", "rb"))
loos["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simple.p", "rb"))
loos["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simplecond.p", "rb"))
```

```{python}
exp1_model_comparison = compare(loos["exp1"], method="BB-pseudo-BMA")
exp2_model_comparison = compare(loos["exp2"], method="BB-pseudo-BMA")

exp1_model_comparison = exp1_model_comparison >> s.mutate(Experiment = "Exp. 1", model = exp1_model_comparison.index)
exp2_model_comparison = exp2_model_comparison >> s.mutate(Experiment = "Exp. 2", model = exp2_model_comparison.index)

model_comparison = exp1_model_comparison.append(exp2_model_comparison)
```

```{r}
py$model_comparison %>% 
  filter(model!="Relative Freq.") %>% 
  mutate(ul=d_loo + dse*2, ll = d_loo - dse*2) %>% 
  ggplot(aes(y=reorder(model,-d_loo), x = d_loo, xmin=ll, xmax=ul)) +
  geom_vline(xintercept=0, linetype="dashed", color="grey") +
  geom_point(size=2) +
  geom_errorbarh(height=.01, size=.5) +
  facet_wrap(~Experiment, scales="free_x") +
  labs(x = TeX("$\\widehat{elpd}_{LOO}$ difference"), y = "Model") +
  theme_bw() +
  theme(panel.grid=element_blank())
```

```{python}
model1 = fits["exp1"]["Unrestricted BS hier."]
model1.posterior["d_prime_pop"] = model1.posterior.d_base_pop + np.exp(model1.posterior.d_delta_pop)
post1 = posterior_to_df_mlm(model1, ["k"], ["d_subj", "d_prime_subj"])

model2 = fits["exp2"]["Unrestricted BS hier."]
model2.posterior["d_prime_pop"] = model2.posterior.d_base_pop + np.exp(model2.posterior.d_delta_pop)
post2 = posterior_to_df_mlm(model2, ["k"], ["d_subj", "d_prime_subj"])

param_posterior1 = model1.posterior.stack(draws=("chain", "draw"))[["d_base_pop", "d_prime_pop"]].to_dataframe()
param_posterior2 = model2.posterior.stack(draws=("chain", "draw"))[["d_base_pop", "d_prime_pop"]].to_dataframe()

param_posterior1 = param_posterior1 >> s.mutate(Experiment = "Exp. 1")
param_posterior2 = param_posterior2 >> s.mutate(Experiment = "Exp. 2")

param_posterior = param_posterior1.append(param_posterior2)
```

```{r}
plt_forest_exp1 <- py$post1 %>% 
  gather(var, val, d_subj, d_prime_subj) %>% 
    group_by(ID, var) %>% 
    summarize(
        M = mean(val),
        ll = quantile(val, .025),
        ul = quantile(val, .975),
    ) %>% 
  mutate(var = if_else(var=="d_subj", "d", "d'")) %>% 
  ggplot(aes(x=reorder(ID,M), y = M, ymin=ll, ymax=ul, color=var)) +
  geom_pointrange(position=position_dodge(width=.5), size=.25) +
  labs(title="Exp. 1", color="Variable", x = "Participant", y = "Estimate") +
  theme_bw() +
  theme(
      panel.grid = element_blank(), 
      legend.position="bottom", 
      axis.text.x = element_text(angle=90, hjust=1, vjust=.5, size=5)
      )

plt_forest_exp2 <- py$post2 %>% 
  gather(var, val, d_subj, d_prime_subj) %>% 
    group_by(ID, var) %>% 
    summarize(
        M = mean(val),
        ll = quantile(val, .025),
        ul = quantile(val, .975),
    ) %>% 
  mutate(var = if_else(var=="d_subj", "d", "d'")) %>% 
  ggplot(aes(x=reorder(ID,M), y = M, ymin=ll, ymax=ul, color=var)) +
  geom_pointrange(position=position_dodge(width=.5),size=.25) +
  labs(title="Exp. 2", color="Variable", x = "Participant", y = "Estimate") +
  theme_bw() +
  theme(
      panel.grid = element_blank(), 
      legend.position="bottom", 
      axis.text.x = element_text(angle=90, hjust=1, vjust=.5, size=5)
      )

library(patchwork)

(plt_forest_exp1 + coord_fixed(ratio=59/.5)) + 
  (plt_forest_exp2 + coord_fixed(ratio=59/.5)) + 
  plot_layout(guides="collect") & theme(legend.position = 'bottom')
```

```{python}
preds1 = fits["exp1"]["Unrestricted BS hier."].posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
preds2 = fits["exp2"]["Unrestricted BS hier."].posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
```

```{r}
plt_pp <- py$df1 %>% 
  mutate(pred = py$preds1, Experiment = "Exp. 1") %>% 
  bind_rows(
    py$df2 %>% 
    mutate(pred = py$preds2, Experiment = "Exp. 2")
  ) %>% 
  ggplot(aes(x=pred, y= estimate)) +
  geom_point(alpha=.25) +
  facet_wrap(~Experiment) +
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(panel.grid=element_blank())

plt_pp
```

```{r}

## calcualte inequalities

plt_ineq_exp2 <- py$df2 %>% 
  mutate(pred = py$preds2) %>% 
  mutate(condition = if_else(condition==0, "{cloudy, windy}", "{cold, rainy}" )) %>% 
  gather(var, val, estimate, pred) %>% 
  select(-querydetail) %>% 
  pivot_wider(names_from = querytype, values_from=val) %>% 
  mutate(
    z1 = A + B - AandB - AorB,
    z2 = A + notAandB - B - AandnotB,
    z3 = A + notAandB - AorB,
    z4 = B + AandnotB - AorB,
    z5 = AandnotB + AandB - A,
    z6 = notAandB + AandB - B,
    z7 = AandnotB + notAandB + AandB - AorB,
    z8 = AandnotB + notAandB + 2*AorB - A -B,
    z9 = AgB*B - BgA*A,
    z10 = AgB*B + AgnotB*notB - A,
    z11 = BgA*A + BgnotA*notA - B,
    z12 = BgA*A + AgnotB*notB - A,
    z13 = AgB*B + BgnotA*notA - B,
    z14 = AgnotB*notB + B - BgnotA*notA - A,
    z15 = AandB - AgB*B,
    z16 = AandB - BgA*A,
    z17 = AandB - A + AgnotB*notB,
    z18 = AandB - B + BgnotA*notA
  ) %>% 
  select(ID, var, condition, starts_with("z")) %>% 
  gather(inequality, val, -ID,-var,-condition) %>% 
  # mutate(inequality = as.numeric(gsub("z","", inequality))) %>% 
  mutate(inequality = ordered(inequality, levels = paste0("z",1:18))) %>% 
  group_by(var, inequality, condition) %>% 
  summarize(
    M = mean(val),
    SD = sd(val),
    se = sd(val)/sqrt(n())
  ) %>% 
  ggplot(aes(x=inequality, y = M, ymin = M-2*se, ymax=M+2*se, color=var)) +
  geom_pointrange(size=.25) +
  geom_hline(yintercept=0, linetype="dashed") +
  facet_wrap(~condition, ncol=1) +
  theme_bw() +
  theme(panel.grid=element_blank())

```

```{python}
df1_all = load_query_avg_data_exp1()
model1_all = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond-all.p", "rb"))
preds1_all = model1_all.posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
```

```{r}
plt_ineq_exp1 <- py$df1_all %>% 
  mutate(pred = py$preds1_all) %>% 
  gather(var, val, estimate, pred) %>% 
  select(-querydetail) %>% 
  pivot_wider(names_from = querytype, values_from=val) %>% 
  mutate(condition = if_else(condition==0, "{icy, frosty}", "{normal, typical}" )) %>% 
  mutate(
    z1 = A + B - AandB - AorB,
    z2 = A + notAandB - B - AandnotB,
    z3 = A + notAandB - AorB,
    z4 = B + AandnotB - AorB,
    z5 = AandnotB + AandB - A,
    z6 = notAandB + AandB - B,
    z7 = AandnotB + notAandB + AandB - AorB,
    z8 = AandnotB + notAandB + 2*AorB - A -B,
    z9 = AgB*B - BgA*A,
    z10 = AgB*B + AgnotB*notB - A,
    z11 = BgA*A + BgnotA*notA - B,
    z12 = BgA*A + AgnotB*notB - A,
    z13 = AgB*B + BgnotA*notA - B,
    z14 = AgnotB*notB + B - BgnotA*notA - A,
    z15 = AandB - AgB*B,
    z16 = AandB - BgA*A,
    z17 = AandB - A + AgnotB*notB,
    z18 = AandB - B + BgnotA*notA
  ) %>% 
  select(ID, var, condition, starts_with("z")) %>% 
  gather(inequality, val, -ID,-var,-condition) %>% 
  # filter(inequality %in% paste0("z",10:13)) %>% 
  # mutate(inequality = as.numeric(gsub("z","", inequality))) %>% 
  mutate(inequality = ordered(inequality, levels = paste0("z",1:18))) %>% 
  group_by(var, inequality, condition) %>% 
  summarize(
    M = mean(val),
    SD = sd(val),
    se = sd(val)/sqrt(n())
  ) %>% 
  ggplot(aes(x=inequality, y = M, ymin = M-2*se, ymax=M+2*se, color=var)) +
  geom_pointrange(size=.25) +
  geom_hline(yintercept=0, linetype="dashed") +
  facet_wrap(~condition, ncol=1) +
  theme_bw() +
  theme(panel.grid=element_blank())

```

```{r}
# In Experiment 1, like participants' responses,  the model's estimates are very slightly positive for {icy, frosty} and very slightly negative for {normal, typical}. This pattern replicates the qualitative pattern reported by Zhu and colleagues.
plt_ineq_exp1 + labs(title="Exp. 1") + 
  plt_ineq_exp2 + labs(title="Exp. 2") + 
  plot_layout(guides="collect") & theme(legend.position="bottom") & labs(color="", y = "Estimate", x = "Identity")
```

```{python}
# az_ptn_exp1 = pickle.load(open("local/az-ptn_simplecond_mlm_trial_level_disc-exp1svi.p", "rb"))
# az_ptn_exp2 = pickle.load(open("local/az-ptn_simplecond_mlm_trial_level_disc-exp2svi.p", "rb"))
# 
# # posterior_ptn1 = posterior_to_df_mlm(
# #     az_ptn_exp1, 
# #     ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
# #     ["d_subj", "d_prime_subj", "k"]
# # )
# # 
# posterior_ptn2 = posterior_to_df_mlm(
#     az_ptn_exp2,
#     ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
#     ["d_base_r", "d_delta_r", "d_subj", "d_prime_subj", "k"]
# )

posterior_ptn1 = pickle.load(open("local/posterior_ptn1.p", "rb"))
posterior_ptn2 = pickle.load(open("local/posterior_ptn2.p", "rb"))
```

```{r}
plt_dk_ptn1 <- py$posterior_ptn1 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_subj),
        d_ll = quantile(d_subj, .05),
        d_ul = quantile(d_subj, .95)
    ) %>% 
#     s.filter(_.k_est < 50.) >>
    filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d", x="k", title="Exp. 1") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())

plt_dk_ptn2 <- py$posterior_ptn2 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_subj),
        d_ll = quantile(d_subj, .05),
        d_ul = quantile(d_subj, .95)
    ) %>% 
#     s.filter(_.k_est < 50.) >>
    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d", x="k", title="Exp. 2") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())

plt_dprimek_ptn1 <- py$posterior_ptn1 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_prime_subj),
        d_ll = quantile(d_prime_subj, .05),
        d_ul = quantile(d_prime_subj, .95)
    ) %>% 
#     s.filter(_.k_est < 50.) >>
    filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d'", x="k") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())

plt_dprimek_ptn2 <- py$posterior_ptn2 %>% 
    group_by(ID) %>% 
    summarize(
        k_est = mean(k),
        k_ll = quantile(k, .05),
        k_ul = quantile(k, .95),
        d_est = mean(d_prime_subj),
        d_ll = quantile(d_prime_subj, .05),
        d_ul = quantile(d_prime_subj, .95)
    ) %>% 
#     s.filter(_.k_est < 50.) >>
    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
    ggplot(aes(x=k_est, y=d_est)) +
    geom_point() +
    geom_errorbar(aes(ymin=d_ll, ymax=d_ul), width=0., alpha=.25) +
    geom_errorbarh(aes(xmin=k_ll, xmax=k_ul), height=0, alpha=.25) +
    # geom_smooth(method="lm") +
    labs(y="d'", x="k") +
    theme_bw() +
    theme(aspect.ratio=1) +
    theme(panel.grid = element_blank())
```

```{r}
plt_dk_ptn1  + plt_dprimek_ptn1 + plt_dk_ptn2 + plt_dprimek_ptn2
```

```{r}
plt_d_avgmodel <- py$param_posterior %>% 
  mutate(
    d = plogis(d_base_pop)/2,
    d_prime = plogis(d_prime_pop)/2
    ) %>% 
  gather(Variable, val, d, d_prime) %>% 
  mutate(Variable = if_else(Variable=="d", "d", "d'")) %>% 
  ggplot(aes(x=val, fill=Variable, color=Variable)) +
  geom_density(position="identity", alpha=.6) +
  facet_wrap(~Experiment) +
  xlim(0,.5) +
  labs(title="Participant-level query-average model")

plt_d_trialmodel <- py$posterior_ptn1 %>% 
  mutate(Experiment = "Exp. 1") %>% 
  bind_rows(
    py$posterior_ptn2 %>% 
      mutate(Experiment = "Exp. 2")
  ) %>% 
  filter(ID==0) %>% 
  select(Experiment, draw, d_base_pop, d_base_sd, d_delta_pop, d_delta_sd) %>% 
  mutate(
    d = plogis(d_base_pop)/2,
    d_prime = plogis(d_base_pop + exp(d_delta_pop))/2
    ) %>% 
  gather(Variable, val, d, d_prime) %>% 
  mutate(Variable = if_else(Variable=="d", "d", "d'")) %>% 
  ggplot(aes(x=val, fill=Variable, color=Variable)) +
  geom_density(position="identity", alpha=.6) +
  facet_wrap(~Experiment) +
  xlim(0,.5) +
  labs(title="Trial-level model")

plt_d_avgmodel + plt_d_trialmodel + plot_layout(ncol=1, guides="collect") & labs(x="Estimate") & theme_bw() & theme(panel.grid=element_blank())
```

### 2022-09-16 paper-rmd/bs_platediagram_mlm.tex
paper edits
created revised manuscript

@@
+\begin{figure}
+  \begin{minipage}{.5\textwidth}
+    \flushright
     \begin{tikzpicture}
@@
         \tikzstyle{plateIJK} = [align=center];
-        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$ijk$} ;
+        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$i$} ;
      }
@@
         \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
-        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$jk$};
+        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$k$};
       }
@@
 \end{tikzpicture}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \flushleft
+    \small
+    \begin{align*}
+  y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
+  \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j')  \\
+  d_j &= \frac{1}{3} \ \text{logistic}(\delta_j) \\
+  d_j' &= \frac{1}{3} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
+  \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
+  \overline{\delta} &\sim Normal(-1, 1) \\
+  \overline{\Delta\delta} &\sim Normal(0, .50) \\
+  log(\sigma_{\delta}) &\sim Normal(-1, 1) \\
+  log(\sigma_{\Delta\delta}) &\sim Normal(-1, 1) \\
+  \delta_j &\sim Normal(\overline{\delta}, \sigma_{\delta}) \\
+  \Delta\delta_j &\sim Normal(\overline{\Delta\delta}, \sigma_{\Delta\delta}) \\
+  \kappa &\sim \mathit{Half \hyph Cauchy(20)} \\
+\end{align*}
+
+  \end{minipage}
+  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications.
+Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters.
+Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
+\end{figure}

### 2022-09-16 paper-rmd/paper-papaja-revision.Rmd
created revised manuscript

--- 2022-07-19 paper-rmd/paper-papaja.Rmd
@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "5088"
+wordcount         : "????"
 
@@
 library("papaja")
-r_refs("references.bib")
+options(encoding = "UTF-8")
+# r_refs("references.bib")
 ```
@@
 library(kableExtra)
+```
+
+```{r get_figures, child='../create-paper-figures.Rmd', include=FALSE}
+
 ```
@@
 In the PT+N model, mental "samples" are drawn from a probabilistic mental model of events and are then "read" with noise, so that some positive examples will be read as negative and some negative examples read as positive with some probability $d$.
-The results are probability judgments reflecting probabilistic credences perturbed by noise.
-In their model, the probability a mental sample for an event $A$ is read as A is the probability that the sample truly is $A$, $p(A)$, and that it is correctly read $(1-d)$, plus the probability that the sample is not $A$, $1-P(A)$ and that it is incorrectly read ($d$), or:
+The end products are probability judgments reflecting probabilistic credences perturbed by noise.
+In their model, the probability a mental sample for an event $A$ is read as $A$ is the probability that the sample truly is $A$, $p(A)$, and that it is correctly read $(1-d)$, plus the probability that the sample is not $A$, $1-P(A)$ and that it is incorrectly read ($d$), or:
 
-\begin{align*}
+\begin{align}
   P(\text{read as A}) &= (1-d)P(A) + d(1-P(A)) \\
-  &= (1-2d)P(A) + d
-\end{align*}
+  &= (1-2d)P(A) + d \tag{1}
+\end{align}
 
@@
 
-$$E[\hat{P}_{PT+N}(A)] = (1-2d)P(A) + d $$
+$$
+E[\hat{P}_{PT+N}(A)] = (1-2d)P(A) + d \tag{2}
+$$
 
-By assumption, a maximum of 50% of samples can be misread, so that $d$ is a number in the range $[0, 1/2]$.
+By assumption, a maximum of 50% of samples can be misread, so $d$ is a number in the range $[0, 1/2]$.
 The PT+N theory provides a unified account for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
@@
 
-$$\hat{P}_{BS}(A) = \frac{S(A)}{N+2\beta} + \frac{\beta}{N+2\beta}$$
+$$\hat{P}_{BS}(A) = \frac{S(A)}{N+2\beta} + \frac{\beta}{N+2\beta} \tag{3}$$
 
@@
 
-$$E[\hat{P}_{BS}(A)] = \frac{N}{N+2\beta}P(A) + \frac{\beta}{N+2\beta}$$
+$$E[\hat{P}_{BS}(A)] = \frac{N}{N+2\beta}P(A) + \frac{\beta}{N+2\beta} \tag{4}$$
 
@@
 
-$$d = \frac{\beta}{N+2\beta}$$
+$$d = \frac{\beta}{N+2\beta} \tag{5}$$
 
@@
 
-### Two accounts of conditional probability judgments
+## Differentiating between the models
+
+These models make subtly different predictions on two levels: First,  the models have distinct accounts of conditional probability judgments that make different predictions in terms of expected values.
+Second, the models present different process-level accounts of probability judgment that entail different predictions about the shape of the distribution of responses across individuals.
+
+### Different accounts of conditional probability judgments
 
@@
 According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments.
-By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models [@chater.etal2020].
+By not assigning any special status to conditional probability judgments, the Bayesian Sampler theory fits neatly into the larger project of Bayesian cognitive science: probability judgments are simply another judgment process applied to the outputs of other (ideally Bayesian) mental models [@chater.etal2020].
 
@@
 According to the PT+N model, conditional probabilities $P(A|B)$ are estimated by a two-stage sampling procedure: first both events $A$ and $B$ are sampled with noise, and then a second noisy process computes the ratio of the events read as $A$ and $B$ over events read as $B$.
-The PT+N model predicts conditional probability estimates using the following equation:
+The probability can be written as:
 
-$$P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}$$
+\begin{align*}
+P_e(A|B) &= P(\textit{read as A}| \textit{read as B}) \\
+& = P(\textit{read as A}|B)P(B|\textit{read as B}) + P(\textit{read as A}| \neg B)P(\neg B|\textit{read as B}) \tag{6}
+\end{align*}
+
+Substituting terms according to the PT+N model and then simplifying, the PT+N model predicts conditional probability estimates using the following equation:
+
+$$P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \tag{7}$$
 
@@
 
-## Comparing the models
+### Different process-level accounts and predicted response distributions
+
+Outside of conditional probability judgments, the primary difference between the models is the psychological process thought to produce bias in probability judgments: sample reading noise in the PT+N model and Bayesian inference in the Bayesian Sampler model.
+These process-level differences imply different response distributions.
+
+In both models, the amount of variability in trial-level responses is related to the number of mental samples drawn, $N$.
+In the Bayesian Sampler model, assuming $\beta$ is relatively small, $N$ should also help to determine the degree to which responses are shrunk toward .50.
+In contrast, in the PT+N model the variance across responses and degree of shrinkage are reasonably considered to be independent.
+Moreover, the Bayesian Sampler model predicts not only shrinkage of probability estimates toward .50, but also a truncation of the response distribution in proportion to $\beta$ and $N$ [@chater.etal2020; @sundh.etal2021].
+Thus, modeling the distributions of raw responses holds promise for disentangling the models.
+
+However, there are at least three challenges to directly modeling raw human response data.
+First, both models are, strictly speaking, discrete and so make a limited set of discrete predictions while assigning zero probability to responses outside that set.
+Second, and similarly, the truncation in the Bayesian Sampler model also assigns zero probability to respones beyond the truncated range.
+And third, from a cursory glance it is clear that a majority of human responses are rounded by some unknown degree, with most seemingly rounded to the nearest 5 or 10%.
+Given the combination of these factors, if fit directly to raw human data the posterior probability of both models is likely to be zero.
+I return to these challenges and my approaches to addressing them in the results.
+
+## Prior comparisons of the models
+
+### Comparison of participant-level query-averaged responses
 
@@
 Unfortunately, their results were somewhat equivocal.
-They fit both "simple" and "complex" versions of each model, where the complex versions of these models introduce additional parameters $d'$ and $N'$ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are crucial to both models’ explanations of conjunction and disjunction fallacies---key findings in the probability judgment literature [@costello.watts2017; @zhu.etal2020].
-They compared the fits of these models to data via Bayesian Information Criteria score.
-Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs [@zhu.etal2020, Supplementary materials].
 
-```{r table1, echo=FALSE, message=F, warning=F}
-t1_df <- read_csv("original-fits.csv")
+Zhu and colleagues [-@zhu.etal2020] measured participants' judgments for each query (e.g.
+"what is the probability that it will be rainy") on three repeated trials.
+Their primary quantitative analysis fit the models separately to participants' average response to each query (averaged over three trials).
+These analyses compare human responses to the models' predictions _in expectation_.
+After fitting, Bayesian Information Criteria (BIC) values were computed for each participant, which were then used to approximate the posterior probability of each model for each participant, assuming a uniform prior.
+The researchers found that a preponderance of participants' responses were best-captured by the Bayesian Sampler model.
+However, a substantial number of participants were instead more strongly fit by the PT+N model.
 
-t1_df %>% 
-  select(-`...1`) %>% 
-  mutate(model = gsub("BS", "Bayesian Sampler", model)) %>% 
-  rename(Model = model) %>% 
-  kbl(caption="Original model fitting results with best-fitting model in bold face.", digits=1, booktabs=T) %>% 
-  kable_classic(full_width=F) %>% 
-  row_spec(1, bold=TRUE)
-```
+Given that these models are proposing quite basic psychological processes, we might expect the same process to be shared across all people.
+But, the authors do not report on the overall posterior probability of each model if one model is assumed to explain all participants' responses.
+Such a comparison with these methods would likely be limited in a few ways.
+First, as they note [@zhu.etal2020], BIC cannot fully account for the differences in the competing models’ complexity [also see @piantadosi2018].
+Further, their "unpooled" analysis likely exagerrates the complexity of the models overall and may therefore affect comparisons between them.
 
-In both experiments, the simple version of the Bayesian Sampler scores best, but the complex version of the PT+N model comes in second (lower BIC scores are better).
-It is not obvious what conclusions should be drawn from these results.
-Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction and disjunction fallacies are extremely robust empirical findings [@mellers.etal2001; @sides.etal2002] and are clearly present in the data collected by Zhu and colleagues [-@zhu.etal2020].
-So we might justifiably rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which could instead favor the PT+N theory.
-For their part, Zhu and colleagues [-@zhu.etal2020] chose to average the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are cautious about drawing strong conclusions in favor of their theory.
-
-### Predictive generalization and model complexity
-
-A chief goal of scientific models is to make accurate predictions about future observations.
-When comparing and selecting models based on their predictive accuracy, it is crucial to consider how the models will perform on new, as-yet-unseen data.
-The traditional approach to this issue is to compute the fit of the model to the observed data, but to then correct for the potential for these models to "overfit" these data.
-Typically this correction comes in the form of a "complexity penalty," penalizing models in proportion to their flexibility for accommodating different patterns of data [@gelman.etal2014].
-Zhu and colleagues [-@zhu.etal2020] warn that BIC, which penalizes models based solely on the number of parameters in the model, cannot fully account for the differences in the competing models’ complexity [also see @piantadosi2018].
+#### Accounting for model complexity
 
@@
 First, the models differ not only in the number of parameters but in the domain of those parameters.
-Zhu and colleagues [-@zhu.etal2020] assume that the Bayesian Sampler model’s prior distribution should reflect ignorance or a lack of information.
-A uniform prior, Beta(1, 1) is the most obvious choice in this case, but theoretical arguments can also be made for Beta(0.5, 0.5) (Jeffrey’s prior), Beta(0,0) (Haldene’s prior), or perhaps other symmetric beta distributions Beta($\beta$,$\beta$) with $\beta \in [0, 1]$ [@jaynes2003].
-Via the bridging conditions, they show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to fall within [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values in $[0, 1/2]$.
-$d$ values in ther ange $[1/3, 1/2]$ imply $\beta$ values greater than 1, corresponding to an informative prior over probabilities.
+Via the bridging conditions, Zhu and colleauges [-@zhu.etal2020] show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to fall within [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values in $[0, 1/2]$.
 
-Second, it is not immediately clear how the models’ structural differences impact their flexibility.
-Zhu and colleagues’ [-@zhu.etal2020] bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
-But, what impact does the PT+N model’s treatment of conditional probabilities have on model complexity?
-Is this component of the model a sort of Ptolemaic epicycle, adding complexity to the model that should be penalized?
-Or, does it constitute a commitment to novel predictions that thereby constrain its flexibility?
-Determining model complexity a priori is not always straightforward when the models being compared differ structurally.
+Second, it is unclear how the PT+N model’s treatment of conditional probabilities impacts its complexity.
+This structural difference could be a sort of Ptolemaic epicycle adding complexity to the model that should be penalized or it could constitute a commitment to novel predictions that thereby constrain its flexibility.
 
-Third, one potential explanation for the relative weakness of the "complex" model variants is Zhu and colleagues’ [-@zhu.etal2020] use of "unpooled" models, where parameters are estimated independently for each individual participant.
-In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
+Finally, Zhu and colleagues [-@zhu.etal2020] decision to fit "unpooled" estimates of all participants individually may exaggerate the complexity of both models.
+This is especially an issue in differentiating between simple and complex variants, which Zhu and colleauges [-@zhu.etal2020] elect to ignore by averaging the model scores.
 If there is limited heterogeneity across individuals, then adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants.
-Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
+In addition, this could also affect comparisons across the PT+N and Bayesian Sampelr theories: if one model must assume greater heterogeneity in parameter estimates across people, then this model is more complex.
+Hierarchical models with partial pooling offer a solution that balances between ignoring individual variation and allowing it to vary freely, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
 
-Finally, it is worth recognizing that formal measures of model complexity cannot be expected to perfectly track notions of simplicity or elegance in scientific explanation [for some related discussions, see @kuhn1977; @sober2002; @piantadosi2018].
-For instance, even if the PT+N model's account of conditional probability judgments constrains its flexibility empirically, it seems clear this added component makes it more complex as a putative scientific explanation.
+### Comparison of distributional model predictions
 
-### The present work
+Rather than computing query-level averages across trials for each participant, examining the models' distributional predictions requires modeling participants' raw trial-by-trial responses.
+As mentioned above, this presents substantial challenges.
+To address these, Zhu and colleages [-@zhu.etal2020] estimated discrete versions of the Bayesian Sampler and PT+N models by minimizing the Wasserstein distance between participants' raw responses and model predictions.
+The use of Wasserstein distance rather than a proper likelihood-based measure of model fit helped to minimize issues created by rounding and out-of-support responses, which could otherwise lead both models to assign probability zero to many observations.
+
+Still, the results of this analysis were largely inconclusive with respect to differentiating the models [@zhu.etal2020].
+Specifically, the quality of the fit for each model depended heavily on the maximum number of samples that is assumed possible.
+For small numbers of samples the Bayesian Sampler model is clearly superior, but for larger numbers of samples, the PT+N model was found to better fit the data.
+Presumably this is because with small numbers of samples the PT+N model is extremely constrained in the distinct discrete responses it can predict.
+For small N, both models predict only a limited set of distinct values are possible.
+However, whereas the size of that set is the same for each model, in the Bayesian Sampler the continuous $\beta$ parameter can shift exactly what those discrete values are, providing it much greater flexibility.
+However, this complexity goes unpunished in comparisons based on Wasserstein distance.
+
+In later work, Sundh et al [-@sundh.etal2022] examined the distributional properties of participants responses using indirect means, by regressing the variance of participants' responses across trials on their mean response for each query type.
+Their findings suggest that earlier fits with Wasserstein distance may have produced biased results [@sundh.etal2022].
+They reported evidence for the truncation of responses and a correlation between variance and shrinkage parameter estimates across participants.
+However, their analysis did not enforce that the underlying probabilities driving participants' judgments be coherent (simply estimating the true probability as the mean across trials), nor did they evaluate how frequently participants gave out-of-support judgments that would be inconsistent with the Bayesian Sampler theory.
+
+## The present work
 
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based newer information criteria, such as the widely-applicable information criterion (${\text{WAIC}}$) and Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$).
-In particular, rather than estimating model fit and then penalizing for model complexity, $\text{PSIS-LOO}$ estimates out-of-sample prediction performance directly by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data [@gelman.etal2014; @vehtari.etal2017].
-From these calculations, an estimate of model complexity ($\hat{p}_{\text{LOO}}$) can also be derived.
+
+First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on modern information criteria, such as Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$).
+Rather than estimating model fit and then penalizing for model complexity, $\text{PSIS-LOO}$ estimates out-of-sample prediction performance directly by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data [@gelman.etal2014; @vehtari.etal2017].
+From these calculations, an estimate of model complexity ($\hat{p}_{\text{LOO}}$) can also be derived.[^2]
+
+[^2]: It is worth recognizing that formal measures of model complexity cannot be expected to perfectly track notions of simplicity or elegance in scientific explanation [for some related discussions, see @kuhn1977; @sober2002; @piantadosi2018].
+For instance, even if the PT+N model's account of conditional probability judgments constrains its flexibility empirically, it seems clear this added component makes it more complex as a putative scientific explanation.
 
@@
 This allows for information about model parameters to be shared across participants, resulting in potential improvements to out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
+
+Finally, I implement new extensions of these models to directly model participants trial-level responses while accounting for rounding and out-of-distribution response errors---allowing for principled probabilistic tests of the distributional predictions of the models.
 
@@
 
-## Modeling
+## Modeling Results: participant-level query-averaged responses
 
@@
 
-### Bayesian implementations of the models
+###  Bayesian implementation of participant-level query-averaged response models
 
@@
   P_e(A\lor B) &= (1-d^\prime)P(A \lor B)+d^\prime \\
-  P_e(A|B) &= \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}
+  P_e(A|B) &= \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \tag{8}
 \end{align*}
@@
   P_{e}(A \lor B) &= \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta} \\
-  P_{e}(A|B) &= \frac{N}{N + 2 \beta}P(A|B) + \frac{\beta}{N+2 \beta}
+  P_{e}(A|B) &= \frac{N}{N + 2 \beta}P(A|B) + \frac{\beta}{N+2 \beta} \tag{9}
 \end{align*}
@@
 
-\begin{figure}
-  \begin{minipage}{.5\textwidth}
-    \flushright
 \input{ptn_platediagram}
-  \end{minipage}%
-  \begin{minipage}{.5\textwidth}
-    \flushleft
-    \small
-    \begin{align*}
-      y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
-      \mu_{ijk} &= f_{PT+N}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
-      d_j &= \frac{1}{2} \ \text{logistic}(\delta_j) \\
-      d_j’ &= \frac{1}{2} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
-      \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
-      \delta_j &\sim Normal(0,1) \\
-      \Delta\delta_j &\sim Normal(0,1) \\
-      \kappa &\sim \mathit{Half \hyph Cauchy}(20)
-    \end{align*}
-  \end{minipage}
-  \caption{Complex unpooled PT+N model diagram and formula specifications.
-Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters.
-Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
-\end{figure}
 
@@
 Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters $d$  and $d'$.
+[^3]
+
+[^3]: Uninformativeness was sought in order to reduce bias in the posterior parameter estimates.
+It should be acknowledged that a uniform prior does not exactly correspond to what the authors of the PT+N theory would predict, as they have frequently assumed $d$ to be a fairly small value [e.g.
+@costello.watts2017]
 
@@
 \begin{align*}
-  \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
+  \mu_{ijk} &= f_{BS}(\overrightarrow{\theta_{jk}}, d_j, d_j)  \\
   d_j &= \frac{1}{3} \ \text{logistic}(\delta_j) \\
-  d_j’ &= \frac{1}{3} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big)
+  d_j’ &= \frac{1}{3} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \tag{10}
 \end{align*}
@@
 
-\begin{figure}
-  \begin{minipage}{.5\textwidth}
-    \flushright
 \input{bs_platediagram_mlm}
-  \end{minipage}%
-  \begin{minipage}{.5\textwidth}
-    \flushleft
-    \small
-    \input{bs_equation_mlm}
-  \end{minipage}
-  \caption{Hierarchical complex Bayesian Sampler model diagram and formula specifications.
-Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters.
-Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
-\end{figure}
 
@@
 
-# Results
+Simulation studies verified that the complex hierarchical PT+N and Bayesian Sampler models can correctly and unbiasedly recover parameters from simulated data (see Supplemental Materials).
+
+### Model comparison
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=6in]{plot_compare.png}
-\caption[]{Model comparison results for data from Experiments 1 and 2.
+<!-- \begin{figure}[ht] -->
+<!-- \centering -->
+<!-- \includegraphics[width=6in]{plot_compare.png} -->
+<!-- \caption[]{Model comparison results for data from Experiments 1 and 2.
 Error bars indicate two standard errors of the estimates.
-Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model (\cite{sivula.etal2020}).}
-\end{figure}
+Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model (\cite{sivula.etal2020}).} -->
+<!-- \end{figure} -->
+
+```{r, warning=F, message=F, fig.cap='Model comparison results for data from Experiments 1 and 2.
+Error bars indicate two standard errors of the estimates.
+Typically, a difference of greater than two standard errors is taken as clear evidence for the superiority of the lower-scoring model [@sivula.etal2020].'}
+py$model_comparison %>% 
+  filter(model!="Relative Freq.") %>% 
+  mutate(ul=d_loo + dse*2, ll = d_loo - dse*2) %>% 
+  ggplot(aes(y=reorder(model,-d_loo), x = d_loo, xmin=ll, xmax=ul)) +
+  geom_vline(xintercept=0, linetype="dashed", color="grey") +
+  geom_point(size=2) +
+  geom_errorbarh(height=.01, size=.5) +
+  facet_wrap(~Experiment, scales="free_x") +
+  labs(x = TeX("$\\widehat{elpd}_{LOO}$ difference"), y = "Model") +
+  theme_bw() +
+  theme(panel.grid=element_blank())
+```
 
@@
 However, there is no clear single winner.
-As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler allows for people to bring informative priors to the probability judgment task, that is, a Bayesian Sampler model allowing $\beta \in [0, \infty]$ (greater values of $\widehat{\text{elpd}}_{\text{LOO}}$ are better).
+As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler that allows for people to bring informative priors to the probability judgment task, that is, a Bayesian Sampler model allowing $\beta \in [0, \infty]$ (greater values of $\widehat{\text{elpd}}_{\text{LOO}}$ are better).
 However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\widehat{\text{elpd}}_{\text{LOO}}$ scores within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
@@
 Figure 4 shows the posterior distributions of the population-level $d$ and $d'$ parameters inferred from the unrestricted Bayesian Sampler model.
-In Experiment 2, population-level estimates of $d'$ are greater than $1/3$, as are a substantial number of participant-level estimates for $d$ (37 of 83).
+In Experiment 2, population-level estimates of $d'$ are greater than $1/3$, as are a substantial number of participant-level estimates for $d$ (37 of 83), as shown in Figure X.
 These values fall outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=4in]{plot_params.png}
-\caption{Posterior density of population-level $d$  and $d'$ parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.
-Dashed line indicates theoretical maximum values for Bayesian Sampler model with uninformative priors.}
-\end{figure}
+<!-- \begin{figure}[ht] -->
+<!-- \centering -->
+<!-- \includegraphics[width=4in]{plot_params.png} -->
+<!-- \caption{Posterior density of population-level $d$  and $d'$ parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.
+Dashed line indicates theoretical maximum values for Bayesian Sampler model with uninformative priors.} -->
+<!-- \end{figure} -->
+
+```{r, fig.cap="Posterior density of population-level $d$  and $d'$ parameters estimated from the unrestricted hierarchical Bayesian Sampler model for data from Experiments 1 and 2.
+Dashed line indicates theoretical maximum values for Bayesian Sampler model with uninformative priors."}
+plt_d_avgmodel + plt_d_trialmodel + plot_layout(ncol=1, guides="collect") & labs(x="Estimate") & theme_bw() & theme(panel.grid=element_blank())
+
+```
+
+```{r, fig.cap = "Participant-level estimated d and d' values across Experiments 1 and 2. Error bars indicate 95% CIs."}
+(plt_forest_exp1 + coord_fixed(ratio=59/.5)) + 
+  (plt_forest_exp2 + coord_fixed(ratio=59/.5)) + 
+  plot_layout(guides="collect") & theme(legend.position = 'bottom')
+```
 
@@
 
+Zhu and colleageus [-@zhu.etal2020] demonstrated that the Bayesian Sampler model can capture a set of probabilistic identities developed by Costello, Watts, and colleagues [@costello.watts2016; @costello.etal2018] that capture some of the incoherence in people's probability judgments.
+Following the design of the present experiments, these identities involve combinations of probability estimates for different combinations of two events $A$ and $B$ that should all be equal to zero according to probability theory.
+Under the Bayesian Sampler and PT+N theories, however, some of these identities should be zero, but some are allowed to take on other values.
+Figure X shows the average prediction of the winning model against the average observed value for each equality.
+Consistent with prior findings, this model captures these identities quite closely.
+
+```{r, fig.cap="Average model predicted and observed values for the 18 identities.
+Note that the Bayesian Sampler but not the PT+N model is capable of predicting non-zero values for identities Z10 through Z13.
+In Experiment 1, like participants' responses, the model's estimates are very slightly positive for \\{icy, frosty\\} and very slightly negative for \\{normal, typical\\}.
+This pattern replicates the qualitative pattern reported by Zhu and colleagues."}
+plt_ineq_exp1 + labs(title="Exp. 1") + 
+  plt_ineq_exp2 + labs(title="Exp. 2") + 
+  plot_layout(guides="collect") & theme(legend.position="bottom") & labs(color="", y = "Estimate", x = "Identity")
+```
+
 Finally, it is worth noting that the best of these models provide quite strong overall fits to the data, not just for the query averages, but also for the query averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
+Figure X shows the correlation between participants' responses across all trials and the best-performing model's predictions.
+
+```{r, fig.cap='Posterior predictions for best-fitting model and participants responses in Experiments 1 and 2.'}
+plt_pp
+```
+
+## Raw trial-level response distributions
+
+Comparisons of the models' predictions in expectation have been revealing, but cannot decisively rule between the different process-level psychological theories behind the models.
+To address this question, response distributions must be considered.
+
+Recall, in contrast to a noise-based account, the Bayesian Sampler theory predicts both a truncation of the response distribution as well as a correlation between the degree of shrinkage in probability estimates and the variability of those estimates trial-to-trial (assuming $\beta$ is relatively small).
+Because the Bayesian Sampler applies a Bayesian adjustment after sampling, it predicts probability judgments will always lie between $d$ and $1 - d$, even when zero positive or negative mental samples are drawn (see Equation X and the bridging condition, Equation X).
+First, it bears noting that the truncated response distribution implied by the Bayesian Sampler model appears at odds with the raw response data.
+However, participants' responses frequently lie outside the range implied by the best estimates of their $d$ parameters: XX% in Experiment 1 and XX% in Experiment 2.
+
+Yet comparing the distributional predictions of the models more rigorously poses three challenges: 1) the discrete nature of the models suggests a limited set of allowable responses, assigning zero probability to all others, 2) Bayesian adjustment implies truncation of the support of the response distribution, again assigning zero probability to other response values and 3) participants routinely round their responses, complicating both of the previous issues.
+
+In the following set of analyses I attempt to lay out a set of reasonble assumptions that permit participants' trial-by-trial responses to be modeled and used to compare the theories' predictions.
+To do so, I first extend the models so as to render them fully continuous in their latent space.
+I then marry them with a specific model of response errors.
+Thus the models compared in the following analyses are not identical to those originally proposed by Zhu et al [-@zhu.etal2020] and Costello and Watts [-@costello.watts2014].
+However, they do provide implementations of the theories' process-level accounts, and thus a means to test the distributional predictions of models based on Bayesian adjustment against models based on sampling noise.
+
+### Continuous extensions of the models
+
+Under both models, the variability of people's responses trial-to-trial is driven by the number of mental samples drawn: more mental samples produce less-variable responses.
+However, if the number of samples is considered to be a truly discrete quantity, then only a limited number of discrete responses are possible.
+As Zhu and colleauges [-@zhu.etal2020] note, this is somewhat implausible on its face and their later work has abandoned this assumption [@zhu.etal2021].
+
+At the same time, from a pragmatic perspective it is highly desirable that all latent parameters within the models be continuous.
+The models would be far more tractable to fit if, rather than including a latent Binomial variable representing the discrete number of samples drawn, we could instead model a continuous proportion of samples using, for instance, a Beta distribution.
+
+Zhu and collauges [-@zhu.etal2020] introduce the possibility of an "autocorrelated" Bayesian Sampler model under which samples are assumed to be autocorrelated (ideas which were advanced further in [@zhu.etal2021]).
+As autocorrelated samples provide less information than i.i.d samples, they should be weighted when computing probability estimates.
+The idea is that people actually draw N autocorrelated samples that approximate some smaller $N_{\text{eff}}$ i.i.d samples.
+Assuming the true number of samples drawn, N, is allowed to vary somewhat noisily, then a model based on autocorrelated samples would no longer be limited to predicting a discrete set of possible responses.
+
+What distribution of mental samples would such a process form?
+Figure X shows the result of a simulation study examining probability estimates drawing on autocorrelated samples with estimated weights.
+Each simulation consists of a simple discrete Markov chain generating autocorrelated samples at $x_t$, returning $x_{t-1}$ with probability $p_{a}$, or else returning a new sample distributed Bernoulli($p$) was simulated.
+The setting of $p_a$ controls the degree of (positive) autocorrelation in the chain.
+Given the degree of autocorrelation in a sample, a weight is initially estimated as $N_{\text{eff}}/N$ and an appropriate number of samples $N$ are drawn to produce $N_{\text{eff}}$ samples.
+A probability estimate is then calculated from these samples.
+As the figure shows, for large-enough autocorrelation values ($p_a$) the resulting distribution becomes quite smooth, with the exception of spikes at zero and one for small to moderate values of $N_{\text{eff}}$.
+This occurs when all samples are zero or all one, because $\sum_i^N0 \cdot w_i = 0$ for any $w_i$ and $\sum_i^N 1 \cdot w_i = 1$ for any weights $\sum_i^N w_i = 1$.
+
+< Figure of simulation study >
+
+Sampling processes producing zero positive or negative examples should produce heaping of responses at zero and one under the PT+N model and at $d$ and $(1-d)$ under the Bayesian Sampler model.
+Compared against human judgment data, it is at least clear that actual human judgments do not equal zero or one as frequently as would be expected under the PT+N model using such a sampling process.
+Visual inspection of histograms of each participants' judgments and their estimated $d$ parameters also make clear that heaping at $d$ and $(1-d)$ is frequently absent (see Supplemental Materials).
+
+One possibility is that there is another source of error that explains these discrepancies, for instance that people just tend to avoid giving "extreme" responses.
+Another possibility is that the mental sampling process somehow enforces sampling of both positive and negative examples.
+A mental sampling process must start somewhere, and these initialization point(s) may influence judgments.
+For example, features of the initializations of mental sampling processes have been proposed to anchoring [@lieder.etal2013] and unpacking effects [@dasgupta.etal2017; also see @sanborn.chater2016].
+In this spirit, one possible explanation for the relative absence of extreme responses is that people draw samples following two initializations, initializing sampling with both a positive and negative instance.
+As the simulations in Figure X show, this produces a predicted response distribution that is  approximately $Beta(1 + p N_{\text{eff}}, 1+(1-p)N_{\text{eff}})$.
+
+Incorporating these assumptions provides a psychologically plausible mechanism to permit the use of a Beta distribution to approximate the latent sampling process, rendering both of the models' latent parameters continuous.
+Note that this process will also itself produce some  shrinkage in the probability estimates, and this shrinkage will be somewhat sensitive to the number of samples drawn in both the Bayesian Sampler and PT+N models.
+Nevertheless the dependency between shrinkage and variance remains stronger for the Bayesian Sampler model and it is only the Bayesian Sampler model that predicts a truncation of the response distribution.
+
+### Mixture modeling: rounding and contaminants
+
+Creating continuous extensions of the models makes their estimation more tractable.
+However, a specific model of participants' response processes and errors is still needed to capture rounded and out-of-support responses.
+To address these challenges, I implemented variants of the PT+N and Bayesian Sampler models within discrete mixture models allowing for varying rounding policies as well as "contaminant" responses generated by noise processes outside the models.
+
+First, it is clear that participants have rounded a majority of their responses.
+This sort of rounding can be modeled by a categorical distribution across the discrete possible rounded responses.
+Each rounded response category corresponds to a set of cut points, $a$ and $b$, with the probability of the categorical response defined by the CDF of the Beta distribution.
+
+$$P([a,b)) = B(a, \mu N, (1-\mu)N) - B(b, \mu N, (1-\mu)N) \tag{11}$$
+
+Where $B$ is the cumulative distribution function of the underlying latent distribution.
+As participants were allowed to respond freely with whole numbers from 0 to 100, the exact rounding policy for each response is unknown.
+Nevertheless these rounding policies can be estimated via mixture modeling.
+For simplicity, rounding to the nearest 5% was enforced for all responses.
+Then, the probability of these categorical responses are computed for 21 and 11 categories (corresponding to rounding to 5% and 10%).
+These probabilities were combined along with a uniform probability representing "contaminants" according to mixing probabilities $\phi$, distributed with a Dirichlet prior (see Appendix for further implemenatation details).
+
+The Bayesian Sampler model predicts a truncated range of possible responses given $\beta$ and $N$ (and consequently, implied $d$).
+Modeling these different rounding processes allows for at least some out-of-bounds responses to be accounted for by rounding processes (e.g.
+when an allowable response of .14 is rounded to the out-of-bounds value of .10).
+However, some participant response still cannot be accounted for by the model.
+Instead, these responses are treated as "contaminants" generated by a random response process.
+Modeling "contaminant" response processes allows the Bayesian Sampler model to be fit in the presence of true outliers.
+Identifying the estimated proportion of "contaminant" responses can also provide a check on the models: if a model can only be fit by assuming a large proportion of contaminant responses, this suggests it is likely not a good model of human behavior.
+
+### Trial-level noise-based model
+
+Compared to the query-averaged model, the trial-level noise-based model adds two features: mixture components for rounding and contaminants and subject-level varying $N$ in place of a fixed $K$ parameter.
+This model's implementation and the implementation of its mixture components is depicted in Figure X.
+However, note that $N$ is allowed to vary independently from $d$, allowing for independence between response shrinkage and variability.
+Moreover, the latent distribution used to calculate categorical cut points is an untransformed Beta distribution.
+
+\input{trial_level_mlm_noise_based}
+
+### Trial-level Bayesian Sampler model
+
+Above we found that an autocorrelated sampling process could be reasonably approximated by a Beta distribution. Taking the original Bayesian Sampler model
+
+$$\hat{P}_{BS}(A) = \frac{S(A)}{N+2\beta} + \frac{\beta}{N+2\beta} \tag{12}$$
+
+We can replace the number of successes $S(A)$ (distributed binomial) with the quantity $\rho(A)N$, where $\rho(A)$ represents the Beta-distributed sample proportions generated by the autocorrelated sampling process outlined above.
+
+$$\hat{P}_{BS}(A) = \frac{\rho(A)N}{N+2\beta} + \frac{\beta}{N+2\beta} \tag{13}$$
+
+Then it is plain that $\hat{P}_{BS}(A)$ is a transformation of $\rho(A)$, and therefore a transformed Beta distribution (see Appendix for derivation).
+Figure X below diagrams the entire Bayesian Sampler model, now parameterized in terms of $\beta$, $N$, and $N'$.
+
+\input{trial_level_mlm_bs}
+
+This model assumes that the number of samples drawn on each trial is fixed as $N$ or $N'$ accordingly (modulo the uncertainty about these parameters).
+However, it also seems reasonable to imagine that the number of samples drawn in fact varies from trial-to-trial.
+This added complexity is likely ignorable for a noise-based model, but could be important for the Bayesian Sampler model, as the number of samples drawn affects the truncation of the response distribution.
+That is, allowing the number of samples drawn to vary across trials may allow the model to capture some respnses that would otherwise be treated as contaminants.
+To capture this, the model can be given one additional extension to allow the number of samples to vary, by adding a new parameter $N_{trial}$.
+This parameter multiplies the number of samples as a fraction of each individual participants' average number of samples drawn, e.g.
+so that a participant might sometimes draw 1.5x or 2x the number of samples they typically draw.
+The appropriate amount of variation in $N_{trial}$ is constrained to be fairly small, but is estimated hierarchically: I assume $log(N_{trial}) \sim N(0,\sigma_{N_{trial}})$ and $log(\sigma_{N_{trial}}) \sim (-1, .3)$.
+
+## Model comparison: Raw trial-level response models
+
+Prior to fitting the models, response data were rounded to the nearest 5%.
+Nearly all responses (XX%) were already divisible by five, and this was necessary to speed model fitting.
+Even still, estimating model posteriors for the trial-level mixture models using MCMC proved intractable.
+Instead, model posteriors were estimated using Stochastic Variational Inference (SVI) with Numpyro [@phan.etal2019] using a multivariate Normal guide [e.g.
+@kucukelbir.etal2015] and 10,000 SVI steps.
+Estimating each model posterior took about 40 minutes on an Nvidia V100 GPU.
+Simulation studies verified that this estimation approach could reliably recover parameters from simulated data (see Supplemental Materials).
+
+Table X shows the scores of each model fit to the trial-level data in Experiments 1 and 2.
+From the quantitative model comparison it is clear that the noise-based model is superior, with substantialy lower $\widehat{\text{elpd}}_{\text{LOO}}$ than both of the competing Bayesian Sampler implementations in both experiments.>
+
+< table X: model comparison results >
+
+The models' performance can be better understood by examining the two main features about which the models' predictions depart: truncation of the response distribution and shrinkage-dependent variance of responses.
+
+The non-varying Bayesian Sampler model fits quite poorly and is estimated to have a large proportion of "contaminant" responses (X% in experiment 1, X% in experiment 2).
+This is likely due to a lack of the predicted truncation of the response distribution.
+As with the trial-average models, the estimates of implied d are quite high, which would predict substantial truncation.
+As mentioned earlier, many of participants' responses fall outside the range implied by their most likely implied d values.
+
+Allowing the effective number of samples drawn to vary trial-by-trial improves the fit of the Bayesian Sampler model substantially and decreases the estimated proportion of contaminant responses (X% in experiment 1, X% in experiment 2).
+Nevertheless, these results still indicate inferior fit compared with the noise-based model, which has substantially lower $\widehat{\text{elpd}}_{\text{LOO}}$ and again attributes fewer responses to the contaminant process (X% in experiment 1, X% in experiment 2).
+
+Second, we can examine the shrinkage-variance relationship.
+In the noise-based model, N and d were allowed to vary freely.
+But if these quantities indeed are actually correlated as predicted by a Bayesian adjustment model, then we should expect to nevertheless see a correlation between the subject-level estimates in these parameters.
+Figure X shows a scatterplot of these estimates.
+Looking across subjects we see they are mostly uncorrelated, against the second BS prediction.
+>
+
+```{r, fig.cap = "Scatterplots showing the relationships between subject-level d and N estimates from the trial-level noise-based model for Experiments 1 and 2.
+Figure for Experiment 1 excludes some outlier paricipants who gave repetitive resposnes that resulted in abnormally high N-values.
+Error bars indicate 95% CI.
+Although there is a slight negative correlation in Experiment 1, this is driven by only a handful of participants with extreme values and in Experiment 2 the pattern appears to if anything be reversed."}
+
+plt_dk_ptn1  + plt_dprimek_ptn1 + plt_dk_ptn2 + plt_dprimek_ptn2
+```
+
+Comparing the d values inferred from the trial-level model we see they are similar to but generally smaller than those estimated in the trial-averaged model.
+This is likely because the hypothesized initialization process accounts for some of the shrinkage in participants' responses, so that less need be attributed to sample read noise.
+However, the numbers estimated in Experiment 2 are still substantially larger than in most prior work [cite].
 
@@
 
-By a fair margin, the models best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] were implementations of the Bayesian Sampler theory, with the best single model being a version of the Bayesian Sampler model without restriction on the range of its $\beta$ parameters.
-Alternatively, this model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
-Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this task.
+Fit to the average of participants' responses over blocks, there is a single clear winner: a model without any special treatment of conditional probability (ala the Bayesian Sampler model) and allowing for an implied $d$ parameter $\in [0,.5]$.
+This model could be interpreted either as a variant of the Bayesian Sampler without restriction on its $\beta$ parameter, or as a variant of the PT+N model that removes its account of conditional probability judgments.
+
+In either case, these findings make clear that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this task.
 In keeping with the larger theoretical framework of Bayesian cognitive science, the Bayesian Sampler theory assumes that subjective probabilities underlie people’s probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process [@chater.etal2020; @zhu.etal2020].
 
-Zhu and colleagues’ experiments cast some doubt on their proposal that the priors of the Bayesian Sampler model should reflect "ignorance priors," symmetric Beta distributions with $\beta \in [0, 1]$.
-As a generic prior that would be used across contexts, this class of uninformative priors has clear appeal.
-Nevertheless, many participants' estimated $d$ and $d'$ parameters fell outside the $[0, 1/3]$ range implied by "ignorance" priors.
-Given the different values of $d$ and $d'$ across Experiments 1 and 2, it seems likely that people bring informative priors to at least some probability judgment tasks.
+The unrestricted BS model and PT+N with Bayesian conditional probability judgments cannot be distinguised by their predictions about expected responses.
+However, they remain substantively different models positing different process-level accounts of probability judgments.
+Both accounts rely on the concept of mental sampling.
+But in the BS model the samples are integrated with a Bayesian prior to produce a judgment, whereas in the PT+N model, noise in the sample-reading process produces bias.
 
-How might these theories explain the difference in implied $d$ parameters between Experiments 1 and 2?
-Under the Bayesian Sampler theory, it could be that people bring informative and domain-specific priors to the probability judgment task, so that participants' priors differed across the different contexts of Experiments 1 and 2.
-However, these experimental contexts are highly similar, so it is not entirely clear why this should be the case.
-Alternatively, it could be that participants held identical informative priors in both contexts, but that they were able to more effectively draw mental samples for the queries presented in Experiment 1 than in Experiment 2 (resulting in lower $d$).
-Similarly, under a noise-based account, it could be that the mental samples are read more noisily in Experiment 2 than Experiment 1.
-Unfortunately, teasing apart these different accounts is likely to prove challenging.
-First, the same sort of factors that could be expected to influence the number of samples could also plausibly influence "noise" (e.g.
-task complexity, expertise).
-Moreover, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and $N$ parameters are not uniquely identifiable from this kind of judgment data alone.
+The Bayesian adjustment process hypothesized by the Bayesian Sampler model makes two clear predictions about the distribution of participants' responses.
+First, this implies a truncated range of possible responses.
+Second, assuming that $\beta$ is constrained to be a relatively small value, then under the Bayesian Sampler model, $N$ influences both the degree to which responses are shrunk toward .50 and the variability of those responses trial-to-trial.
+Thus, participant's inferred $N$ values should be somehow correlated with the noise in their trial-level responses.
+In contrast, under the noise-based account of the PT+N model, there is no truncation of responses and no predicted correlation between shrinkage and response variability.
 
-In addition, the implied $d$ and $d'$ noise parameters also varied across individuals.
-Of note, some participants' implied $d$ and $d'$ parameters _were_ consistent with the class of ignorance priors, even in Experiment 2.
-Further research should explore this heterogeneity within and across individuals.
+Directly fitting the models to these trial-level human response data presents several challenges addressed here by creating continuous extensions of these models and incorporating additional modeling of response processes.
+Though not identical to the models as originally proposed, this exercise allowed for tests of the key distributional predictions of the theories.
+Neither of the Bayesian Sampler's predictions appear to be borne out by the data.
+First, participants' responses frequently fall outside the truncated range implied by the parameters estimated under the Bayesian Sampler model.
+Fit to the raw data, this requires treating an unreasonably large proportion of responses as "contaminants".
+Second, the degree of shrinkage in participants' responses and the variability in those responses appear largely uncorrelated.
 
-One thing this model comparison has not decided, and likely *cannot* decide, is whether distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given the two models' tight connections via the bridging condition [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
-Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that "noise" might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].
+Altogether then, the distributions of participants' responses are more consistent with the PT+N theory's account of sampling noise than the Bayesian adjustment implied by the Bayesian Sampler theory.
+In the end, I find the best overall account of participants' probability judgments is a continuous extension of the PT+N theory without any special treatment of conditional probability judgments.
 
-Finally, some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020].
+## Remaining questions and limitations
+
+Despite the model's quantitative success, some more qualitative questions remain.
+First, it bears asking: how plausible are the parameter values inferred from the model?
+Many participants' estimated $d$ and $d'$ parameters were quite high---potentially against the spirit of the original PT+N model.
+This model bounds $d$ at .50 in principle, but a sample-reading process with such high error-rates may or may not be plausible.
+In prior work simulations have often assumed values of $d$ around .10 (Howe & costello, 2020; Costello & Watts, 2017)].
+Further research examining what factors might affect the mental sampling and reading processes (e.g.
+task complexity, distractions, prior experience) might help to shed light on the plausibility of $d$ estimates in different contexts.
+
+High estimates of $d$ parameters might also call into question arguments for the rational utility of such a process.
+Zhu and colleagues argue that the regularizing effect of bayesian adjustment should be seen as adaptive.
+They also consider that "noise" might give an algorithmic-level solution to the computation-level goals defined by the Bayesian Sampler [-@zhu.etal2020].
+Even high implied $d$ values might be consistent with rational inference in cases where the number of effective mental samples is very low.
+For instance, a beta(2,2) prior is only modestly informative, but could produce $d = .40$ if $N=1$.
+However, if more samples are drawn then high $d$ values would correspond to potentially inflexible and suboptimal priors (i.e.
+too large $\beta$ values).
+From the results, it is clear there are some individuals with both relatively high $d$ and $N$ values, which may press somewhat against the rational justification for the desirability of sampling noise.
+
+As noted, the comparisons of these models using trial-level data rests on a number of elaborating assumptions to support fitting of the models.
+In particular, the models' implementations incorporate additional assumptions and approximations of the autocorrelated nature of mental samples and of ancillary error processes (e.g.
+rounding).
+It should be recognized that different assumptions may have produced different results, and other error processes remain possible.
+Although other indirect analysis approaches might be designed to avoid these concerns [e.g.
+@sundh.etal2021], ultimately it seems crucial that cognitive models at some point be fit to the actual human behaviors of interest.
+
+Finally, one important direction for future elaborations of sampling-based theories are more rigorous theories of realistic mental sampling processes, including details of the initialization and autocorrelation among samples.
+Here, I proposed a relatively-highly autocorrelated sampling process initialized with a positive and negative example to account for qualitative features of participants' response distributions and to support a simple approximation.
+Future work might explore the details of these processes and incorporate them into models of probability judgments in more rigorous ways.
+
+## Conclusions
+
+The implications of sampling-based models like the Bayesian Sampler and PT+N theory go well-beyond the probability judgment task itself: these models esssentially rescue probabilistic theories of cognition by rejecting the explicit representation of probabilistic credences.
+Instead, credences are emergent properties of mental models that probabilistically generate samples [@chater.etal2020].
+
+Clear models of these processes offer the potential to support a measurement model for probabilistic beliefs.
 Indeed, by representing the "true" subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
@@
 @franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006].
-For instance, people's responses in various reasoning tasks are often explicitly related to inferred subjective mental probabilities (e.g.
-@jern.etal2014), so accounting for biases in those reports may permit more rigorous model testing.
+For instance, people's responses in various reasoning tasks are often explicitly related to inferred subjective mental probabilities [e.g.
+@jern.etal2014], so accounting for biases in those reports may permit more rigorous model testing.
 One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes [e.g.
 @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].
+
+\pagebreak
+
+# Appendix
+
+For the trial-level response models participants' rounded responses are modeled as discrete responses with a categorical (multinomial) distribution.
+For $i \in \{0,1, ...,m\}$ where $m=20$ possible responses, define a set of cut points $a_i =  \frac{i}{m}-\frac{1}{2m}$ and $b_i = \frac{i}{m}+\frac{1}{2m}$.
+Using $x|_{[0,1]}$ to denote that $x$ is restricted to the domain $[0,1]$, the probability of each response given $\mu$ and $N$ is:
+
+$$
+\begin{aligned}
+p_{i,5} &= P([a_{i,5},b_{i,5})) \\
+& = B(a_{i,5}|_{[0,1]}, \mu N, (1-\mu)N) - B(b_{i,5}|_{[0,1]}, \mu N, (1-\mu)N)
+\end{aligned}
+$$
+
+Where $B$ is the appropriate cumulative distribution function.
+To capture rounding to 10 we define $a_{i,10} =  \frac{2i}{m}-\frac{1}{m}$ and $b_i = \frac{2i}{m}+\frac{1}{m}$, so that the probability of each response is:
+
+$$
+p_{i,10} = \begin{cases}
+  P([a_{i,10},b_{i,10})) & \text{i is even} \\
+  0  & \text{i is odd}
+\end{cases}
+$$
+
+Next defined a vector of mixture probabilities $\overrightarrow{\phi}$, with the zeroth index indicating a "contaminant" process.
+Combining these response processes, we can define the marginal probability of each response as:
+
+$$
+p_i = \frac{1}{21} \phi_0 + p_{i,5} \phi_1 + p_{i,10} \phi_2
+$$
+
+And then responses themselves are distributed Categorical:
+
+$$
+y_i \sim Categorical(\overrightarrow{p})
+$$
+
+For the noise-based model, $B$ is the incomplete Beta function, the CDF of the Beta distribution.
+The computations for the Bayesian Sampler model response probabilities are identical save that instead of $B(x, \alpha, \beta)$ we have $B(f_{BS}^{-1}(x), \alpha, \beta)$ when computing the probability of each response $p_i$, and we use $N$ and $N'$ where appropriate.
+To see this, let $X$ be the Beta distribution success proportions from the mental sampling operations, $\rho(A)$, and let $Y$ be the distribution of resulting probabilities from the Bayesian sampler model.
+Then  $Y = g(X)$ where $g$ is the function defined in equation 13 from the manuscript.
+
+$$\hat{P}_{BS}(A) = \frac{\rho(A)N}{N+2\beta} + \frac{\beta}{N+2\beta} \tag{13}$$
+
+Letting $F_X$ and $F_Y$ be the CDF of $X$ and $Y$ respectively, we have that:
+
+$$
+F_Y(y)= P(Y\leq y) = P(g(X)\leq y) = P(X\leq g^{-1}(y)) = F_X( g^{-1}(y))
+$$
+
+Putting this all together, define $Z_{NB}$ as the function which calculates the probability of each categorical response under the noise-based model given the inputs of $\mu_{ijk}, d_j, d'_j$ and $\phi$.
+Here, $\mu_{ijk} = f_{\text{NB}}(\overrightarrow{\theta_{jk}}, d_j, d'_j, x_{ijk})$ computes the expected probability according to the PT+N theory except that it treats conditional probability judgments like simple probability judgments.
+
+Finally, define $Z_{BS}$ as the function which calculate the probability of each categorical response under the Bayesian Sampler model.
+Note that, here, $\mu_{ijk} = f_0(\overrightarrow{\theta_{jk}}, x_{ijk})$ where the value of $\mu$ depends only on the underlying probabilities and query asked on a specific trial.
 

### 2022-09-16 paper-rmd/paper-papaja.Rmd
created revised manuscript

@@
 library("papaja")
+options(encoding = "UTF-8")
 r_refs("references.bib")
@@
 
-\pagebreak
+# References
 
-# Declarations
-
-### Funding
-
-No funding was obtained for this study.
-
-### Conflict of Interests
-
-The author declares no conflicts of interest.
-
-### Availability of Data and Materials
-
-This paper presents secondary analyses of data. The datasets generated and/or analysed during the current study are available at https://osf.io/mgcxj/files/.
-
-### Code availability
-
-All analysis code is available at https://github.com/derekpowell/bayesian-sampler and at https://osf.io/bpkjf/.
-
-### Author contributions
-
-Derek Powell is the sole author of this manuscript.
-
-### Ethical approval
-
-Not applicable.
-
-### Consent to participate
-
-Not applicable.
-
-### Consent to publish
-
-Not applicable.
+::: {#refs custom-style="Bibliography"}
+:::

### 2022-09-16 paper-rmd/paper-papaja.tex
created revised manuscript

@@
 \documentclass[
-  english,
   man,floatsintext]{apa6}
@@
 \usepackage{lmodern}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Comparing probabilistic accounts of probability judgments},
-  pdfauthor={Derek Powell1},
-  pdflang={en-EN},
-  pdfkeywords={probability judgments, Bayesian cognitive science, heuristics and biases},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 \mathchardef\hyph="2D
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
-\ifluatex
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
@@
 \addbibresource{references.bib}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Comparing probabilistic accounts of probability judgments},
+  pdfauthor={Derek Powell1},
+  pdflang={en-EN},
+  pdfkeywords={probability judgments, Bayesian cognitive science, heuristics and biases},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 
-\pagebreak
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-\hypertarget{declarations}{%
-\section{Declarations}\label{declarations}}
-
-\hypertarget{funding}{%
-\subsubsection{Funding}\label{funding}}
-
-No funding was obtained for this study.
-
-\hypertarget{conflict-of-interests}{%
-\subsubsection{Conflict of Interests}\label{conflict-of-interests}}
-
-The author declares no conflicts of interest.
-
-\hypertarget{availability-of-data-and-materials}{%
-\subsubsection{Availability of Data and Materials}\label{availability-of-data-and-materials}}
-
-This paper presents secondary analyses of data. The datasets generated and/or analysed during the current study are available at \url{https://osf.io/mgcxj/files/}.
-
-\hypertarget{code-availability}{%
-\subsubsection{Code availability}\label{code-availability}}
-
-All analysis code is available at \url{https://github.com/derekpowell/bayesian-sampler} and at \url{https://osf.io/bpkjf/}.
-
-\hypertarget{author-contributions}{%
-\subsubsection{Author contributions}\label{author-contributions}}
-
-Derek Powell is the sole author of this manuscript.
-
-\hypertarget{ethical-approval}{%
-\subsubsection{Ethical approval}\label{ethical-approval}}
-
-Not applicable.
-
-\hypertarget{consent-to-participate}{%
-\subsubsection{Consent to participate}\label{consent-to-participate}}
-
-Not applicable.
-
-\hypertarget{consent-to-publish}{%
-\subsubsection{Consent to publish}\label{consent-to-publish}}
-
-Not applicable.
+\hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\end{CSLReferences}
 

### 2022-09-16 paper-rmd/paper.Rmd
created revised manuscript

@@
 The PT+N theory provides a unified accounts for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
-
-Meanwhile, Zhu, Sanborn, & Chater [-@zhu.etal2020] have proposed a Bayesian model of probability judgment they call the “Bayesian Sampler.” Under this model, probability judgment is itself seen as a process of Bayesian inference.
-To judge the probability of an event, a limited number of samples are again drawn from a mental model of the event.
-Then, those “observed” samples are integrated with a prior over probabilities to produce a probability judgment.
-The effect of this prior is to regularize the judgments, pushing them away from extreme values.
-They argue this regularization of judgments is adaptive, in that it pushes judgments and decisions away from extremes that might, e.g.
-encourage taking large risks based on a small number of mental samples.
-However, this hedging comes at the cost of systematic incoherence and biases.
-Like the PT+N model, their proposed model accounts for a wide array of biases in probability judgments, including the novel biases identified by Costello and Watts [@costello.watts2014; @costello.watts2016].
-
-Under the simplest version of the Bayesian Sampler model, the expected value of a probability judgment is given by the following formula:
-
-$$E[\hat{P}_{BS}(A)] = \frac{N}{N+2\beta}P(A) + \frac{\beta}{N+2\beta}$$
-
-where *N* represents the number of mental samples drawn.
-The mentally-observed samples are integrated with a Bayesian prior over probabilities, represented with a symmetric beta distribution parameterized by $\beta$.
-Zhu and colleagues [-@zhu.etal2020] show that the *N* and $\beta$ parameters of their model can be related to the *d* parameter of the PT+N model via the following bridging formula:
-
-$$d = \frac{\beta}{N+2\beta}$$
-
-### Two accounts of conditional probability judgments
-
-By explaining the incoherence of human probability judgments using coherent mental probabilities, both models have the potential to rescue the larger project of Bayesian cognitive science as applied to everyday beliefs [@chater.etal2020].
-However, because they diverge in their treatment of conditional probability judgments, the two theories fit very differently into the larger Bayesian project.
-According to the Bayesian sampler model, conditioning is something that happens in the mental model of the events, not as part of the process of rendering probability judgments.
-Thus, the Bayesian sampler model does not assign any special status to conditional probability judgments and so fits neatly into the larger Bayesian project of cognitive science: probability judgments are simply another judgment process applied to the outputs of other Bayesian mental models [@chater.etal2020].
-
-In contrast, the PT+N model presents a constructive account of conditional probability judgments that is fundamentally non-Bayesian [@costello.watts2016].
-According to the PT+N model, conditional probabilities P(A|B) are estimated by a two-stage sampling procedure: first both events A and B are sampled with noise, and then a second noisy process computes the ratio of the events read as A and B over events read as B.
-The PT+N model predicts conditional probability estimates using the following equation:
-
-$$ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} $$                                            
-Given that conditional inference and belief updating lies at the core of Bayesian cognitive theory, PT+N’s account of conditional probability judgment separates it from the Bayesian sampler theory not just quantitatively, but quite fundamentally.
-First, it denies any central importance of conditional probability judgments by pulling conditioning out of the mental model of an event and placing it into the probability judgment process.
-And perhaps more importantly, this non-bayesian account of conditional inference can no longer be considered “rational” at any stage under the PT+N model.
-
-## Comparing the models
-
-Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model against Costello & Watts’ (2014; 2016; 2017; 2018) PT+N model as explanations for human probability judgments in two experiments.
-Unfortunately, their results were somewhat equivocal.
-They fit both “simple” and “complex” versions of each model and computed a Bayesian Information Criteria score (BIC) for all models.
-The complex versions of these models introduce additional parameters d’ and N’ that allow for different patterns of judgments for conjunctive and disjunctive judgments as compared with simple probability judgments.
-These additional parameters are needed in order to explain conjunction fallacies (Costello & Watts, 2017; Zhu et al., 2020).
-Table 1 below presents the total BIC scores computed for each model as originally fit, using the authors’ original code and saved model outputs (Zhu et al, 2020, Supplementary materials).
-
-```{r table1, echo=FALSE, message=F, warning=F}
-t1_df <- read_csv("original-fits.csv")
-
-t1_df %>% 
-  select(-X1) %>% 
-  mutate(model = gsub("BS", "Bayesian Sampler", model)) %>% 
-  rename(Model = model) %>% 
-  kbl(caption="Original model fitting results", digits=1, booktabs=T) %>% 
-  kable_classic(full_width=F)
-
-```
-
-In both experiments, the simple version of the Bayesian Sampler scores best, but the complex version of the PT+N model comes in second (lower BIC scores are better).
-It is not obvious what conclusions should be drawn from these results.
-Of course, the simple Bayesian Sampler model appears to win by the numbers.
-But conjunction fallacies are an extremely robust empirical finding (e.g.
-cite) and are clearly present in the data collected by Zhu and colleagues (2020).
-So we might rule out the simple variants of the models on the grounds that they will fail to capture important qualitative features of the data, which would instead favor the PT+N theory.
-For their part, Zhu and colleagues (2020) elect to split the baby by averaging the simple and complex model scores together.
-This approach somewhat favors the Bayesian Sampler theory overall, though they are rightfully cautious about drawing any strong conclusions.
-
-### Accounting for Model Complexity
-
-Zhu and colleagues (2020) warn that BIC, which penalizes models based only on the number of parameters in the model, cannot fully account for the differences in the models’ complexity.
-There are at least three challenges to accounting for model complexity in the comparison of these models.
-First, the models differ not only in the number of parameters but in the domain of those parameters.
-Zhu and colleagues (2020) restrict the range of the $\beta$ parameter in their Bayesian Sampler model to fall within the interval [0,1].
-This restriction constrains the Bayesian Sampler model’s prior distribution to the class of “ignorance priors”, symmetric beta distributions where $\beta < 1$.
-Via the bridging conditions (eq X), they show that this restricts the “noise level” for the Bayesian Sampler,  represented in implied d under the PT+N model, to [0, $\frac{1}{3}$] whereas the PT+N model permits noise values between [0, 0.50].
-
-Second, it is not immediately clear how the structural differences between the models impact their flexibility.
-For the purposes of model comparison, the “complexity penalty” that models truly ought to pay is determined by their relative flexibility in accommodating different patterns of data, not simply the number or domain of their parameters (cite Gelman).
-To be sure, Zhu and colleagues’ (2020) bridging condition makes it clear that the PT+N model is more flexible when it comes to predicting unconditional probabilities.
-But, what impact does the PT+N model’s treatment of conditional probabilities have on model complexity?
-Is this component of the model a sort of Ptolemaic epicycle, adding complexity to the model that should be penalized?
-Or, does it constitute a commitment to novel predictions that thereby constrain its flexibility?
-Determining model complexity a priori is not always straightforward when the models being compared differ structurally.
-
-Finally, one potential explanation for the relative weakness of the complex model versions is Zhu and colleagues’ (2020) use of unpooled models, where parameters are estimated independently for each individual participant.
-Adding a parameter for each participant may effectively over-penalize the “complex” variants relative to the simple variants, especially if there is actually limited heterogeneity across participants.
-In contrast, a comparison of fully pooled variants of the simple and complex models (where a single population parameter is used for all participants) would require adding only one extra parameter to the penalty term.
-Partial pooling is a solution that balances between these extreme approaches, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
-
-### The present work
-
-Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that permits a more decisive (or at least alternative) comparison.
-A Bayesian analysis allows issues of model complexity to be addressed through comparisons of model fit based on information criteria, such as WAIC and LOO~ic~.
-These indices allow for complexity penalties that directly capture differences in the flexibility of the models, rather than being estimated based solely on the number of distinct parameters [cite gelman].
-In addition, this framework also supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in an overall reduction in model complexity and a more realistic test of the models.
-
-# Methods
-
-## Data selection
-
-Zhu, Sanborn, & Chater (2020) conducted two experiments to compare the PT+N and Bayesian sampler theories.
-These experiments asked participants to judge the probability of different events in various combinations.
-Following prior work by Costello and Watts [e.g.
--@costello.watts2016; -@costello.watts2018], both experiments focused on the everyday events of different kinds of weather.
-
-Experiment 1 asked about the events {icy, frosty} and {normal, typical} (e.g.
-“what is the probability that the weather in London is normal and not typical?”).
-The authors’ goal was to ask about highly correlated events, but the events used are perhaps nearly perfectly correlated.
-Because the terms used to describe these events are nearly synonymous, there is a concern about the interpretation of the statements evaluated in this experiment.
-This is especially clear, as the authors note, for disjunctive trials such as “normal or typical,” where “or typical” might not be read as a disjunction but rather an elaborative clause.
-In light of these concerns, I excluded the disjunctive trials from Experiment 1 from my analyses.
-Experiment 2 focused on more moderately correlated events, {cold, rainy} and {windy, cloudy}, that do not admit these misinterpretations.
-In addition, a third experimental condition asking about {warm, snowy} was also included in the experiment, but was dropped from the analyses reported in the paper.
-Exploring the raw responses from this condition reveals a substantial fraction of “zero” and “one” responses for certain trials.
-This may reflect a different response process than was intended.
-For instance, some participants may have engaged in deductive reasoning to judge that it is not possible for the weather to at once be warm and snowy, and therefore responded with zero—failing to properly consider that it is possible (at least logically) for it to be warm and snowy at different times within the same day.
-Given these potentially aberrant responses, I followed Zhu and colleagues (2020) in ignoring data from this condition.
-
-## Modeling
-
-I implement both the Bayesian Sampler and PT+N models as Bayesian models in the probabilistic programming language Numpyro.
-All code and results are available as supplemental materials (https://github.com/derekpowell/bayesian-sampler).
-
-### Bayesian implementations of the models
-
-The PT+N model defines probability judgments as
-
-$$P_{e}(A) = (1-2d)P(A) + d$$
-$$P_e(A\land B) = (1-d^\prime)P(A \land B)+d^\prime$$
-$$P_e(A\lor B) = (1-d^\prime)P(A \lor B)+d^\prime$$
-$$P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}$$
-
-In contrast, the Bayesian Sampler model defines probability judgments as
-
-$$P_{e}(A) = \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta}$$
-$$P_{e}(A \land B) = \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta}$$
-$$P_{e}(A \lor B) = \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta}$$
-
-Fixing d and d’ or N and N’ equal yields the “simple” variant of each of the models, which treat conjunctive and disjunctive probability judgments identically to simple probability judgments.
-
-Notice that for each model the probability judgments depend on underlying subjective probabilities, derived from a mental sampling process.
-These subjective probabilities are unobserved, and must be estimated as a latent variable.
-Here, they are estimated as a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events (A&B, \~A&B, A&\~B, \~A&\~B).
-
-Zhu, Sanborn & Chater (2020) implement completely unpooled models with separate d, d’, N, N’, and $\beta$ parameters for each participant.
-Translated into a Bayesian Framework, we can write the PT+N model as:
-
-$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{PT+N}(p_i, d_j, d_j) $$
-$$ d_j = \frac{expit(\delta_j)}{2}$$
-$$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
-$$ p_i \sim Dirichlet(\overrightarrow{1})$$
-$$ \delta_j \sim Normal(0,1)$$
-$$ \Delta\delta_j \sim Normal(0,1)$$
-$$ k \sim HalfCauchy(20)$$
-
-The function $f_{PT+N}$ computes the expected probability estimate using the underlying subjective probability p, the noise parameters d and d’, and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
-Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters d and d’.
-
-Recall that Zhu and colleagues (2020) identified a bridging condition relating b and N in the Bayesian Sampler model to the d parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied d and d’, rather than directly according to its $\beta$, N, and N’ parameters.
-I constrain d to [0, $\frac{1}{3}$] for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
-This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
-
-The Bayesian Sampler model is therefore:
-
-$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{BS}(p_i, d_j, d_j’) $$
-$$ d_j = \frac{expit(\delta_j)}{2}$$
-$$ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}$$
-$$ p_i \sim Dirichlet(\overrightarrow{1})$$
-$$ \delta_j \sim Normal(0,1)$$
-$$ \Delta\delta_j \sim Normal(0,1)$$
-$$ k \sim HalfCauchy(20)$$
-
-Where the function $f_{BS}$ computes the expected probability estimate as prescribed by the Bayesian Sampler theory.
-
-### Hierarchical implementations of the models
-
-Both of these models can also be implemented as hierarchical models with partial pooling for the d and d’ parameters (implicitly, for N and N’ in the case of the Bayesian Sampler).
-This should prevent over-penalizing complex versions of the model, as well as potentially improving their overall predictive fit.
-The hierarchical implementation adds parameters for the population-level d and d’ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
-For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency (citation?).
-
-$$y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)$$
-$$\mu_{ij} = f_{BS}(p_i, d_j, d_j’) $$
-$$d_j = \frac{ expit(\delta_j )}{3}$$
-$$ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}$$
-$$ p_i \sim Dirichlet(\overrightarrow{1})$$
-$$ \delta_{pop} \sim Normal(-1, 1)$$
-$$ \Delta\delta_{pop} \sim Normal(0, .50)$$
-$$ log(\sigma_{\delta}) \sim Normal(-1, 1)$$
-$$ log(\sigma_{\Delta\delta}) \sim Normal(-1, 1)$$
-$$\delta_j \sim Normal(\delta_{pop}, \sigma_{\delta})$$
-$$\Delta\delta_j \sim Normal(\Delta\delta_{pop}, \sigma_{\Delta\delta})$$
-$$k \sim HalfCauchy(20)$$
-
-Finally, I also explored fitting versions of the Bayesian Sampler model that allowed values of $\beta > 1$.
-Restricting $\beta$ to [0,1] restricts the prior distribution of the Bayesian sampler to the class of “ignorance priors” (Zhu et al., 2020).
-However, it is also possible that people bring informative priors to the probability judgment task.
-Indeed, Zhu and colleagues (2020) acknowledge there are situations where an informative prior may be warranted (see e.g., Fennel & Baddeley, 2012).
-If $\beta$ is unrestricted, allowed to fall in the domain [0, $\infty$] then the Bayesian Sampler model becomes more flexible, allowing for equivalent “noise” levels in the same [0, .50] range as the PT+N model.
-Though it would seem a more fundamental change, this model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
-Thus, fitting this additional “hybrid” model allows for a complete comparison of the models along both of their differing dimensions.
-
-# Results
-
-I fit each of the models specified above to data from Zhu et al’s [-@zhu.etal2020] Experiment 1 and 2 and computed LOO\_ic scores for each combination.
-Model posteriors were estimated using the Numpyro implementation of the No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler (cite).
-For each model, four MCMC chains of 2000 iterations were sampled after 2000 iterations of warmup and all passed convergence tests according to $\hat{R}$.
-Figure X below displays the estimated differences in LOO\_ic scores for each of the models as compared to the best-scoring model.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=5in]{plot_compare.png}
-\end{figure}
-
-Data from Experiment 1 are only modestly informative with respect to discriminating among the models.
-However, data from Experiment 2 reveal a clearer winner: a hierarchical implementation of the “hybrid” Bayesian Sampler model that removes the restriction on $\beta$, allowing it to range from [0, $\infty$].
-
-Recall, this model may also be seen as a version of the PT+N theory that removes the original theory’s constructive account of conditional probability judgments.
-Comparing predictions for conditional probability judgments from the hybrid model and the best-fitting PT+N model, we see that the hybrid model better captures these judgments from Experiment 2 (trial-average-level $R^2$ = .85 vs .73; participant-level $R^2$ = .51 vs .45).
-These findings suggest that conditioning is better seen as part of the mental model than the probability judgment process.
-
-Figure X shows the posterior distributions of the population-level d and d’ parameters inferred from the “hybrid” model.
-Using data from experiment 2, population-level estimates of d and d’ are greater than $\frac{1}{3}$, outside the range implied by the assumption of “ignorance priors” in the Bayesian Sampler model.
-Parameters fit to the data from Experiment 1 are more consistent with this assumption, although a substantial proportion of of individual participants’ d and d’ estimates also lie outside this range (13 of 59 for d, 26 of 59 for d’).
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=4in]{plot_params.png}
-\end{figure}
-
-Among the remaining models, hierarchical implementations with partial pooling outperform the unpooled corresponding models—suggesting that unpool variants of the complex models were over-penalized in the original analysis by @zhu.etal2020.
-In both experiments, the simple version of the PT+N model is clearly outperformed by the other competing models.
-However, differences among the remaining three models are approximately within the error of the LOO\_ic estimates.
-As expected, the simple variants of the models have smaller penalty terms than the complex variants.
-
-Compared to the hybrid model, the PT+N model with its constructive account of conditional probability judgments has a smaller penalty term despite having the same parameterization in terms of d and d’.
-Evidently this structural component isn’t a complexifying epicycle, instead it is a prediction that constrains flexibility.
-However, it is one that ultimately leads to a worse-fitting model.
-
-Perhaps surprisingly, the Bayesian Sampler with unrestricted beta actually receives a smaller penalty term than the more restricted version of the model.
-This is at first unintuitive, but it illustrates how model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
-Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is estimated from data [-@gelman.etal2014].
-If the estimated value is some very large positive number, then the constraint won’t have been very informative.
-But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model’s penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied d and d’ estimated for this data under this model are both very near $\frac{1}{3}$, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
-
-The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria.
-Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives [e.g.
-@gronau.wagenmakers2019; @vehtari.etal2019].
-Fortunately, conclusions from the comparisons here do not rest solely on differences between the models’ complexity.
-
-```{r table2, echo=FALSE, message=F, warning=F}
-t2_df <- read_csv("model-comparison-table.csv")
-
-t2_df %>% 
-  select(-X1) %>% 
-  gather(measure, value, -Experiment, -model) %>% 
-  pivot_wider(names_from = c(Experiment, measure), values_from=value,names_sort = T) %>% 
-  rename(Model = model) %>%
-  kbl(
-    caption="Bayesian model comparison results", 
-    digits=3, 
-    booktabs=T,
-    col.names = c("Model", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{trial}$", "$LOO_{ic}$", "Penalty","$r_{resp}$", "$r_{trial}$"),
-    escape = F
-    ) %>%
-  add_header_above(c(" ", "Experiment 1"=4, "Experiment 2"=4)) %>% 
-  kable_classic(full_width=F)
-
-```
-
-Finally, it is worth noting that these models provide quite strong overall fits to the data, not just for the query averages, but also for the trial averages across individual participants as seen from the correlations between predicted and observed responses in Table X.
-
-# Discussion
-
-By a fair margin, the model best accounting for the experimental data from Zhu and colleagues [-@zhu.etal2020] was a version of the Bayesian Sampler model without restriction on the range of its Beta parameters.
-Alternatively, this “hybrid” model can also be seen as a variant of the PT+N model that removes its account of conditional probability judgments.
-Thus, what these findings indicate most clearly is that the Bayesian Sampler theory provides a superior account of conditional probability judgments in this experimental task.
-In keeping with the larger theoretical framework of Bayesian cognitive science, this theory assumes that subjective probabilities underlie people’s probability judgments, and that conditional probability judgments are produced by Bayesian conditioning occurring in their mental models of the events in question, rather than as arising from the probability judgment process [@chater.etal2020; @zhu.etal2020].
-
-Findings from  Zhu and colleagues’ Experiment 2 do cast doubt on their proposal that the priors of the Bayesian Sampler model should reflect “ignorance priors,” symmetric beta distributions with $\beta$ < 1.
-As a generic prior that would be used across contexts, this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people bring informative rather than “ignorance” priors to the probability judgment task, indicated by estimated d and d’ parameters outside the [1, $\frac{1}{3}$] range implied by the bridging conditions when beta is restricted < 1.
-One possibility is that people bring domain-specific priors to many judgment tasks, meaning that the most appropriate priors might be dictated by the context in which they make their judgments.
-If so, other experimental contexts might see parameter estimates consistent with an uninformative prior.
-In addition, the noise parameters also varied across individuals.
-Exploring the participant-level parameters reveals that some individuals’ implied d and d’ parameters were consistent with the class of ignorance priors (see supplemental materials).
-Further research should explore this heterogeneity within and across individuals.
-Unfortunately, pinning down specific components of the Bayesian Sampler model is a challenge, as the $\beta$ and N parameters are not uniquely identifiable from judgment data of the sort examined here.
-
-One thing this model comparison has not decided, and likely *cannot* decide, is whether the distortions of probability judgments are products of mental noise or of further reasoning processes.
-Given their tight connections via the bridging conditions [@zhu.etal2020], it may not be possible to draw decisive conclusions here.
-Moreover, the theories may not be in any real competition over this point: Zhu and colleagues consider that “noise” might give an algorithmic-level solution to the computation-level task defined by the Bayesian Sampler [-@zhu.etal2020].
-
-Some of the most interesting implications of these models go well beyond the probability judgment task itself: the models both support a probabilistic account of beliefs [@chater.etal2020].
-Indeed, by representing the true subjective probabilities as a latent variable, the Bayesian modeling approach employed here allows those underlying subjective probabilities to be inferred.
-Examining the model posteriors reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
-Of course, Zhu and colleagues’ (2020) experiments were never designed for this purpose.
-Future research could explore how these estimates might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g.
-@jern.etal2014; @griffiths.tenenbaum2006; @franke.etal2016].
-One particularly promising direction could be to integrate these models with formal models of belief revision, which could shed new light on fundamental these cognitive processes [e.g.
-@jern.etal2014; @powell.etal2018; @cook.lewandowsky2016].

### 2022-09-16 paper-rmd/paper.tex
created revised manuscript

@@
 \usepackage{microtype}      % microtypography
-\usepackage{lipsum}
 \usepackage{graphicx}
@@
 
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 % Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
 \newlength{\csllabelwidth}
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
 % for Pandoc 2.8 to 2.10.1
@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
   % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
   % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
+  \setlength{\parskip}{#2\cslentryspacingunit}
  }%
  {}
-\usepackage{calc} % for calculating minipage widths
+\usepackage{calc}
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \usepackage{xcolor}
-
 \begin{document}
 \maketitle
-
-\def\tightlist{}
 
@@
 
-Meanwhile, Zhu, Sanborn, \& Chater (2020) have proposed a Bayesian model
-of probability judgment they call the ``Bayesian Sampler.'' Under this
-model, probability judgment is itself seen as a process of Bayesian
-inference. To judge the probability of an event, a limited number of
-samples are again drawn from a mental model of the event. Then, those
-``observed'' samples are integrated with a prior over probabilities to
-produce a probability judgment. The effect of this prior is to
-regularize the judgments, pushing them away from extreme values. They
-argue this regularization of judgments is adaptive, in that it pushes
-judgments and decisions away from extremes that might, e.g.~encourage
-taking large risks based on a small number of mental samples. However,
-this hedging comes at the cost of systematic incoherence and biases.
-Like the PT+N model, their proposed model accounts for a wide array of
-biases in probability judgments, including the novel biases identified
-by Costello and Watts (Costello and Watts 2014, 2016).
-
-Under the simplest version of the Bayesian Sampler model, the expected
-value of a probability judgment is given by the following formula:
-
-\[E[\hat{P}_{BS}(A)] = \frac{N}{N+2\beta}P(A) + \frac{\beta}{N+2\beta}\]
-
-where \emph{N} represents the number of mental samples drawn. The
-mentally-observed samples are integrated with a Bayesian prior over
-probabilities, represented with a symmetric beta distribution
-parameterized by \(\beta\). Zhu and colleagues (2020) show that the
-\emph{N} and \(\beta\) parameters of their model can be related to the
-\emph{d} parameter of the PT+N model via the following bridging formula:
-
-\[d = \frac{\beta}{N+2\beta}\]
-
-\hypertarget{two-accounts-of-conditional-probability-judgments}{%
-\subsubsection{Two accounts of conditional probability
-judgments}\label{two-accounts-of-conditional-probability-judgments}}
-
-By explaining the incoherence of human probability judgments using
-coherent mental probabilities, both models have the potential to rescue
-the larger project of Bayesian cognitive science as applied to everyday
-beliefs (Chater et al. 2020). However, because they diverge in their
-treatment of conditional probability judgments, the two theories fit
-very differently into the larger Bayesian project. According to the
-Bayesian sampler model, conditioning is something that happens in the
-mental model of the events, not as part of the process of rendering
-probability judgments. Thus, the Bayesian sampler model does not assign
-any special status to conditional probability judgments and so fits
-neatly into the larger Bayesian project of cognitive science:
-probability judgments are simply another judgment process applied to the
-outputs of other Bayesian mental models (Chater et al. 2020).
-
-In contrast, the PT+N model presents a constructive account of
-conditional probability judgments that is fundamentally non-Bayesian
-(Costello and Watts 2016). According to the PT+N model, conditional
-probabilities P(A\textbar B) are estimated by a two-stage sampling
-procedure: first both events A and B are sampled with noise, and then a
-second noisy process computes the ratio of the events read as A and B
-over events read as B. The PT+N model predicts conditional probability
-estimates using the following equation:
-
-\[ P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d} \]\\
-Given that conditional inference and belief updating lies at the core of
-Bayesian cognitive theory, PT+N's account of conditional probability
-judgment separates it from the Bayesian sampler theory not just
-quantitatively, but quite fundamentally. First, it denies any central
-importance of conditional probability judgments by pulling conditioning
-out of the mental model of an event and placing it into the probability
-judgment process. And perhaps more importantly, this non-bayesian
-account of conditional inference can no longer be considered
-``rational'' at any stage under the PT+N model.
-
-\hypertarget{comparing-the-models}{%
-\subsection{Comparing the models}\label{comparing-the-models}}
-
-Zhu, Sanborn, and Chater (2020) compared their Bayesian Sampler model
-against Costello \& Watts' (2014; 2016; 2017; 2018) PT+N model as
-explanations for human probability judgments in two experiments.
-Unfortunately, their results were somewhat equivocal. They fit both
-``simple'' and ``complex'' versions of each model and computed a
-Bayesian Information Criteria score (BIC) for all models. The complex
-versions of these models introduce additional parameters d' and N' that
-allow for different patterns of judgments for conjunctive and
-disjunctive judgments as compared with simple probability judgments.
-These additional parameters are needed in order to explain conjunction
-fallacies (Costello \& Watts, 2017; Zhu et al., 2020). Table 1 below
-presents the total BIC scores computed for each model as originally fit,
-using the authors' original code and saved model outputs (Zhu et al,
-2020, Supplementary materials).
-
-\begin{table}
-
-\caption{\label{tab:table1}Original model fitting results}
-\centering
-\begin{tabular}[t]{lrr}
-\toprule
-Model & Exp. 1 & Exp. 2\\
-\midrule
-Bayesian Sampler simple & 956.9 & -5371.9\\
-Bayesian Sampler complex & 1174.4 & -5099.3\\
-PT+N simple & 1257.5 & -4901.1\\
-PT+N complex & 1039.7 & -5159.6\\
-Bayesian Sampler avg. & 1065.6 & -5235.6\\
-\addlinespace
-PT+N avg. & 1148.6 & -5030.4\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-In both experiments, the simple version of the Bayesian Sampler scores
-best, but the complex version of the PT+N model comes in second (lower
-BIC scores are better). It is not obvious what conclusions should be
-drawn from these results. Of course, the simple Bayesian Sampler model
-appears to win by the numbers. But conjunction fallacies are an
-extremely robust empirical finding (e.g.~cite) and are clearly present
-in the data collected by Zhu and colleagues (2020). So we might rule out
-the simple variants of the models on the grounds that they will fail to
-capture important qualitative features of the data, which would instead
-favor the PT+N theory. For their part, Zhu and colleagues (2020) elect
-to split the baby by averaging the simple and complex model scores
-together. This approach somewhat favors the Bayesian Sampler theory
-overall, though they are rightfully cautious about drawing any strong
-conclusions.
-
-\hypertarget{accounting-for-model-complexity}{%
-\subsubsection{Accounting for Model
-Complexity}\label{accounting-for-model-complexity}}
-
-Zhu and colleagues (2020) warn that BIC, which penalizes models based
-only on the number of parameters in the model, cannot fully account for
-the differences in the models' complexity. There are at least three
-challenges to accounting for model complexity in the comparison of these
-models. First, the models differ not only in the number of parameters
-but in the domain of those parameters. Zhu and colleagues (2020)
-restrict the range of the \(\beta\) parameter in their Bayesian Sampler
-model to fall within the interval {[}0,1{]}. This restriction constrains
-the Bayesian Sampler model's prior distribution to the class of
-``ignorance priors,'' symmetric beta distributions where \(\beta < 1\).
-Via the bridging conditions (eq X), they show that this restricts the
-``noise level'' for the Bayesian Sampler, represented in implied d under
-the PT+N model, to {[}0, \(\frac{1}{3}\){]} whereas the PT+N model
-permits noise values between {[}0, 0.50{]}.
-
-Second, it is not immediately clear how the structural differences
-between the models impact their flexibility. For the purposes of model
-comparison, the ``complexity penalty'' that models truly ought to pay is
-determined by their relative flexibility in accommodating different
-patterns of data, not simply the number or domain of their parameters
-(cite Gelman). To be sure, Zhu and colleagues' (2020) bridging condition
-makes it clear that the PT+N model is more flexible when it comes to
-predicting unconditional probabilities. But, what impact does the PT+N
-model's treatment of conditional probabilities have on model complexity?
-Is this component of the model a sort of Ptolemaic epicycle, adding
-complexity to the model that should be penalized? Or, does it constitute
-a commitment to novel predictions that thereby constrain its
-flexibility? Determining model complexity a priori is not always
-straightforward when the models being compared differ structurally.
-
-Finally, one potential explanation for the relative weakness of the
-complex model versions is Zhu and colleagues' (2020) use of unpooled
-models, where parameters are estimated independently for each individual
-participant. Adding a parameter for each participant may effectively
-over-penalize the ``complex'' variants relative to the simple variants,
-especially if there is actually limited heterogeneity across
-participants. In contrast, a comparison of fully pooled variants of the
-simple and complex models (where a single population parameter is used
-for all participants) would require adding only one extra parameter to
-the penalty term. Partial pooling is a solution that balances between
-these extreme approaches, allowing for an accounting of heterogeneity
-without over-penalizing in cases where heterogeneity is low.
-
-\hypertarget{the-present-work}{%
-\subsubsection{The present work}\label{the-present-work}}
-
-Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian
-data analysis framework that permits a more decisive (or at least
-alternative) comparison. A Bayesian analysis allows issues of model
-complexity to be addressed through comparisons of model fit based on
-information criteria, such as WAIC and LOO\textsubscript{ic}. These
-indices allow for complexity penalties that directly capture differences
-in the flexibility of the models, rather than being estimated based
-solely on the number of distinct parameters {[}cite gelman{]}. In
-addition, this framework also supports straightforward implementation of
-hierarchical versions of these models with partial pooling. This allows
-for information about model parameters to be shared across participants,
-resulting in an overall reduction in model complexity and a more
-realistic test of the models.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{data-selection}{%
-\subsection{Data selection}\label{data-selection}}
-
-Zhu, Sanborn, \& Chater (2020) conducted two experiments to compare the
-PT+N and Bayesian sampler theories. These experiments asked participants
-to judge the probability of different events in various combinations.
-Following prior work by Costello and Watts (e.g. 2016, 2018), both
-experiments focused on the everyday events of different kinds of
-weather.
-
-Experiment 1 asked about the events \{icy, frosty\} and \{normal,
-typical\} (e.g.~``what is the probability that the weather in London is
-normal and not typical?''). The authors' goal was to ask about highly
-correlated events, but the events used are perhaps nearly perfectly
-correlated. Because the terms used to describe these events are nearly
-synonymous, there is a concern about the interpretation of the
-statements evaluated in this experiment. This is especially clear, as
-the authors note, for disjunctive trials such as ``normal or typical,''
-where ``or typical'' might not be read as a disjunction but rather an
-elaborative clause. In light of these concerns, I excluded the
-disjunctive trials from Experiment 1 from my analyses. Experiment 2
-focused on more moderately correlated events, \{cold, rainy\} and
-\{windy, cloudy\}, that do not admit these misinterpretations. In
-addition, a third experimental condition asking about \{warm, snowy\}
-was also included in the experiment, but was dropped from the analyses
-reported in the paper. Exploring the raw responses from this condition
-reveals a substantial fraction of ``zero'' and ``one'' responses for
-certain trials. This may reflect a different response process than was
-intended. For instance, some participants may have engaged in deductive
-reasoning to judge that it is not possible for the weather to at once be
-warm and snowy, and therefore responded with zero---failing to properly
-consider that it is possible (at least logically) for it to be warm and
-snowy at different times within the same day. Given these potentially
-aberrant responses, I followed Zhu and colleagues (2020) in ignoring
-data from this condition.
-
-\hypertarget{modeling}{%
-\subsection{Modeling}\label{modeling}}
-
-I implement both the Bayesian Sampler and PT+N models as Bayesian models
-in the probabilistic programming language Numpyro. All code and results
-are available as supplemental materials
-(\url{https://github.com/derekpowell/bayesian-sampler}).
-
-\hypertarget{bayesian-implementations-of-the-models}{%
-\subsubsection{Bayesian implementations of the
-models}\label{bayesian-implementations-of-the-models}}
-
-The PT+N model defines probability judgments as
-
-\[P_{e}(A) = (1-2d)P(A) + d\]
-\[P_e(A\land B) = (1-d^\prime)P(A \land B)+d^\prime\]
-\[P_e(A\lor B) = (1-d^\prime)P(A \lor B)+d^\prime\]
-\[P_e(A|B) = \frac{(1-2d)^2P(A \land B) + d(1-2d)\big(P(A)+P(B)\big)+d^2}{(1-2d)P(B)+d}\]
-
-In contrast, the Bayesian Sampler model defines probability judgments as
-
-\[P_{e}(A) = \frac{N}{N + 2 \beta}P(A) + \frac{\beta}{N+2 \beta}\]
-\[P_{e}(A \land B) = \frac{N’}{N’ + 2 \beta}P(A \land B) + \frac{\beta}{N’+2 \beta}\]
-\[P_{e}(A \lor B) = \frac{N’}{N’ + 2 \beta}P(A \lor B) + \frac{\beta}{N’+2 \beta}\]
-
-Fixing d and d' or N and N' equal yields the ``simple'' variant of each
-of the models, which treat conjunctive and disjunctive probability
-judgments identically to simple probability judgments.
-
-Notice that for each model the probability judgments depend on
-underlying subjective probabilities, derived from a mental sampling
-process. These subjective probabilities are unobserved, and must be
-estimated as a latent variable. Here, they are estimated as a
-four-dimensional dirichlet distribution for each subject, representing
-the probability of the elementary events (A\&B, \textasciitilde A\&B,
-A\&\textasciitilde B, \textasciitilde A\&\textasciitilde B).
-
-Zhu, Sanborn \& Chater (2020) implement completely unpooled models with
-separate d, d', N, N', and \(\beta\) parameters for each participant.
-Translated into a Bayesian Framework, we can write the PT+N model as:
-
-\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{PT+N}(p_i, d_j, d_j) \]
-\[ d_j = \frac{expit(\delta_j)}{2}\]
-\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
-\[ p_i \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_j \sim Normal(0,1)\] \[ \Delta\delta_j \sim Normal(0,1)\]
-\[ k \sim HalfCauchy(20)\]
-
-The function \(f_{PT+N}\) computes the expected probability estimate
-using the underlying subjective probability p, the noise parameters d
-and d', and the relevant equation as defined by the PT+N theory (see
-supplemental materials for implementation details). Prior predictive
-checks were conducted for all models to select priors that would be
-uninformative or minimally informative on the scale of the model
-parameters d and d'.
-
-Recall that Zhu and colleagues (2020) identified a bridging condition
-relating b and N in the Bayesian Sampler model to the d parameter of the
-PT+N model. To support direct comparisons of the models, I parameterize
-the Bayesian Sampler model according to the implied d and d', rather
-than directly according to its \(\beta\), N, and N' parameters. I
-constrain d to {[}0, \(\frac{1}{3}\){]} for the Bayesian Sampler model
-to reflect the assumption that \(\beta \in [0, 1]\). This allows the
-same priors to be used for the corresponding Bayesian Sampler and PT+N
-models, simplifying their comparison.
-
-The Bayesian Sampler model is therefore:
-
-\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{BS}(p_i, d_j, d_j’) \]
-\[ d_j = \frac{expit(\delta_j)}{2}\]
-\[ d_j’ = \frac{expit\big(\delta_j + \exp(\Delta\delta_j)\big)}{2}\]
-\[ p_i \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_j \sim Normal(0,1)\] \[ \Delta\delta_j \sim Normal(0,1)\]
-\[ k \sim HalfCauchy(20)\]
-
-Where the function \(f_{BS}\) computes the expected probability estimate
-as prescribed by the Bayesian Sampler theory.
-
-\hypertarget{hierarchical-implementations-of-the-models}{%
-\subsubsection{Hierarchical implementations of the
-models}\label{hierarchical-implementations-of-the-models}}
-
-Both of these models can also be implemented as hierarchical models with
-partial pooling for the d and d' parameters (implicitly, for N and N' in
-the case of the Bayesian Sampler). This should prevent over-penalizing
-complex versions of the model, as well as potentially improving their
-overall predictive fit. The hierarchical implementation adds parameters
-for the population-level d and d' as well as a parameter controlling the
-standard deviation of the distribution for the subject-level effects.
-For ease of interpretation, the centered parameterization is shown
-below, although the actual models used a non-centered parameterization
-to improve sampling efficiency (citation?).
-
-\[y_{ij} \sim beta\big(\mu_{ij} k, (1-\mu_{ij})k\big)\]
-\[\mu_{ij} = f_{BS}(p_i, d_j, d_j’) \]
-\[d_j = \frac{ expit(\delta_j )}{3}\]
-\[ d_j’ = \frac{ expit\big( \delta_j  + \exp(\Delta\delta_j )\big) }{3}\]
-\[ p_i \sim Dirichlet(\overrightarrow{1})\]
-\[ \delta_{pop} \sim Normal(-1, 1)\]
-\[ \Delta\delta_{pop} \sim Normal(0, .50)\]
-\[ log(\sigma_{\delta}) \sim Normal(-1, 1)\]
-\[ log(\sigma_{\Delta\delta}) \sim Normal(-1, 1)\]
-\[\delta_j \sim Normal(\delta_{pop}, \sigma_{\delta})\]
-\[\Delta\delta_j \sim Normal(\Delta\delta_{pop}, \sigma_{\Delta\delta})\]
-\[k \sim HalfCauchy(20)\]
-
-Finally, I also explored fitting versions of the Bayesian Sampler model
-that allowed values of \(\beta > 1\). Restricting \(\beta\) to {[}0,1{]}
-restricts the prior distribution of the Bayesian sampler to the class of
-``ignorance priors'' (Zhu et al., 2020). However, it is also possible
-that people bring informative priors to the probability judgment task.
-Indeed, Zhu and colleagues (2020) acknowledge there are situations where
-an informative prior may be warranted (see e.g., Fennel \& Baddeley,
-2012). If \(\beta\) is unrestricted, allowed to fall in the domain {[}0,
-\(\infty\){]} then the Bayesian Sampler model becomes more flexible,
-allowing for equivalent ``noise'' levels in the same {[}0, .50{]} range
-as the PT+N model. Though it would seem a more fundamental change, this
-model may also be seen as a version of the PT+N theory that jettisons
-its constructive account of conditional probability judgment. Thus,
-fitting this additional ``hybrid'' model allows for a complete
-comparison of the models along both of their differing dimensions.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-I fit each of the models specified above to data from Zhu et al's (2020)
-Experiment 1 and 2 and computed LOO\_ic scores for each combination.
-Model posteriors were estimated using the Numpyro implementation of the
-No-U-Turn Hamiltonian Markov chain Monte Carlo (MCMC) sampler (cite).
-For each model, four MCMC chains of 2000 iterations were sampled after
-2000 iterations of warmup and all passed convergence tests according to
-\(\hat{R}\). Figure X below displays the estimated differences in
-LOO\_ic scores for each of the models as compared to the best-scoring
-model.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=5in]{plot_compare.png}
-\end{figure}
-
-Data from Experiment 1 are only modestly informative with respect to
-discriminating among the models. However, data from Experiment 2 reveal
-a clearer winner: a hierarchical implementation of the ``hybrid''
-Bayesian Sampler model that removes the restriction on \(\beta\),
-allowing it to range from {[}0, \(\infty\){]}.
-
-Recall, this model may also be seen as a version of the PT+N theory that
-removes the original theory's constructive account of conditional
-probability judgments. Comparing predictions for conditional probability
-judgments from the hybrid model and the best-fitting PT+N model, we see
-that the hybrid model better captures these judgments from Experiment 2
-(trial-average-level \(R^2\) = .85 vs .73; participant-level \(R^2\) =
-.51 vs .45). These findings suggest that conditioning is better seen as
-part of the mental model than the probability judgment process.
-
-Figure X shows the posterior distributions of the population-level d and
-d' parameters inferred from the ``hybrid'' model. Using data from
-experiment 2, population-level estimates of d and d' are greater than
-\(\frac{1}{3}\), outside the range implied by the assumption of
-``ignorance priors'' in the Bayesian Sampler model. Parameters fit to
-the data from Experiment 1 are more consistent with this assumption,
-although a substantial proportion of of individual participants' d and
-d' estimates also lie outside this range (13 of 59 for d, 26 of 59 for
-d').
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=4in]{plot_params.png}
-\end{figure}
-
-Among the remaining models, hierarchical implementations with partial
-pooling outperform the unpooled corresponding models---suggesting that
-unpool variants of the complex models were over-penalized in the
-original analysis by Zhu, Sanborn, and Chater (2020). In both
-experiments, the simple version of the PT+N model is clearly
-outperformed by the other competing models. However, differences among
-the remaining three models are approximately within the error of the
-LOO\_ic estimates. As expected, the simple variants of the models have
-smaller penalty terms than the complex variants.
-
-Compared to the hybrid model, the PT+N model with its constructive
-account of conditional probability judgments has a smaller penalty term
-despite having the same parameterization in terms of d and d'. Evidently
-this structural component isn't a complexifying epicycle, instead it is
-a prediction that constrains flexibility. However, it is one that
-ultimately leads to a worse-fitting model.
-
-Perhaps surprisingly, the Bayesian Sampler with unrestricted beta
-actually receives a smaller penalty term than the more restricted
-version of the model. This is at first unintuitive, but it illustrates
-how model complexity depends not only on the model and priors, but also
-the observed data (see Gelman, Hwang, and Vehtari 2014). Gelman and
-colleagues consider a case where a parameter is constrained to be
-positive and its value is estimated from data (2014). If the estimated
-value is some very large positive number, then the constraint won't have
-been very informative. But, if the estimated value is very close to
-zero, then the constraint that the parameter is positive will provide
-substantial information and the model's penalty term will therefore be
-smaller. Here, it seems reasonable to conjecture that because the
-implied d and d' estimated for this data under this model are both very
-near \(\frac{1}{3}\), the restriction results in posterior estimates of
-the linear parameters that are relatively far from the prior, which can
-result in a greater penalty.
-
-The dependence of complexity penalties on observed data may strike some
-as an undesirable feature of model comparison through information
-criteria. Indeed, it is worth acknowledging that principled model
-comparison is still an area of active inquiry, with differing
-perspectives (e.g. Gronau and Wagenmakers 2019; Vehtari et al. 2019).
-Fortunately, conclusions from the comparisons here do not rest solely on
-differences between the models' complexity.
-
-\begin{table}
-
-\caption{\label{tab:table2}Bayesian model comparison results}
-\centering
-\begin{tabular}[t]{lrrrrrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Experiment 1} & \multicolumn{4}{c}{Experiment 2} \\
-\cmidrule(l{3pt}r{3pt}){2-5} \cmidrule(l{3pt}r{3pt}){6-9}
-Model & $LOO_{ic}$ & Penalty & $r_{resp}$ & $r_{trial}$ & $LOO_{ic}$ & Penalty & $r_{resp}$ & $r_{trial}$\\
-\midrule
-Unrestricted BS MLM & -1178.1 & 149.2 & 0.725 & 0.821 & -3953.1 & 368.7 & 0.688 & 0.878\\
-Bayesian Sampler complex MLM & -1167.7 & 145.6 & 0.723 & 0.816 & -3815.4 & 399.9 & 0.675 & 0.852\\
-PT+N complex & -1163.5 & 133.6 & 0.722 & 0.821 & -3768.5 & 398.2 & 0.667 & 0.840\\
-Bayesian Sampler complex & -1159.5 & 149.3 & 0.723 & 0.815 & -3733.2 & 453.6 & 0.679 & 0.848\\
-PT+N complex MLM & -1157.0 & 147.1 & 0.722 & 0.820 & -3796.7 & 356.1 & 0.658 & 0.835\\
-\addlinespace
-Bayesian Sampler simple MLM & -1149.5 & 141.0 & 0.719 & 0.810 & -3794.7 & 381.0 & 0.667 & 0.839\\
-Bayesian Sampler simple & -1138.1 & 146.8 & 0.718 & 0.806 & -3722.8 & 418.9 & 0.667 & 0.831\\
-PT+N simple & -1051.8 & 132.6 & 0.695 & 0.793 & -3643.0 & 319.5 & 0.619 & 0.777\\
-PT+N simple MLM & -1044.7 & 137.9 & 0.695 & 0.793 & -3643.8 & 305.3 & 0.617 & 0.772\\
-Relative Freq. & -746.6 & 146.8 & 0.664 & 0.737 & -1287.7 & 424.6 & 0.516 & 0.639\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Finally, it is worth noting that these models provide quite strong
-overall fits to the data, not just for the query averages, but also for
-the trial averages across individual participants as seen from the
-correlations between predicted and observed responses in Table X.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-By a fair margin, the model best accounting for the experimental data
-from Zhu and colleagues (2020) was a version of the Bayesian Sampler
-model without restriction on the range of its Beta parameters.
-Alternatively, this ``hybrid'' model can also be seen as a variant of
-the PT+N model that removes its account of conditional probability
-judgments. Thus, what these findings indicate most clearly is that the
-Bayesian Sampler theory provides a superior account of conditional
-probability judgments in this experimental task. In keeping with the
-larger theoretical framework of Bayesian cognitive science, this theory
-assumes that subjective probabilities underlie people's probability
-judgments, and that conditional probability judgments are produced by
-Bayesian conditioning occurring in their mental models of the events in
-question, rather than as arising from the probability judgment process
-(Chater et al. 2020; Zhu, Sanborn, and Chater 2020).
-
-Findings from Zhu and colleagues' Experiment 2 do cast doubt on their
-proposal that the priors of the Bayesian Sampler model should reflect
-``ignorance priors,'' symmetric beta distributions with \(\beta\)
-\textless{} 1. As a generic prior that would be used across contexts,
-this class of uninformative priors has an appealing rational basis.
-Nevertheless, the data suggest that people bring informative rather than
-``ignorance'' priors to the probability judgment task, indicated by
-estimated d and d' parameters outside the {[}1, \(\frac{1}{3}\){]} range
-implied by the bridging conditions when beta is restricted \textless{}
-1. One possibility is that people bring domain-specific priors to many
-judgment tasks, meaning that the most appropriate priors might be
-dictated by the context in which they make their judgments. If so, other
-experimental contexts might see parameter estimates consistent with an
-uninformative prior. In addition, the noise parameters also varied
-across individuals. Exploring the participant-level parameters reveals
-that some individuals' implied d and d' parameters were consistent with
-the class of ignorance priors (see supplemental materials). Further
-research should explore this heterogeneity within and across
-individuals. Unfortunately, pinning down specific components of the
-Bayesian Sampler model is a challenge, as the \(\beta\) and N parameters
-are not uniquely identifiable from judgment data of the sort examined
-here.
-
-One thing this model comparison has not decided, and likely
-\emph{cannot} decide, is whether the distortions of probability
-judgments are products of mental noise or of further reasoning
-processes. Given their tight connections via the bridging conditions
-(Zhu, Sanborn, and Chater 2020), it may not be possible to draw decisive
-conclusions here. Moreover, the theories may not be in any real
-competition over this point: Zhu and colleagues consider that ``noise''
-might give an algorithmic-level solution to the computation-level task
-defined by the Bayesian Sampler (2020).
-
-Some of the most interesting implications of these models go well beyond
-the probability judgment task itself: the models both support a
-probabilistic account of beliefs (Chater et al. 2020). Indeed, by
-representing the true subjective probabilities as a latent variable, the
-Bayesian modeling approach employed here allows those underlying
-subjective probabilities to be inferred. Examining the model posteriors
-reveals these estimates often come with considerable uncertainty, but at
-least for some participants they can be estimated with useful levels of
-precision. Of course, Zhu and colleagues' (2020) experiments were never
-designed for this purpose. Future research could explore how these
-estimates might be made more reliable, and how inferences about these
-mental probabilities might be integrated with other Bayesian models of
-reasoning (e.g. Jern, Chang, and Kemp 2014; Griffiths and Tenenbaum
-2006; Franke et al. 2016). One particularly promising direction could be
-to integrate these models with formal models of belief revision, which
-could shed new light on fundamental these cognitive processes (e.g.
-Jern, Chang, and Kemp 2014; Powell, Weisman, and Markman 2018; Cook and
-Lewandowsky 2016).
-
 \hypertarget{refs}{}
@@
 \url{https://doi.org/10.1177/0963721420954801}.
-
-\leavevmode\hypertarget{ref-cook.lewandowsky2016}{}%
-Cook, John, and Stephan Lewandowsky. 2016. {``Rational {Irrationality}:
-{Modeling Climate Change Belief Polarization Using Bayesian
-Networks}.''} \emph{Topics in Cognitive Science} 8 (1): 160--79.
-\url{https://doi.org/10.1111/tops.12186}.
 
@@
 
-\leavevmode\hypertarget{ref-franke.etal2016}{}%
-Franke, Michael, Fabian Dablander, Anthea Scholler, Erin Bennett, Judith
-Degen, Michael Henry Tessler, Justine Kao, and Noah D Goodman. 2016.
-{``What Does the Crowd Believe? {A} Hierarchical Approach to Estimating
-Subjective Beliefs from Empirical Data,''} 6.
-
-\leavevmode\hypertarget{ref-gelman.etal2014}{}%
-Gelman, Andrew, Jessica Hwang, and Aki Vehtari. 2014. {``Understanding
-Predictive Information Criteria for {Bayesian} Models.''}
-\emph{Statistics and Computing} 24 (6): 997--1016.
-\url{https://doi.org/10.1007/s11222-013-9416-2}.
-
-\leavevmode\hypertarget{ref-griffiths.tenenbaum2006}{}%
-Griffiths, Thomas L., and Joshua B. Tenenbaum. 2006. {``Optimal
-{Predictions} in {Everyday Cognition}.''} \emph{Psychological Science}
-17 (9): 767--73. \url{https://doi.org/10.1111/j.1467-9280.2006.01780.x}.
-
-\leavevmode\hypertarget{ref-gronau.wagenmakers2019}{}%
-Gronau, Quentin F., and Eric-Jan Wagenmakers. 2019. {``Limitations of
-{Bayesian Leave}-{One}-{Out Cross}-{Validation} for {Model
-Selection}.''} \emph{Computational Brain \& Behavior} 2 (1): 1--11.
-\url{https://doi.org/10.1007/s42113-018-0011-7}.
-
-\leavevmode\hypertarget{ref-jern.etal2014}{}%
-Jern, Alan, Kai-min K. Chang, and Charles Kemp. 2014. {``Belief
-Polarization Is Not Always Irrational.''} \emph{Psychological Review}
-121 (2): 206--24. \url{https://doi.org/10.1037/a0035941}.
-
-\leavevmode\hypertarget{ref-powell.etal2018}{}%
-Powell, Derek, Kara Weisman, and Ellen M Markman. 2018. {``Articulating
-Lay Theories Through Graphical Models: {A} Study of Beliefs Surrounding
-Vaccination Decisions,''} 6.
-
-\leavevmode\hypertarget{ref-vehtari.etal2019}{}%
-Vehtari, Aki, Daniel P. Simpson, Yuling Yao, and Andrew Gelman. 2019.
-{``Limitations of {`{Limitations} of {Bayesian Leave}-One-Out
-{Cross}-{Validation} for {Model Selection}'}.''} \emph{Computational
-Brain \& Behavior} 2 (1): 22--27.
-\url{https://doi.org/10.1007/s42113-018-0020-6}.
-
 \leavevmode\hypertarget{ref-zhu.etal2020}{}%

### 2022-09-16 paper-rmd/ptn_platediagram.tex
paper edits
created revised manuscript

@@
        \tikzstyle{plateIJK} = [align=center];
-       \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$ijk$} ;
+       \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$i$} ;
     }
@@
        \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
-       \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$jk$};
+       \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$k$};
      }
@@
 \end{tikzpicture}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \flushleft
+    \small
+    \begin{align*}
+      y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
+      \mu_{ijk} &= f_{PT+N}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
+      d_j &= \frac{1}{2} \ \text{logistic}(\delta_j) \\
+      d_j’ &= \frac{1}{2} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
+      \overrightarrow{p_{jk}} &\sim Dirichlet(\overrightarrow{1}) \\
+      \delta_j &\sim Normal(0,1) \\
+      \Delta\delta_j &\sim Normal(0,1) \\
+      \kappa &\sim \mathit{Half \hyph Cauchy}(20)
+    \end{align*}
+  \end{minipage}
+  \caption{Complex unpooled PT+N model diagram and formula specifications.
+Circular nodes are parameters, shaded nodes are observations, and squared nodes are deterministic functions of parameters.
+Plates signify values defined for $i$ trials, $j$ participants, and $k$ conditions.}
+\end{figure}

### 2022-09-16 paper-rmd/bs_platediagram_mlm.tex
full revision draft

@@
       \node[det, above=of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
+      \node[obs, right=.5cm of mu]  (x) {$x_{ijk}$};
       \node[latent, above=of d]  (delta) {$\delta_j$};
@@
       \node[latent, right=2cm of y]  (kappa) {$\kappa$};
-      \node[latent, left=1cm of mu]            (p) {$\vec{p_{jk}}$};
+      \node[latent, left=1cm of mu]            (p) {$\vec{\theta_{jk}}$};
       \node[latent, above=of delta, xshift=-.5cm]  (delta_pop) {$\overline{\delta}$};
@@
       \edge {mu, kappa} {y} ; %
-      \edge {d, d_prime} {mu}
+      \edge {d, d_prime, x} {mu}
       \edge {delta} {d, d_prime}
@@
         \tikzstyle{plateIJK} = [align=center];
-        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$i$} ;
+        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)(x)} {$i$} ;
      }
@@
         \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
-        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)} {$j$} ;
+        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)(x)} {$j$} ;
       }
@@
         \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
-        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$k$};
+        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)(x)} {$k$};
       }
@@
   y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
-  \mu_{ijk} &= f_{BS}(\overrightarrow{p_{jk}}, d_j, d_j')  \\
+  \mu_{ijk} &= f_{BS}(\overrightarrow{\theta_{jk}}, x_{ijk}, d_j, d_j')  \\
   d_j &= \frac{1}{3} \ \text{logistic}(\delta_j) \\

### 2022-09-16 paper-rmd/paper-papaja-revision.Rmd
full revision draft

@@
 These theories fit quite differently into the larger project of Bayesian cognitive science, but their many similarities complicate comparisons of their predictive accuracy.
-In particular, comparing the models demands a careful accounting of model complexity.
+In particular, the models diverge most clearly in their distributional predictions, requiring a careful treatment of errors and external response processes.
 Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
-Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by a modified version of the Bayesian Sampler model under which people may hold informative priors about probabilities.
+Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by an account of biases based on "noise" in the sample-reading process.
 
@@
 
-\begin{align}
+\begin{align*}
   P(\text{read as A}) &= (1-d)P(A) + d(1-P(A)) \\
   &= (1-2d)P(A) + d \tag{1}
-\end{align}
+\end{align*}
 
@@
 
-By assumption, a maximum of 50% of samples can be misread, so $d$ is a number in the range $[0, 1/2]$.
+By assumption, a maximum of 50% of samples can be misread on average, so $d$ is a number in the range $[0, 1/2]$.
+The overall consequence of the sample-reading noise will be to shrink probability estimates toward .50 in proportion to $d$.
 The PT+N theory provides a unified account for a wide variety of biases in probability judgment that were previously attributed to different types of heuristics, as well as novel biases identified based on the model's predictions [@costello.watts2014; @costello.watts2016; @costello.watts2017; @costello.watts2018].
+For example, the PT+N theory offers an explanation for many instances of "conservatism" [@costello.watts2014]---people's tendency to shy away from extreme probability judgments near 0 and 1, even when strong evidence warrants such judgments [e.g.
+@edwards.1968; @erev.etal1994].
 
@@
 
-In later work, Sundh et al [-@sundh.etal2022] examined the distributional properties of participants responses using indirect means, by regressing the variance of participants' responses across trials on their mean response for each query type.
-Their findings suggest that earlier fits with Wasserstein distance may have produced biased results [@sundh.etal2022].
+In later work, Sundh et al [-@sundh.etal2021] examined the distributional properties of participants responses using indirect means, by regressing the variance of participants' responses across trials on their mean response for each query type.
+Their findings suggest that earlier fits with Wasserstein distance may have produced biased results [@sundh.etal2021].
 They reported evidence for the truncation of responses and a correlation between variance and shrinkage parameter estimates across participants.
@@
 
-Figure 4 shows the posterior distributions of the population-level $d$ and $d'$ parameters inferred from the unrestricted Bayesian Sampler model.
-In Experiment 2, population-level estimates of $d'$ are greater than $1/3$, as are a substantial number of participant-level estimates for $d$ (37 of 83), as shown in Figure X.
+Figure 4 (top) shows the posterior distributions of the population-level $d$ and $d'$ parameters inferred from the unrestricted Bayesian Sampler model.
+In Experiment 2, population-level estimates of $d'$ are greater than $1/3$, as are a substantial number of participant-level estimates for $d$ (37 of 83), as shown in Figure 5.
 These values fall outside the range implied by the assumption of "ignorance priors" in the Bayesian Sampler model.
@@
 So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
-Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 2).
+Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 1).
 This is true for both query-level averages as well as for individual participants' trial responses.
@@
 
-Table 2 also presents estimates of $\hat{p}_{\text{LOO}}$, the estimate of the effective number of parameters for each model.
+Table 1 also presents estimates of $\hat{p}_{\text{LOO}}$, the estimate of the effective number of parameters for each model.
 Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of $d$  and $d'$.
@@
   kbl(
-    caption="Bayesian model comparison results with best scoring model in bold face.", 
+    caption="Bayesian model comparison results with best scoring model in bold face.
+Models are compared based on X and the correlation between their predictions for each participants' query-level average ($r_{\\text{query}}$) as well as overall query-level averages $r_{\\text{query-avg}}$.",
     digits=3, 
     booktabs=T,
-    col.names = c("Model", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$","$r_{resp}$", "$r_{query}$", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$","$r_{resp}$", "$r_{query}$"),
+    col.names = c("Model", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$","$r_{\\text{query}}$", "$r_{\\text{query-avg}}$", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$","$r_{\\text{query}}$", "$r_{\\text{query-avg}}$"),
     escape = F
@@
 
-Zhu and colleageus [-@zhu.etal2020] demonstrated that the Bayesian Sampler model can capture a set of probabilistic identities developed by Costello, Watts, and colleagues [@costello.watts2016; @costello.etal2018] that capture some of the incoherence in people's probability judgments.
+Zhu and colleageus [-@zhu.etal2020] demonstrated that the Bayesian Sampler model can capture a set of probabilistic identities developed by Costello and Watts [@costello.watts2016; @costello.watts2018] that capture some of the incoherence in people's probability judgments.
 Following the design of the present experiments, these identities involve combinations of probability estimates for different combinations of two events $A$ and $B$ that should all be equal to zero according to probability theory.
 Under the Bayesian Sampler and PT+N theories, however, some of these identities should be zero, but some are allowed to take on other values.
-Figure X shows the average prediction of the winning model against the average observed value for each equality.
+Figure 6 shows the average prediction of the winning model against the average observed value for each equality.
 Consistent with prior findings, this model captures these identities quite closely.
@@
 Finally, it is worth noting that the best of these models provide quite strong overall fits to the data, not just for the query averages, but also for the query averages across individual participants as seen from the correlations between predicted and observed responses in Table 2.
-Figure X shows the correlation between participants' responses across all trials and the best-performing model's predictions.
+Figure 7 shows the correlation between participants' responses across all trials and the best-performing model's predictions.
 
@@
 First, it bears noting that the truncated response distribution implied by the Bayesian Sampler model appears at odds with the raw response data.
-However, participants' responses frequently lie outside the range implied by the best estimates of their $d$ parameters: XX% in Experiment 1 and XX% in Experiment 2.
+However, participants' responses frequently lie outside the range implied by the best estimates of their $d$ parameters (41% in Experiment 1 and 60% in Experiment 2).
 
@@
 What distribution of mental samples would such a process form?
-Figure X shows the result of a simulation study examining probability estimates drawing on autocorrelated samples with estimated weights.
+Figure 8 shows the result of a simulation study examining probability estimates drawing on autocorrelated samples with estimated weights.
 Each simulation consists of a simple discrete Markov chain generating autocorrelated samples at $x_t$, returning $x_{t-1}$ with probability $p_{a}$, or else returning a new sample distributed Bernoulli($p$) was simulated.
@@
 
-< Figure of simulation study >
+```{r, warning=F, message=F, fig.cap="Simulated proportions estimated from autocorrelated binary Markov Chain with single random initialization (left) or with initializations at 1 and 0 (right)."}
+source("../autocorr-sim.R")
+
+set.seed(12345)
+
+ps <- c(.25, .5, .9)
+Ns <- c(5, 10, 25)
+
+dsim <- expand_grid(
+  p = ps,
+  N = Ns
+) %>% 
+  mutate(
+    samps = map2(p, N, ~rBinEssInit(1000, .x, .y, .8)),
+    approx = map2(p, N, ~rbeta(1000, 1+ .x*.y, 1+(1-.x)*.y)),
+    )
+
+plt_init <- dsim %>% 
+  unnest(samps, approx) %>% 
+  gather(var, val, samps, approx) %>% 
+  mutate(var = if_else(var=="samps", "True", "Approximate")) %>% 
+  ggplot(aes(x=val, fill=var)) +
+  geom_histogram(position="identity", alpha=.6, binwidth=.025) +
+  facet_grid(p~N, scales="free_y") +
+  theme_bw() +
+  theme(panel.grid = element_blank()) +
+  labs(title = "Initialized autocorrelated sampling")
+
+dsim1 <- expand_grid(
+  p = ps,
+  N = Ns
+) %>% 
+  mutate(
+    samps = map2(p, N, ~rBinEss(1000, .x, .y)),
+    approx = map2(p, N, ~rbeta(1000, .x*.y, (1-.x)*.y)),
+    )
+
+plt_raw <- dsim1 %>% 
+  unnest(samps, approx) %>% 
+  gather(var, val, samps, approx) %>% 
+  mutate(var = if_else(var=="samps", "True", "Approximate")) %>% 
+  ggplot(aes(x=val, fill=var)) +
+  geom_histogram(position="identity", alpha=.6, binwidth=.025) +
+  # scale_fill_viridis_d() +
+  # facet_wrap(N + p~., scales="free") +
+  facet_grid(p~N, scales="free_y") +
+  theme_bw() +
+  theme(panel.grid=element_blank()) +
+  labs(title = "Raw autocorrelated sampling")
+
+(plt_raw + plt_init) + plot_layout(guides="collect") & theme(legend.position = 'bottom') & labs(fill = "Samples", x="p", y="Count") & theme(axis.text.x = element_text(angle=90, vjust=.5))
+```
 
@@
 In this spirit, one possible explanation for the relative absence of extreme responses is that people draw samples following two initializations, initializing sampling with both a positive and negative instance.
-As the simulations in Figure X show, this produces a predicted response distribution that is  approximately $Beta(1 + p N_{\text{eff}}, 1+(1-p)N_{\text{eff}})$.
+As the simulations in Figure 8 (bottom) show, this produces a predicted response distribution that is  approximately $Beta(1 + p N_{\text{eff}}, 1+(1-p)N_{\text{eff}})$.
 
@@
 Compared to the query-averaged model, the trial-level noise-based model adds two features: mixture components for rounding and contaminants and subject-level varying $N$ in place of a fixed $K$ parameter.
-This model's implementation and the implementation of its mixture components is depicted in Figure X.
+This model's implementation and the implementation of its mixture components is depicted in Figure 9.
 However, note that $N$ is allowed to vary independently from $d$, allowing for independence between response shrinkage and variability.
@@
 Then it is plain that $\hat{P}_{BS}(A)$ is a transformation of $\rho(A)$, and therefore a transformed Beta distribution (see Appendix for derivation).
-Figure X below diagrams the entire Bayesian Sampler model, now parameterized in terms of $\beta$, $N$, and $N'$.
+Figure 10 diagrams the entire Bayesian Sampler model, now parameterized in terms of $\beta$, $N$, and $N'$.
 
@@
 Prior to fitting the models, response data were rounded to the nearest 5%.
-Nearly all responses (XX%) were already divisible by five, and this was necessary to speed model fitting.
+Nearly all responses (Exp.
+1: 93% and Exp.
+2: 89%) were already divisible by five, and this was necessary to speed model fitting.
 Even still, estimating model posteriors for the trial-level mixture models using MCMC proved intractable.
@@
 
-Table X shows the scores of each model fit to the trial-level data in Experiments 1 and 2.
+Table 2 shows the scores of each model fit to the trial-level data in Experiments 1 and 2.
 From the quantitative model comparison it is clear that the noise-based model is superior, with substantialy lower $\widehat{\text{elpd}}_{\text{LOO}}$ than both of the competing Bayesian Sampler implementations in both experiments.>
 
-< table X: model comparison results >
+```{r, echo=FALSE, message=F, warning=F}
+texp1 <- read_csv("exp1-trial-level-comparison.csv") %>% mutate(Experiment = c("Exp. 1", "", ""))
+texp2 <- read_csv("exp1-trial-level-comparison.csv") %>% mutate(Experiment = c("Exp. 2", "", ""))
+
+t_trial_comp <- bind_rows(texp1, texp2)
+t_trial_comp %>% 
+  rename(Model = `...1`) %>% 
+  mutate(Model = case_when(
+    Model=="ptn" ~ "Noise-based",
+    Model=="bs" ~ "Bayesian Sampler",
+    TRUE ~ "BS: Varying N"
+  )) %>% 
+  select(Experiment, Model, loo, p_loo, d_loo, dse) %>%
+
+#   gather(measure, value, -Experiment, -model) %>% 
+#   pivot_wider(names_from = c(Experiment, measure), values_from=value,names_sort = T) %>% 
+  kbl(
+    caption="Bayesian model comparison results for trial-level models with best scoring model in bold face.",
+    digits=1,
+    booktabs=T,
+    col.names = c("Experiment", "Model", "$\\widehat{\\text{elpd}}_{\\text{LOO}}$", "$\\hat{p}_{\\text{LOO}}$", "$\\widehat{\\Delta\\text{elpd}}_{\\text{LOO}}$", "$\\text{SE}_{\\Delta \\text{LOO}}$"),
+    escape = F
+    ) %>% 
+  kable_classic(full_width=F) %>%
+  kable_styling(latex_options="scale_down",font_size = 10) %>%
+  row_spec(c(1,4), bold=TRUE)
+```
 
@@
 
-The non-varying Bayesian Sampler model fits quite poorly and is estimated to have a large proportion of "contaminant" responses (X% in experiment 1, X% in experiment 2).
+The non-varying Bayesian Sampler model fits quite poorly and is estimated to have a large proportion of "contaminant" responses (34% in Experiment 1, 25% in Experiment 2).
 This is likely due to a lack of the predicted truncation of the response distribution.
@@
 
-Allowing the effective number of samples drawn to vary trial-by-trial improves the fit of the Bayesian Sampler model substantially and decreases the estimated proportion of contaminant responses (X% in experiment 1, X% in experiment 2).
-Nevertheless, these results still indicate inferior fit compared with the noise-based model, which has substantially lower $\widehat{\text{elpd}}_{\text{LOO}}$ and again attributes fewer responses to the contaminant process (X% in experiment 1, X% in experiment 2).
+Allowing the effective number of samples drawn to vary trial-by-trial improves the fit of the Bayesian Sampler model substantially and decreases the estimated proportion of contaminant responses (28% in Experiment 1, 12% in Experiment 2).
+Nevertheless, these results still indicate inferior fit compared with the noise-based model, which has substantially lower $\widehat{\text{elpd}}_{\text{LOO}}$ and again attributes fewer responses to the contaminant process (14% in Experiment 1, 4% in Experiment 2).
 
@@
 But if these quantities indeed are actually correlated as predicted by a Bayesian adjustment model, then we should expect to nevertheless see a correlation between the subject-level estimates in these parameters.
-Figure X shows a scatterplot of these estimates.
-Looking across subjects we see they are mostly uncorrelated, against the second BS prediction.
->
+Figure 11 shows a scatterplot relating these estimates.
+Looking across subjects we see they are mostly uncorrelated.
+Although there is a slight negative correlation in Experiment 1, this is driven by only a handful of participants with extreme values and in Experiment 2 the pattern appears to if anything be reversed.
+These findings again run counter to the predictions of the Bayesian Sampler theory.
 
@@
 Error bars indicate 95% CI.
-Although there is a slight negative correlation in Experiment 1, this is driven by only a handful of participants with extreme values and in Experiment 2 the pattern appears to if anything be reversed."}
+"}
 
@@
 This is likely because the hypothesized initialization process accounts for some of the shrinkage in participants' responses, so that less need be attributed to sample read noise.
-However, the numbers estimated in Experiment 2 are still substantially larger than in most prior work [cite].
 
@@
 This model bounds $d$ at .50 in principle, but a sample-reading process with such high error-rates may or may not be plausible.
-In prior work simulations have often assumed values of $d$ around .10 (Howe & costello, 2020; Costello & Watts, 2017)].
+In prior work simulations have often assumed values of $d$ around .10 [e.g.
+@howe.costello2020; @costello.watts2017].
 Further research examining what factors might affect the mental sampling and reading processes (e.g.

### 2022-09-16 paper-rmd/ptn_platediagram.tex
full revision draft

@@
      \node[det, above=of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
+     \node[obs, right=.75cm of mu]  (x) {$x_{ijk}$};
      \node[latent, above=of d]  (delta) {$\delta_j$};
@@
      \node[latent, right=2cm of y]  (kappa) {$\kappa$};
-     \node[latent, left=1cm of mu]            (p) {$\vec{p_{jk}}$};
+     \node[latent, left=.75cm of mu]            (p) {$\vec{\theta_{jk}}$};
 
@@
      \edge {mu, kappa} {y} ; %
-     \edge {d, d_prime} {mu}
+     \edge {d, d_prime, x, p} {mu}
      \edge {delta} {d, d_prime}
      \edge {delta_delta} {d_prime}
-     \edge {p} {mu}
 
@@
        \tikzstyle{plateIJK} = [align=center];
-       \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)} {$i$} ;
+       \plate[inner sep=0.125cm] {plateIJK} {(mu)(y)(x)} {$i$} ;
     }
@@
        \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
-       \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)} {$j$} ;
+       \plate[inner sep=0.125cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)(x)} {$j$} ;
      }
@@
        \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
-       \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)} {$k$};
+       \plate[inner sep=0.125cm] {plateIJ} {(p)(mu)(y)(x)} {$k$};
      }
@@
       y_{ijk} &\sim Beta\big(\mu_{ijk} \kappa, (1-\mu_{ijk}) \kappa \big) \\
-      \mu_{ijk} &= f_{PT+N}(\overrightarrow{p_{jk}}, d_j, d_j)  \\
+      \mu_{ijk} &= f_{PT+N}(\overrightarrow{\theta_{jk}}, x_{ijk}, d_j, d_j)  \\
       d_j &= \frac{1}{2} \ \text{logistic}(\delta_j) \\

### 2022-09-16 autocorr-sim.R

@@ -1,0 +1,93 @@
+library(tidyverse)
+
+Rcpp::cppFunction('NumericVector rcpp_clip( NumericVector x, double a, double b){
+    return clamp( a, x, b ) ;
+}')
+sample_corr_binary <- function(N, p, autocor_p){
+  samps <- rbernoulli(1, p)
+
+  for (i in 2:N){
+      samps <- c(samps, ifelse(rbernoulli(1, autocor_p), samps[length(samps)], rbernoulli(1,p)))
+  }
+  return(as.numeric(samps))
+}
+
+sample_corr_binary_2init <- function(N, p, autocor_p){
+  samps0 <- 0
+  samps1 <- 1
+
+  for (i in 1:N){
+    if (rbernoulli(1, .5)){
+      samps0 <- c(samps0, ifelse(rbernoulli(1, autocor_p), samps0[length(samps0)], rbernoulli(1,p)))
+    } else {
+      samps1 <- c(samps1, ifelse(rbernoulli(1, autocor_p), samps1[length(samps1)], rbernoulli(1,p)))
+    }
+
+    samps <- c(samps0[1:length(samps0)], samps1[1:length(samps1)])
+
+  }
+  return(samps)
+}
+
+est_weight <- function(p, p_auto, n=100){
+  x <- sample_corr_binary(n, p, p_auto)
+  weight <- rcpp_clip(coda::effectiveSize(x)[[1]]/n,.005,1)
+
+  return(weight)
+}
+
+
+rapprox <- function(n, p, N_eff){
+  # create a mixture distribution ?
+  x <- sample_corr_binary(100, p, .7)
+  weight <- rcpp_clip(coda::effectiveSize(x)[[1]]/100,.005,1)
+
+  # N <- round(N_eff/weight)
+  p1 <- p^N_eff
+  p0 <- (1-p)^N_eff
+
+  ones <- rep(1, round(p1*n))
+  zeroes <- rep(0, round(p0*n))
+
+  output <- rbeta(n-round(p1*n)-round(p0*n), p*N_eff, (1-p)*N_eff)
+
+  return(c(ones, zeroes, output))
+}
+
+sample_autocorrelated_ess <- function(p, N_eff){
+  ## Draw autocorrelated samples to achieve effective sample size
+  x <- sample_corr_binary(100, p, .7)
+  weight <- rcpp_clip(coda::effectiveSize(x)[[1]]/100,.005,1)
[…]

### 2022-09-17 create-paper-figures.Rmd
supplement revisions

@@
   plt_ineq_exp2 + labs(title="Exp. 2") + 
-  plot_layout(guides="collect") & theme(legend.position="bottom") & labs(color="", y = "Estimate", x = "Identity")
+  plot_layout(guides="collect") & 
+  theme(legend.position="bottom", axis.text.x = element_text(angle=90, hjust=1, vjust=.5)) & 
+  labs(color="", y = "Estimate", x = "Identity")
 ```
@@
     # geom_smooth(method="lm") +
-    labs(y="d", x="k", title="Exp. 1") +
+    labs(y="d", x="N", title="Exp. 1") +
     theme_bw() +
@@
     # geom_smooth(method="lm") +
-    labs(y="d", x="k", title="Exp. 2") +
+    labs(y="d", x="N", title="Exp. 2") +
     theme_bw() +
@@
     # geom_smooth(method="lm") +
-    labs(y="d'", x="k") +
+    labs(y="d'", x="N") +
     theme_bw() +

### 2022-09-17 paper-rmd/paper-papaja-revision.Rmd
supplement revisions

@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "????"
+wordcount         : "9195"
 
@@
 For example, the PT+N theory offers an explanation for many instances of "conservatism" [@costello.watts2014]---people's tendency to shy away from extreme probability judgments near 0 and 1, even when strong evidence warrants such judgments [e.g.
-@edwards.1968; @erev.etal1994].
+@edwards1968; @erev.etal1994].
 
@@
 
-These models make subtly different predictions on two levels: First,  the models have distinct accounts of conditional probability judgments that make different predictions in terms of expected values.
+The model's predictions can be distinguished on two levels: First, the models have distinct accounts of conditional probability judgments that make different predictions in terms of expected values.
 Second, the models present different process-level accounts of probability judgment that entail different predictions about the shape of the distribution of responses across individuals.
@@
 According to the PT+N model, conditional probabilities $P(A|B)$ are estimated by a two-stage sampling procedure: first both events $A$ and $B$ are sampled with noise, and then a second noisy process computes the ratio of the events read as $A$ and $B$ over events read as $B$.
-The probability can be written as:
+Schematically, the estimated probability can be written as:
 
@@
 
-Outside of conditional probability judgments, the primary difference between the models is the psychological process thought to produce bias in probability judgments: sample reading noise in the PT+N model and Bayesian inference in the Bayesian Sampler model.
-These process-level differences imply different response distributions.
+Although the model's predictions for unconditional probability judgments are identical in expectation (as seen via the bridging condition), the models posit different psychological processes underlying those judgments: sample reading noise in the PT+N model and Bayesian inference in the Bayesian Sampler model.
+These process-level differences imply different predictions about the distributions of people's judgments.
 
+The models make qualitatively different distributional predictions on two fronts.
+First, the Bayesian Sampler predicts a clear relationship between the degree to which responses are shrunk toward .50 and the trial-by-trial variability in those responses.
 In both models, the amount of variability in trial-level responses is related to the number of mental samples drawn, $N$.
@@
 In contrast, in the PT+N model the variance across responses and degree of shrinkage are reasonably considered to be independent.
-Moreover, the Bayesian Sampler model predicts not only shrinkage of probability estimates toward .50, but also a truncation of the response distribution in proportion to $\beta$ and $N$ [@chater.etal2020; @sundh.etal2021].
-Thus, modeling the distributions of raw responses holds promise for disentangling the models.
+Second, because the Bayesian Sampler describes a process of adjustment after the sampling process, in which people report the mean of their mental posterior over probabilities, the model also predicts a truncation of the response distribution in proportion to $\beta$ and $N$ [@chater.etal2020; @sundh.etal2021].
+That is, even when zero positive or negative samples are drawn, the mean of the posterior is drawn away from extreme responses of zero and one.
 
+Modeling the distributions of raw responses holds clear promise for disentangling the models.
 However, there are at least three challenges to directly modeling raw human response data.
@@
 Finally, Zhu and colleagues [-@zhu.etal2020] decision to fit "unpooled" estimates of all participants individually may exaggerate the complexity of both models.
-This is especially an issue in differentiating between simple and complex variants, which Zhu and colleauges [-@zhu.etal2020] elect to ignore by averaging the model scores.
+This is especially likely to pose an issue in differentiating between simple and complex variants, which Zhu and colleauges [-@zhu.etal2020] elect to ignore by averaging the model scores.
 If there is limited heterogeneity across individuals, then adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants.
@@
 However, whereas the size of that set is the same for each model, in the Bayesian Sampler the continuous $\beta$ parameter can shift exactly what those discrete values are, providing it much greater flexibility.
-However, this complexity goes unpunished in comparisons based on Wasserstein distance.
+Yet, this additional model flexibility goes unpunished in comparisons based on Wasserstein distance.
 
@@
 These subjective probabilities are unobserved, and must be estimated as a latent variable.
-Here, they are represented with a four-dimensional dirichlet distribution for each subject, representing the probability of the elementary events ($A \land B$, $\neg A \land B$, $A \land \neg B$, $\neg A \land \neg B$).
+Here, they are represented with a four-dimensional dirichlet distribution for each subject, $\vec{\theta}$, representing the probability of the elementary events ($A \land B$, $\neg A \land B$, $A \land \neg B$, $\neg A \land \neg B$).
 
@@
 
-The function $f_{PT+N}$ computes the expected probability estimate using the underlying subjective probability $p$, the noise parameters $d$  and $d'$, and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
+The function $f_{PT+N}$ computes the expected probability estimate using the underlying subjective probability computed from $\vec{\theta}$ and the query, the noise parameters $d$  and $d'$, and the relevant equation as defined by the PT+N theory (see supplemental materials for implementation details).
 Prior predictive checks were conducted for all models to select priors that would be uninformative or minimally informative on the scale of the model parameters $d$  and $d'$.
@@
 
-[^3]: Uninformativeness was sought in order to reduce bias in the posterior parameter estimates.
+[^3]:Uninformativeness was sought in order to reduce bias in the posterior parameter estimates.
 It should be acknowledged that a uniform prior does not exactly correspond to what the authors of the PT+N theory would predict, as they have frequently assumed $d$ to be a fairly small value [e.g.
@@
 Recall that Zhu and colleagues [-@zhu.etal2020] identified a bridging condition relating $\beta$ and $N$ in the Bayesian Sampler model to the $d$ parameter of the PT+N model.
-To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and $N'$ parameters.^[Strictly speaking, under the original form of the Bayesian sampler model, $N$ and $N'$ are discrete parameters representing the number of distinct independent samples drawn.
+To support direct comparisons of the models, I parameterize the Bayesian Sampler model according to the implied $d$  and $d'$, rather than directly according to its $\beta$, $N$, and $N'$ parameters.[^4] I constrain $d$ to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
+This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
+
+[^4]:Strictly speaking, under the original form of the Bayesian sampler model, $N$ and $N'$ are discrete parameters representing the number of distinct independent samples drawn.
 Given a particular implied $d$, this could create constraints on the possible values of $d'$, assuming $\beta$ is held constant.
 However, Zhu and colleagues [-@zhu.etal2020] also consider the possibility that people draw non-independent mental samples, in which case $N$ and $N'$ would represent the _effective number of samples_, accounting for their autocorrelation.
-In this case, we could treat this effective number of samples as a continuous quantity, and therefore imagine there are no clear constraints on $d$ and $d'$ except the stipulation that $d \leq d'$.] I constrain $d$ to $[0, 1/3]$ for the Bayesian Sampler model to reflect the assumption that $\beta \in [0, 1]$.
-This allows the same priors to be used for the corresponding Bayesian Sampler and PT+N models, simplifying their comparison.
+In this case, we could treat this effective number of samples as a continuous quantity, and therefore imagine there are no clear constraints on $d$ and $d'$ except the stipulation that $d \leq d'$.
+These ideas will be developed further in the trial-level analyses.
 
@@
 \begin{align*}
-  \mu_{ijk} &= f_{BS}(\overrightarrow{\theta_{jk}}, d_j, d_j)  \\
+  \mu_{ijk} &= f_{BS}(\overrightarrow{\theta_{jk}}, x_{ijk}, d_j, d'_j)  \\
   d_j &= \frac{1}{3} \ \text{logistic}(\delta_j) \\
@@
 Recall, in contrast to a noise-based account, the Bayesian Sampler theory predicts both a truncation of the response distribution as well as a correlation between the degree of shrinkage in probability estimates and the variability of those estimates trial-to-trial (assuming $\beta$ is relatively small).
-Because the Bayesian Sampler applies a Bayesian adjustment after sampling, it predicts probability judgments will always lie between $d$ and $1 - d$, even when zero positive or negative mental samples are drawn (see Equation X and the bridging condition, Equation X).
+Because the Bayesian Sampler applies a Bayesian adjustment after sampling, it predicts probability judgments will always lie between $d$ and $1 - d$, even when zero positive or negative mental samples are drawn (see Equation 12 and the bridging condition, Equation 5).
 First, it bears noting that the truncated response distribution implied by the Bayesian Sampler model appears at odds with the raw response data.
@@
 In the following set of analyses I attempt to lay out a set of reasonble assumptions that permit participants' trial-by-trial responses to be modeled and used to compare the theories' predictions.
-To do so, I first extend the models so as to render them fully continuous in their latent space.
-I then marry them with a specific model of response errors.
+To do so, I first extend the models so as to render them fully continuous in their latent space and then marry them with a specific model of response errors.
 Thus the models compared in the following analyses are not identical to those originally proposed by Zhu et al [-@zhu.etal2020] and Costello and Watts [-@costello.watts2014].
@@
 
-Zhu and collauges [-@zhu.etal2020] introduce the possibility of an "autocorrelated" Bayesian Sampler model under which samples are assumed to be autocorrelated (ideas which were advanced further in [@zhu.etal2021]).
+Zhu and collauges [-@zhu.etal2020] introduce the possibility of an "autocorrelated" Bayesian Sampler model under which samples are assumed to be autocorrelated [ideas which were advanced further in [@zhu.etal2021].
 As autocorrelated samples provide less information than i.i.d samples, they should be weighted when computing probability estimates.
@@
 As the figure shows, for large-enough autocorrelation values ($p_a$) the resulting distribution becomes quite smooth, with the exception of spikes at zero and one for small to moderate values of $N_{\text{eff}}$.
-This occurs when all samples are zero or all one, because $\sum_i^N0 \cdot w_i = 0$ for any $w_i$ and $\sum_i^N 1 \cdot w_i = 1$ for any weights $\sum_i^N w_i = 1$.
+This occurs when all samples are zero or all one.
 
@@
 A mental sampling process must start somewhere, and these initialization point(s) may influence judgments.
-For example, features of the initializations of mental sampling processes have been proposed to anchoring [@lieder.etal2013] and unpacking effects [@dasgupta.etal2017; also see @sanborn.chater2016].
+For example, features of the initializations of mental sampling processes have been proposed to anchoring [@lieder.etal2012] and unpacking effects [@dasgupta.etal2017; also see @sanborn.chater2016].
 In this spirit, one possible explanation for the relative absence of extreme responses is that people draw samples following two initializations, initializing sampling with both a positive and negative instance.
@@
 
-Incorporating these assumptions provides a psychologically plausible mechanism to permit the use of a Beta distribution to approximate the latent sampling process, rendering both of the models' latent parameters continuous.
+Assuming such an autocorrelated and initialized sampling process provides a psychologically plausible mechanism to permit the use of a Beta distribution to approximate the latent sampling process, rendering both of the models' latent parameters continuous.
 Note that this process will also itself produce some shrinkage in the probability estimates, and this shrinkage will be somewhat sensitive to the number of samples drawn in both the Bayesian Sampler and PT+N models.
@@
 when an allowable response of .14 is rounded to the out-of-bounds value of .10).
-However, some participant response still cannot be accounted for by the model.
+However, some responses still cannot be accounted for by the model.
 Instead, these responses are treated as "contaminants" generated by a random response process.
@@
 However, note that $N$ is allowed to vary independently from $d$, allowing for independence between response shrinkage and variability.
-Moreover, the latent distribution used to calculate categorical cut points is an untransformed Beta distribution.
 
@@
 Table 2 shows the scores of each model fit to the trial-level data in Experiments 1 and 2.
-From the quantitative model comparison it is clear that the noise-based model is superior, with substantialy lower $\widehat{\text{elpd}}_{\text{LOO}}$ than both of the competing Bayesian Sampler implementations in both experiments.>
+From the quantitative model comparison it is clear that the noise-based model is superior, with substantialy lower $\widehat{\text{elpd}}_{\text{LOO}}$ than both of the competing Bayesian Sampler implementations in both experiments.
 
@@
 Second, we can examine the shrinkage-variance relationship.
-In the noise-based model, N and d were allowed to vary freely.
+In the noise-based model, $N$ and $d$ were allowed to vary freely.
 But if these quantities indeed are actually correlated as predicted by a Bayesian adjustment model, then we should expect to nevertheless see a correlation between the subject-level estimates in these parameters.
@@
 
-Comparing the d values inferred from the trial-level model we see they are similar to but generally smaller than those estimated in the trial-averaged model.
+Comparing the $d$ values inferred from the trial-level model we see they are similar to but generally smaller than those estimated in the trial-averaged model.
 This is likely because the hypothesized initialization process accounts for some of the shrinkage in participants' responses, so that less need be attributed to sample read noise.
@@
 
-Fit to the average of participants' responses over blocks, there is a single clear winner: a model without any special treatment of conditional probability (ala the Bayesian Sampler model) and allowing for an implied $d$ parameter $\in [0,.5]$.
+Fit to the average of participants' responses over blocks, there is a single clear winner among the competing models: a model without any special treatment of conditional probability (ala the Bayesian Sampler model) and allowing for an implied $d$ parameter $\in [0,.5]$.
 This model could be interpreted either as a variant of the Bayesian Sampler without restriction on its $\beta$ parameter, or as a variant of the PT+N model that removes its account of conditional probability judgments.
@@
 Fit to the raw data, this requires treating an unreasonably large proportion of responses as "contaminants".
-Second, the degree of shrinkage in participants' responses and the variability in those responses appear largely uncorrelated.
+Second, the degree of shrinkage in participants' responses and the variability in those responses are not correlated in the ways predicted by the Bayesian Sampler theory.
 
@@
 Even high implied $d$ values might be consistent with rational inference in cases where the number of effective mental samples is very low.
-For instance, a beta(2,2) prior is only modestly informative, but could produce $d = .40$ if $N=1$.
+For instance, a Beta(2,2) prior is only modestly informative, but could produce $d = .40$ if $N=1$.
 However, if more samples are drawn then high $d$ values would correspond to potentially inflexible and suboptimal priors (i.e.
@@
 
-Finally, one important direction for future elaborations of sampling-based theories are more rigorous theories of realistic mental sampling processes, including details of the initialization and autocorrelation among samples.
-Here, I proposed a relatively-highly autocorrelated sampling process initialized with a positive and negative example to account for qualitative features of participants' response distributions and to support a simple approximation.
+Finally, one important direction for future elaborations of sampling-based theories are more rigorous theories of realistic mental sampling processes, including details of their initialization, autocorrelation, and amortization [@gershman.goodman2016].
+Here, I proposed an autocorrelated sampling process initialized with a positive and negative example to account for qualitative features of participants' response distributions and to support a simple approximation.
 Future work might explore the details of these processes and incorporate them into models of probability judgments in more rigorous ways.

### 2022-09-17 paper-rmd/trial_level_mlm_bs.tex
cleaning out files

\begin{figure}
  \begin{minipage}{.5\textwidth}

\begin{tikzpicture}
      % Define nodes
      \node[det] (theta) {$\vec{p_{ijk}}$};
      \node[det, left=.75cm of theta] (mu) {${\mu_{ijk}}$};
      \node[obs, below=1cm of theta]  (y) {$y_{ijk}$};
      \node[obs, below=1cm of mu]  (x) {$x_{ijk}$};
      \node[det, above=of theta, xshift=-1cm]  (N_prime) {$N_j'$};
      \node[det, left=.75cm of N_prime]  (N) {$N_j$};
      \node[latent, above=of N]  (eta) {$\eta_j$};
      \node[latent, above=of N_prime, xshift=.5cm]  (delta_eta) {$\Delta\eta_j$};
      \node[latent, right=.75cm of N_prime]  (beta) {$\beta_j$};
      \node[latent, right=of beta, yshift=-.5cm]  (beta_pop) {$\overline{\beta}$};
      \node[latent, right=of beta, yshift=.5cm]  (beta_sd) {$\sigma_{\beta}$};
      \node[latent, below=1cm of mu, xshift=-1.1cm]            (p) {$\vec{\theta_{jk}}$};
      \node[latent, above=of eta, xshift=-.5cm]  (eta_pop) {$\overline{\eta}$};
      \node[latent, above=of delta_eta, xshift=-.5cm]  (delta_eta_pop) {$\overline{\Delta\eta}$};
      \node[latent, above=of eta, xshift=.5cm]  (sigma_eta) {$\sigma_{\eta}$};
      \node[latent, above=of delta_eta, xshift=.5cm]  (sigma_delta_eta) {$\sigma_{\Delta\eta}$};
      \node[latent, right=1.5cm of y]  (phi) {$\phi$};
      % Connect the nodes
      \edge {theta} {y} ; %
      \edge {x,p} {mu}
      \edge {N, N_prime, beta, mu, phi} {theta}
      \edge {eta} {N, N_prime}
      \edge {delta_eta} {N_prime}
      \edge {eta_pop} {eta}
      \edge {delta_eta_pop} {delta_eta}
      \edge {sigma_eta} {eta}
      \edge {sigma_delta_eta} {delta_eta}
      \edge {beta_pop, beta_sd} {beta}

      % Plates
      {
        \tikzset{plate caption/.append style={below=15pt of #1.south east}};
        \tikzstyle{plateIJK} = [align=center];
        \plate[inner sep=0.25cm, xshift=.2cm] {plateIJK} {(mu)(y)(x)} {$i$} ;
     }
      {
        \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
        \plate[inner sep=0.25cm, xshift=0cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(N)(N_prime)(eta)(delta_eta)(beta)} {$j$} ;
      }
      {
        \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
        \plate[inner sep=0.25cm, xshift=.125cm] {plateIJ} {(p)(mu)(y)(x)} {$k$};
      }

\end{tikzpicture}

  \end{minipage}%
  \begin{minipage}{.5\textwidth}
    \centering
    $
\begin{aligned}[c]
  y_{ij} &\sim Categorical(\theta) \\
  \overrightarrow{\theta_{ijk}} &= Z_{BS}(\mu_{ijk}, N_j, N'_j, \beta_j, \phi) \\
  \mu_{ijk} &= f_0(\overrightarrow{\theta_{jk}}, x_{ijk})  \\
  N_j' &= 1 + exp(\eta_j) \\
  N_j &= 1 + exp(\eta_j) + \exp(\Delta\eta_j) \\
  log(\beta_j) &\sim Normal(\overline{\beta}, \sigma_{\beta}) \\
  \eta_j &\sim Normal(\overline{\eta}, \sigma_{\eta}) \\ 
  \Delta\eta_j &\sim Normal(\overline{\Delta\eta}, \sigma_{\Delta\eta}) \\
  \overline{\eta'} &\sim Normal(1, 1) \\
  \overline{\Delta\eta} &\sim Normal(0, .1) \\
  \overline{\beta} &\sim Normal(-.5, .4) \\
  log(\sigma_{\eta}) &\sim Normal(-.5, .5) \\
  log(\sigma_{\Delta\eta}) &\sim Normal(-.5, -.5) \\
  log(\sigma_{\beta}) &\sim Normal(-1, .5) \\
  \overrightarrow{p_i} &\sim Dirichlet(\overrightarrow{1}) \\
  \overrightarrow{\phi} &\sim Dirichlet(\overrightarrow{1}) \\
\end{aligned}
$
  \end{minipage}
  \caption{Hierarchical complex trial-level Bayesian Sampler model diagram and formula specifications. $Z_{BS}$ and $f_{0}$ are functions that compute the probability of each categorical response and the expected proportion of mental samples given underlying mental probabilities before Bayesian adjustment. See Appendix for further descriptions of these details.}
\end{figure}

### 2022-09-17 paper-rmd/trial_level_mlm_noise_based.tex
cleaning out files

\begin{figure}
  \begin{minipage}{.5\textwidth}

\begin{tikzpicture}
      % Define nodes
      \node[obs]                               (y) {$y_{ijk}$};
      \node[det, above=2.5cm of y] (mu) {$\mu_{ijk}$};
      \node[det, above=2.5cm of mu, xshift=-1.2cm]  (d) {$d_j$};
      \node[det, above=2.5cm of mu, xshift=1.2cm]  (d_prime) {$d_j'$};
      \node[latent, above=of d]  (delta) {$\delta_j$};
      \node[latent, above=of d_prime]  (delta_delta) {$\Delta\delta_j$};
      \node[latent, left=1cm of mu]            (p) {$\vec{\theta}_{jk}$};
      \node[latent, above=1cm of y]  (theta) {$\vec{p}_{ijk}$};
      \node[latent, left=1cm of theta]  (kappa) {$\kappa_j$};
      \node[latent, above=of delta, xshift=-.5cm]  (delta_pop) {$\overline{\delta}$};
      \node[latent, above=of delta_delta, xshift=-.5cm]  (delta_delta_pop) {$\overline{\Delta\delta}$};
      \node[latent, above=of delta, xshift=.5cm]  (sigma_delta) {$\sigma_{\delta}$};
      \node[latent, above=of delta_delta, xshift=.5cm]  (sigma_delta_delta) {$\sigma_{\Delta\delta}$};
      \node[latent, right=2cm of theta]  (phi) {$\phi$};
      \node[obs, above= of mu]  (x) {$x_{ijk}$};
      % Connect the nodes
      \edge {theta} {y} ; %
      \edge {d, d_prime} {mu}
      \edge {delta} {d, d_prime}
      \edge {delta_delta} {d_prime}
      \edge {p, x} {mu}
      \edge {delta_pop} {delta}
      \edge {delta_delta_pop} {delta_delta}
      \edge {sigma_delta} {delta}
      \edge {sigma_delta_delta} {delta_delta}
      \edge {mu, phi, kappa} {theta}

      % Plates
      {
        \tikzset{plate caption/.append style={below=15pt of #1.south east}};
        \tikzstyle{plateIJK} = [align=center];
        \plate[inner sep=0.25cm, xshift=-.0625cm] {plateIJK} {(mu)(y)(theta)(x)} {$i$} ;
     }
      {
        \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)(kappa)} {$j$} ;
      }
      {
        \tikzset{plate caption/.append style={above right=0pt and 0pt of #1.north west}};
        \plate[inner sep=0.25cm] {plateIJ} {(p)(mu)(y)(x)} {$k$};
      }

\end{tikzpicture}

  \end{minipage}%
  \begin{minipage}{.5\textwidth}
    \centering
    $
\begin{aligned}[c]
  y_{ijk} &\sim Categorical(\overrightarrow{\theta_{ijk}}) \\
  \overrightarrow{p_{ijk}} &= Z_{NB}(\mu_{ijk},\kappa_j, \phi) \\
  \mu_{ijk} &= f_{NB}(\overrightarrow{\theta_{jk}}, d_j, d'_j, x_{ijk})  \\
  d_j &= \frac{1}{2} \ \text{logistic}(\delta_j) \\
  d_j’ &= \frac{1}{2} \ \text{logistic}\big(\delta_j + \exp(\Delta\delta_j)\big) \\
  \delta_j &\sim Normal(0,1) \\
  \Delta\delta_j &\sim Normal(0,1) \\
  log(\sigma_{\eta}) &\sim Normal(-.5, .5) \\
  log(\sigma_{\Delta\eta}) &\sim Normal(-.5, -.5) \\
  log(\sigma_{\beta}) &\sim Normal(-1, .5) \\
  \overrightarrow{\theta_{jk}} &\sim Dirichlet_4(\overrightarrow{1}) \\
  \overrightarrow{\phi} &\sim Dirichlet_3(\overrightarrow{1}) \\
\end{aligned}
$
  \end{minipage}
  \caption{Hierarchical complex trial-level noise-based model diagram and formula specifications. $Z_{NB}$ and $f_{NB}$ are functions that compute the probability of each categorical response and the expected proportion of read-out mental samples given underlying mental probabilities and sample reading noise. See Appendix for further descriptions of these details.}
\end{figure}

### 2022-09-17 fit-gpu-bs-discrete-exp2.py

@@ -1,235 +1,0 @@
-import numpyro
-numpyro.enable_x64()
-numpyro.set_platform('gpu')
-
-import jax
-print(jax.devices())
-
-import os
-os.environ['XLA_PYTHON_CLIENT_PREALLOCATE']='false'
-
-import numpyro.distributions as dist
-from numpyro.infer import MCMC, NUTS, Predictive, init_to_median, init_to_value, init_to_sample
-
-
-import tensorflow_probability.substrates.jax as tfp
-
-from jax import numpy as jnp
-from jax import random, vmap
-
-
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import arviz as az
-
-import siuba as s
-from siuba import _
-
-from dfply import *
-
-from plotnine import *
-
-import pickle
-
-## --- run local scripts
-from model_helpers import *
-from models import *
-
-## -- load data
-
-def load_data_exp2_trials():
-    df = load_raw_data(2)
-    
-    df["condition"] = np.select(
-    [
-        df.querydetail.str.contains("windy|cloudy"), 
-        df.querydetail.str.contains("cold|rainy"),
-        df.querydetail.str.contains("warm|snowy")
-
-    ], 
-    [
-        0,
-        1,
-        2
-    ], 
-    default=0 )
-    
-    original_ids = list(np.unique(df.ID))
-    fix_id_dict = {original_ids[i]:i for i in range(0, len(original_ids))}
-    
-    df = df.assign(ID = df.ID.apply(lambda x: fix_id_dict[x]))
-
-    return df
-
-df = load_data_exp2_trials() # see data_helpers.py
-df = df[df["condition"]!=2] # filter out "warm/snowy" as per paper
-# df = df.sort_values(by=["ID","block","condition"]) # don't think I need to sort?
-
-df = (df >> 
-      s.mutate(block = _.block-1) >> 
-#       s.filter(_.block==0, _.ID < 20) >>
-      # s.group_by(_.ID, _.condition, _.querytype, _.querydetail) >> 
-      # s.summarize(estimate = _.estimate.mean()) >>
-      s.mutate(estimate =  np.round(_.estimate*20))
-     )
-df.estimate = df.estimate.astype("int64")
-
[…]

### 2022-09-17 fit-gpu-ptn-discrete-exp2.py

@@ -1,216 +1,0 @@
-import numpyro
-numpyro.enable_x64()
-numpyro.set_platform('gpu')
-
-import jax
-print(jax.devices())
-
-import os
-os.environ['XLA_PYTHON_CLIENT_PREALLOCATE']='false'
-
-import numpyro.distributions as dist
-from numpyro.infer import MCMC, NUTS, Predictive, init_to_median, init_to_value, init_to_sample
-
-
-import tensorflow_probability.substrates.jax as tfp
-
-from jax import numpy as jnp
-from jax import random, vmap
-
-
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import arviz as az
-
-import siuba as s
-from siuba import _
-
-from dfply import *
-
-from plotnine import *
-
-import pickle
-
-## --- run local scripts
-from model_helpers import *
-from models import *
-
-## -- load data
-
-def load_data_exp2_trials():
-    df = load_raw_data(2)
-    
-    df["condition"] = np.select(
-    [
-        df.querydetail.str.contains("windy|cloudy"), 
-        df.querydetail.str.contains("cold|rainy"),
-        df.querydetail.str.contains("warm|snowy")
-
-    ], 
-    [
-        0,
-        1,
-        2
-    ], 
-    default=0 )
-    
-    original_ids = list(np.unique(df.ID))
-    fix_id_dict = {original_ids[i]:i for i in range(0, len(original_ids))}
-    
-    df = df.assign(ID = df.ID.apply(lambda x: fix_id_dict[x]))
-
-    return df
-
-df = load_data_exp2_trials() # see data_helpers.py
-df = df[df["condition"]!=2] # filter out "warm/snowy" as per paper
-# df = df.sort_values(by=["ID","block","condition"]) # don't think I need to sort?
-
-df = (df >> 
-      s.mutate(block = _.block-1) >> 
-#       s.filter(_.block==0, _.ID < 20) >>
-      # s.group_by(_.ID, _.condition, _.querytype, _.querydetail) >> 
-      # s.summarize(estimate = _.estimate.mean()) >>
-      s.mutate(estimate =  np.round(_.estimate*20))
-     )
-df.estimate = df.estimate.astype("int64")
-
-X_data, y_data = make_model_data(df) # see data_helpers.py
[…]

### 2022-09-17 fit-gpu-ptn-discrete-simdata.py

@@ -1,233 +1,0 @@
-import numpyro
-numpyro.enable_x64()
-numpyro.set_platform('gpu')
-
-import jax
-print(jax.devices())
-
-import os
-os.environ['XLA_PYTHON_CLIENT_PREALLOCATE']='false'
-
-import numpyro.distributions as dist
-from numpyro.infer import MCMC, NUTS, Predictive, init_to_median, init_to_value, init_to_sample
-
-
-import tensorflow_probability.substrates.jax as tfp
-
-from jax import numpy as jnp
-from jax import random, vmap
-
-
-import numpy as np
-import pandas as pd
-import arviz as az
-
-import siuba as s
-from siuba import _
-
-from plotnine import *
-
-import pickle
-
-from lib.helpers import *
-from lib.models import *
-from lib.simdata import make_sim_data_trials
-
-## -- load data
-
-n_participants, n_blocks, n_conditions = 50, 3, 2
-
-np.random.seed(346248) ## old school but works here
-
-params = {
-    "N_base": {k: np.exp(1 + np.random.normal(1, .5)) for k in range(0,n_participants)},
-    "N_delta": {k: np.exp(np.random.normal(1, .5)) for k in range(0,n_participants)},
-    "beta": {k: np.random.beta(5,1) for k in range(0, n_participants)},
-}
-
-sim_data = make_sim_data_trials(n_participants, n_blocks, n_conditions, params)
-sim_data["estimate"] = sim_data.estimate_ptn
-
-df = (sim_data >>
-      s.mutate(estimate =  np.round(_.estimate*20))
-     )
-
-df.estimate = df.estimate.astype("int64")
-
-X_data, y_data = make_model_data(df) # see data_helpers.py
-print(len(y_data), "observations")
-
-## --- 
-
-def prob_judge_BS(theta, X_num, X_denom, N, beta):
-
-    pi = calc_prob(theta, X_num, X_denom)
-    p_bs = pi * N / (N + 2.*beta) + beta/(N + 2.*beta)
-    
-    return p_bs
-
-def prob_judge_BS_d(theta, X_num, X_denom, d):
-
-    pi = calc_prob(theta, X_num, X_denom)
-    return calc_ptn_prob(pi, d)
-
-
-## rounding stuff
-from jax import vmap
-
-def spread_vec(x, step_size): # this works without static arguments
-    base_steps = x.shape[0]
[…]

### 2022-09-17 fit-models-exp1.py

@@ -1,74 +1,0 @@
-import numpyro
-numpyro.util.set_host_device_count(4)
-import numpyro.distributions as dist
-from numpyro.infer import MCMC, NUTS, Predictive, DiscreteHMCGibbs
-
-from jax import numpy as jnp
-from jax import random
-
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import arviz as az
-
-from dfply import *
-
-import pickle
-
-from lib.helpers import *
-from lib.models import *
-from lib.icc import *
-
-
-# load data
-
-df = load_query_avg_data_exp1() # see data_helpers.py
-df = df[~df.querytype.str.contains('or')]
-
-X_data, y_data = make_model_data(df) # see data_helpers.py
-print(len(y_data), "observations")
-
-# fit models
-
-models = {
-    "rf": relative_freq,
-    "bs_simple": bayesian_sampler_simple,
-    "bs": bayesian_sampler_complex,
-    "ptn_simple": PTN_simple,
-    "ptn": PTN_complex,
-    "bs_mlm_simple": bayesian_sampler_simple_mlm_d,
-    "bs_mlm": bayesian_sampler_complex_mlm_d,
-    "ptn_mlm_simple": PTN_simple_mlm,
-    "ptn_mlm": PTN_complex_mlm,
-    "ptn_mlm_simplecond": PTN_complex_mlm_simplecond
-}
-
-targ_accept_arg = {
-    "rf": .80,
-    "bs_simple": .80,
-    "bs": .80,
-    "ptn_simple": .80,
-    "ptn": .80,
-    "bs_mlm_simple": .95,
-    "bs_mlm": .95,
-    "ptn_mlm_simple": .95,
-    "ptn_mlm": .95,
-    "ptn_mlm_simplecond": .95
-}
-
-for name in models.keys():
-    
-    model = models[name]
-    kernel = NUTS(model, target_accept_prob=targ_accept_arg[name])
-    mcmc = MCMC(kernel, 
-               num_warmup=2_000, 
-               num_samples=2_000, 
-               num_chains=4)
-
-    mcmc.run(random.PRNGKey(0), X_data, y_data)
-    
-    loo = model_reloo(mcmc, kernel, k_thresh=.70)
-    
-    az_data = make_arviz_data(mcmc, model, X_data) 
-    pickle.dump(az_data, open("local/exp1/az_data_" + name + ".p", "wb"))
-    pickle.dump(loo, open("local/exp1/loo_" + name + ".p", "wb"))

### 2022-09-17 fit-models-exp2.py

@@ -1,75 +1,0 @@
-import numpyro
-numpyro.util.set_host_device_count(4)
-import numpyro.distributions as dist
-from numpyro.infer import MCMC, NUTS, Predictive, DiscreteHMCGibbs
-
-from jax import numpy as jnp
-from jax import random
-
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import arviz as az
-
-from dfply import *
-
-import pickle
-
-from lib.helpers import *
-from lib.models import *
-from lib.icc import *
-
-
-# load data experiment 2
-
-df = load_query_avg_data_exp2() # see data_helpers.py
-df = df[df["condition"]!=2] # filter out "warm/snowy" as per paper
-df.head()
-
-X_data, y_data = make_model_data(df) # see data_helpers.py
-print(len(y_data), "observations")
-
-# fit models
-
-models = {
-    "rf": relative_freq,
-    "bs_simple": bayesian_sampler_simple,
-    "bs": bayesian_sampler_complex,
-    "ptn_simple": PTN_simple,
-    "ptn": PTN_complex,
-    "bs_mlm_simple": bayesian_sampler_simple_mlm_d,
-    "bs_mlm": bayesian_sampler_complex_mlm_d,
-    "ptn_mlm_simple": PTN_simple_mlm,
-    "ptn_mlm": PTN_complex_mlm,
-    "ptn_mlm_simplecond": PTN_complex_mlm_simplecond
-}
-
-targ_accept_arg = {
-    "rf": .80,
-    "bs_simple": .80,
-    "bs": .80,
-    "ptn_simple": .80,
-    "ptn": .80,
-    "bs_mlm_simple": .95,
-    "bs_mlm": .95,
-    "ptn_mlm_simple": .95,
-    "ptn_mlm": .95,
-    "ptn_mlm_simplecond": .95
-}
-
-for name in models.keys():
-    
-    model = models[name]
-    kernel = NUTS(model, target_accept_prob=targ_accept_arg[name])
-    mcmc = MCMC(kernel, 
-               num_warmup=2_000, 
-               num_samples=2_000, 
-               num_chains=4)
-
-    mcmc.run(random.PRNGKey(0), X_data, y_data)
-    
-    loo = model_reloo(mcmc, kernel, k_thresh=.70)
-    
-    az_data = make_arviz_data(mcmc, model, X_data) 
[…]

### 2022-09-17 ic-compare.py

@@ -1,421 +1,0 @@
-from arviz.stats.stats_utils import ELPDData # modify to use relative import
-
-import warnings
-from copy import deepcopy
-from typing import List, Optional, Tuple, Union, Mapping, cast, Callable
-
-import numpy as np
-import pandas as pd
-import scipy.stats as st
-import xarray as xr
-from scipy.optimize import minimize
-from typing_extensions import Literal
-
-## set these back to relative imports
-from arviz.data import InferenceData, convert_to_dataset, convert_to_inference_data
-from arviz.rcparams import rcParams, ScaleKeyword, ICKeyword
-
-from arviz import loo, waic
-
-NO_GET_ARGS: bool = False
-try:
-    from typing_extensions import get_args
-except ImportError:
-    NO_GET_ARGS = True
-    
-
-def get_ic_type(ic_input):
-
-    try:
-        ic_input.loo
-        ic = "loo"
-    except:
-        try:
-            ic_input.waic
-            ic = "waic"
-        except:
-            ic = "unknown"
-    return(ic)
-
-
-def _ic_matrix(ics, ic_i):
-    """Store the previously computed pointwise predictive accuracy values (ics) in a 2D matrix."""
-    cols, _ = ics.shape
-    rows = len(ics[ic_i].iloc[0])
-    ic_i_val = np.zeros((rows, cols))
-
-    for idx, val in enumerate(ics.index):
-        ic = ics.loc[val][ic_i]
-
-        if len(ic) != rows:
-            raise ValueError("The number of observations should be the same across all models")
-
-        ic_i_val[:, idx] = ic
-
-    return rows, cols, ic_i_val
-
-
-def compare(
-    dataset_dict: Mapping[str, InferenceData],
-    ic: Optional[ICKeyword] = None,
-    method: Literal["stacking", "BB-pseudo-BMA", "pseudo-MA"] = "stacking",
-    b_samples: int = 1000,
-    alpha: float = 1,
-    seed=None,
-    scale: Optional[ScaleKeyword] = None,
-    var_name: Optional[str] = None,
-):
-    r"""Compare models based on PSIS-LOO `loo` or WAIC `waic` cross-validation.
-    LOO is leave-one-out (PSIS-LOO `loo`) cross-validation and
-    WAIC is the widely applicable information criterion.
-    Read more theory here - in a paper by some of the leading authorities
-    on model selection dx.doi.org/10.1111/1467-9868.00353
-    Parameters
-    ----------
[…]

### 2022-09-17 model_helpers.py

@@ -1,367 +1,0 @@
-import jax.numpy as jnp
-import numpy as np
-import arviz as az
-
-import re
-
-import siuba as s
-from siuba import _
-
-### ------ Data processing
-
-def load_raw_data(x=1):
-    import glob
-    df = pd.concat([pd.read_csv(f) for f in glob.glob('osfstorage-archive/Experiment {}/*.csv'.format(x))], ignore_index = True)
-
-    original_ids = list(np.unique(df.ID))
-    fix_id_dict = {original_ids[i]:i for i in range(0, len(original_ids))}
-
-#     df = (df >> 
-#           mutate(estimate = X.estimate/100.) >>
-#           mutate(ID = X.ID.apply(lambda x: fix_id_dict[x]))
-#          )
-    df = (df.assign(estimate = df.estimate/100., ID=df.ID.apply(lambda x: fix_id_dict[x]))
-         )
-    
-    return df
-
-
-# def load_data_exp1():
-#     df = load_raw_data(1)
-#     df["condition"] = np.select([df.querydetail.str.contains("icy|frosty")],[0], default=0)
-# 
-#     return df
-
-def load_data_exp1():
-    df = load_raw_data(1)
-    # df["condition"] = np.select([df.querydetail.str.contains("icy|frosty")],[0], default=0)
-    df = df >> s.mutate(condition = s.if_else(df.querydetail.str.contains("icy|frosty"), 0, 1))
-
-    return df
-
-def load_data_exp2():
-    df = load_raw_data(2)
-    
-    df["condition"] = np.select(
-    [
-        df.querydetail.str.contains("windy|cloudy"), 
-        df.querydetail.str.contains("cold|rainy"),
-        df.querydetail.str.contains("warm|snowy")
-
-    ], 
-    [
-        0,
-        1,
-        2
-    ], 
-    default=0 )
-
-    return df
-
-def load_data_exp2_trials():
-    df = load_raw_data(2)
-    
-    df["condition"] = np.select(
-    [
-        df.querydetail.str.contains("windy|cloudy"), 
-        df.querydetail.str.contains("cold|rainy"),
[…]

### 2022-09-17 models.py

@@ -1,717 +1,0 @@
-## -------------------------------------
-## Relative Frequency
-## -------------------------------------
-
-def calc_prob(theta, X_num, X_denom):
-
-    numerator = jnp.sum(theta*X_num, axis=-1)
-    denom = jnp.sum(theta*X_denom, axis=-1)
-    
-    return jnp.divide(numerator, denom)
-
-def relative_freq(data, y=None):
-
-    # Data processing
-    trial, subj, cond = data["trial"], data["subj"], data["cond"]
-    n_Ps, n_conds = np.unique(subj).shape[0], np.unique(cond).shape[0] 
-    
-    # setup "design matrix" (of sorts)
-    X_num, X_denom = jnp.stack([num_vecs[i] for i in trial]), jnp.stack([denom_vecs[i] for i in trial])
-
-    # population level parameters/priors
-#     sigma = numpyro.sample("sigma", dist.HalfCauchy(.1)) # for Normal() response noise version
-    k = numpyro.sample("k", dist.HalfNormal(50)) # noise parameter
-
-    # need a theta per person-query
-    with numpyro.plate("cond", n_Ps*n_conds):
-        thetas = numpyro.sample("theta", dist.Dirichlet(jnp.ones(4)))
-    
-    theta_ind = ((subj*n_conds)+cond)
-    theta = thetas[theta_ind,:]
-        
-    pi = calc_prob(theta, X_num, X_denom)
-
-    with numpyro.plate("data", len(trial)):
-#         yhat = numpyro.sample("yhat", dist.Normal(pi, sigma), obs = y) # wrong but replicates paper
-        yhat = numpyro.sample("yhat", dist.Beta(pi*k, (1-pi)*k), obs=y)
-    
-    return yhat
-
-## -------------------------------------
-## Bayesian Sampler
-## -------------------------------------
-
-def prob_judge_BS(theta, X_num, X_denom, N, beta):
-
-    pi = calc_prob(theta, X_num, X_denom)
-    p_bs = pi * N / (N + 2.*beta) + beta/(N + 2.*beta)
-    
-    return p_bs
-
-def prob_judge_BS_d(theta, X_num, X_denom, d):
-
-    pi = calc_prob(theta, X_num, X_denom)
-    return calc_ptn_prob(pi, d)
-
-
-def bayesian_sampler_simple(data, y=None):
-
-    # Data processing
[…]

### 2022-09-17 reloo-compare.py

@@ -1,433 +1,0 @@
-from arviz.stats.stats_utils import ELPDData # modify to use relative import
-
-import warnings
-from copy import deepcopy
-from typing import List, Optional, Tuple, Union, Mapping, cast, Callable
-
-import numpy as np
-import pandas as pd
-import scipy.stats as st
-import xarray as xr
-from scipy.optimize import minimize
-from typing_extensions import Literal
-
-## set these back to relative imports
-from arviz.data import InferenceData, convert_to_dataset, convert_to_inference_data
-from arviz.rcparams import rcParams, ScaleKeyword, ICKeyword
-
-from arviz import loo, waic
-
-NO_GET_ARGS: bool = False
-try:
-    from typing_extensions import get_args
-except ImportError:
-    NO_GET_ARGS = True
-    
-
-def get_ic_type(ic_input):
-
-    try:
-        ic_input.loo
-        ic = "loo"
-    except:
-        try:
-            ic_input.waic
-            ic = "waic"
-        except:
-            ic = "unknown"
-    return(ic)
-
-
-def _ic_matrix(ics, ic_i):
-    """Store the previously computed pointwise predictive accuracy values (ics) in a 2D matrix."""
-    cols, _ = ics.shape
-    rows = len(ics[ic_i].iloc[0])
-    ic_i_val = np.zeros((rows, cols))
-
-    for idx, val in enumerate(ics.index):
-        ic = ics.loc[val][ic_i]
-
-        if len(ic) != rows:
-            raise ValueError("The number of observations should be the same across all models")
-
-        ic_i_val[:, idx] = ic
-
-    return rows, cols, ic_i_val
-
-
-def compare(
-    dataset_dict: Mapping[str, InferenceData],
-    ic: Optional[ICKeyword] = None,
-    method: Literal["stacking", "BB-pseudo-BMA", "pseudo-MA"] = "stacking",
-    b_samples: int = 1000,
-    alpha: float = 1,
-    seed=None,
-    scale: Optional[ScaleKeyword] = None,
-    var_name: Optional[str] = None,
-):
-    r"""Compare models based on PSIS-LOO `loo` or WAIC `waic` cross-validation.
-    LOO is leave-one-out (PSIS-LOO `loo`) cross-validation and
-    WAIC is the widely applicable information criterion.
-    Read more theory here - in a paper by some of the leading authorities
-    on model selection dx.doi.org/10.1111/1467-9868.00353
-    Parameters
-    ----------
[…]

### 2022-09-17 costello_watts_condprob_test.R
readme tweaks

@@ -1,197 +1,0 @@
-
-# reproduce the data from Zhao
-zhao <- tribble(
-  ~query, ~true_prob, ~exp1, ~exp2, ~sim,
-  "p_B", .3, .31, .29, .32,
-  "p_B", .6, .58, .59, .59,
-  "p_B", .9, .86, .85, .86,
-  "p_AandB", .1, .19, .27, .15,
-  "p_AandB", .4, .50, .51, .42,
-  "p_AandB", .8, .80, .78, .77,
-  "p_AgB", .33, .35, .33, .36,
-  "p_AgB", .67, .63, .63, .65,
-  "p_AgB", .89, .82, .83, .84,
-  "ratio", .33, .70, .65, .44,
-  "ratio", .67, .93, 1.00, .71,
-  "ratio", .89, 1.03, 1.15, .90
-) %>% 
-  mutate(
-    conj = ifelse(query=="p_AandB", 1,0),
-    avg = (exp1+exp2)/2
-    )
-
-calc_ptn <- function(p, d, delta_d=0){
-  d = d + delta_d
-  
-  (1-2*d)*p + d
-}
-
-ptn_loss <- function(data,par){
-  with(
-    data,
-    {
-    d_par <- par[1] + conj*par[2]
-    preds <- calc_ptn(true_prob, d_par)
-    return(sum( (avg-preds)^2 ))
-    }
-  )
-}
-
-res <- optim(c(.001,.001), ptn_loss %>% filter(query!="ratio"), data=zhao, method="SANN") # this was being grumpy
-
-zhao <- zhao %>% 
-  mutate(preds = calc_ptn(true_prob, res$par[1] + conj*res$par[2]))
-
-zhao$preds[10:12] <- c(.553, .716, .872) # manually insert ratio predictions calc by hand
-
-## compare results of their model vs mine
-zhao %>% 
-  select(exp1, exp2, true_prob, sim, preds) %>% 
-  cor()
-
-## not enough data points to compare statistically, but seems my version is 
-## at least as good as theirs.
-
-## ---------
-
-
-p_AB <- .8
-p_AnotB <- .1
-p_notAB <- .05
-p_notAnotB <- .05
-
-p_A <- p_AB + p_AnotB
-p_B <- p_AB + p_notAB
[…]

### 2022-09-20 paper-rmd/paper-papaja-revision.Rmd
minor updates

@@
 Two recent proposals, the "Probability Theory plus Noise" [@costello.watts2014] and "Bayesian Sampler" [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
-These theories fit quite differently into the larger project of Bayesian cognitive science, but their many similarities complicate comparisons of their predictive accuracy.
-In particular, the models diverge most clearly in their distributional predictions, requiring a careful treatment of errors and external response processes.
-Here, I cast these theories into a Bayesian data analysis framework that supports principled model comparison using information criteria.
+These models differ in their averaged predictions about people's conditional probability judgments and in their distributional predictions about their overall patterns of judgments.
+In particular, the Bayesian sampler's Bayesian adjustment process predicts a truncated range of responses as well as a correlation between the average degree of bias and variability trial-to-trial.
+However, exploring these distributional predictions with participants' raw responses requires a careful treatment of rounding errors and exogenous response processes.
+Here, I cast these theories into a Bayesian data analysis framework that supports the treatment of these issues along with principled model comparison using information criteria.
 Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by an account of biases based on "noise" in the sample-reading process.
@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "9195"
+wordcount         : "9202"
 
@@
 
-This non-Bayesian account of conditional probabiliity judgments separates the PT+N theory quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
+This non-Bayesian account of conditional probability judgments separates the PT+N theory quite fundamentally from the Bayesian Sampler and the larger project of Bayesian cognitive science.
 
@@
 Another possibility is that the mental sampling process somehow enforces sampling of both positive and negative examples.
-A mental sampling process must start somewhere, and these initialization point(s) may influence judgments.
+Any sampling process must start somewhere, and these initialization point(s) may influence judgments.
 For example, features of the initializations of mental sampling processes have been proposed to anchoring [@lieder.etal2012] and unpacking effects [@dasgupta.etal2017; also see @sanborn.chater2016].
@@
 
-### Mixture modeling: rounding and contaminants
+### Mixture modeling: Rounding and contaminants
 
@@
 texp1 <- read_csv("exp1-trial-level-comparison.csv") %>% mutate(Experiment = c("Exp. 1", "", ""))
-texp2 <- read_csv("exp1-trial-level-comparison.csv") %>% mutate(Experiment = c("Exp. 2", "", ""))
+texp2 <- read_csv("exp2-trial-level-comparison.csv") %>% mutate(Experiment = c("Exp. 2", "", ""))
 
@@
 The Bayesian adjustment process hypothesized by the Bayesian Sampler model makes two clear predictions about the distribution of participants' responses.
-First, this implies a truncated range of possible responses.
+First, it implies a truncated range of possible responses.
 Second, assuming that $\beta$ is constrained to be a relatively small value, then under the Bayesian Sampler model, $N$ influences both the degree to which responses are shrunk toward .50 and the variability of those responses trial-to-trial.
@@
 Despite the model's quantitative success, some more qualitative questions remain.
-First, it bears asking: how plausible are the parameter values inferred from the model?
+First, the plausibility of the parameter values inferred from the model bears consideration.
 Many participants' estimated $d$ and $d'$ parameters were quite high---potentially against the spirit of the original PT+N model.
@@
 Further research examining what factors might affect the mental sampling and reading processes (e.g.
-task complexity, distractions, prior experience) might help to shed light on the plausibility of $d$ estimates in different contexts.
+task complexity, distractions, prior experience) might help to shed light on the most plausible range of $d$ values in different contexts.
 
 High estimates of $d$ parameters might also call into question arguments for the rational utility of such a process.
-Zhu and colleagues argue that the regularizing effect of bayesian adjustment should be seen as adaptive.
+Zhu and colleagues argue that the regularizing effect of Bayesian adjustment should be seen as adaptive.
 They also consider that "noise" might give an algorithmic-level solution to the computation-level goals defined by the Bayesian Sampler [-@zhu.etal2020].
@@
 
-As noted, the comparisons of these models using trial-level data rests on a number of elaborating assumptions to support fitting of the models.
+Finally as noted, the comparisons of these models using trial-level data rests on a number of elaborating assumptions to support fitting of the models.
 In particular, the models' implementations incorporate additional assumptions and approximations of the autocorrelated nature of mental samples and of ancillary error processes (e.g.
@@
 @sundh.etal2021], ultimately it seems crucial that cognitive models at some point be fit to the actual human behaviors of interest.
-
-Finally, one important direction for future elaborations of sampling-based theories are more rigorous theories of realistic mental sampling processes, including details of their initialization, autocorrelation, and amortization [@gershman.goodman2016].
+An important direction for future elaborations of sampling-based theories are more rigorous theories of realistic mental sampling processes, including details of their initialization, autocorrelation, and amortization [@gershman.goodman2016].
 Here, I proposed an autocorrelated sampling process initialized with a positive and negative example to account for qualitative features of participants' response distributions and to support a simple approximation.
@@
 
-The implications of sampling-based models like the Bayesian Sampler and PT+N theory go well-beyond the probability judgment task itself: these models esssentially rescue probabilistic theories of cognition by rejecting the explicit representation of probabilistic credences.
-Instead, credences are emergent properties of mental models that probabilistically generate samples [@chater.etal2020].
+Probability judgments have proven a fruitful testing ground for sampling-based theories of cognition.
+But, the implications of sampling-based models like the Bayesian Sampler and PT+N theory go well-beyond the probability judgment task itself: these models have the potential to extend the success of Bayesian theories of cognition to develop a probabilistic science of everyday beliefs.
+Under such an account, beliefs are not explicitly represented, stored, or even computed as probabilities, but rather they are emergent properties of mental models generating probabilistic samples [@sanborn.chater2016; @chater.etal2020].
 
-Clear models of these processes offer the potential to support a measurement model for probabilistic beliefs.
-Indeed, by representing the "true" subjective probabilities as a latent variable, Bayesian data analysis allows those underlying credences to be inferred.
-Examining the model posteriors here reveals these estimates often come with considerable uncertainty, but at least for some participants they can be estimated with useful levels of precision.
-Of course, Zhu and colleagues’ [-@zhu.etal2020] experiments were never designed for this purpose.
+Nevertheless we might still use the logic of Bayesian models to understand the operation of these beliefs and how they respond to evidence.
+Indeed, by representing the "true" subjective probabilities as a latent variable in the models used here, Bayesian data analysis allows those underlying credences to be inferred.
 Future research could explore how estimates of people's credences might be made more reliable, and how inferences about these mental probabilities might be integrated with other Bayesian models of reasoning [e.g.
 @franke.etal2016; @jern.etal2014; @griffiths.tenenbaum2006].
-For instance, people's responses in various reasoning tasks are often explicitly related to inferred subjective mental probabilities [e.g.
-@jern.etal2014], so accounting for biases in those reports may permit more rigorous model testing.
+For instance, people's responses in various reasoning tasks are often explicitly related to inferred subjective mental probabilities, so accounting for biases in those reports may permit more rigorous model testing.
 One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes [e.g.
-@cook.lewandowsky2016; @jern.etal2014; @powell.etal2018].
-
+@edwards1968; @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018; @powell2022].
 \pagebreak

### 2022-09-20 paper-rmd/paper-papaja-revision.Rmd
submitted

@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "9202"
+wordcount         : "9374"
 
@@
 output:
-  papaja::apa6_pdf:
-    citation_package: biblatex
+  papaja::apa6_pdf
 
@@
 Note that the Bayesian Sampler but not the PT+N model is capable of predicting non-zero values for identities Z10 through Z13.
+Error bars represent 95% CI.
 In Experiment 1, like participants' responses, the model's estimates are very slightly positive for \\{icy, frosty\\} and very slightly negative for \\{normal, typical\\}.
@@
 
-```{r, fig.cap='Posterior predictions for best-fitting model and participants responses in Experiments 1 and 2.'}
+```{r, fig.width=6, fig.height=3, fig.cap='Posterior predictions for best-fitting model and participants responses in Experiments 1 and 2.'}
 plt_pp
@@
 This sort of rounding can be modeled by a categorical distribution across the discrete possible rounded responses.
-Each rounded response category corresponds to a set of cut points, $a$ and $b$, with the probability of the categorical response defined by the CDF of the Beta distribution.
+Each rounded response category corresponds to a set of cut points, $a$ and $b$, with the probability of the categorical response defined by the cumulative distribution function of the underlying latent distribution.
 
@@
 
-Where $B$ is the cumulative distribution function of the underlying latent distribution.
 As participants were allowed to respond freely with whole numbers from 0 to 100, the exact rounding policy for each response is unknown.
@@
 One particularly promising direction could be to integrate these models with formal models of belief revision, which might then shed new light on these fundamental cognitive processes [e.g.
-@edwards1968; @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018; @powell2022].
+@cook.lewandowsky2016; @jern.etal2014; @powell.etal2018; @powell2022].
 \pagebreak
@@
 
-Where $B$ is the appropriate cumulative distribution function.
+where $B$ is the appropriate cumulative distribution function.
 To capture rounding to 10 we define $a_{i,10} =  \frac{2i}{m}-\frac{1}{m}$ and $b_i = \frac{2i}{m}+\frac{1}{m}$, so that the probability of each response is:

### 2022-12-02 paper-rmd/paper-papaja-revision.Rmd
added revised paper

@@
 output:
-  papaja::apa6_pdf
-
+  papaja::apa6_word
+keep_tex: yes
+always_allow_html: true
 header-includes:

### 2022-12-02 paper-rmd/paper-papaja-revision2.Rmd
2nd revision begun

--- 2022-12-02 paper-rmd/paper-papaja-revision.Rmd
@@
 First, as they note [@zhu.etal2020], BIC cannot fully account for the differences in the competing models’ complexity [also see @piantadosi2018].
-Further, their "unpooled" analysis likely exagerrates the complexity of the models overall and may therefore affect comparisons between them.
-
-#### Accounting for model complexity
-
-There are at least three challenges to accounting for model complexity in the comparison of the PT+N and Bayesian Sampler models.
-First, the models differ not only in the number of parameters but in the domain of those parameters.
-Via the bridging conditions, Zhu and colleauges [-@zhu.etal2020] show that assuming $\beta \in [0, 1]$ restricts the "noise level" for the Bayesian Sampler, represented in implied $d$ under the PT+N model, to fall within [0, $1/3$] (approaching $1/3$ as $N$ approaches 1), whereas the PT+N model permits noise values in $[0, 1/2]$.
-
-Second, it is unclear how the PT+N model’s treatment of conditional probabilities impacts its complexity.
-This structural difference could be a sort of Ptolemaic epicycle adding complexity to the model that should be penalized or it could constitute a commitment to novel predictions that thereby constrain its flexibility.
-
-Finally, Zhu and colleagues [-@zhu.etal2020] decision to fit "unpooled" estimates of all participants individually may exaggerate the complexity of both models.
-This is especially likely to pose an issue in differentiating between simple and complex variants, which Zhu and colleauges [-@zhu.etal2020] elect to ignore by averaging the model scores.
-If there is limited heterogeneity across individuals, then adding a parameter for each participant may effectively over-penalize the complex variants relative to the simple variants.
-In addition, this could also affect comparisons across the PT+N and Bayesian Sampelr theories: if one model must assume greater heterogeneity in parameter estimates across people, then this model is more complex.
-Hierarchical models with partial pooling offer a solution that balances between ignoring individual variation and allowing it to vary freely, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
+Further, their "unpooled" analysis likely exaggerates the complexity of the models overall and may therefore affect comparisons between them.
+In contrast, hierarchical models with partial pooling offer a solution that balances between ignoring individual variation and allowing it to vary freely, allowing for an accounting of heterogeneity without over-penalizing in cases where heterogeneity is low.
 
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
+<<revise: add that I will take two approaches, first examining their predictions in expectation to test their different accounts of conditional probability judgments.
+Then testing their distributional predictions to test their process-level accounts of probability judgments >>
 
-First, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on modern information criteria, such as Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$).
-Rather than estimating model fit and then penalizing for model complexity, $\text{PSIS-LOO}$ estimates out-of-sample prediction performance directly by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data [@gelman.etal2014; @vehtari.etal2017].
-From these calculations, an estimate of model complexity ($\hat{p}_{\text{LOO}}$) can also be derived.[^2]
+The Bayesian framework supports straightforward implementation of hierarchical versions of these models with partial pooling.
+This allows for information about model parameters to be shared across participants, resulting in potential improvements to out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
+In addition, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on modern information criteria, such as Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$; [@gelman.etal2014; @vehtari.etal2017]).[^2]
 
-[^2]: It is worth recognizing that formal measures of model complexity cannot be expected to perfectly track notions of simplicity or elegance in scientific explanation [for some related discussions, see @kuhn1977; @sober2002; @piantadosi2018].
-For instance, even if the PT+N model's account of conditional probability judgments constrains its flexibility empirically, it seems clear this added component makes it more complex as a putative scientific explanation.
-
-In addition, the Bayesian framework supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in potential improvements to out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
+[^2]: Rather than estimating model fit and then penalizing for model complexity, $\text{PSIS-LOO}$ estimates out-of-sample prediction performance directly by estimating the expected log predictive density ($\widehat{\text{elpd}}$) of the model, or the expected probability of new unseen data [@gelman.etal2014; @vehtari.etal2017].
+From these calculations, an estimate of model complexity ($\hat{p}_{\text{LOO}}$) can also be derived.
+However, it is worth recognizing that formal measures of model complexity will not always track notions of simplicity or elegance in scientific explanation [for some related discussions, see @kuhn1977; @sober2002; @piantadosi2018].
 
@@
 
-## Modeling Results: participant-level query-averaged responses
+## Modeling Results: Participant-level query-averaged responses
+
+<< this first set of analyses compares the models' ability to capture participant's proability judgments in expectation, averaged over the three blocks on which they made judgments about each query.
+These analyses will test the first points of differentiation between the models, their different predictions with respect to conditional probability judgmetns and their specific parameterizations.>>
 
@@
 This partial pooling can help to regularize parameter estimates and improve out-of-sample predictive performance.
-In addition, partial pooling effectively reduces model complexity, and could support more realistic comparison between the "simple" and "complex" variants of the models.
-
-The hierarchical implementation adds parameters for the population-level $d$ and $d'$ as well as a parameter controlling the standard deviation of the distribution for the subject-level effects.
+In addition, partial pooling effectively reduces model complexity, and could support more realistic comparison between the "simple" and "complex" variants of the models.Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency [@papaspiliopoulos.etal2007].
-Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 
@@
 
-Data from Experiment 1 favor "complex" variants of the Bayesian Sampler model compared with the "simple" variants and all versions of the PT+N model.
+Data from Experiment 1 favor "complex" variants of the Bayesian Sampler model compared with the "simple" variants and all versions of the PT+N model (greater values of $\widehat{\text{elpd}}_{\text{LOO}}$ are better).
 However, there is no clear single winner.
-As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler that allows for people to bring informative priors to the probability judgment task, that is, a Bayesian Sampler model allowing $\beta \in [0, \infty]$ (greater values of $\widehat{\text{elpd}}_{\text{LOO}}$ are better).
-However, the complex and complex hierarchical implementations of the Bayesian Sampler assuming uninformative priors (i.e., restricting $\beta \in [0,1]$) have $\widehat{\text{elpd}}_{\text{LOO}}$ scores within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
+As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler that allows for people to bring informative priors to the probability judgment task (i.e.
+allowing $\beta \in [0, \infty]$.
+However, the difference between this and the next best-fitting models are within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
+Data from Experiment 2 more decisively reveal a single winning model: the hierarchical "unrestricted" implementation of the Bayesian Sampler model allowing for informative priors.
 
-Data from Experiment 2 more decisively reveal a single winning model: the hierarchical "unrestricted" implementation of the Bayesian Sampler model allowing for informative priors.
+<< This unrestricted BS model differs from the PT+N model only in its treatment of conditional probability judgments.
+Thus, it can also be seen as a variant of the PT+N model that jettisons its constructive account of conditional probability judgments.
+From its superior fit, we can infer that the Bayesian Sample theory provides a better account of human conditional probability judgments.>>
 
@@
 ```
-
-Recall, the unrestricted Bayesian Sampler model may also be seen as a version of the PT+N theory that excises its constructive account of conditional probability judgments.
-So, comparing the PT+N models to the unrestricted Bayesian Sampler model offers a test of the PT+N's constructive account of conditional probability jugments.
-Comparing predictions from the unrestricted Bayesian Sampler model and the best-fitting PT+N model, we see that the Bayesian Sampler model better captures these judgments from both experiments (see Table 1).
-This is true for both query-level averages as well as for individual participants' trial responses.
-And comparing their predictions specifically for conditional probability judgments, the unrestricted Bayesian Sampler again provides a better fit, especially for the modestly-correlated events of Experiment 2 (Exp 1: response-level $r$ = .91 vs.
-.89 (_p_ < .05); query-level $r$ = .98 vs.
-.97; Exp 2: response-level $r$ = .72 vs.
-.67 (_p_ < .01); query-level $r$ = .92 vs.
-.85) These findings suggest that conditioning is better seen as part of a mental model of the events than as part of the probability judgment process.
-
-Table 1 also presents estimates of $\hat{p}_{\text{LOO}}$, the estimate of the effective number of parameters for each model.
-Compared to the Bayesian Sampler model, the PT+N model with its constructive account of conditional probability judgments has a similar penalty term estimate when fit to Experiment 1, but has a smaller penalty term when fit to Experiment 2, despite having the same parameterization in terms of $d$  and $d'$.
-Although its special treatment of conditional probability judgments makes it more complex as a putative scientific explanation, this structural component appears to actually constrain its predictive flexibility.
-However, this constraint leads to a worse-fitting model.
-
-Perhaps surprisingly, the Bayesian Sampler with unrestricted $\beta$ actually receives a smaller penalty term than the more "restricted" version of the model for the data in Experiment 2.
-This is at first counterintuitive.
-However, model complexity depends not only on the model and priors, but also the observed data [see @gelman.etal2014].
-To illustrate, Gelman and colleagues consider a case where a parameter is constrained to be positive and its value is then estimated from data [-@gelman.etal2014].
-If the estimated value is some very large positive number, then the constraint won’t have been very informative.
-But, if the estimated value is very close to zero, then the constraint that the parameter is positive will provide substantial information and the model’s penalty term will therefore be smaller.
-Here, it seems reasonable to conjecture that because the implied $d$  and $d'$ estimated from Experiment 2's data are both very near $1/3$ under this model, the restriction results in posterior estimates of the linear parameters that are relatively far from the prior, which can result in a greater penalty.
-
-The dependence of complexity penalties on observed data may strike some as an undesirable feature of model comparison through information criteria like $\text{PSIS-LOO}$.
-Indeed, it is worth acknowledging that principled model comparison is still an area of active inquiry, with differing perspectives [e.g.
-@gronau.wagenmakers2019; @vehtari.etal2019].
-Fortunately, in this case, comparisons between the models do not rest solely on differences in estimated model complexity.
 
@@
 
-Comparisons of the models' predictions in expectation have been revealing, but cannot decisively rule between the different process-level psychological theories behind the models.
+The best-fitting model capturing participant's predictions in expectation can be seen either as a variant of the Bayesian Sampler theory allowing for informative priors or as a variant of the PT+N theory without any special treatment of conditional versus uncondtional probability judgments.
+That is to say, these prior analyses have not decisively ruled between the different process-level psychological theories behind the models.
 To address this question, response distributions must be considered.
@@
 
-In the following set of analyses I attempt to lay out a set of reasonble assumptions that permit participants' trial-by-trial responses to be modeled and used to compare the theories' predictions.
+In the following set of analyses I attempt to lay out a set of reasonable assumptions that permit participants' trial-by-trial responses to be modeled and used to compare the theories' predictions.
 To do so, I first extend the models so as to render them fully continuous in their latent space and then marry them with a specific model of response errors.
@@
 
-What distribution of mental samples would such a process form?
-Figure 8 shows the result of a simulation study examining probability estimates drawing on autocorrelated samples with estimated weights.
-Each simulation consists of a simple discrete Markov chain generating autocorrelated samples at $x_t$, returning $x_{t-1}$ with probability $p_{a}$, or else returning a new sample distributed Bernoulli($p$) was simulated.
-The setting of $p_a$ controls the degree of (positive) autocorrelation in the chain.
-Given the degree of autocorrelation in a sample, a weight is initially estimated as $N_{\text{eff}}/N$ and an appropriate number of samples $N$ are drawn to produce $N_{\text{eff}}$ samples.
-A probability estimate is then calculated from these samples.
-As the figure shows, for large-enough autocorrelation values ($p_a$) the resulting distribution becomes quite smooth, with the exception of spikes at zero and one for small to moderate values of $N_{\text{eff}}$.
-This occurs when all samples are zero or all one.
-
-```{r, warning=F, message=F, fig.cap="Simulated proportions estimated from autocorrelated binary Markov Chain with single random initialization (left) or with initializations at 1 and 0 (right)."}
-source("../autocorr-sim.R")
-
-set.seed(12345)
-
-ps <- c(.25, .5, .9)
-Ns <- c(5, 10, 25)
-
-dsim <- expand_grid(
-  p = ps,
-  N = Ns
-) %>% 
-  mutate(
-    samps = map2(p, N, ~rBinEssInit(1000, .x, .y, .8)),
-    approx = map2(p, N, ~rbeta(1000, 1+ .x*.y, 1+(1-.x)*.y)),
-    )
-
-plt_init <- dsim %>% 
-  unnest(samps, approx) %>% 
-  gather(var, val, samps, approx) %>% 
-  mutate(var = if_else(var=="samps", "True", "Approximate")) %>% 
-  ggplot(aes(x=val, fill=var)) +
-  geom_histogram(position="identity", alpha=.6, binwidth=.025) +
-  facet_grid(p~N, scales="free_y") +
-  theme_bw() +
-  theme(panel.grid = element_blank()) +
-  labs(title = "Initialized autocorrelated sampling")
-
-dsim1 <- expand_grid(
-  p = ps,
-  N = Ns
-) %>% 
-  mutate(
-    samps = map2(p, N, ~rBinEss(1000, .x, .y)),
-    approx = map2(p, N, ~rbeta(1000, .x*.y, (1-.x)*.y)),
-    )
-
-plt_raw <- dsim1 %>% 
-  unnest(samps, approx) %>% 
-  gather(var, val, samps, approx) %>% 
-  mutate(var = if_else(var=="samps", "True", "Approximate")) %>% 
-  ggplot(aes(x=val, fill=var)) +
-  geom_histogram(position="identity", alpha=.6, binwidth=.025) +
-  # scale_fill_viridis_d() +
-  # facet_wrap(N + p~., scales="free") +
-  facet_grid(p~N, scales="free_y") +
-  theme_bw() +
-  theme(panel.grid=element_blank()) +
-  labs(title = "Raw autocorrelated sampling")
-
-(plt_raw + plt_init) + plot_layout(guides="collect") & theme(legend.position = 'bottom') & labs(fill = "Samples", x="p", y="Count") & theme(axis.text.x = element_text(angle=90, vjust=.5))
-```
-
-Sampling processes producing zero positive or negative examples should produce heaping of responses at zero and one under the PT+N model and at $d$ and $(1-d)$ under the Bayesian Sampler model.
-Compared against human judgment data, it is at least clear that actual human judgments do not equal zero or one as frequently as would be expected under the PT+N model using such a sampling process.
-Visual inspection of histograms of each participants' judgments and their estimated $d$ parameters also make clear that heaping at $d$ and $(1-d)$ is frequently absent (see Supplemental Materials).
-
-One possibility is that there is another source of error that explains these discrepancies, for instance that people just tend to avoid giving "extreme" responses.
-Another possibility is that the mental sampling process somehow enforces sampling of both positive and negative examples.
-Any sampling process must start somewhere, and these initialization point(s) may influence judgments.
-For example, features of the initializations of mental sampling processes have been proposed to anchoring [@lieder.etal2012] and unpacking effects [@dasgupta.etal2017; also see @sanborn.chater2016].
-In this spirit, one possible explanation for the relative absence of extreme responses is that people draw samples following two initializations, initializing sampling with both a positive and negative instance.
-As the simulations in Figure 8 (bottom) show, this produces a predicted response distribution that is  approximately $Beta(1 + p N_{\text{eff}}, 1+(1-p)N_{\text{eff}})$.
-
-Assuming such an autocorrelated and initialized sampling process provides a psychologically plausible mechanism to permit the use of a Beta distribution to approximate the latent sampling process, rendering both of the models' latent parameters continuous.
-Note that this process will also itself produce some shrinkage in the probability estimates, and this shrinkage will be somewhat sensitive to the number of samples drawn in both the Bayesian Sampler and PT+N models.
-Nevertheless the dependency between shrinkage and variance remains stronger for the Bayesian Sampler model and it is only the Bayesian Sampler model that predicts a truncation of the response distribution.
+<< say a bit more on this assumption / revise to clarify a bit >>
 
@@
 
-The Bayesian Sampler model predicts a truncated range of possible responses given $\beta$ and $N$ (and consequently, implied $d$).
+The Bayesian Sampler model predicts a truncated range of possible responses given $\beta$ and effective $N$ (and consequently, implied $d$).
 Modeling these different rounding processes allows for at least some out-of-bounds responses to be accounted for by rounding processes (e.g.
@@
 
-Above we found that an autocorrelated sampling process could be reasonably approximated by a Beta distribution. Taking the original Bayesian Sampler model
+Above we found that an autocorrelated sampling process could be reasonably approximated by a Beta distribution. Taking the original Bayesian Sampler model,
 
@@
 
-We can replace the number of successes $S(A)$ (distributed binomial) with the quantity $\rho(A)N$, where $\rho(A)$ represents the Beta-distributed sample proportions generated by the autocorrelated sampling process outlined above.
+we can replace the number of successes $S(A)$ (distributed binomial) with the quantity $\rho(A)N$, where $\rho(A)$ represents the Beta-distributed sample proportions generated by the autocorrelated sampling process outlined above.
 
@@
 This added complexity is likely ignorable for a noise-based model, but could be important for the Bayesian Sampler model, as the number of samples drawn affects the truncation of the response distribution.
-That is, allowing the number of samples drawn to vary across trials may allow the model to capture some respnses that would otherwise be treated as contaminants.
+That is, allowing the effective number of samples drawn to vary across trials may allow the model to capture some responses that would otherwise be treated as contaminants.
 To capture this, the model can be given one additional extension to allow the number of samples to vary, by adding a new parameter $N_{trial}$.
-This parameter multiplies the number of samples as a fraction of each individual participants' average number of samples drawn, e.g.
-so that a participant might sometimes draw 1.5x or 2x the number of samples they typically draw.
+This parameter multiplies the number of effective samples as a fraction of each individual participants' average number of samples drawn, e.g.
+so that a participant might sometimes draw 1.5x or 2x the number of effective samples they typically draw.
 The appropriate amount of variation in $N_{trial}$ is constrained to be fairly small, but is estimated hierarchically: I assume $log(N_{trial}) \sim N(0,\sigma_{N_{trial}})$ and $log(\sigma_{N_{trial}}) \sim (-1, .3)$.
@@
 Table 2 shows the scores of each model fit to the trial-level data in Experiments 1 and 2.
-From the quantitative model comparison it is clear that the noise-based model is superior, with substantialy lower $\widehat{\text{elpd}}_{\text{LOO}}$ than both of the competing Bayesian Sampler implementations in both experiments.
+From the quantitative model comparison it is clear that the noise-based model is superior, with substantially lower $\widehat{\text{elpd}}_{\text{LOO}}$ than both of the competing Bayesian Sampler implementations in both experiments.
 
@@
 In the noise-based model, $N$ and $d$ were allowed to vary freely.
-But if these quantities indeed are actually correlated as predicted by a Bayesian adjustment model, then we should expect to nevertheless see a correlation between the subject-level estimates in these parameters.
+But if these quantities are actually correlated as predicted by a Bayesian adjustment model, then we should expect to nevertheless see a correlation between the subject-level estimates in these parameters.
 Figure 11 shows a scatterplot relating these estimates.
@@
 
-The unrestricted BS model and PT+N with Bayesian conditional probability judgments cannot be distinguised by their predictions about expected responses.
-However, they remain substantively different models positing different process-level accounts of probability judgments.
-Both accounts rely on the concept of mental sampling.
-But in the BS model the samples are integrated with a Bayesian prior to produce a judgment, whereas in the PT+N model, noise in the sample-reading process produces bias.
-
-The Bayesian adjustment process hypothesized by the Bayesian Sampler model makes two clear predictions about the distribution of participants' responses.
+At the psychological process-level, the Bayesian adjustment process hypothesized by the Bayesian Sampler model makes two clear predictions about the distribution of participants' responses.
 First, it implies a truncated range of possible responses.
@@
 Finally as noted, the comparisons of these models using trial-level data rests on a number of elaborating assumptions to support fitting of the models.
-In particular, the models' implementations incorporate additional assumptions and approximations of the autocorrelated nature of mental samples and of ancillary error processes (e.g.
-rounding).
 It should be recognized that different assumptions may have produced different results, and other error processes remain possible.
@@
 An important direction for future elaborations of sampling-based theories are more rigorous theories of realistic mental sampling processes, including details of their initialization, autocorrelation, and amortization [@gershman.goodman2016].
-Here, I proposed an autocorrelated sampling process initialized with a positive and negative example to account for qualitative features of participants' response distributions and to support a simple approximation.
-Future work might explore the details of these processes and incorporate them into models of probability judgments in more rigorous ways.
 
@@
 @cook.lewandowsky2016; @jern.etal2014; @powell.etal2018; @powell2022].
+
 \pagebreak

### 2022-12-14 create-paper-figures.Rmd
updated rmd and avg code

@@
 library(tidyverse)
+library(patchwork)
 library(latex2exp)
@@
 
-fits["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/az_data_rf.p", "rb"))
-fits["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/az_data_bs_simple.p", "rb"))
-fits["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/az_data_bs.p", "rb"))
-fits["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/az_data_bs_mlm_simple.p", "rb"))
-fits["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/az_data_bs_mlm.p", "rb"))
-fits["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/az_data_ptn.p", "rb"))
-fits["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/az_data_ptn_simple.p", "rb"))
-fits["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm.p", "rb"))
-fits["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simple.p", "rb"))
-fits["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond.p", "rb"))
+# fits["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/az_data_rf.p", "rb"))
+# fits["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/az_data_bs_simple.p", "rb"))
+# fits["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/az_data_bs.p", "rb"))
+# fits["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/az_data_bs_mlm_simple.p", "rb"))
+# fits["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/az_data_bs_mlm.p", "rb"))
+# fits["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/az_data_ptn.p", "rb"))
+# fits["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/az_data_ptn_simple.p", "rb"))
+# fits["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm.p", "rb"))
+# fits["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simple.p", "rb"))
+# fits["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond.p", "rb"))
+fits["exp1"]["Unrestricted BS hier."] = az.from_netcdf("local/exp1/az_data_ptn_mlm_simplecond.nc")
 
-fits["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/az_data_rf.p", "rb"))
-fits["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/az_data_bs_simple.p", "rb"))
-fits["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/az_data_bs.p", "rb"))
-fits["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/az_data_bs_mlm_simple.p", "rb"))
-fits["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/az_data_bs_mlm.p", "rb"))
-fits["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/az_data_ptn.p", "rb"))
-fits["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/az_data_ptn_simple.p", "rb"))
-fits["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm.p", "rb"))
-fits["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simple.p", "rb"))
-fits["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simplecond.p", "rb"))
+# fits["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/az_data_rf.p", "rb"))
+# fits["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/az_data_bs_simple.p", "rb"))
+# fits["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/az_data_bs.p", "rb"))
+# fits["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/az_data_bs_mlm_simple.p", "rb"))
+# fits["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/az_data_bs_mlm.p", "rb"))
+# fits["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/az_data_ptn.p", "rb"))
+# fits["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/az_data_ptn_simple.p", "rb"))
+# fits["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm.p", "rb"))
+# fits["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simple.p", "rb"))
+# fits["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/az_data_ptn_mlm_simplecond.p", "rb"))
+fits["exp2"]["Unrestricted BS hier."] = az.from_netcdf("local/exp2/az_data_ptn_mlm_simplecond.nc")
 
-loos = dict()
-loos["exp1"] = dict()
-loos["exp2"] = dict()
-
-loos["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/loo_rf.p", "rb"))
-loos["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/loo_bs_simple.p", "rb"))
-loos["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/loo_bs.p", "rb"))
-loos["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/loo_bs_mlm_simple.p", "rb"))
-loos["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/loo_bs_mlm.p", "rb"))
-loos["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/loo_ptn.p", "rb"))
-loos["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/loo_ptn_simple.p", "rb"))
-loos["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/loo_ptn_mlm.p", "rb"))
-loos["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simple.p", "rb"))
-loos["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simplecond.p", "rb"))
-
-loos["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/loo_rf.p", "rb"))
-loos["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/loo_bs_simple.p", "rb"))
-loos["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/loo_bs.p", "rb"))
-loos["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/loo_bs_mlm_simple.p", "rb"))
-loos["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/loo_bs_mlm.p", "rb"))
-loos["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/loo_ptn.p", "rb"))
-loos["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/loo_ptn_simple.p", "rb"))
-loos["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/loo_ptn_mlm.p", "rb"))
-loos["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simple.p", "rb"))
-loos["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simplecond.p", "rb"))
+# loos = dict()
+# loos["exp1"] = dict()
+# loos["exp2"] = dict()
+# 
+# loos["exp1"]["Relative Freq."] = pickle.load(open("local/exp1/loo_rf.p", "rb"))
+# loos["exp1"]["Bayesian Sampler simple"] = pickle.load(open("local/exp1/loo_bs_simple.p", "rb"))
+# loos["exp1"]["Bayesian Sampler complex"] = pickle.load(open("local/exp1/loo_bs.p", "rb"))
+# loos["exp1"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp1/loo_bs_mlm_simple.p", "rb"))
+# loos["exp1"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp1/loo_bs_mlm.p", "rb"))
+# loos["exp1"]["PT+N complex"] = pickle.load(open("local/exp1/loo_ptn.p", "rb"))
+# loos["exp1"]["PT+N simple"] = pickle.load(open("local/exp1/loo_ptn_simple.p", "rb"))
+# loos["exp1"]["PT+N complex hier."] = pickle.load(open("local/exp1/loo_ptn_mlm.p", "rb"))
+# loos["exp1"]["PT+N simple hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simple.p", "rb"))
+# loos["exp1"]["Unrestricted BS hier."] = pickle.load(open("local/exp1/loo_ptn_mlm_simplecond.p", "rb"))
+# 
+# loos["exp2"]["Relative Freq."] = pickle.load(open("local/exp2/loo_rf.p", "rb"))
+# loos["exp2"]["Bayesian Sampler simple"] = pickle.load(open("local/exp2/loo_bs_simple.p", "rb"))
+# loos["exp2"]["Bayesian Sampler complex"] = pickle.load(open("local/exp2/loo_bs.p", "rb"))
+# loos["exp2"]["Bayesian Sampler simple hier."] = pickle.load(open("local/exp2/loo_bs_mlm_simple.p", "rb"))
+# loos["exp2"]["Bayesian Sampler complex hier."] = pickle.load(open("local/exp2/loo_bs_mlm.p", "rb"))
+# loos["exp2"]["PT+N complex"] = pickle.load(open("local/exp2/loo_ptn.p", "rb"))
+# loos["exp2"]["PT+N simple"] = pickle.load(open("local/exp2/loo_ptn_simple.p", "rb"))
+# loos["exp2"]["PT+N complex hier."] = pickle.load(open("local/exp2/loo_ptn_mlm.p", "rb"))
+# loos["exp2"]["PT+N simple hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simple.p", "rb"))
+# loos["exp2"]["Unrestricted BS hier."] = pickle.load(open("local/exp2/loo_ptn_mlm_simplecond.p", "rb"))
 ```
@@
 ```{python}
-exp1_model_comparison = compare(loos["exp1"], method="BB-pseudo-BMA")
-exp2_model_comparison = compare(loos["exp2"], method="BB-pseudo-BMA")
+# exp1_model_comparison = compare(loos["exp1"], method="BB-pseudo-BMA")
+# exp2_model_comparison = compare(loos["exp2"], method="BB-pseudo-BMA")
 
-exp1_model_comparison = exp1_model_comparison >> s.mutate(Experiment = "Exp. 1", model = exp1_model_comparison.index)
-exp2_model_comparison = exp2_model_comparison >> s.mutate(Experiment = "Exp. 2", model = exp2_model_comparison.index)
+exp1_model_comparison = pd.read_csv("local/exp1/model-comparison.csv")
+exp2_model_comparison = pd.read_csv("local/exp2/model-comparison.csv")
+
+exp1_model_comparison = exp1_model_comparison >> s.mutate(Experiment = "Exp. 1", model = exp1_model_comparison["Unnamed: 0"])
+exp2_model_comparison = exp2_model_comparison >> s.mutate(Experiment = "Exp. 2", model = exp1_model_comparison["Unnamed: 0"])
 
@@
 
-library(patchwork)
-
 (plt_forest_exp1 + coord_fixed(ratio=59/.5)) + 
@@
   ggplot(aes(x=pred, y= estimate)) +
-  geom_point(alpha=.25) +
+  geom_abline(slope=1, intercept=0, linetype="dashed", color="grey") +
+  geom_point(alpha=.25, shape=16) +
   facet_wrap(~Experiment) +
@@
   theme(aspect.ratio=1) +
-  theme(panel.grid=element_blank())
+  labs(x="Predicted", y = "Observed") +
+  theme(panel.grid=element_blank()) +
+  scale_x_continuous(labels=c("0",".25",".50",".75","1")) +
+  scale_y_continuous(labels=c("0",".25",".50",".75","1"))
 
@@
   mutate(inequality = ordered(inequality, levels = paste0("z",1:18))) %>% 
+  mutate(
+    # condition = if_else(condition=="estimate", "Human data", "Model Pred."),
+    var = if_else(var=="estimate","Observed", "Predicted")
+    ) %>% 
   group_by(var, inequality, condition) %>% 
@@
 df1_all = load_query_avg_data_exp1()
-model1_all = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond-all.p", "rb"))
+# model1_all = pickle.load(open("local/exp1/az_data_ptn_mlm_simplecond-all.p", "rb"))
+model1_all = az.from_netcdf("local/exp1/az_data_ptn_mlm_simplecond-all.nc")
 preds1_all = model1_all.posterior_predictive.mean(dim=['chain', 'draw']).yhat.to_numpy()
@@
 # az_ptn_exp2 = pickle.load(open("local/az-ptn_simplecond_mlm_trial_level_disc-exp2svi.p", "rb"))
-# 
-# # posterior_ptn1 = posterior_to_df_mlm(
-# #     az_ptn_exp1, 
-# #     ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
-# #     ["d_subj", "d_prime_subj", "k"]
-# # )
-# # 
-# posterior_ptn2 = posterior_to_df_mlm(
-#     az_ptn_exp2,
-#     ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
-#     ["d_base_r", "d_delta_r", "d_subj", "d_prime_subj", "k"]
-# )
 
-posterior_ptn1 = pickle.load(open("local/posterior_ptn1.p", "rb"))
-posterior_ptn2 = pickle.load(open("local/posterior_ptn2.p", "rb"))
+az_ptn_exp1 = az.from_netcdf("local/az-ptn_simplecond_mlm_trial_level_disc-exp1svi.nc") # fix filenames
+az_ptn_exp2 = az.from_netcdf("local/az-ptn_simplecond_mlm_trial_level_disc-exp2svi.nc")
+
+posterior_ptn1 = posterior_to_df_mlm(
+    az_ptn_exp1,
+    ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
+    ["d_subj", "d_prime_subj", "k"]
+)
+
+posterior_ptn2 = posterior_to_df_mlm(
+    az_ptn_exp2,
+    ["d_base_pop","d_delta_pop","d_base_sd","d_delta_sd"],
+    ["d_base_r", "d_delta_r", "d_subj", "d_prime_subj", "k"]
+)
+
+# posterior_ptn1 = pickle.load(open("local/posterior_ptn1.p", "rb"))
+# posterior_ptn2 = pickle.load(open("local/posterior_ptn2.p", "rb"))
 ```
@@
     ) %>% 
-#     s.filter(_.k_est < 50.) >>
-    filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
+    filter(k_est < 100) %>%
+    # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>%
     ggplot(aes(x=k_est, y=d_est)) +
@@
     ) %>% 
-#     s.filter(_.k_est < 50.) >>
+    # s.filter(_.k_est < 50.) %>% 
     # filter(ID!=51, ID!=45, ID!=29, ID!=28) %>% 
@@
 ```{r}
+plt_dprimek_ptn1 + geom_smooth(method="lm")
+```
+
+```{r}
 plt_d_avgmodel <- py$param_posterior %>% 
@@
   geom_density(position="identity", alpha=.6) +
+  geom_vline(xintercept=.3333, linetype="dashed", color="grey") +
   facet_wrap(~Experiment) +

### 2022-12-14 paper-rmd/paper-papaja-revision2.Rmd
updated rmd and avg code

@@
  Bayesian theories of cognitive science hold that cognition is fundamentally probabilistic, but people’s explicit probability judgments often violate the laws of probability.
-Two recent proposals, the "Probability Theory plus Noise" [@costello.watts2014] and "Bayesian Sampler" [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
+Two recent proposals, the "Probability Theory plus Noise" [PT+N; @costello.watts2014] and "Bayesian Sampler" [@zhu.etal2020] theories of probability judgments, both seek to account for these biases while maintaining that mental credences are fundamentally probabilistic.
 These models differ in their averaged predictions about people's conditional probability judgments and in their distributional predictions about their overall patterns of judgments.
-In particular, the Bayesian sampler's Bayesian adjustment process predicts a truncated range of responses as well as a correlation between the average degree of bias and variability trial-to-trial.
+In particular, the Bayesian Sampler's Bayesian adjustment process predicts a truncated range of responses as well as a correlation between the average degree of bias and variability trial-to-trial.
 However, exploring these distributional predictions with participants' raw responses requires a careful treatment of rounding errors and exogenous response processes.
 Here, I cast these theories into a Bayesian data analysis framework that supports the treatment of these issues along with principled model comparison using information criteria.
-Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020] I find the data are best explained by an account of biases based on "noise" in the sample-reading process.
+Comparing the fits of both models on data collected by Zhu and colleagues [-@zhu.etal2020], I find these data are best explained by an account of biases based on "noise" in the sample-reading process but in which conditional probability judgments are produced by a process of conditioning in the mental model of the events, rather than in a two-stage mental sampling process as proposed by the PT+N model.
 
@@
 keywords          : "probability judgments, Bayesian cognitive science, heuristics and biases"
-wordcount         : "9374"
+wordcount         : "8484"
 
@@
 output:
-  papaja::apa6_word
+  papaja::apa6_pdf
 keep_tex: yes
@@
 According to Bayesian theories of cognition and epistemology [@jaynes2003], the degree to which people believe in various propositions, or their credences, should reflect subjective mental probabilities.
-So asking people to express beliefs in terms of probability seems only natural.
+So, asking people to express beliefs in terms of probability seems only natural.
 
@@
 Other demonstrations of the incoherence of probability judgments include disjunction fallacies, subadditivity or "unpacking" effects [@tversky.koehler1994], and a number of others [for an accessible review, see @kahneman2013].
-Altogether these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
+Altogether, these findings have led many researchers to abandon the notion that degrees of belief are represented as probabilities.
 
@@
 The end products are probability judgments reflecting probabilistic credences perturbed by noise.
-In their model, the probability a mental sample for an event $A$ is read as $A$ is the probability that the sample truly is $A$, $p(A)$, and that it is correctly read $(1-d)$, plus the probability that the sample is not $A$, $1-P(A)$ and that it is incorrectly read ($d$), or:
+In their model, the probability that a mental sample for an event $A$ is correctly read as $A$ is the probability that the sample truly is $A$, $p(A)$, and that it is correctly read $(1-d)$, plus the probability that the sample is not $A$, $1-P(A)$ and that it is incorrectly read ($d$), or:
 
@@
 Thus, in many cases the effect of a Bayesian prior is identical to the effect of noise in the PT+N model (at least in expectation).
-But, rather than merely perturbing people's probability judgments, this prior can be seen as regularizing these judgments away from extreme values.
+A caveat to this is that the Bayesian Sampler theory restricts the parameterization of the equivalent $d$ parameter compared to the PT+N model.
+Whereas PT+N assumes $d \in [0, 1/2]$, the Bayesian Sampler theory assumes an uninformatative prior parameter $\beta \in [0, 1]$, which in turn restricts the equivalent parameterization of $d \in [0, 1/3]$.
+But beyond this subtle difference of parameterization, there is a larger difference in interpretation: rather than merely perturbing people's probability judgments, this prior can be seen as regularizing these judgments away from extreme values.
 Zhu and colleagues [-@zhu.etal2020] argue that such regularization can be adaptive in cases where only a small number of mental samples can be drawn.
@@
 The model's predictions can be distinguished on two levels: First, the models have distinct accounts of conditional probability judgments that make different predictions in terms of expected values.
-Second, the models present different process-level accounts of probability judgment that entail different predictions about the shape of the distribution of responses across individuals.
+Second, the models present different process-level accounts of probability judgment that entail different predictions about the shape of the distribution of responses across trials.
 
@@
 
-In later work, Sundh et al [-@sundh.etal2021] examined the distributional properties of participants responses using indirect means, by regressing the variance of participants' responses across trials on their mean response for each query type.
+In later work, Sundh et al [-@sundh.etal2021] examined the distributional properties of participants' responses using indirect means, by regressing the variance of participants' responses across trials on their mean response for each query type.
 Their findings suggest that earlier fits with Wasserstein distance may have produced biased results [@sundh.etal2021].
@@
 Here, I cast both the Bayesian Sampler and PT+N models into a Bayesian data analysis framework that may permit a more decisive comparison.
-<<revise: add that I will take two approaches, first examining their predictions in expectation to test their different accounts of conditional probability judgments.
-Then testing their distributional predictions to test their process-level accounts of probability judgments >>
+Two sets of analyses compared different aspects of the models' predictions: First, the model's were examined based on their predictions in expectation.
+This analysis allows for a test of their different accounts of conditional probability judgments and their parameterizations (i.e.
+the restriction of $d$ to $[0, 1/2]$ versus $[0, 1/3]$).
+To preview the results, these analyses revealed the Bayesian Sampler's account of probability judgments to be superior, but could not distinguish between the different psychological processes proposed by the two models.
+A second set of analyses examined the model's distributional predictions to test their process-level accounts of probability judgments.
+As will be described, modeling response distributions directly presents a number of challenges, and so this second set of analyses required some minor additions and modifications to the models to permit their fitting to experimental data.
 
-The Bayesian framework supports straightforward implementation of hierarchical versions of these models with partial pooling.
-This allows for information about model parameters to be shared across participants, resulting in potential improvements to out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
-In addition, Bayesian data analysis allows issues of model complexity to be addressed through comparisons of model fit based on modern information criteria, such as Pareto smoothed importance sampling approximate leave-one-out cross validation ($\text{PSIS-LOO}$; [@gelman.etal2014; @vehtari.etal2017]).[^2]
+Both sets of analyses are supported by the use of a Bayesian Framework.
+First, Bayesian analyses allow issues of model complexity to be addressed through comparisons of model fit based on modern information criteria, such as Pareto smoothed importance sampling approximate leave-one-out cross validation [$\text{PSIS-LOO}$; @gelman.etal2014; @vehtari.etal2017].[^2] Second, the Bayesian framework supports straightforward implementation of hierarchical versions of these models allowing for information about model parameters to be shared across participants, resulting in potential improvements to out-of-sample prediction, reductions in model complexity, and a more realistic test of the models.
+Finally, a Bayesian framework also supports new extensions of these models to directly model participants trial-level responses while accounting for rounding and out-of-distribution response errors.
+These extensions allow for principled probabilistic tests of the distributional predictions of the models.
 
@@
 However, it is worth recognizing that formal measures of model complexity will not always track notions of simplicity or elegance in scientific explanation [for some related discussions, see @kuhn1977; @sober2002; @piantadosi2018].
-
-Finally, I implement new extensions of these models to directly model participants trial-level responses while accounting for rounding and out-of-distribution response errors---allowing for principled probabilistic tests of the distributional predictions of the models.
 
@@
 
-<< this first set of analyses compares the models' ability to capture participant's proability judgments in expectation, averaged over the three blocks on which they made judgments about each query.
-These analyses will test the first points of differentiation between the models, their different predictions with respect to conditional probability judgmetns and their specific parameterizations.>>
+This first set of analyses compares the models' ability to capture participant's probability judgments in expectation, averaged over the three blocks on which they made judgments about each query.
+These analyses will test the first points of differentiation between the models: their different predictions with respect to conditional probability judgments and their specific parameterizations.
 
@@
 This partial pooling can help to regularize parameter estimates and improve out-of-sample predictive performance.
-In addition, partial pooling effectively reduces model complexity, and could support more realistic comparison between the "simple" and "complex" variants of the models.Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
+In addition, partial pooling effectively reduces model complexity, and could support more realistic comparison between the "simple" and "complex" variants of the models.
+Figure 2 displays the translation of a hierarchical implementation of the Bayesian Sampler model into the Bayesian framework, along with a plate diagram representing the dependencies among parameters.
 For ease of interpretation, the centered parameterization is shown below, although the actual models used a non-centered parameterization to improve sampling efficiency [@papaspiliopoulos.etal2007].
@@
 That is, through the bridging condition, the implied $d$ approaches $1/2$ in the limit as $N \to 1$ and $\beta \to \infty$.
-Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its constructive account of conditional probability judgment.
+Though it would seem a more fundamental change, this same model may also be seen as a version of the PT+N theory that jettisons its two-stage process of conditional probability judgment.
 Thus, fitting this additional unrestricted model allows for a complete comparison of the models along both of their differing dimensions.
@@
 Data from Experiment 1 favor "complex" variants of the Bayesian Sampler model compared with the "simple" variants and all versions of the PT+N model (greater values of $\widehat{\text{elpd}}_{\text{LOO}}$ are better).
-However, there is no clear single winner.
 As shown in Figure 3, the best-scoring model is an unrestricted variant of the Bayesian Sampler that allows for people to bring informative priors to the probability judgment task (i.e.
 allowing $\beta \in [0, \infty]$.
-However, the difference between this and the next best-fitting models are within two standard errors of the difference, indicating that these models are also plausible [@sivula.etal2020].
 Data from Experiment 2 more decisively reveal a single winning model: the hierarchical "unrestricted" implementation of the Bayesian Sampler model allowing for informative priors.
 
-<< This unrestricted BS model differs from the PT+N model only in its treatment of conditional probability judgments.
-Thus, it can also be seen as a variant of the PT+N model that jettisons its constructive account of conditional probability judgments.
-From its superior fit, we can infer that the Bayesian Sample theory provides a better account of human conditional probability judgments.>>
+This unrestricted BS model differs from the PT+N model only in its treatment of conditional probability judgments and so, from its superior fit, we can infer that the Bayesian Sample theory provides a better account of human conditional probability judgments.
 
@@
 Dashed line indicates theoretical maximum values for Bayesian Sampler model with uninformative priors."}
-plt_d_avgmodel + plt_d_trialmodel + plot_layout(ncol=1, guides="collect") & labs(x="Estimate") & theme_bw() & theme(panel.grid=element_blank())
+plt_d_avgmodel + plt_d_trialmodel + plot_layout(ncol=1, guides="collect") & labs(y = "Density", x="Estimate") & theme_bw() & theme(panel.grid=element_blank())
 
@@
     caption="Bayesian model comparison results with best scoring model in bold face.
-Models are compared based on X and the correlation between their predictions for each participants' query-level average ($r_{\\text{query}}$) as well as overall query-level averages $r_{\\text{query-avg}}$.",
+Models are compared based on $\\widehat{\text{elpd}}_{\text{LOO}}$ and the correlation between their predictions for each participants' query-level average ($r_{\\text{query}}$) as well as overall query-level averages $r_{\\text{query-avg}}$.",
     digits=3, 
@@
 
-## Raw trial-level response distributions
+## Results: Raw trial-level response distributions
 
-The best-fitting model capturing participant's predictions in expectation can be seen either as a variant of the Bayesian Sampler theory allowing for informative priors or as a variant of the PT+N theory without any special treatment of conditional versus uncondtional probability judgments.
-That is to say, these prior analyses have not decisively ruled between the different process-level psychological theories behind the models.
-To address this question, response distributions must be considered.
+ The best-fitting model capturing participant's predictions in expectation can be seen either as a variant of the Bayesian Sampler theory allowing for informative priors or as a variant of the PT+N theory without any special treatment of conditional versus unconditional probability judgments.
+Thus, these prior analyses have not decisively ruled between the different process-level psychological theories behind the models.
+To evaluate the Bayesian Sampler and PT+N theories' competing accounts of the psychological processes behind probability judgments, a second set of analyses focused on the distribution of participant's trial-by-trial responses was conducted.
 
@@
 Because the Bayesian Sampler applies a Bayesian adjustment after sampling, it predicts probability judgments will always lie between $d$ and $1 - d$, even when zero positive or negative mental samples are drawn (see Equation 12 and the bridging condition, Equation 5).
-First, it bears noting that the truncated response distribution implied by the Bayesian Sampler model appears at odds with the raw response data.
-However, participants' responses frequently lie outside the range implied by the best estimates of their $d$ parameters (41% in Experiment 1 and 60% in Experiment 2).
+First, it bears noting that the truncated response distribution implied by the Bayesian Sampler model appears at odds with the raw response data: participants' responses frequently lie outside the range implied by the best estimates of their $d$ parameters (41% in Experiment 1 and 60% in Experiment 2).
 
@@
 As Zhu and colleauges [-@zhu.etal2020] note, this is somewhat implausible on its face and their later work has abandoned this assumption [@zhu.etal2021].
-
-At the same time, from a pragmatic perspective it is highly desirable that all latent parameters within the models be continuous.
+At the same time, from a pragmatic perspective it is highly desirable that all latent parameters within the models be continuous rather than discrete.
 The models would be far more tractable to fit if, rather than including a latent Binomial variable representing the discrete number of samples drawn, we could instead model a continuous proportion of samples using, for instance, a Beta distribution.
 
-Zhu and collauges [-@zhu.etal2020] introduce the possibility of an "autocorrelated" Bayesian Sampler model under which samples are assumed to be autocorrelated [ideas which were advanced further in [@zhu.etal2021].
+Zhu and collauges [-@zhu.etal2020] introduce the possibility of an "autocorrelated" Bayesian Sampler model under which samples are assumed to be autocorrelated [ideas which were advanced further in @zhu.etal2021].
 As autocorrelated samples provide less information than i.i.d samples, they should be weighted when computing probability estimates.
-The idea is that people actually draw N autocorrelated samples that approximate some smaller $N_{\text{eff}}$ i.i.d samples.
-Assuming the true number of samples drawn, N, is allowed to vary somewhat noisily, then a model based on autocorrelated samples would no longer be limited to predicting a discrete set of possible responses.
-
-<< say a bit more on this assumption / revise to clarify a bit >>
+The idea is that people actually draw N autocorrelated samples that approximate some smaller effective number of i.i.d samples ($N_{\text{eff}}$).
+Assuming the actual number of autocorrelated samples drawn, $N$, is allowed to vary somewhat noisily, then a model based on autocorrelated samples would no longer be limited to predicting a discrete set of possible responses.
+To approximate this as part of a wholly continuous model, I model the proportions calculated from such a hypothetical autocorrelated sampling process using a Beta distribution.
 
@@
 This sort of rounding can be modeled by a categorical distribution across the discrete possible rounded responses.
-Each rounded response category corresponds to a set of cut points, $a$ and $b$, with the probability of the categorical response defined by the cumulative distribution function of the underlying latent distribution.
+Each rounded response category corresponds to a set of cut points, $a$ and $b$, with the probability of the categorical response defined by the cumulative distribution function of the underlying latent distribution ($B$).
 
@@
 Then, the probability of these categorical responses are computed for 21 and 11 categories (corresponding to rounding to 5% and 10%).
-These probabilities were combined along with a uniform probability representing "contaminants" according to mixing probabilities $\phi$, distributed with a Dirichlet prior (see Appendix for further implemenatation details).
+These probabilities were combined along with a uniform probability representing "contaminants" according to mixing probabilities $\phi$, distributed with a Dirichlet prior (see Appendix for further implementation details).
 
@@
 
-Above we found that an autocorrelated sampling process could be reasonably approximated by a Beta distribution. Taking the original Bayesian Sampler model,
+Approximating an autocorrelated sampling process using a Beta distribution and starting from the original Bayesian Sampler model,
 
@@
 However, it also seems reasonable to imagine that the number of samples drawn in fact varies from trial-to-trial.
-This added complexity is likely ignorable for a noise-based model, but could be important for the Bayesian Sampler model, as the number of samples drawn affects the truncation of the response distribution.
-That is, allowing the effective number of samples drawn to vary across trials may allow the model to capture some responses that would otherwise be treated as contaminants.
+For the Bayesian Sampler model, this could substantially impact the model's fit.
+As the number of samples drawn affects the truncation of the response distribution, this may allow the model to capture some responses that would otherwise be treated as contaminants.
 To capture this, the model can be given one additional extension to allow the number of samples to vary, by adding a new parameter $N_{trial}$.
@@
 Instead, model posteriors were estimated using Stochastic Variational Inference (SVI) with Numpyro [@phan.etal2019] using a multivariate Normal guide [e.g.
-@kucukelbir.etal2015] and 10,000 SVI steps.
-Estimating each model posterior took about 40 minutes on an Nvidia V100 GPU.
+@kucukelbir.etal2015].
+Estimating each model posterior took between approximately 20 to 90 minutes on an Nvidia V100 GPU.
 Simulation studies verified that this estimation approach could reliably recover parameters from simulated data (see Supplemental Materials).
@@
 
-The non-varying Bayesian Sampler model fits quite poorly and is estimated to have a large proportion of "contaminant" responses (34% in Experiment 1, 25% in Experiment 2).
+The non-varying Bayesian Sampler model fits quite poorly and is estimated to have a large proportion of "contaminant" responses (43% in Experiment 1, 26% in Experiment 2).
 This is likely due to a lack of the predicted truncation of the response distribution.
@@
 
-Allowing the effective number of samples drawn to vary trial-by-trial improves the fit of the Bayesian Sampler model substantially and decreases the estimated proportion of contaminant responses (28% in Experiment 1, 12% in Experiment 2).
-Nevertheless, these results still indicate inferior fit compared with the noise-based model, which has substantially lower $\widehat{\text{elpd}}_{\text{LOO}}$ and again attributes fewer responses to the contaminant process (14% in Experiment 1, 4% in Experiment 2).
+Allowing the effective number of samples drawn to vary trial-by-trial improves the fit of the Bayesian Sampler model substantially and somewhat decreases the estimated proportion of contaminant responses (30% in Experiment 1, 28% in Experiment 2).
+Nevertheless, these results still indicate inferior fit compared with the noise-based model, which has substantially lower $\widehat{\text{elpd}}_{\text{LOO}}$ and attributes far fewer responses to the contaminant process (11% in Experiment 1, 4% in Experiment 2).
 
@@
 But if these quantities are actually correlated as predicted by a Bayesian adjustment model, then we should expect to nevertheless see a correlation between the subject-level estimates in these parameters.
-Figure 11 shows a scatterplot relating these estimates.
-Looking across subjects we see they are mostly uncorrelated.
-Although there is a slight negative correlation in Experiment 1, this is driven by only a handful of participants with extreme values and in Experiment 2 the pattern appears to if anything be reversed.
+Figure 11 shows scatterplots relating these estimates.
+Although there is a slight negative correlation in Experiment 1, this is driven by only a handful of participants with extreme values.
+In Experiment 2 the correlation appears to if anything be positive rather than negative.
 These findings again run counter to the predictions of the Bayesian Sampler theory.
@@
 ```{r, fig.cap = "Scatterplots showing the relationships between subject-level d and N estimates from the trial-level noise-based model for Experiments 1 and 2.
-Figure for Experiment 1 excludes some outlier paricipants who gave repetitive resposnes that resulted in abnormally high N-values.
+Figure for Experiment 1 excludes some outlier paricipants who gave repetitive responses that resulted in abnormally high N-values.
 Error bars indicate 95% CI.
@@
 Comparing the $d$ values inferred from the trial-level model we see they are similar to but generally smaller than those estimated in the trial-averaged model.
-This is likely because the hypothesized initialization process accounts for some of the shrinkage in participants' responses, so that less need be attributed to sample read noise.
+It is unclear exactly why this is, though it could owe at least in part to the mixture component preventing "contaminant" responses from affecting the estimates of these parameters.
 
@@
 
-Directly fitting the models to these trial-level human response data presents several challenges addressed here by creating continuous extensions of these models and incorporating additional modeling of response processes.
-Though not identical to the models as originally proposed, this exercise allowed for tests of the key distributional predictions of the theories.
-Neither of the Bayesian Sampler's predictions appear to be borne out by the data.
+The distributions of participants' responses are more consistent with the PT+N theory's account of sampling noise than the Bayesian adjustment implied by the Bayesian Sampler theory as neither of the Bayesian Sampler's predictions appear to be borne out by the data.
 First, participants' responses frequently fall outside the truncated range implied by the parameters estimated under the Bayesian Sampler model.
@@
 
-Altogether then, the distributions of participants' responses are more consistent with the PT+N theory's account of sampling noise than the Bayesian adjustment implied by the Bayesian Sampler theory.
-In the end, I find the best overall account of participants' probability judgments is a continuous extension of the PT+N theory without any special treatment of conditional probability judgments.
+In the end, I find the best overall account of participants' probability judgments is a modification of the PT+N theory without its two-stage process of conditional probability judgments.
+Like the PT+N theory, this model accounts for distortions in probability judgments via a process of noisy sampling.
+But like the Bayesian Sampler theory, under this theory, conditional probability judgments are produced by a process of conditioning in the mental model of the events, rather than as part of the mental sampling process itself.
+
+Outside of probability judgments, Bayesian conditioning is a key aspect of the sorts of mental models imagined by Bayesian cognitive scientists, where cognitive models are conditioned on information as part of learning, prediction, and inference.
+Within this framework, it seems only natural to imagine that a similar conditioning process would subserve probability judgments, rather than conditional probability judgments being made by a distinct two-stage sampling process.
+
+To illustrate the distinction, consider being asked to judge the conditional probability that it will rain tomorrow in London given it rained today in London.
+Now, compare that with first being told that it rained today in London, and then being asked to make the simple probability judgment that it will rain tomorrow in London.
+The original PT+N theory would draw a distinction between the two tasks: The first task would invoke the two-stage sampling process, whereas presumably the second would involve some change in the mental model to reflect learning about the day's weather followed by only a single stage of sampling.
+In contrast, the present findings suggest these tasks would invoke the same set of mental processes---conditioning of the mental model followed by the drawing of samples from that model.
 
@@
 This model bounds $d$ at .50 in principle, but a sample-reading process with such high error-rates may or may not be plausible.
-In prior work simulations have often assumed values of $d$ around .10 [e.g.
+In prior work, simulations have often assumed values of $d$ around .10 [e.g.
 @howe.costello2020; @costello.watts2017].
@@
 For instance, a Beta(2,2) prior is only modestly informative, but could produce $d = .40$ if $N=1$.
-However, if more samples are drawn then high $d$ values would correspond to potentially inflexible and suboptimal priors (i.e.
-too large $\beta$ values).
+However, in cases where more samples are drawn, high $d$ values would correspond to potentially inflexible and suboptimal priors.
 From the results, it is clear there are some individuals with both relatively high $d$ and $N$ values, which may press somewhat against the rational justification for the desirability of sampling noise.
+For instance, some participants are inferred to have  $N \approx 20$ and $d \approx .30$, implying $\beta \approx 15$.
 

### 2022-12-14 paper-rmd/trial_level_mlm_noise_based.tex
updated rmd and avg code

@@
       \node[latent, above=1cm of y]  (theta) {$\vec{p}_{ijk}$};
-      \node[latent, left=1cm of theta]  (kappa) {$\kappa_j$};
+      \node[latent, left=1cm of theta]  (N) {$N_j$};
       \node[latent, above=of delta, xshift=-.5cm]  (delta_pop) {$\overline{\delta}$};
@@
       \edge {sigma_delta_delta} {delta_delta}
-      \edge {mu, phi, kappa} {theta}
+      \edge {mu, phi, N} {theta}
 
@@
         \tikzset{plate caption/.append style={above right=0pt of #1.north east}};
-        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)(kappa)} {$j$} ;
+        \plate[inner sep=0.25cm, xshift=-.12cm, yshift=-.12cm] {plateJ} {(mu)(y)(p)(d)(d_prime)(delta)(delta_delta)(N)} {$j$} ;
       }
@@
   y_{ijk} &\sim Categorical(\overrightarrow{\theta_{ijk}}) \\
-  \overrightarrow{p_{ijk}} &= Z_{NB}(\mu_{ijk},\kappa_j, \phi) \\
+  \overrightarrow{p_{ijk}} &= Z_{NB}(\mu_{ijk},N_j, \phi) \\
   \mu_{ijk} &= f_{NB}(\overrightarrow{\theta_{jk}}, d_j, d'_j, x_{ijk})  \\
@@
   \overrightarrow{\phi} &\sim Dirichlet_3(\overrightarrow{1}) \\
+  N_j &\sim Cauchy(20) \\
 \end{aligned}
